### 2017-09-14 paper/biblio/scale-out.tex
paper: add outline and keyspace graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\documentclass{article}
\usepackage{graphicx,color,setspace,hyperref}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{listings}
\lstset{language=C}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\(\checkmark\)}%
\newcommand{\xmark}{\ding{55}}%

\begin{document}

\title{Scale-out Bibliography}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Cloud Computing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cloud Computing}

\noindent\cite{zhang:journal2010-cloud-challenges} Challenges of cloud computing adn resource managements; Differences: SLOs and migration units; Similarities: map cloud resoruces to CPU and memory. \\

\noindent\cite{vilutis:ITI2012-cloud-load-balancing} Resource Provisioning - postpone tasking during peak traffic. \\

\noindent\cite{quiroz:grid2009-cloud-workload-provisioning} dynamic resource rescheduling in virtual machine environments - according to data center wide availability, load balancing, and constraints.\\

\noindent\cite{ranjan:cloud-computing2010-peer-to-peer} overview of current resource distribution mechanisms (CloudWatch, Eucalyptus, etc.).\\

\noindent\cite{gulati:hotcloud2011-cloud-resource-management} VMWare DRS, Challenges - discusses the modifiable ``knobs", challenges, and solutions in a cloud resource manager.\\

\noindent\cite{gulati:osdi2010-mclock} mClock: schedule IO for VMs with constraints and reservations. This is more difficult because storage is shared, locality is difficult, and because throughput varies. The solution is to logically interleave a constraint-based scheduler (overloaded) and a weight-based scheduler (priority).\\

\noindent\cite{xen-wlb} Xen WLB -  migrates VMs to balance load by maximizing performance or density; similare to SandPiper.\\

\noindent\cite{zaharia:hotcloud2011-datacenter-OS} Datacenter needs an OS - there are so many resources in a distributed setting that we need mechanisms for resource sharing (Mesos), data sharing (Spark), programming abstractions (RDDs, Mesos, MapReduce, Chubby), and debugging/monitoring (XTrace). \\

\section{Self-Tuning Systems}
\noindent\cite{padala:eurosys2009-autocontrol} AutoControl - online model estimator measures applications's resource usage and multi-in/output resource controller allocates resources.\\

\noindent\cite{wood:nsdi07-sandpiper} Sandpiper - migrates VMs to underloaded PMs; (physical resources \(\mapsto\) VMs)
Hotspot: if aggregate usage of any resource \(>\) thresh/SLA for ``sustained" period
\begin{itemize}
	\item profiler: tracks resources (CPU, network, and memory utilization)
	\begin{itemize}
		\item ``nucleus": server profiles (resource utilization over a sliding window)
		\begin{itemize}
			\item pr. distr. of the resource usage over W (i.e. usage variation; helps migration manager estimate peak resource requirements)
			\item time series of all observations in W (i.e. temporal fluctuations; helps hotspot detector to spot increasing utilization)
		\end{itemize}
	\end{itemize}
	\item hotspot detection: ``when to migrate VMs"
	\begin{itemize}
		\item trigger: (1) sustained utilization, (2) next predicted value\footnote{captures rising trends} \(>\) thresh
		\item grey box approaches can get more accurate peak utilization by reading application logs and perf counters (e.g., queueing theory for applications). 
	\end{itemize}
	\item migrator: ``what/where to migrate; how much to allocate after migration"
	\begin{itemize}
		\item orders available serves and assigns VMs, given the predicted utilization
		\item heuristic: minimize migration overhead (i.e. data transferred)
		\item volume = \(\Pi\frac{1}{1 - \text{rsrc}}\); ordered based on \(\frac{\text{vol}}{\text{size}}\) ratio to move the most load per byte
		\item match high VSR VM to low VSR PM; only move it CPU/network/memory can handle it
	\end{itemize}
\end{itemize}
Profiler
\begin{itemize}
	\item[\xmark] bottleneck at interval \(I\) for the control plane
	\item[\xmark] overhead of monitoring CPU?
\end{itemize}
Profiler
\begin{itemize}
	\item[\xmark] bytes per load is a good heuristic... what about resource per load?
\end{itemize}
Results: single server hotspots in \(< 20 s\); multi-server hotspots in minutes
\begin{itemize}
	\item load per byte is a good heuristic
	\item swaps have a lot of overhead but help mitigate hotspots
	\item responds to individual resources (switch network-bound w/ CPU-bound)
	\item black box can incorrectly move VMs
	\item microbenchmarks: overhead of computing placement (i.e. computing a new mapping of VMs to PMs is a NP-hard, so they approximate it with the VSR heuristic). 
\end{itemize}

\noindent\cite{} Cloud fault resilience - injected faults to different services at different points and observed (1) fault tolerance is not enabled by default, (2) timeouts are not supported, (3) it is difficult for layers to talk, and (4) external libraries can commingle.\\

%\noindent\cite{} IO Virtualization (waldspurger,rosenblum)- representative I/O system challenges. Conclusion is: virtual IO is a great tool with lotsa potential but it adds lotsa complexity, ambiguity, abstraction
%\begin{itemize}
%	\item benefits: (1) time/space multiplexing (less devices, higher utilization, device aggregation), (2) seamless migration/portability (across devices and OSSs), (3) flexibility (improve availability and balance load) and new features (transforming I/O requests, e.g., encryption)
%	\item challenges: performance (overhead traversing layers), quantifying ``nontrivial" penalties, how do we multiplex physical hardware across VMs of varying importance?, new algorithms to schedule requests fairly/efficiently (e.g., breaking TCP layer), security, customer expects good performance
%	\item approaches+: pluggable structure for back-end implementations, lower emulation cost (with traps, pass-thru modes), 
%\end{itemize}

%\noindent\cite{} Virtio: An I/O virtualization framework for Linux (tim jones) - interface for paravirtualized entities (i.e., modified guest OS and hypervisor) to communicate. \\
%Problem: full virtualization requires trap and emulate, which is inefficient\\
%Soln: virtio is an abstraction for common emulated devices in paravirtualizatyion, where you can implement block, network, pic, balloon and console drivers
%\begin{itemize}
%	\item QEMU: system emulator for the entire system
%	\item virtual queue interface: attach fe to be
%	\begin{itemize}
%		\item virtio device: configure the device
%		\item virtio queue: buffers for transmitting requests and data
%	\end{itemize}
%\end{itemize}

\noindent\cite{nadav:atc2011-viommu} vIOMMU: Efficient IOMMU emulation (*amit, yahuda, tasfiri, shcuster) - achieves direct device assignment for fully virtualized environments. \\
Motivation: we want DMA (access RAM independently of CPU) and direct device assignment for virtualized environments.\\
Problems: legacy DMA can't be used natively in virtualization, it lacks memory protection, and drivers cannot access entire memory; IOMMU is a piece of hardware that translates DMA addresses \(\rightarrow\) physical address\\
Problem (again): this has a number of shortcomings: (1) can't dynamically change RAM,  (2) no intra-guest protection between hypervisor and guest, (3) no redirection of DMA transactions
Soln: introduce a vIOMMU that multiplexes the IOMMU to remap pages in the hypervisor
\begin{enumerate}
	\item guest - IO device driver requests IO virtual address and updates mapping in PTEs and IOMMU TLB (which traps to IOMMU)
	\item IOMMU - reads emulated PTEs and translates guest to host address, then it pins the pages, and maps the physical pages, then it points to IO device driver
Advantages: finer control over memory pages (don't pin everything), intra-guest protection provided by IOMMU, inter-guest protection maintained
\end{enumerate}
Optimizations: VM exit and remapping is costly, so do side core emulation and save mappings optimistically.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MapReduce Tweaks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MapReduce Tweaks}

\noindent\cite{buck:hpc2011-scihadoop} SciHadoop - improves the performance of MapReduce for array-based data and queries. Extensions allow the user to specify more information about the data so that the system can partition and retrieve data more efficiently.\\

\noindent\cite{bu:vldb2010-haloop} HaLoop - accommodates iterative tasks by introducing a new API, loop-aware task scheduling, caching loop-invariant data, and fixpoint evaluation caching.\\

\noindent\cite{ekanayake:hpdc2010-twister} Twister - accomodates iterative tasks by introducing a distributed in-memory cache, it extends the programming model to support ``broadcast" and ``scatter" type data transfers.\\

\noindent\cite{peng:osdi2010-percolator} Percolator - accomodates incremental updates and random access by implementing a library for BigTable operators that use observers, a timestamp oracle, lightweight locks, and notifications to leverage the indexing we've done previously.\\

\noindent\cite{ekanayake:escience2008-eglmapreduce} CGL-MapReduce - a streaming-based implementation the lets teh runtime use map/reduce tasks multiple times and can stream intermediate data to other components. They also show that with more data, the overhead of a the runtime diminishes.\\

\noindent\cite{gkantsidis:nsdi2013-rhea} Rhea - filters unstructured data so that less data is passed out on the wire. Achieved by doing some compile-time sorcery and determining what information is important for the computation.\\

\noindent\cite{cho:socc2013-natjam} Natjam - implements priority based eviction policies so that late-arriving high priority jobs can evict other jobs; employs CPU techniques.\\

\noindent\cite{mihailescu:hotstorage2012-mixapart} MixApart - introduces a cache layer so that compute tasks can be scheduled to overlay with data ingest; useful for the MapReduce workload characterizations (i.e. data reuse, CPU-bound input, predictable IO demands, etc.).\\

\subsection{Query Processing}

\noindent\cite{buck:sc2013-scidr} SIDR -  intelligently partitions and routes intermediate data, allowing it to: remove Hadoop's global barrier and execute Reduce tasks prior to all Map tasks completing; minimize intermediate key skew; and produce early, correct results.\\

\noindent\cite{balmin:sigmod2012-clydesdale} Clydesdale - accomodates structured data processing using high dimension star schemes, column-oriented layouts to interact with HDFS, and query processing sharing data structures.\\

\noindent\cite{elmore:sigmod2013-pythia} Pythia - database migration by learning tenant features and resource consumption. Templates are constructed and rules learned about which perform well together.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Other Distributed Runtimes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Other Distributed Runtimes}

\noindent\cite{zaharia:nsdi2012-spark} Spark - accomodates iterative/interactive tasks by forcing the user to interact with RDDs (read-only, partitioned records) to manipulate and cache working-set data. The runtime manages coarse-grained lineage to control partitioning, ttrack lineage, schedule, compute on, and persist intermediate results.\\

\subsection{Query Processing}

\noindent\cite{engle:sigmod2012-shark} Shark - a query physical plan generator in Spark converts the Hive operator tree to operators that perform transformations on RDDs, which allows the runtime to optimize for	 inter/intra-quer locality.\\

\noindent\cite{melnik:vldb2010-dremel} Dremel - multi-level execution tree that stores relational data in a nested repreesentation, using a nested data model, server trees, and column-striped storage. This allows record-oriented storage, instead of record reconstruction, which can be run in parallel.\\

\section{Streaming}
\noindent\cite{rixner:2002-stream} Stream processing - introduces double buffering, shows the applicability for MapReduce.\\

\noindent\cite{neumeyer:icdmw2010-s4} S4 - creates a programing flow through the system, where processing elements (vertices) listen for events, forward to computation, and publish events; it processes stream data with a stream runtime that pipelines operations between computations\\

\noindent\cite{isard:EuroSys2007-dryad} Dryad - extended API for communciation DAG and per-vertex algorithms. The user iterativel manipulates the input and communication graphs to optimize perforamnce. It is more flexible than MapReduce because user can define data dependencies and many algorithms.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Performance/Benchmarking
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Performance}

\noindent\cite{fadika:cloud2012-evaluating-hadoop} Hadoop evaluation - ran HPC experiments on Hadoop and determined that tradeoffs for system configuration (file system, networking, replication, etc.) make a difference.\\

\noindent\cite{huang:icde2010-hibench} Hadoop Benchmark - set of Hadoop benchmarks (micro-benchmarks and real-world applications) to characterize speed, throughput, bandwidth, and system resources.\\

\noindent\cite{wisniewski:2007europar-commercial-scale-out} HPC Performance - scale-out performance is contingent on complex execution stack and many processes/threads per processing elem.\\

\noindent\cite{chen:2012ccpe-distributed-db} Distributed Databases - compares centralized, partitioned, and replicated databases.\\

\noindent\cite{chen:2011mascots-suites} Workload Suites - contends that benchmarks do not accurately simulate a real workload. They suggest sampling a trace to build a more realistic workload.\\

\noindent\cite{chen:2012vldb-cross-industry} Cross-Industry MR Study - in-depth study of MapReduce workloads. 90\% of all jobs are small; clusters are bursty and unpredictable and database functionalities are common.\\

\noindent\cite{wang:mascots2009-mrperf} MRPerf - modeling MapReduce using statistical analysis.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Distributed Systems vs. Single Node
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Distributed Systems vs. Single Node}

\noindent\cite{appuswamy:socc2013-hadoop-vs-single-node} Scale-up vs. Scale-out for Hadoop - they compare single-node and distributed Hadoop because it (1) provides backwards compatibility, (2) keeps the applications portable, and (3) allows an apples to apples comparison.\\

\noindent\cite{michael:2007pdps-scale-up-x-scale-out} scaling comparison -  scale-out has a cost (resource utilization) and performance (4\(\times\)) advantage. \\

\noindent\cite{gigaspaces:whitepaper2011-su-vs-so} scaling comparison - we need to consider other scale-out properties and calls for an automated comparison process.\\

\noindent\cite{talkington:2002journal-scaling}	 scale-up is limited by processor/memory speeds and shared resource contention. Scale-out is limited by intra-node communication and workload management.\\

\noindent\cite{rowstron:hotcdp2012-hadoop-vs-single-node} Nobody ever got fired... - examines AdPredictor as a way to measure memory costs, job input sizes, implementation complexity, and performance.\\

\noindent\cite{appuswamy:techreport2013-hadoop-vs-single-node} Nobody ever got fired... - makes the claim that MapReduce jobs are often better served by a scale-up server for performance, \$, power, and server density. \\

\noindent\cite{schwarzkopf:hotcloud2012-7-sins} 7 Deadly Sins - common simplifications/shortcuts in cloud research are ``clouding" our analysis: unnecessary distributed data parallelism  (not all jobs are parallel), assuming performance homogeneity (poor methodology/result presentation), picking the low-hanging fruit (speedup over general Hadoop is not useful), forcing the abstraction (workload not big enough), unrepresentative workloads (should be running many jobs at once, not batch), perfect elasticity (limits to scalability/parallelism), ignoring fault tolerance (there is a cost).\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% HPC
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent\cite{perez:jctc20150parsplice} ParSplice - molecular dynamics
simulator that uses a hybrid MD/AMD approach to get the best of both worlds:
fast simulation for uneventful persiods and exact simulations for rapid
transitions to new states. The novelty is speculation and segment splicing in
the AMD phase, and a model for determining when transitions occur.

%1. What is the problem the authors are trying to solve?
%
%Simulating atomic interactions in long-time dynamics with heterogenous
%barriers. Energy barriers are thresholds for when the system changes state;
%these transisitions have interesting events but the lull between causes
%problems for existing techniques.
%
%2. What other approaches or solutions existed at the time that this work was done?
%
%MD and AMD are two competing approaches to simulating atoms. MD is exact but
%only runs for short periods because the error explodes while AMD is highly
%parallel because it relaxes accuracy, instead opting for simulating
%state-to-state trajectory instead of per-molecule trajectories. ParRep is an
%AMD technique that uses statistics and parallelization by fuzzing the accuracy
%of simulations with similiar barriers (and thus longer state occupations).
%
%3. What was wrong with the other approaches or solutions?
%
%For running simulations with different barrier types, neither technique is
%sufficient. MD is too precise and the error nullifies the validity of the
%simulation. AMD does not offer enough accuracy for the interesting periods.`
%
%4. What is the authors' approach or solution?
%
%Using MD for state transitions and AMD for periods of innactivity.
%- inactivity modeled with segments, which are spliced from smaller segments
%- model for long term dynamics that pinpoints state transitions
%- meta-simulation to predict transitions to increase parallelism 
%
%Implementation
%- splicer: tries to make biggest segment
%  - coalesce existing segments, speculatively generate offset for P
%  - label states with connecitvity graphs to get a hash
%- producers: create segments given offset and act as a cache for requests
%  - use connective graphs to make DB rd/wr efficient
%
%5. Why is it better than the other approaches or solutions?
%
%ParSplice supports different accuracy (i.e. arbitrarily) for unstable states
%and performance can be optimioed if the metastable states/correlation times are
%known. It defeat the low barrier problem for long running dynamic simluations.
%
%6. How does it perform?
%
%Good?
%
%7. Why is this work important?
%
%Hybrid approach to MD and AMD that lets us use the resources of HPC via
%speculation, asynchornity, and fault model. The approach makes it feasible
%(correctness wise) and accurate enough to see what interesting events.
%
%8. 3+ comments/questions

\bibliographystyle{abbrv}
\bibliography{references}

\addcontentsline{toc}{section}{References}

\newpage
\onecolumn
\appendix
%Appendix A
%\section{1}

\end{document}

### 2017-09-14 paper/biblio/scale-up.tex
paper: add outline and keyspace graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\documentclass{article}
\usepackage{graphicx,color,setspace,hyperref}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\(\checkmark\)}%
\newcommand{\xmark}{\ding{55}}%

\begin{document}

\title{Scale-up Bibliography}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Supercomputers
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Super Computers}

\noindent\cite{schroeder:2006icdsn-HPC-failures} LANLs failures - looks at failure rates of a 9 year trace of LANL SMP/NUMA clusters. Determines that failures can be modeled with the Weibull curve.\\

\noindent\cite{bryant:2000lsc-linux-smp} Linux SMP - shows that Linux SMP support greatly improves between 2.2 and 2.4.\\

\noindent\cite{gara:bim2005-blue-gene} IBM BlueGene/Q - attempts to address the gap between supercomputers and dedicated systems with special links/networks, core/floating point enhancements, distributed memory, and fault tolerance techniques. \\

\noindent\cite{feldman:website2011-cray} CrayXK6 - using GPUs in supercomputers. \\

\noindent\cite{van-essen:DISCS2012-DI-MMAP} DI-MMAP - LKM for memory management \& page policies.  Provides out-of-core performance for data-intensive applications. Provides: 
\begin{itemize}
	\item fixed size page buffer
	\item minimal dynamic memory allocator
	\item FIFO buffer replacement
	\item preferential page caching 
\end{itemize}

\noindent Common themes at SC '13:
\begin{itemize}
	\item scaling: computation scales, I/O does not
	\item checkpointing: needs to be re-thought
	\item in-situ/workflows: we can't store our data
		\begin{itemize}
			\item streaming ({\it e.g.,  FastBit\footnote{process chunks; determine when/where to index; threshold for diminishing returns over network}}): code injection is encouraging
			\item rewrite code focused on data movement instead of arithmetic
			\item move from file data to stream data
		\end{itemize}
	\item metadata ({\it e.g., ADIOS}\footnote{abstract data at rest/in motion}): elastic, statistical
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Experimental OSs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Parallel Programming
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parallel Programming}

\noindent\cite{ranger:2007hpca-phoenix} Phoenix - converts MapReduce programs to multi-core/processor systems (including shared memory systems) by spawning parallel threads for each core.\\

\noindent\cite{yoo:2009iiswc-phoenix2} Phoenix 2 - optimized for \{large-scale NUMA, Linux x86\_64\}\\

\noindent\cite{talbot:2011mapreduce-phoenix++} Phoenix++ - optimized for \{modularity, extensibility, workloads\}\\

\noindent\cite{manferdelli:ieee2008-parallelism} many core parallelism - enumerates the challenges/solutions to many-core computing: (1) brick wall for serial performance = power wall + memory wall + ILP wall, (2) software must accommodate parallelism, and (3) OSs are limited by their isolation of programs, services, and devices; hypervisors must assign resources and manage environments.\\

\noindent\cite{gordon:osdi2012-comet} COMET - a runtime that lets parts of an unmodified parallel programs, spawned by a mobile device, run on a powerful server using DSM techniques and virtual machine synchronization. \\

\noindent\cite{liu:sosp2011-dthreads} DTHREADS - a library that ensures determinism between threads by adding a new API.\\

\noindent\cite{czechowski:ipdps2013-codesign} Co-design - a model for designing/reasoning about high level; where they try to relate parallelism with cores, structure of memory hierarchy, and network topology.\\

\noindent\cite{david:sosp2013-synch} Synchronization - to scale, synchronization should be limited to a single core; crossing sockets is a killer, message passing shines when contention is high, there is no dominant locking scheme.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Experimental OSs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Operating Systems}

\noindent\cite{scheupback:manycore2008-barrelfish} Barrelfish - proposes a new OS architecture which facilitates heterogenous hardware and application demands by maintaing a system knowledge base and making execution decisions at runtime.\\

\noindent\cite{wentzlaff:acm2009-fos} FOS - proposes a new OS architecture which distributes kernel system services spatially instead of time-sharing resources.\\

\noindent\cite{boyd-wickizer:osdi2008-corey} Corey - address multiprocessor scalability by giving applications control of shared kernel data structures. Introduces address ranges, kernel cores, and shares to facilitate user-space control via system calls (corey).\\

\noindent\cite{appavoo:acm2007-k42} To be read - K42 (prevent applications from contending for resources).\\

\noindent\cite{gamsa:osdi1999-tornado} To be read - Tornado (prevent applications from contending for resources).\\

\noindent\cite{bugnion:acm1997-disco} To be read - DISCO (aims to avoid kernel bottlenecks with a small kernel that minimizes shared data).\\

\noindent\cite{saha:eurosys2007-cmp}  To be read - McRT (Intel's manycore runtime; addresses the scalability of an application runtime, including some notion of inter- and intra- machine heterogeneity at the application level).\\

\noindent\cite{agarwal:dac2007-killmulticore} Cores per Chip - is an interesting article about the tradeoffs of scaling up (adding more resources to a core) and scaling out (adding more cores to a chip). Probably more useful because it predicts the size and scope of cores.\\

\noindent\cite{mao:eurosys2012-masstree} Masstree - is a key-value database store that fits into memory and is persistent. This may be a useful benchmark to run on the SAP machines to see if performance increases or to see what latencies are actually present.\\

\noindent\cite{song:eurosys2011-cerberus} Cerberus - based on the observation that commodity operating systems can't scale for a large number of CPU cores, this system introduces OS-clustering to provide a single OS image (via a SuperProcess) to the user while virtualizing resources with Xen (share address space, file system, file contents, NICs) via shared memory and a static resource allocation (i.e. 8 cores per OS).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Big Data Case Studies
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Big Data Case Studies}

\noindent\cite{jacobs:acm2009-big-data-article} Big Data - explains the big data problem and the limitations of hardware, database architectures (relational), and distributed computing when processing big data. \\

\noindent\cite{oreilly:2012-big-data-now} Big Data Now - describes the volume, variety, velocity, and veracity of big data. Also has a useful Apache distributed systems table. \\

\noindent\cite{zong:DISCS2012-satellite} EROS - processing and distribution of satellite images; discusses workload effects, challenges for the system (caching, memory, threads), and optimizations (prefetching, state tracking, migration). \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Performance, Benchmarking, Tracing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Performance, Benchmarking, Tracing}

\noindent\cite{boyd:osdi2010-scalability} OProfile - iteratively adds cores, identifies Linux bottlenecks @48 core threshold, fixes problems (macro/system)\\

\noindent\cite{sharma:lasci2005-ltt} LTT - looks at an individual system call; presents functions they call, how often they are called, how long they take (micro/system call)\\

\noindent\cite{abeni:rtas2002-linuxperf} Latencies - talks about latencies but does not say where they came from; measures given sleep time vs. observed sleep time; eliminates jitters and pre-emption.\\

\noindent\cite{pesterev:acm2010-cacheperf} Cache Tracing - looks at L1/L2 cache hits using OProfile + memecached.\\

\noindent\cite{shende:elw1999-profiling} Tracing vs. Profiling - explains the difference between profiling (summary statistics of performance metrics) and tracing (when and where an event occurred).\\

\noindent\cite{desnoyers:linuxsymp2008-hypervisor} Hypervisor Tracing - discusses hypervisor tracing (not that I need it). \\

\noindent\cite{desnoyers:LWNnet2012-lttng1} LTTng-2.0 1 - LTTng-2.0 features, advantages, and tradeoffs; kernel activity overview with tracepoints; dynamic probes\\

\noindent\cite{desnoyers:LWNnet2012-lttng2} LTTng-2.0 2 - LTTng-2.0 examples; user space and kernel tracing; performance counters and kreteprobes; LTTngTop.\\

\noindent\cite{kuz:usenix2011-benchmarks} Mixed workloads - why single applications (like MOSBENCH) are not accurate benchmarks for multi-core OSs.\\

\noindent\cite{attariyan:osdi2012-xray} X-ray - a runtime that identifies the root cause of slow performance (usually configuration or input). It has an online phase, which tracks performance data/system calls, and an offline phase, which replays code with binary instrumentation (using \textcolor{magenta}{\href{http://download-software.intel.com/sites/default/files/article/256675/cgo2013.pdf}{Pin}}) to determine which inputs/configurations caused slowdowns.\\

\noindent\cite{tan:na-gdbstub} gdbstub - explains the requirements, necessities, and internal workings (mainly the CPU responses) of gdbstub and a remote GDServer.  This should allow us to debug on GRUB by stepping through instructions as we look for an access violation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Memory Usage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Memory Usage}

\noindent\cite{wulf:sigarch1995-memory-wall} memory wall - memory transfer rate (speed) and onboard connections (physical space) limit the performance of the system as a whole; result is that system speeds were not doubling according to Moore's law anymore.\\

\noindent\cite{olschanowsky:ICPPW2010-PIR} PIR - tool for HPC application source code which finds/outputs memory references so a programmer can optimize. We can use this tool to find excessive memory references and figure out how the OS reacts to it. \\

\noindent\cite{luk:PLDI2005-pin} Pin - tool for adding \emph{instrumentation} (or extra code) into an application at run-time for optimization and transparency; allows us to see contents of registers, memory, and control flow.\\

\noindent\cite{movall:atec2005-physical} Physical Memory Visualization - tool suite for analyzing what the physical memory looks like (ex: allocations to processes, free space, cpu, page allocations, mapping to virtual memory, etc.).\\

\noindent\cite{corbet:LWNnet2012-NUMA} Sched/numa vs. AutoNUMA - compares Zijlstra and Arcangeli's approach to NUMA optimization (process migration, page migration, process decay, etc.)\\

\noindent\cite{novakovic:asplos2014-soNUMA} Scale-out NUMA - circumvents layers (OS, Network, I/O Bus) to provide fast RDMA. Uses (1) stateless request/reply protocol over NUMA interconnects and (2) adds RMC controller in node's local coherency hierarchy to bypass PCIe. \\

\noindent\cite{} Positional Effects of RAM - fault rate declines over time, vendor matters, SRAM faults are dominated by transient faults

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Fault Tolerance
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fault Tolerance}
\noindent\cite{swift:sosp2003-nook, swift:osdi2004-shadow-drivers} memory management hardware - encapsulate DD in protected domain (a.k.a ``hard-wall"). When faults occur, enable address-space isolation around the driver~\cite{swift:sosp2003-nook} and microreboot the DD~\cite{swift:osdi2004-shadow-drivers} to stop errant code from writing to the kernel. \\

\noindent\cite{zhou:osdi2006-safedrive} SafeDrive - compile assertion code into DD that checks segfaults and restarts driver\\

\noindent\cite{david:osdi2008-curios} microkernel techniques - create a new protection domain to house session state.\\

\noindent\cite{shye:DSN2007-plr} redundant processes - per application process, compare output to ensure correct execution. 16.9\% overhead for single threaded applications.\\

\noindent\cite{ansel:ipdps2009-dmtcp} DMTCP - program that checkpoints/restarts an application execution context to/from disk.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 2/4/2014
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Lampsons paper: goals for system design and how
	- many STEADY (simplicity, concepts are against each other
	- how: approximate, divide and conquer, paralleliize
	(1) simple, (2) abstract with interface, (3) write a specification
A scalable conflict free replicated data set (distributed eventually consistent set)
	- operations are commutative
	- change from a state operation propagation to a delta propagating operation

\bibliographystyle{abbrv}
\bibliography{references}

\addcontentsline{toc}{section}{References}

\newpage
\onecolumn
\appendix
%Appendix A
%\section{1}

\end{document}

### 2017-09-14 paper/biblio/security.tex
paper: add outline and keyspace graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\documentclass{article}
\usepackage{graphicx,color,setspace,hyperref}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}

\begin{document}

\title{Security}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Motivation
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}

\noindent\cite{garfinkel:NDSSS2003-VM-introspection} VM IDSs - the paper provides good motivation for the use of virtual machine IDSs and their abilities contribute to security. They prove that deviations from the model and reality can allow evasion.\\

\noindent\cite{laureano:euromicro2004-idsInVMs} is the pioneering work and motivation for virtual machine IDSs. Discusses VMs, hypervisors, \& introspection (monitor \& respond).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Introspection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introspection}

\noindent\cite{dolan-gavitt:ieeesp2011-virtuoso} Virtuoso - a system that edits QEMU assembly translations, outputs trace logs, analyzes/consolidates the trace logs, \& creates a host python simulation (including IO).\\

\noindent\cite{bahram:IEEErds2012-DKSM} DKSM - a LKM that changes in-kernel data structures (using QEMU + IL) to thwart introspection. \\

\noindent\cite{payne:acsac2007-secure-flexible-vm-monitoring} XenAccess -  provides the foundation introspection tools viw virtual {memory introspection, virtual disk monitoring}.\\

\noindent\cite{fu:sp2012-space-traveling} VMST - extension to Virtuoso the *automatically* bridges the semantic gap to generate VMI tools - (user doesn't specify tools); system wide instruction monitoring to redirect important kernel data. \\

\noindent\cite{dolan-gavitt:techreport2011-leveraging-forensic-introspection} FMA tools can be applied to live VMI-introspection. \\

\noindent\cite{toldinas:kaunas2009-VM-rootkits} memory snapshots - viability and drawbacks. They varied VM memory size, and, in addition to other affects, concluded that VM memory size is correlated to the root kit detection time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Alternatives to Introspection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternate to introspection}

\noindent\cite{vasisht:micro2008-SHARK} SHARK - proposes micro architectural support for a synergistic hardware/software system to detect root kits - it completely distrusts the OS.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Alternatives to Introspection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Class Readings}

\noindent\cite{gentry:acm2010-homomorphic} talks about homomorphic encryption schemes which may be helpful for interactions with the cloud.\\

\noindent\cite{rivers:acm1978-onewayf} the fundamental basis for encryption - a function that is easy to compute but difficult to reverse.\\

\noindent\cite{chen:sp2010-side-channels} side channels - how user clicks or keystrokes can reveal information.\\

\noindent\cite{dingledine:ssym2004-tor} Tor - system that pads data and encrypts messages with relays.\\

\bibliographystyle{abbrv}
\bibliography{references}

\addcontentsline{toc}{section}{References}

\newpage
\onecolumn
\appendix
%Appendix A
%\section{1}

\end{document}

### 2017-09-14 paper/biblio/storage.tex
paper: add outline and keyspace graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\documentclass{article}
\usepackage{graphicx,color,setspace,hyperref}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\centcmark}{\centering\checkmark}

\lstset{language=C}

\begin{document}

\title{ Storage Bibliography }

\maketitle

%NOAH NODB: instant gateway to data
%- get rid of the load time (either before or within queries)
%1. selective tokenizing: doesn't need to parse more than column 4
%2. selective parsing: don't need column 4, don't convert
%3. selective tuple formation: don't allocate if you don't need it
%4. grow index over where the data is in the file (into byte stream)
%5. caching data that I parsed
%
%Dimitris: channels
%how do we use flash in a cloud distributed
%- use flash and hods
%
%Ivo: AscotDB
%- byte wrappers around scidb
%- change the logical coordinates of scientific data
%
%Carlos: Making the Web Faster with HTTP 2.0
%- extensibility in storage by thinking of RESTful architectures
%- put a stream layer between http and text
%- fix multiplexing everything in a  TCP connection with streams
%- content headers (1) streaming (2) multiplexing in a  stream
%
%Adam:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% File Systems
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{File Systems}

\noindent\cite{mckusick:acm1984-FFS} FFS - establishes variable sized blocks for optimal storage, layout awareness to exploit locality of r eference, and a new way to organize data.\\

\noindent\cite{mckusick:fast2015-FFS} History of BSD Filesystem - decoupling policy from mechanism in the block allocator. \\

\noindent\cite{hitz:wtec1994-WAFL} WAFL - focuses on write requests (buffered, parallel, anywhere) instead of read requests so that we can change the layout the data and snapshots.\\

\noindent\cite{lanyue:fast2013-FS-evolution} FS evolution - studies where most of the bugs come from, what they mean, where most of the development time is spent, and what affects performance; useful for the pointers to these FSs.\\

\noindent\cite{rosenblum:acm1992-LFS} LFS - writes will be bottleneck and we have larger caches/faster CPUs; (1) buffer file changes, (2), i-node map, (3) segment cleaning.\\

\noindent\cite{rumble:fast2014-lfs-for-dram} LFS for RAMCloud - storage workloads incur 2X the memory space, so use LFS techniques for memory allocation - 2 level incremental cleaning solution with concurrent, minimal memory copy, no global cleaning.\\

\noindent\cite{aghayev:inflow2013-lfs-for-android} LFS for Android - HTTP browser cache incurs many random writes, which perform poorly on flash; they trade hit rate for performance. Implementation: allocate a large file, divide into segments, evict many entries \(\in\) segment; avoids segment cleaning because there is no copying involved.\\

\noindent\cite{kang:fast2014-multilanes} MultiLanes - ext4, xfs, btrfs are poor  file systems for the host/hypervisor when housing VMs, so deliver IO requests to driver directly.\\

\noindent\cite{he:fast2015-chopper} Chopper - discovers high latency operations using Latin Hypercube Sampling. They expose problematic block allocation algorithms in ext4 for corner cases and fix them. \\

\noindent\cite{wildani:systor2011-block-locality} Working sets in block IO streams - finds temporal and spatial locality for blocks by (1) calculating offset differences biased with time and (2) partitioning the accesses into distinct sets; but this is for data and does not reflext what the application is doing. \\

\noindent\cite{nightingale:sosp2005-speculator} Speculator - overlaps computation by executing with stale data; when the asynchronous reply returns, computation may be rolled back if the data is invalid.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RAID
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RAID}

\noindent\cite{stodolsky:techreport1994-redundant} RAID Overview - shows that different RAID levels utilize a variety of striping, mirroring, and parity techniques to store data. \\

\noindent\cite{stodolsky:comparch1993-paritylogging} Small Writes - describes the small-write problem - four accesses are needed for each small-write instead of two.\\

\noindent\cite{wacha:venue2010-RAID4s} RAID4S - adds an SSD as the parity drive in RAID4 schemes in order to increase the throughput of small writes.\\

\noindent\cite{mao:ipsps2010-HPDA} HPDA - proposes the HPDA solution, which uses SSDs as the data disks and two hard disks drives for the parity drive and a RAID1-style write buffer.\\

\noindent\cite{hitz:wtec1994-WAFL} WAFL - to get more space/performance via the network, they change  writes: (1) buffered, (2) performed in parallel, and (3) placed anywhere. This work shows that writes are slower than reads, that disk layout matters (tree of files, metadata + file), and that snapshots cans be fast/efficient.\\

\noindent\cite{im:msst2010-delayedparity} Delayed Parity - proposes a delayed parity techniques to log the parity updates, instead of the small-writes, on a partial parity cache.\\
f
\noindent\cite{park:cit2009-ssdRAID} Hetero SSDs - presents heterogenous SSD based RAID4, which injects SSDs into a RAID4 setup by replacing all the data drives with SSDs.\\

\noindent\cite{chen:acm1994-RAID} Parity Overview - shows that RAID schemes based on parity enhance storage reliability by managing a recovery parity disk, but parity calculations degrade performance.\\

\noindent\cite{XDD} XDD - users guide.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Distributed Storage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Distributed Object Stores}
\noindent\cite{weil:osdi2006-ceph} Ceph - distributed file system decouples metadata/data, dynamically distributes metadata, and leverages OSDs so they can do most of the system maintenance on their own.\\

\noindent\cite{beaver:osdi2010-haystack} Haystack - custom object store that avoids metadata operations by keeping logical \(\mapsto\) physical mappings in a Directory and volume maps in memory on the Store devices. Leverages the fact that pictures/videos are large, immutable objects that don't ``seek" functionality.\\

\noindent\cite{muralidhar:osdi2014-f4} f4 - erasure coded storage system for warm data that requires less throughput. They start by demonstrating that temperature zones (unchanging content with a low request rate) do exist.  Pushing computation to the storage nodes not always a good idea, as sometimes you want them to just serve data.\\

\noindent\cite{toor:nas2012-swift} Swift - looked at the applicability of Swift for CERN experiments; non-filesystem work shows that scalability is of the data/metadata is good, when using a database.\\

\noindent\cite{thusoo:sigmod2010-facebook-infrastructure} Facebook - multiple compute silos; limitations in the namenode (\# of files, memory) have them looking to distribute data across many clusters. Workloads invoke many small files, which have inspired concatenation and compression. \\

\noindent\cite{nightingale:osdi2012-fds} FDS - locality-oblivious blog store that exposes a flat namespace.
\begin{itemize}
	\item blobs, which are broken into tracts (R/W granularity), are stored on disks managed by tractservers (i.e. OSD). Tracts are striped across the system.
	\item MDS stores/distributes maps of hardware 
	\item each client can do a map-based table lookup by hashing the blobID and track number \(\rightarrow\) one hop everywhere \(\rightarrow\) equal I/O bandwidth everywhere \(\rightarrow\) no need for data locality
	\item table has \(n^2\) rows to store every pair of disks \(\rightarrow\) reduce TLT size \(\rightarrow\) push work to OSD
	\item \(\rightarrow\) flat \(\rightarrow\) no namespace \(\rightarrow\)  hash objects \(\rightarrow\) fast lookups \(\rightarrow\)  metadata with data
	\item exposes full disk bandwidth to all processors \(\rightarrow\) (1) recovery is fast (total disk contents on \(n\) disks) (2) not restrict to data locality programming models, e.g., MapReduce so no stragglers because you can dynamically assign work, and (3) hardware heterogeneity because I/O and compute can be purchased/managed separately.
	\item evenly distribute replicas (disk has replicas on \(n\) disks) \(\rightarrow\) use bandwidth of entire cluster to recover
	\item recovery: all disks participate in recovery by striping blob tracts among all disks in the cluster.
	\item results: \(\frac{1}{2}\) bandwidth of local disk, peaks at cluster bandwidth
	\item[\xmark] no file system locality
	\item[\xmark] no stripe blobs
	\item[\xmark] why always comparing to GFS?
	\item[\xmark] every node has network bandwidth equal to disk bandwidth, they each have as much network bandwidth as they need I/O bandwidth		
\end{itemize}

\noindent\cite{mickens:ndsi2014-blizzard} Blizzard - cloud-scale block store (1200 MB/s) that uses FDS to avoid performance tiers. Reduces write barrier and supports small random IO with:
\begin{itemize}
	\item nested striping: fine-grain virtual disk partitioning; avoids IOP convoy dilation (i.e. increased inter-request latency due to network jitter) and increases disk parallelism. 
	\item asynchronous flushes: immediately acknowledged while data flushed in background - lengthens vulnerability period but allows higher write performance. 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Distributed Key-Value Stores
\section{Distributed Key-Value Stores}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent\cite{chang:osdi2006-bigtable} BigTable - distributed key-value store tailored for performance that accepts variable data and multiple data attributes (infinite columns). Similar rows are grouped into tablets (put on servers), similar columns are grouped into locality groups (put into memory), and everything is managed by a Master/Chubby.\\

\noindent\cite{decandia:sosp2007-dynamo} Dynamo - distributed ``eventually consistent" key-value store tailored for reliability/availability that uses consistent hashing and quorums (allow the programmer to control R, W, N replication).\\

\noindent\cite{} Megastore - ACID compliance using commit logs and PAXOS to agree on a timestamp for a transaction.\\

\noindent\cite{} Spanner - SQL query language, global consistency/transactions, better write performance, lock free read-only transactions.\\

\noindent\cite{} F1 - SQL semantics with noSQL scalability. Takes Spanner's global transactions, consistency, and transparent sharding and adds distributed SQL queries, transactionally consistent secondary indexes, asynchronous schema changes including database reorganizations, optimistic transactions, and automatic change history recording and publishing.\\

\noindent\cite{greenberg:hotstorage2015-mdhim} MDHIM - multi-D hierarchical
indexing middleware: key-value store designed for HPC and multi-dimensional
data.

\begin{itemize}
  \item mechanisms: cursor types, bulk operations
  \item policies: partitioning, server allocation, secondary indices
  \item hpc tailored: infiniband, dynamic servers, pluggable DB, library
\end{itemize}

%Problem: cloud key-value stores don't perform well for tightly coupled and
%parallel supercomputers and large data sets; HPC has simulation data,
%structured (indexed in multiple dimensions), PB scale, infiniband/RDMA,
%resource manager (no deamones)
%
%Existing: cloud kv-stores are diverse/unstructured, usually partitioned
%w/consistent hash
%
%Motivation: PLFS can materialize reads on compute cluster; leverage big data
%capabilities within an HPC environment, e.g., coordinates = key enhances
%locality, access rate, semantically meaningfulG
%
%Solution: MDHIM 
%  - HPC: infiniband, dynamic servers, pluggable DB, leverage resources, library
%  - partitioned by range (not hash): # of spawned servers affects performance
%  - cursor type operations and bulk ops: off-load query decision making to clients
%  - multiple dimensions: users create secondary index instances
%
%Why does PLFS need a file system back end in the first place?  How does ingest
%work if they are spawned at job start?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Metadata Management
\section{Metadata Management}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent namespace = hierarchy, record changes to files or properties\\

\subsection{Single MDS}
\noindent\cite{ghemawat:sosp2003-gfs} GFS - single node authority (Master) manages metadata
\begin{itemize}
	\item metadata = namespace, ACLs, files \(\mapsto\) chunks, chunk \(\mapsto\) chunk server
	\item management = chunk lease, garbage collection, chunk migration
	\item facilities many failures, huge files, and sequential access using a metadata server and chunk servers for housing large byte sequences, which reduce metadata/network overhead and reliance on master); first to separate data/metadata I/O.
\end{itemize}

\noindent\cite{mckusick:acm2010-gfs-evolution} GFS metadata server is overwhelmed.\\

\noindent\cite{shvachko:login2012-hdfs-scalability} HDFS - single node authority (Namenode) manages metadata
\begin{itemize}
	\item metadata = namespace, access control, files \(\mapsto\) blocks, block \(\mapsto\) data node
	\item management = replication based on heartbeats/blockreports, rebalancing not implemented
	\item scalability - good overview and explains why the system is not scalable (bad for writes because of journaling). Namenode holds file system index and file system name node.
\end{itemize}

\subsection{Hashing/Table-based Mapping}
\noindent\cite{schmuck:fast2002-gpfs} GPFS - hash filename to find disk, metanode updates metadata
\begin{itemize}
	\item metadata = files \(\mapsto\) blocks, size, ACLs
	\item token manager = determines metanode (synchronize access to metadata)
	\item distributed locking for data, synchronization for metadata; clients only write to parts of a file
\end{itemize}

\noindent\cite{brandt:mss2003-lh} Lazy Hybrid - hashes pathname to locate metadata but maintains permissions and file system nameespace using tables (lookup MDS, access control), invalidation (broadcast renames/removes), lazy update/migration (add servers, delete files), and consistency semantics. \\

\noindent\cite{hildebrand:msst2005-pnfs} PVFSv2/pNFS - hash filename to find server, which holds a Berkeley DB
\begin{itemize}
	\item NFSv4 binds one network endpoint to all files
	\item allows parallel access by passing layout from storage system to clients
\end{itemize}

\noindent\cite{li:msst2006-dynamic} Dynamic Hashing - adds a metadata cache, directory hashing, and dynamic formulas. Formulas account for a forgetting factor, access information, and the number of MDSs leaving to account for load balancing and elastic clusters.\\

\noindent\cite{xing:sc2009-skyfs} skyFS - hashes filename to find server
\begin{itemize}
	\item (1) directory fragmentation for large directories, (2) parallelize within directories, (3) cache metadata to improve locality, (4) consistent hashing minimizes the amount of data that needs to be moved when servers are added/deleted (servers contact masters if overloaded, migrate tree partition with a hash)
\end{itemize}

\noindent\cite{zhu:pds2008-hba} Hierarchical Bloom Filter Arrays - distributed randomly, lookup location in replicated table
\begin{itemize}
	\item metadata: namespace, access control, files \(\mapsto\) blocks, block \(\mapsto\) node
	\item management: synchronize metadata tables
	\item insight: temporal locality is small
	\item 1st level gives accurate hot data, 2nd level gives inaccurate all data
\end{itemize}

\noindent\cite{sinnamohideen:atc2010-ursa} Ursa Minor - single node authority (root metadata server) manages metadata location; metadata distributed by enumerating according to namespace; multiple servers manage metadata requests
\begin{itemize}
	\item metadata = size, link count, attributes, permissions, files \(\mapsto\) nodes
	\item metadata servers access metadata objects in lower storage system; namespace servers manage directories
	\item good locality for multi-object operations, transactions
\end{itemize}

\noindent\cite{patil:pdsw2007-giga+} GIGA+ - incremental growth and minimal shared state is motivation for 2-level metadata and inconsistent metadata.\\

\noindent\cite{patil:fast2011-giga+} GIGA+ - splits directories when the entries reach a threshold by exercising lazy synchronization and redirecting queries to servers with histories of the partition splits.\\

\noindent\cite{ren:atc2013-tablefs} TableFS - stores small metadata (directories, inodes) and files in a LSM database so that they are written to the local FS as large objects (write-ahead logs, SSTables, large files). \\

\noindent\cite{ren:sc2014-indexfs} IndexFS - hashes directories, moves whole directories to different nodes, clients cache paths and permissions, logs stored in LSM for fast insertion and lookup.

	\begin{enumerate}
		\item dyn.nspace partition: migrate dirfrags (round-robin)
		\item stateless dir caching: MDSs pin pathnames/permissions
		\begin{itemize}
			\item clients cache server names, block until lease expires
		\end{itemize}
		\item LSM tree for metadata \(\rightarrow\) sort, index, write-ahead log
		\begin{itemize}
			\item TableFS: embed inodes/small files, column-style table 
		\end{itemize}
		\item Metadata bulk insert: clients cache all creates, lock dir
	\end{enumerate}
	CephFS Advantage
	\begin{itemize}
		\item production-level file system
		\item integrated into RADOS and built from the ground up
		\item monitoring or reporting system
		\item more of a general purpose file system 
	\end{itemize}

\noindent\cite{zheng:pdsw2014-batchfs} BatchFS (non-POSIX) - push computation to the clients (``lazy namespace synchronization": local operations merge with global namespace image) to avoid MDS synchronization problems ({\it e.g.,} directory lock contention, serialized transactions, and RPCs).\\

\noindent\cite{thomson:fast2015-calvinfs} CalvinFS - hashes filename to locate server with metadata. Handles many small files and fully linearizable random writes using the feature rich Calvin database, which has support for WAN/LAN replication, OLLP for mid-commit commits, and a sophisticated logging subsystem. The authors opt for balanced metadata distribution by hashing the filename to identify the metadata server, but admit to losing locality for creating files, deleting files, and changing permissions.
\subsection{Subtree Partitioning}

\noindent\cite{welch:fast2008-panasas} Panasas - each MDS manages a subtree bounded by the volume that it is introduced in. Each directory and its files within must be managed by the same MDS, which allows separate fault domains (i.e. no random failures and outages). There is no directory fragmenting or dynamic balancing.\\

\noindent\cite{weil:sc2004-dyn-metadata} Dynamic Subtree Partitioning (DynSP) - describes current approaches to managing metadata (subtree partitioning/hashing) and motivates the need for dynamic metadata (authority/collaborative caching, hierarchies, load balancing, traffic control, directory locality, storage). \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Programmable Storage Systems
\section{Programmable Storage Systems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%s%%

\noindent\cite{grawinkel:pdsw2012-lua} Scripted pNFS Layouts - application matches file type to layout algorithm, such as placement/striping, using Lua.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Metadata Studies
\section{Metadata Studies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%s%%
\noindent\cite{roselli:atec2000-FS-workloads} FS Workloads - as much as 50\% of all workloads is composed of metadata operations (this is important). \\

\noindent\cite{alam:pdsw2011-metadata-scaling} Metadata Wall - theoretical ops/second are not realized for the metadata on Lustre/GPFS; bottlenecks are the simultaneous creates and fast creates/deletes, so performance cannot be improved by deploying a large volume of hardware.\\

\noindent\cite{konstantinos:pdsw2014-lustre-metadata} Lustre's MDS - performance decreases with more sockets because of the thread limitations in the Linux kernel. Also, the back end is not a bottleneck.\\

\noindent\cite{abad:techreport2012-fstrace, abad:ucc2012-mimesis} Mimesis - limited metadata traces and workload models motivates work in statically modeling workloads; useful for the limitations of metadata benchmarks. \\

\noindent\cite{leung:atc2008-nfs-trace} - surveys traces from papers; discovers that traces are old and not repressentative of true metadata workloads.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Accessing Metadata
\section{Accessing Metadata}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent\cite{hua:sc2009-smartstore} SmartStore - applies correlation to metadata to make search fast and responsive. \\

\noindent\cite{leung:fast2009-spyglass} Spyglass - makes file system metadata search faster by introducing a smart crawler that extracts metadata and an index that stores/serves metadata.\\

\noindent\cite{cipar:eurosys2012-lazybase} LazyBase - pipelines metadata updates through database transformations allowing arbitrary injection to satisfy different amounts of freshness and performance.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Dynamic Load Balancing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dynamic Load Balancing}

\noindent\cite{anderson:fast2001-hippodrome} Hippodrome - iterative design tool that tries to provision enough storage nodes to achieve optimal throughput, balance, and utilization by analyzing the workload, modeling the performance, finding a solution, and migrating data. They show that the solution converges and allocates a reasonable amount of resources.\\

\noindent\cite{brandt:rtss2003-rad} RAD/RBED - real-time scheduling model that separates resource allocation from timing of delivery so that we can service different real-time and non-real-time-scheduling classes; RBED is a scheduler that assigns a target rate of progress and period to each process. Rates can dynamically adjusts for processes not making enough progress or processes coming into and out of the system. \\

\noindent\cite{povzer:eurosys2008-fahrrad} Fahrrad - use reservations for utilization and period (based off of RAD/RBED with the limitation that I/O cannot be preempted); achieves better hard guarantees while maintaining performance best-efforts I/O schedulers. \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Request-Type Locality
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dynamic Load Balancing}
\noindent\cite{pai:asplos1998-lard} LARD - algorithm for a proxy to distribute requests that emphasizes locality by sending requests to back-end nodes based on their content and the node's utilization (connections). \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Fault Tolerance
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fault Tolerance}

\noindent\cite{sundararaman:fast2010-membrane} Membrane - on file system crash, resume execution via logging and snapshots. \\

\noindent\cite{swift:sosp2003-nook, swift:osdi2004-shadow-drivers} memory management hardware - encapsulate DD in protected domain (a.k.a ``hard-wall"). When faults occur, enable address-space isolation around the driver~\cite{swift:sosp2003-nook} and microreboot the DD~\cite{swift:osdi2004-shadow-drivers} to stop errant code from writing to the kernel. \\

\noindent\cite{zhou:osdi2006-safedrive} SafeDrive - compile assertion code into DD that checks segfaults and restarts driver.\\

\noindent\cite{david:osdi2008-curios} microkernel techniques - create a new protection domain to house session state.\\

\noindent\cite{lu:osdi2014-icefs} IceFS - disentangles file system into cubes to provide localized reaction to faults, fast recovery, and concurrent updates. Logical entities physically separated to provide localization, specialization, and isolation. Achieved with no shared physical resources ({\it e.g.}, blocks), access dependencies ({\it e.g.}, metadata path traversal) and bundled transactions ({\it e.g.}, group updates into a global transaction).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Archival
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Archival}
\noindent\cite{adams:techreport2012-archival} Scientific/Historical Repositories - Showed that tertiary data is stored on disk, is highly dynamic, and subject to crawls.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Programmable Storage Systems
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Programmable Storage Systems}

\noindent\cite{spahn:osdi2014-pebbles} Pebbles: data management tools built on the notion of objects (both at the storage system and application level). \\
Problem: modern OSs lack the appropriate abstractions to manage data (only acts at file level; as a result, applications must manage data. Applications don't do well managing data because objects are mapped to files in the OS, which is application-specific, complex, and too coarse-grained. As a result, many applications don't do what they say the will (e.g., delete an object completely). It would be nice if we could teach the OS how to manage data.
\begin{enumerate}
	\item Pebbles Android: modified SQLite, XML, FS to use TaintDroid to track data flow through the system; feeds access data to...
	\item Pebbles Object Manager: counstruct uni-directional relationships based on access
\end{enumerate}
Advantages: users mark objects to audit in Pebbles application, then use regular application to delete/modify object
\begin{itemize}
	\item encryption/access/object manager improves flexibility/manageability of objects across software stacks
	\item no application user-level objects by leveraging the structure in databases
	\item automatically discovers objects by leveraging database constructs
\end{itemize}
Conclusion: finding these objects is possible and these abstractions can provide new functionality for objects.\\

\noindent\cite{seshadri:osdi2014-willow} Willow: expose APIs for both the application and the SSD-side CPUs.\\
Motivation: SSDs already have software-defined storage. This is analogous to GPUs; initially they had rigid interface, but inside there was hidden, programmable firmware. As they become more flexible, the began exposing partial programmability (shaders) and then full blown programmability (Tesla).
\begin{itemize}
	\item SSDs had a fixed interface (SATA/NVMe), but inside they have multi-core processors, firmware
	\item Goals: data dependent accesses, semantic extension, offload privilege execution
\end{itemize}
Architecture: custom code on the host and SPUs, designed as SSDApps by programmer
\begin{itemize}
	\item system installs/deploys code that (1) runs on SPU, (2) runs in Application, (3) runs in kernel
	\item system uses RPCs to communicate and do access control with process ID taint that flows through the kernel
\end{itemize}
Example application: write-ahead logging system based on an editable atomic write (log writes in SSD Willow application, metadata file points to log file and data file; data flushed on commit)
\begin{itemize}
	\item re-implemented Verilog application that uses a sophisticated data structure over direct-IO
	\item advantage: 50\% improvement in performance because we can simplify write-ahead logging scheme and push it to SSD
\end{itemize}
Advantage: improve performance with kernel-bypass, data dependent access, and SSD-side permission checks. Improve flexibility with semantic extension.\\

\noindent\cite{jenkins:ipdsw17-mochi} Mochi - memoization (computation caching) as a service used multiscale physics (CoEVP).

\bibliographystyle{abbrv}
\bibliography{references}

\addcontentsline{toc}{section}{References}

\newpage
\onecolumn
\appendix
%Appendix A
%\section{1}

\end{document}

### 2017-09-15 experiments/keyspace-locality/viz/helpers.py
paper: fail at machine learning and edge detection
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,81 @@
+import warnings; warnings.filterwarnings('ignore')
+import matplotlib, tarfile, shutil, os, glob
+import matplotlib.pyplot as plt
+import matplotlib.patches as patches
+import pandas as pd
+import numpy as np
+
+font = {'size'   : 12}
+
+matplotlib.rc('font', **font)
+
+names = {"0": "Splicer", "1": "Worker", "2": "PersistentDB",
+         "3": "InMemoryDB", "4": "WorkManager", -1: "all ranks"}
+
+def untar(fname):
+    try: shutil.rmtree("tmp")
+    except: pass
+    os.mkdir("tmp")
+    tar = tarfile.open(fname)
+    tar.extractall()
+    tar.close()
+d = "../results/"
+def plot_times(ax, label, results):
+    #colors = ['#8c510a', '#d8b365', '#f6e8c3', '#c7eae5', '#5ab4ac', '#01665e']
+    if os.path.exists(results + "/times.out"):
+        df = pd.read_csv(results + "/times.out", delimiter=" ", index_col=False, names=["wc", "traj"])
+    else:
+        untar(results + "/times.tar.gz")
+        shutil.move("times.out", "tmp/times.out")                
+        df = pd.read_csv("tmp/times.out", delimiter=" ", index_col=False, names=["wc", "traj"])
+    ax.plot(df["wc"], df["traj"], label=label, lw=2)
+    ax.set_ylabel("Trajectory Length (pico seconds)")
+    ax.set_xlabel("Wall Clock Time (seconds)")
+    ax.legend()
+### Boring Parsing Info
+# The data structure we read into is a dictionary, where the key is a tuple that identifies the job and the 
+# value is a tuple with parsed data in Python data frames. The `parse()` function reads the raw data and 
+# puts it into the dictionary:
+#
+# dfs[(name, op, rank)] => {KEY TIMES, KEY COUNT}
+#
+# where `name` is a human readable string that describes the job, `op` is the type of database operation,
+# `rank` is the MPI rank number, `KEY TIMES` has keys and timestamps, and `KEY COUNT` has keys and counts (i.e. # 
+# of occurences). We save `countDF` because it takes forever to parse this data.
+def parse(dfs, name, results):
+    stats = []
+    for rank in ["0", "1", "2", "3", "4"]:
+        untar(results + "/out/semanticPerf." + rank + ".tar.gz")
+        shutil.move("semanticPerf." + rank, "tmp/semanticPerf." + rank)
[…]

### 2017-09-27 experiments/keyspace-locality/viz/jupyter.sh
paper: add nicer figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -2,4 +2,4 @@
 
 set -e -x
 cd ../
-docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c33a7dc0eece
+docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook #:c33a7dc0eece

### 2017-09-27 paper/sections/1_introduction.tex
paper: new outline based on SSIO feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Introduction}
ParSplice~\cite{perez:jctc20150parsplice}

### 2017-09-27 paper/sections/2_parsplice-keyspace-analysis.tex
paper: new outline based on SSIO feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{ParSplice Keyspace Analysis}

\subsection{Structured Access Regimes}

Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
%shows ``access regimes", or groups of accesses
%to the same subset of keys:
\begin{itemize}
  \item change immediately, different sizes
  \item monotonically increasing
  \item random access to a single key
  \item early keys accessed more
\end{itemize}

\noindent\textbf{Conclusion}: unique patterns of a real HPC application

\subsection{Cache Size Analysis}

Figure~\ref{fig:keyspace-analysis_cachesize}.

\noindent\textbf{Conclusion}: workload phases (high request rate to a small
number of keys and then low request rate to a large number keys) need a dynamic
load balancing policy.

\subsection{Overfitting Policies}

Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.

\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
sensitive to: max usage or usage over time.\\

\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
request rate for a 2.5 hour run, but it is infeasible to do this for every
combination of system setup, job lengths, parsplice parameters.

\subsubsection*{Discussion}
Why don't we just an LRU cache:
\begin{itemize}
  \item we can do better (workload is structured and has locality)
  \item finding the size of the cache is hard
  \item hotspots dissipate too quickly
\end{itemize}

### 2017-09-27 paper/sections/3_methodology.tex
paper: new outline based on SSIO feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Mantle Engine: Methodology}

\subsection{Experimental Setup}

\subsection{Mantle: Dynamic Load Balancing Policies}

\begin{itemize}
  \item motivation: Mochi load balancer microservice
  \item background: CephFS implementation
  \item library architecture, callbacks, environment
\end{itemize}

\subsection{Integrating Mantle into ParSplice}

\begin{itemize}
  \item providing environment of metrics
  \item identifying where policies are made
\end{itemize}

### 2017-09-27 paper/sections/4_dynamic-policies.tex
paper: new outline based on SSIO feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Mantle Brains: Tools We Plug In}

Requirements: run online, be fast enough to run as often as we want to detect
regimes

\subsection{Failed, Overcomplicated Brains}
These techniques proliferated more, less transparent knobs
\begin{itemize}
  \item Statistics
  \item Calculus
  \item K-Means
  \item DBScan
  \item Anomaly Detection
\end{itemize}

\subsection{System Specific Knowledge}

We tried looking at system-specific metrics to design policies, like request
rate, unique keys in a sliding window, bandwidth capabilties. For example, we
know that LevelDB cannot handle high IO request rates.

\subsubsection{when: request rate}

\subsubsection{when: Belady's Min}

\subsection{Domain Specific Knowledge}

We tried looking at domain-specific metrics to design policies, like ParSplice
key access locality.

\subsubsection{howmuch: regime detection}
\subsubsection{when: trajectory length goodness}
\subsubsection{howmuch: cache policy from past}

\subsection{Cloud Techniques: Elastic Search}

What if we re-provision resource in response to events outside the
application's control, such as a slow Lustre.

### 2017-09-27 paper/sections/5_file-systems.tex
paper: new outline based on SSIO feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Relation to File Systems}

\begin{itemize}
  \item Lustre Trace
  \item LinkedIn Trace
  \item Nathan's Trace
\end{itemize}

\subsection{Using File System Balancers for ParSplice}
\subsection{Using ParSplice Balancers for File Systems}
\subsection{Visualizing File System Traces like ParSplice Keyspace Traces}

### 2017-09-27 paper/sections/6_conclusion.tex
paper: new outline based on SSIO feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Conclusion}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_size.png}\\
  \caption{\label{fig:keyspace-analysis_size}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_locality.png}\\
  \caption{\label{fig:keyspace-analysis_locality}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_throughput.png}\\
  \caption{\label{fig:keyspace-analysis_throughput}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize.png}\\
  \caption{\label{fig:keyspace-analysis_cachesize}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize-dynamic.png}\\
  \caption{\label{fig:keyspace-analysis_cachesize-dynamic}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize-caveats.png}\\
  \caption{\label{fig:keyspace-analysis_cachesize-caveats}}
\end{figure}

### 2017-09-27 jupyter.sh

@@ -1,5 +1,4 @@
 #!/bin/bash
 
 set -e -x
-cd ../
 docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook #:c33a7dc0eece

### 2017-09-27 paper/figures/copy.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+set -x
+DIR="$HOME/Documents/parsplice-popper/paper/figures"
+cp $DIR/motivation-regimes.png keyspace-analysis_throughput.png
+cp $DIR/methodology-tradeoff.png keyspace-analysis_cachesize.png
+cp $DIR/methodology-tradeoff-dynamic.png keyspace-analysis_cachesize-dynamic.png
+cp $DIR/methodology-trajectory.png keyspace-analysis_cachesize-caveats.png
+cp $DIR/methodology-keyspace.png keyspace-analysis_size.png
+

### 2017-10-30 paper/sections/1_introduction.tex
paper: add code snippets for FS comparison
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{Introduction}
-ParSplice~\cite{perez:jctc20150parsplice}
+
+% What is the problem we are trying to solve
+[DESCRIBE PROBLEM WE ARE TRYING TO SOLVE - CUT/PASTE FROM PDSW PAPER]
+
+% What is Mantle
+To solve this problem, we integrate Mantle into
+ParSplice~\cite{perez:jctc20150parsplice}. Mantle is a dynamic policy engine
+that decomposes complicated services into manageable pieces so developers
+unfamiliar with a code-base can reason about its behavior. Administrators use
+the framework to describe ``when" they want data moved, ``where" they want data
+moved, and ``how much" of the data to move.  This abstraction helps developers
+unfamiliar with the domain quickly reason about, develop, and deploy new
+policies that control temporal and spatial locality. It has already proven to
+be a critical control plane for improving file system metadata load
+balancing~\cite{sevilla:sc15-mantle} and in this work we show its usefulness in
+cache management.
+
+% What are we going to show
+We show how Mantle:
+\begin{itemize}
+
+  \item decomposes cache management into independent policies, making the
+  problem more manageable and facilitating rapid development. We quickly implement
+  the ``how much" policy and focus on the ``when" policy for the majority of this paper.
+
+  \item has useful primitives that, while designed for file systems, are
+  crucial for effective cache management. This finding shows how the Mantle
+  engine generalizes to a different domain and code-base.
+
+  \item can be used to quickly deploy a variety of cache management strategies,
+  ranging from basic algorithms and heuristics to statistical models and machine
+  learning. This mixing and matching may not outperform a hard-coded solution but
+  can more adeptly change to different workloads and hardware.
+
+\end{itemize}
+
+We demonstrate these contributions by changing the input to our molecular
+dynamics simulation, which changes the keyspace access patterns.
+
+%First, we implement the ``how much" abstraction as an LRU and then we focus on
+%the ``when" abstraction.
+
+%Manageable: abstracts away complexities of the system (pass around to others,
+%use different strategies)

### 2017-10-30 paper/sections/4_dynamic-policies.tex
paper: add code snippets for FS comparison
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\section{Mantle Brains: Tools We Plug In}
+\section{Mantle Brains: Tools We Plug In to Detect ``when"}
 
@@
 regimes
+
+\subsection{System Specific Knowledge}
+
+{\it e.g.}, request rate, unique keys in a sliding window, bandwidth
+capabilties. For example, we know that LevelDB cannot handle high IO request
+rates.
+
+\subsubsection{Request Rate}
+
+\subsubsection{Belady's Min}
+
+\subsection{Domain Specific Knowledge}
+
+{\it e.g.}, ParSplice key access locality.
+
+\subsubsection{Regime Detection}
+\subsubsection{Trajectory Length}
 
@@
 
-\subsection{System Specific Knowledge}
-
-We tried looking at system-specific metrics to design policies, like request
-rate, unique keys in a sliding window, bandwidth capabilties. For example, we
-know that LevelDB cannot handle high IO request rates.
-
-\subsubsection{when: request rate}
-
-\subsubsection{when: Belady's Min}
-
-\subsection{Domain Specific Knowledge}
-
-We tried looking at domain-specific metrics to design policies, like ParSplice
-key access locality.
-
-\subsubsection{howmuch: regime detection}
-\subsubsection{when: trajectory length goodness}
-\subsubsection{howmuch: cache policy from past}
-
 \subsection{Cloud Techniques: Elastic Search}
 
-What if we re-provision resource in response to events outside the
+What if we re-provision resources in response to events outside the
 application's control, such as a slow Lustre.
+
+\subsection{Future work}
+
+How Much: cache policy from past, regime detection

### 2017-10-30 paper/sections/5_file-systems.tex
paper: add code snippets for FS comparison
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{Relation to File Systems}
+
+The code snippets in Figures~\ref{lua:parsplice} and~\ref{lua:cephfs} are the
+policies used in ParSplice and CephFS, respectively.  ParSplice uses policies
+to manage its caches and CephFS uses policies to control load balancing, but
+they both can be expressed with the Mantle API.  From a high-level the
+ParSplice policy trims the cache if the cache reaches a certain size {\it and}
+if it has already absorbed the initial burstiness of the workload; the CephFS
+policy migrates load if the metadata load is higher than the average load {\it
+and} the current load has been overloaded for more than two iterations.
+
+\textbf{Condition for ``Overloaded"} (Fig.~\ref{lua:parsplice}:Line 2;
+Fig.~\ref{lua:cephfs}:Line 2) - these lines detect whether the node is
+overloaded using the ``load" calculated in the load callback; while the load
+calculations and thresholds are different, the actual logic is exactly the
+same.  Recall that this decision is made locally because there is no global
+scheduler or centralized intelligence.
+
+\textbf{State Persisted Across Decisions} (Fig.~\ref{lua:parsplice}:Lines 5,6;
+Fig~\ref{lua:cephfs}:Lines 3,4,10) - these lines use the Mantle API to
+write/read state from previous decisions.  For ParSplice, we save a boolean
+that indicates whether we have absorbed the workload's initial burstiness. For
+CephFS, we save the number of consecutive instances that the server has been
+overloaded. We also clear the count (Line 10) if the server is no longer
+overloaded. The underlying implementation saves the values to local disk.
+
+\textbf{Condition that Switches Policy} (Fig.~\ref{lua:parsplice}:Line 3;
+Fig.~\ref{lua:cephfs}:Line 5) - these lines switch the policies using
+information from previous decisions. ParSplice trims its cache once it eclipses
+the ``absorb" threshold while CephFS allows balancing when overloaded for more
+than two iterations. The persistent state is essential for both of these
+policy-switching conditions.
+
+\begin{figure}[tb]
+\begin{minted}[linenos]{lua}
+local function when()
+  if servers[whoami]["load"] > target then
+    if servers[whoami]["load"] > absorb_target then
+      WRstate(1)
+    end
+    if RDstate() == 1 then
+      return true
+    end
+  end
+  return false
+end
+\end{minted}
+\caption{ParSplice cache management policy.\label{lua:parsplice}}
+\end{figure}
+
+\begin{figure}[tb]
+\begin{minted}[linenos]{lua}
+local function when()
+  if servers[whoami]["load"] > target then
+    overloaded = RDstate() + 1
+    WRstate(overloaded)
+    if overloaded > 2 then
+      return true
+    end
+  end
+  else then
+    WRstate(0)
+  end
+  return false
+end
+\end{minted}
+\caption{CephFS file system metadta load balancer.\label{lua:cephfs}}
+\end{figure}
 

### 2017-11-01 paper/sections/1_introduction.tex
paper: edits, viz: add Dockerfile
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 ParSplice~\cite{perez:jctc20150parsplice}. Mantle is a dynamic policy engine
-that decomposes complicated services into manageable pieces so developers
-unfamiliar with a code-base can reason about its behavior. Administrators use
-the framework to describe ``when" they want data moved, ``where" they want data
-moved, and ``how much" of the data to move.  This abstraction helps developers
-unfamiliar with the domain quickly reason about, develop, and deploy new
-policies that control temporal and spatial locality. It has already proven to
-be a critical control plane for improving file system metadata load
+that decomposes complicated services into manageable pieces, where developers
+write policies for ``when" they want data moved, ``where" they want data moved,
+and ``how much" of the data to move.  The framework then executes these
+policies whenever a decision needs to be made.  This abstraction helps
+developers unfamiliar with the domain quickly reason about, develop, and deploy
+new policies that control temporal and spatial locality. It has already proven
+to be a critical control plane for improving file system metadata load
 balancing~\cite{sevilla:sc15-mantle} and in this work we show its usefulness in

### 2017-11-01 paper/sections/4_dynamic-policies.tex
paper: edits, viz: add Dockerfile
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \subsubsection{Regime Detection}
+
+At each time step, we find the lowest ID and compare against the local minima,
+which is the smallest ID we have seen thus far. If we move left to right, the
+local minima never changes because the local minima will start small. If we
+move from right to left, the local minima changes at each access regime.  For
+points \(z\), \(y\), and \(x\), if the local minimum is the same we are in a
+regime. Processing \(y\), we set the local minima to be \(min(y, m_l)\), where
+\(m_l\) is the local minima of the previous time step of \(z\).
+
+This
+raises false positives for regime changes
+
+\begin{figure}[tb]
+\begin{minted}[linenos]{lua}
+local function when()
+
+  if servers[whoami]["load"] > target then
+    if servers[whoami]["load"] > absorb_target then
+      WRstate(1)
+    end
+    if RDstate() == 1 then
+      return true
+    end
+  end
+  return false
+end
+\end{minted}
+\caption{ParSplice cache management policy.\label{lua:parsplice}}
+\end{figure}
+
 \subsubsection{Trajectory Length}

### 2017-11-01 jupyter.sh

@@ -1,4 +1,4 @@
 #!/bin/bash
 
 set -e -x
-docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook #:c33a7dc0eece
+docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:xkcd

### 2017-11-02 paper/sections/1_introduction.tex
paper: integrate Carlos feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-09-27 paper/sections/1_introduction.tex
@@
 \section{Introduction}
-ParSplice~\cite{perez:jctc20150parsplice}
+
+The fine-grained data annotation capabilities provided by key-value storage is
+a natural match for many types of scientific simulation. Simulations relying on
+a mesh-based decomposition of a physical region may result in millions or
+billions of mesh cells. Each cell contains materials, pressures, temperatures
+and other characteristics that are required to accurately simulate phenomena of
+interest. In our target application, the
+ParSplice~\cite{perez:jctc20150parsplice} molecular dynamics simulation, a
+hierarchy of cache nodes and a single node key-value store are used to store
+both observed minima across a molecule's equation of motion (EOM) and the
+hundreds or thousands of partial trajectories calculated each second during a
+parallel job. Unfortunately, if we scale the system the IO to the storage
+hierarchy will quickly saturate both the storage and bandwidth capacity of a
+single node.
+
+In this paper we present a detailed analysis of how the ParSplice application
+accesses key-value pairs over the course of a long running simulation across a
+variety of initial conditions. We reason that limiting the size of a cache on a
+single node saves memory and sacrifices negligible performance. This type of
+analysis (1) shows the capacity and resource requirements of a single node and
+(2) will help inform our load balancing policies for when we switch to a
+distributed key-value store back-end to store EOM minima. We need to know when
+and how to partition the keyspace: a smaller cache hurts performance because
+key-value pairs need to be retrieved from other nodes while a larger cache has
+higher memory pressure.
+
+% What is Mantle
+To explore the effects of different cache management strategies, we link the
+Mantle policy engine into ParSplice~\cite{perez:jctc20150parsplice}.
+Developers write policies for ``when" they want data moved, ``where" they want
+data moved, and ``how much" of the data to move and the framework executes
+these policies whenever a decision needs to be made.  This abstraction helps
+developers unfamiliar with the domain quickly reason about, develop, and deploy
+new policies that control temporal and spatial locality.  It has already proven
+to be a critical control plane for improving file system metadata load
+balancing~\cite{sevilla:sc15-mantle} and in this work we show its usefulness in
+cache management for the changing key-value workloads generated by ParSplice.
+
+% What are we going to show
+We show how Mantle:
+\begin{itemize}
+
+  \item decomposes cache management into independent policies that can be
+  dynamically changed, making the problem more manageable and facilitating rapid
+  development. Changing the policy in use is critical in applications such as
+  ParSplice that have alternating stable and chaotic simulation ``access regimes"
+  over the course of a long-running simulation.
+
+  \item has useful primitives that, while designed for file systems, turn
+  out to also be effective for cache management. This finding shows how the Mantle
+  engine generalizes to a different domain and code-base.
+
+  \item can be used to quickly deploy a variety of cache management strategies,
+  ranging from basic algorithms and heuristics to statistical models and machine
+  learning.
+
+\end{itemize}
+
+% this gives us many policies that are effective across disciplines
+% - reuse: eases burden of writing policies
+% - autonomic: lays groundwork for an adaptable policy that mixes/matches policies
+
+This last contribution is explored in Section~\S\ref{sec:mantle-brains}, where
+we try a range of policies from different disciplines; but more importantly, in
+Section~\S\ref{sec:relation-to-file-systems}, we conclude that the collection
+of policies we designed for ParSplice's cache management are very similar to
+the policies implemented in the Ceph file system metadata load balancer. This
+lays the foundation for future work, where we will focus on formalizing a
+collection of general data management policies that can be used across domains
+and services. The value of such a collection eases the burden of policy
+development and paves the way for solutions that remove the administrator from
+the development cycle, such as (1) adaptable policies that automatically switch
+to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
+making no progress, etc.), and (2) policy generation, where new policies are
+constructed automatically by examining the collection of existing policies.
+Such work is made possible with Mantle's ability to dynamically change
+policies.
+
+%First, we implement the ``how much" abstraction as an LRU and then we focus on
+%the ``when" abstraction. 
+%Manageable: abstracts away complexities of the system (pass around to others,
+%use different strategies)

### 2017-11-02 paper/sections/4_dynamic-policies.tex
paper: integrate Carlos feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-10-30 paper/sections/4_dynamic-policies.tex
@@
 \section{Mantle Brains: Tools We Plug In to Detect ``when"}
+\label{sec:mantle-brains}
 
@@
 \subsubsection{Regime Detection}
+
+At each time step, we find the lowest ID and compare against the local minimum,
+which is the smallest ID we have seen thus far. If we move left to right, the
+local minimum never changes because the local minimum will start small. If we
+move from right to left, the local minimum changes at each access regime.  For
+points \(z\), \(y\), and \(x\), if the local minimum is the same we are in a
+regime. Processing \(y\), we set the local minimum to be \(min(y, m_l)\), where
+\(m_l\) is the local minimum of the previous time step of \(z\). The algorithm
+incorrectly detects a regime change if the local minimum of \(y\) is lower than
+local minimum of \(z\), since \(y\) may have points {\it within} \(z\); recall
+that we are trying to detect the whole fan, not just the bottom edge of each
+fan.
+
+\begin{figure}[tb]
+\begin{minted}[linenos]{lua}
+
+-- assume that (ts, keyid) are in a table
+local function when()
+
+  if servers[whoami]["load"] > target then
+    if servers[whoami]["load"] > absorb_target then
+      WRstate(1)
+    end
+    if RDstate() == 1 then
+      return true
+    end
+  end
+  return false
+end
+\end{minted}
+\caption{ParSplice cache management policy.\label{lua:parsplice}}
+\end{figure}
+
 \subsubsection{Trajectory Length}

### 2017-11-02 paper/sections/1_introduction.tex
paper: new intro, title, abstract
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 To explore the effects of different cache management strategies, we link the
-Mantle policy engine into ParSplice~\cite{perez:jctc20150parsplice}.
+Mantle policy engine into ParSplice.  Mantle~\cite{sevilla:sc15-mantle} has
+already proven to be a critical control plane for improving file system
+metadata load balancing  and in this work we show its usefulness in cache
+management for the changing key-value workloads generated by ParSplice.
 Developers write policies for ``when" they want data moved, ``where" they want
@@
 developers unfamiliar with the domain quickly reason about, develop, and deploy
-new policies that control temporal and spatial locality.  It has already proven
-to be a critical control plane for improving file system metadata load
-balancing~\cite{sevilla:sc15-mantle} and in this work we show its usefulness in
-cache management for the changing key-value workloads generated by ParSplice.
+new policies that control temporal and spatial locality. We show that Mantle:
 
-% What are we going to show
-We show how Mantle:
 \begin{itemize}
@@
 of policies we designed for ParSplice's cache management are very similar to
-the policies implemented in the Ceph file system metadata load balancer. This
-lays the foundation for future work, where we will focus on formalizing a
+the policies in the Ceph file system that are used to load balance metadata.
+This lays the foundation for future work, where we will focus on formalizing a
 collection of general data management policies that can be used across domains

### 2017-11-06 jupyter.sh
viz: cleanup algorithm with Carlos comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,4 +1,5 @@
 #!/bin/bash
 
 set -e -x
-docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:xkcd
+#docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:xkcd
+docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c33a7dc0eece

### 2017-11-06 paper/sections/1_introduction.tex
paper: switch to IEEE template, hack on intro, move future work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
+\begin{abstract}
+
+Our analysis of the key-value activity generated by the ParSplice molecular
+dynamics simulation demonstrates the need for more complex cache management
+strategies. Baseline measurements show clear keyspace access patterns and hot
+spots that offer significant opportunity for optimization. We use the Mantle
+policy engine to dynamically explore a variety of techniques, ranging from
+basic algorithms and heuristics to statistical models, calculus, and machine
+learning. While Mantle was originally designed for distributed file systems, we
+show how it effectively decomposes the problem into manageable policies for a
+different domain and service (in this case, cache management).  Our exploration
+of this space results in a two policy scheme that achieves 96\% efficiency
+while using only 7.6\% of the memory resources required by the base case.
+
+\end{abstract}
+
 \section{Introduction}
@@
   development. Changing the policy in use is critical in applications such as
-  ParSplice that have alternating stable and chaotic simulation ``access regimes"
+  ParSplice that have alternating stable and chaotic keyspace access patterns
   over the course of a long-running simulation.
-
-  \item has useful primitives that, while designed for file systems, turn
-  out to also be effective for cache management. This finding shows how the Mantle
-  engine generalizes to a different domain and code-base.
 
@@
   learning.
+
+  \item has useful primitives that, while designed for file system metadata
+  load balancing, turn out to also be effective for cache management. This
+  finding shows how the policy engine generalizes to different domains and
+  enables control of policies by machines instead of administrators.
 
@@
 % - autonomic: lays groundwork for an adaptable policy that mixes/matches policies
+% FUTURE WORK
 
@@
 of policies we designed for ParSplice's cache management are very similar to
-the policies in the Ceph file system that are used to load balance metadata.
-This lays the foundation for future work, where we will focus on formalizing a
-collection of general data management policies that can be used across domains
-and services. The value of such a collection eases the burden of policy
-development and paves the way for solutions that remove the administrator from
-the development cycle, such as (1) adaptable policies that automatically switch
-to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
-making no progress, etc.), and (2) policy generation, where new policies are
-constructed automatically by examining the collection of existing policies.
-Such work is made possible with Mantle's ability to dynamically change
-policies.
+the policies in the Ceph file system that are used to load balance metadata,
+suggesting that there is potential for automatically adapting and generating
+policies dynamically.
 

### 2017-11-06 paper/sections/2_parsplice-keyspace-analysis.tex
paper: integrate Carlos feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: switch to IEEE template, hack on intro, move future work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
+%\section{ParSplice Keyspace Analysis}
+%
+%We demonstrate these contributions by changing the input to our molecular
+%dynamics simulation, which changes the keyspace access patterns. 
+%
+%\subsection{Structured Access Regimes}
+%
+%Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
+%%shows ``access regimes", or groups of accesses
+%%to the same subset of keys:
+%\begin{itemize}
+%  \item change immediately, different sizes
+%  \item monotonically increasing
+%  \item random access to a single key
+%  \item early keys accessed more
+%\end{itemize}
+%
+%\noindent\textbf{Conclusion}: unique patterns of a real HPC application
+%
+%\subsection{Cache Size Analysis}
+%
+%Figure~\ref{fig:keyspace-analysis_cachesize}.
+%
+%\noindent\textbf{Conclusion}: workload phases (high request rate to a small
+%number of keys and then low request rate to a large number keys) need a dynamic
+%load balancing policy.
+%
+%\subsection{Overfitting Policies}
+%
+%Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.
+%
+%\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
+%sensitive to: max usage or usage over time.\\
+%
+%\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
+%request rate for a 2.5 hour run, but it is infeasible to do this for every
+%combination of system setup, job lengths, parsplice parameters.
+%
+%\subsubsection*{Discussion}
+%Why don't we just an LRU cache:
+%\begin{itemize}
+%  \item we can do better (workload is structured and has locality)
+%  \item finding the size of the cache is hard
+%  \item hotspots dissipate too quickly
+%\end{itemize}
+
 \section{ParSplice Keyspace Analysis}
+\label{sec:parsplice-keyspace-analysis}
 
-\subsection{Structured Access Regimes}
+ParSplice~\cite{perez:jctc20150parsplice} is an accelerated molecular dynamics
+(MD) simulation package developed at LANL. It is part of the Exascale Computing
+Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
+Materials for the Future initiative. Its phases are:
 
-Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
-%shows ``access regimes", or groups of accesses
-%to the same subset of keys:
-\begin{itemize}
-  \item change immediately, different sizes
-  \item monotonically increasing
-  \item random access to a single key
-  \item early keys accessed more
-\end{itemize}
+\begin{enumerate}
 
-\noindent\textbf{Conclusion}: unique patterns of a real HPC application
+  \item a splicer tells workers to generate segments (short MD trajectory) for
+  specific states
 
-\subsection{Cache Size Analysis}
+  \item workers read initial coordinates for their assigned segment from data
+  store; the key-value pair is (state ID, coordinate)
 
-Figure~\ref{fig:keyspace-analysis_cachesize}.
+  \item upon completion, workers insert final coordinates for each segment into
+  data store, and wait for new segment assignment
 
-\noindent\textbf{Conclusion}: workload phases (high request rate to a small
-number of keys and then low request rate to a large number keys) need a dynamic
-load balancing policy.
+\end{enumerate}
 
-\subsection{Overfitting Policies}
+The computation can be parallelized by adding more workers or by adding worker
+tasks to parallelize individual workers.  The workers are stateless and read
+initial coordinates from the data store each time they begin generating
+segments. Since worker tasks do not maintain their own history, they can end up
+reading the same coordinates repeatedly. To mitigate the consequences of these
+repeated reads, ParSplice provisions a hierarchy of nodes to act as caches that
+sit in front of a single node persistent database.  Values are written to each
+tier and reads traverse up the hierarchy until they find the data.
 
-Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.
+We use ParSplice to simulate the evolution of metallic nanoparticles that grow
+from the vapor phase. As the run progresses, the energy landscape of the system
+becomes more complex. Two domain factors control the number of entries in the
+data store: the growth rate and the temperature. The growth rate controls how
+quickly new atoms are added to the nanoparticle: fast growth rates lead to
+non-equilibrium conditions, and hence increase the number of states that can be
+visited. However, as the particle grows, the simulation slows down because the
+calculations become more expensive, limiting the rate at which new states are
+visited. On the other hand, the temperature controls how easily a trajectory
+can jump from state to state; higher temperatures lead to more frequent
+transitions.
 
-\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
-sensitive to: max usage or usage over time.\\
+The nanoparticle simulation stresses the data store architecture of ParSplice.
+It visits more states than other input decks because the system uses a cheap
+potential, has a small number of atoms, and operates in a complex energy
+landscape with many accessible states. Changing growth rates and temperature
+alters the size, shape, and locality of the data store keyspace. Lower
+temperatures and smaller growth rates create hotter keys with smaller keyspaces
+as many segments are generated in the same set of states before the trajectory
+can escape to a new region of state space.
 
-\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
-request rate for a 2.5 hour run, but it is infeasible to do this for every
-combination of system setup, job lengths, parsplice parameters.
+Our evaluation uses the total ``trajectory length" as the goodness metric. This
+value is the duration of the overall trajectory produced by ParSplice. At
+ideal efficiency, the trajectory length should increase with the square root of
+the wall-clock time, since the wall-clock cost of time-stepping the system by
+one simulation time unit increases in proportion of the total number of atoms.
 
-\subsubsection*{Discussion}
-Why don't we just an LRU cache:
-\begin{itemize}
-  \item we can do better (workload is structured and has locality)
-  \item finding the size of the cache is hard
-  \item hotspots dissipate too quickly
-\end{itemize}
+\subsection{ParSplice Keyspace Analysis}
+\label{sec:parsplice-keyspace-analysis}
+%IT HAS 8K keys!  How did we take these measurements
+
+We instrumented ParSplice with performance counters and keyspace counters.  The
+performance counters track ParSplice progress while keyspace counters track
+which keys are being accessed by the ParSplice ranks. Because the keyspace
+counters have high overhead we only turn them on for the keyspace analysis.
+The cache hierarchy is unmodified but for the persistent database node, we
+replace BerkeleyDB on NFS with LevelDB on Lustre. Original ParSplice
+experiments showed that BerkeleyDB's syncs caused reads/writes to bottleneck on
+the persistent database. We also use Riak's customized
+LevelDB\footnote{https://github.com/basho/leveldb} version, which comes
+instrumented with its own set of performance counters.
+
+All experiments ran on Trinitite, a Cray XC40 with 32 Intel Haswell 2.3GHz
+cores per node.  Each node has 128GB of RAM and our goal is to limit the size
+of the cache to 3\% of RAM\footnote{Empirically, this is a threshold that we
+find to work well for most applications}. Note that this is an addition to the
+30GB that ParSplice uses to manage other ranks on the same node.  A single Cray
+node produced trajectories that are \(5\times\) times longer than our 10 node
+CloudLab clusters and \(25\times\) longer than UCSC's 10 node cluster. As
+a result, it reaches different job phases faster and gives us a more
+comprehensive view of the workload. The performance gains compared to the
+commodity clusters have more to do with memory/PCI bandwidth than network.
+
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\
+  \caption{The keyspace size is small but must satisfy many reads as workers
+  calculate new segments. It is likely that we will need
+  more than one node to manage segment coordinates when we scale the system or jobs up.
+  \label{fig:methodology-keyspace}}
+\end{figure}
+
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keys.png}\\
+  \caption{The keyspace imbalance is due to workers generating deep
+  trajectories and reading the same coordinates. Over time, the accesses get
+  dispersed across different coordinates resulting in some keys being more
+  popular than others.\label{fig:methodology-keys}}
+\end{figure}
+
+\subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
+keyspace size (black annotations) and request load (bars) after a one hour run
+with a different number of workers (\(x\) axis). While the keyspace size and
+capacity is relatively modest the memory usage scales linearly with the number
+of workers. This is a problem if we want to scale to Trinitite's 6000 cores.
+Furthermore, the size of the keyspace also increases linearly with the length
+of the run.  Extrapolating these results puts an 8 hour run across all 100
+Trinitite nodes at 20GB for the cache.  This memory utilization easily eclipses
+the 3\% threshold we set earlier, even without factoring in the memory usage
+from other workers.
+
+\subsubsection*{An active but small keyspace} The bars in
+Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many reads
+(\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same key
+for extended periods because the trajectory can remain stuck in so-called
+superbasins composed of tightly connected sets of states. In this case, many
+trajectory segments with the same coordinates are needed before the trajectory
+moves on.  Writes only occur for the final state of segments generated by
+worker tasks; their magnitude is smaller than reads because the caches ignore
+redundant write requests. The number of read and write requests are highest at
+the beginning of the run when worker tasks generate segments for the same
+state, which is computationally cheap (this motivates
+Section~\S\ref{sec:static-load-balancing}).
+
+\subsubsection*{Entropy increases over time} The reads per second in
+Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
+and the number of active keys increases over time. The resulting key access
+imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
+shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
+unique state, or key, along the \(x\) axis. Keys are more popular than others
+(up to \(5\times\)) because worker tasks start generating states with different
+coordinates later in the run (this motivates
+Section~\S\ref{sec:the-need-for-dynamic-load-balancing-policies}).  The growth
+rate, temperature, and number of workers have a predictable effect on the
+structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that the
+number of reads changes with different growth rates, but that the spatial
+locality of key accesses is similar ({\it e.g.}, some keys are still
+\(5\times\) more popular than others).  Figure~\ref{fig:motivation-regimes}
+shows how entropy for different growth rates has temporal locality, as the
+reads per second for \(\Delta_2\) looks like the reads per second for
+\(\Delta_1\) stretched out along the time axis.  Trends also exist for
+temperature and number of workers but are omitted here for space. This
+structure means that we can learn the regimes and adapt the storage system
+(this motivates Section~\S\ref{sec:ml-for-the-keyspace}).

### 2017-11-06 paper/sections/6_conclusion.tex
paper: switch to IEEE template, hack on intro, move future work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
+\section{Future Work}
+
+This lays the foundation for future work, where we will focus on formalizing a
+collection of general data management policies that can be used across domains
+and services. The value of such a collection eases the burden of policy
+development and paves the way for solutions that remove the administrator from
+the development cycle, such as (1) adaptable policies that automatically switch
+to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
+making no progress, etc.), and (2) policy generation, where new policies are
+constructed automatically by examining the collection of existing policies.
+Such work is made possible with Mantle's ability to dynamically change
+policies.
+
 \section{Conclusion}

### 2017-11-06 jupyter.sh

@@ -2,4 +2,4 @@
 
 set -e -x
 #docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:xkcd
-docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c33a7dc0eece
+docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:ea9927c3f043

### 2017-11-08 paper/sections/1_introduction.tex
paper: cleanup intro, pull in evaluation/related work, re-org evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
+  \caption{The keyspace activity for ParSplice using two different growth
+rates.  The line shows the rate that EOM minima values are retrieved from the
+key-value store (\(y1\) axis) and the points along the bottom show the number
+of keys accessed in a 1 second sliding window (\(y2\) axis).
+\label{fig:motivation-regimes}}
+\end{figure}
+
 In this paper we present a detailed analysis of how the ParSplice application
 accesses key-value pairs over the course of a long running simulation across a
-variety of initial conditions. We reason that limiting the size of a cache on a
-single node saves memory and sacrifices negligible performance. This type of
-analysis (1) shows the capacity and resource requirements of a single node and
-(2) will help inform our load balancing policies for when we switch to a
-distributed key-value store back-end to store EOM minima. We need to know when
-and how to partition the keyspace: a smaller cache hurts performance because
-key-value pairs need to be retrieved from other nodes while a larger cache has
-higher memory pressure.
+variety of initial conditions.  This type of analysis (1) shows the capacity
+and resource requirements of a single node and (2) will help inform our load
+balancing policies for when we switch to a distributed key-value store back-end
+to store EOM minima. We need to know when and how to partition the keyspace: a
+smaller cache hurts performance because key-value pairs need to be retrieved
+from other nodes while a larger cache has higher memory pressure.
+
+For example, Figure~\ref{fig:motivation-regimes} shows the
+cache activity of a ParSplice node and demonstrates that small changes to the
+rates at which new atoms enter the simulation (\(\Delta\)) can have a strong
+effect on the timing and frequency with which new EOM minima are discovered and
+referenced.  The ParSplice caches on each node are trimmed when memory pressure
+reaches a threshold but the number of unique keys accessed (black points
+along bottom of both graphs) in Figure~\ref{fig:motivation-regimes} suggests
+that a small cache of EOM minima should be sufficient.
 
@@
 
-%First, we implement the ``how much" abstraction as an LRU and then we focus on
-%the ``when" abstraction. 
 %Manageable: abstracts away complexities of the system (pass around to others,

### 2017-11-08 paper/sections/4_dynamic-policies.tex
paper: cleanup intro, pull in evaluation/related work, re-org evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\section{Mantle Brains: Tools We Plug In to Detect ``when"}
-\label{sec:mantle-brains}
+\section{Cache Management Using System Architecture Knowledge}
+\label{sec:cache-management-using-system-archiecture-knowledge}
 
-Requirements: run online, be fast enough to run as often as we want to detect
-regimes
+Using the Mantle policy engine, we test a variety of cache management tools and
+algorithms on a single in-memory database node using the keyspace analysis in
+Section~\S\ref{sec:parsplice-keyspace-analysis}. These strategies are
+implemented as the ``when" and ``how much" policies from the Mantle API; the
+``where" callback does not make sense for a single node (segments are either in
+the cache or they are not). The evaluation metric is the accuracy and runtime
+of each strategy; the strategy should be accurate enough so as to sacrifice
+negligible performance and fast tenough to run as often as we want to detect
+regimes.
 
-\subsection{System Specific Knowledge}
+First we sized the cache according to our system specific knowledge. By
+``system", we mean the hardware and software of the storage hierarchy. We look
+at request rate, unique keys in a sliding window, and bandwidth capabilties. For
+example, we know that LevelDB cannot handle high IO request rates.
 
-{\it e.g.}, request rate, unique keys in a sliding window, bandwidth
-capabilties. For example, we know that LevelDB cannot handle high IO request
-rates.
+% technical details
+In the original ParSplice implementation, each cache node uses an unlimited
+amount of memory to store segment coordinates. We limit the size of the cache
+using an LRU eviction policy, where the penalty for a cache miss is retrieving
+the data from the persistent database.  We evict keys (if necessary) at every
+operation instead of when segments complete because the cache fills up too
+quickly otherwise.
 
-\subsubsection{Request Rate}
+% results: cache size trade-offs
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
 
-\subsubsection{Belady's Min}
+  \caption{The performance and resource utilization trade-off for different
+cache sizes. ``Baseline" is
+ParSplice unmodified and the ``Policy: Constrain Cache Size" graph limits the
+size of the cache to save memory.  \label{fig:methodology-tradeoff}}
 
-\subsection{Domain Specific Knowledge}
+\end{figure}
 
-{\it e.g.}, ParSplice key access locality.
+The results for different cache sizes for a growth rate of \(\Delta_1\) over a
+2.5 hour run across 256 workers is shown in
+Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
+unmodified ParSplice  measured in trajectory duration (\(y\) axis) and
+utilization is measured with memory footprint of just the cache (\(y2\) axis).
+The other graph shares the \(y\) axis and shows the trade-off of constraining
+the cache to different sizes.  The error bars are the standard deviation of 3
+runs.
 
-\subsubsection{Regime Detection}
+% results: raw numbers
+Although the keyspace grows to 150K, a 100K key cache achieves 99\% of the
+performance. Decreasing the cache degrades performance and predictability.
+While this result is not unexpected, it nonetheless achieves our goal of
+showing the benefits of load balancing keys across nodes and that smaller
+caches on each node are an effective way to save memory without completely
+sacrificing performance.
+
+% why is the performance lower for smaller caches?
+Despite the memory savings, our results suggest that dynamic load balancing
+policies could save even more memory.  Figure~\ref{fig:methodology-tradeoff}
+show that a 100K key cache is sufficient as a static policy but the top graph
+in Figure~\ref{fig:motivation-regimes} indicates that the cache size could be
+much smaller. That graph shows that the beginning of the run is characterized
+by many reads to a small set of keys and the end sees much lower reads per
+second to a larger keyspace. Specifically, it shows only about 100 keys as
+active in the latter half of the run.
+
+After analyzing traces, we see that the 100 key cache is insufficient because
+the persistent database cannot service the read-write traffic.  According to
+Figure~\ref{fig:motivation-regimes}, the read requests arrive at 750 reads per
+second in addition to the writes that land in each tier (about 300 puts/second,
+some redundant). This traffic triggers a LevelDB compaction and reads block,
+resulting in very slow progress.  Traces verify this hypothesis and show reads
+getting backed up as the read/write ratio increases. To recap, small caches
+incur too much load on the persistent database  at the beginning of the run but
+should suffice after the initial read flash crowd passes because
+the keyspace is far less active. This suggests a two-part load balancing
+policy.
+
+% what is mantle
+To explore dynamic load balancing policies ({\it i.e.} policies that change
+during the run), we use the Mantle approach.  Mantle is a framework built on the
+Ceph file system that lets administrators control file system metadata load
+balancing policies. The basic premise is that load balancing policies can be
+expressed with a simple API consisting of ``when", ``where", and ``how much".
+The succinctness of the API lets users inject multiple, dynamic policies.
+
+% Why is this a good idea
+Although ParSplice does not use a distributed file system, its workload is very
+similar because the minima key-value store responds to small and frequent
+requests, which results in hot spots and flash crowds.  Modern distributed file
+systems have found efficient ways to measure, migrate, and partition metadata
+load and have shown large performance gains and
+better scalability~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs,
+grider:pdsw2015-marfs, ren:sc2014-indexfs, patil:fast2011-giga+,
+brandt:msst2003-lh}.  Previous work quantified the speedups achieved with
+Mantle and formalized balancers that were good for file systems.
+
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff-dynamic.png}\\
+  \caption{The performance and resource utilization trade-off of using a
+  dynamic load balancing policy that switches to a constrained cache policy after absorbing
+  the initial burstiness of the workload. The sizes of these smaller caches are
+  on the \(x\) axis.  \label{fig:methodology-tradeoff-dynamic}}
+\end{figure}
+
+Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using the
+Mantle API to program a dynamic load balancing policy into ParSplice that
+switches between two policies:
+
+\begin{itemize}
+  \item unlimited growth policy: cache increases on every write
+  \item \(n\) key limit policy: cache constrained to \(n\) keys
+\end{itemize}
+
+We trigger the policy switch at 100K keys to absorb the flash crowd at the
+beginning of the run. Once triggered, keys are evicted to bring the size of the
+cache down to the threshold.  In the bar chart, the cache sizes for the \(n\)
+key limit policy are along the \(x\) axis.
+
+% results: same level of performance can be achieved 
+The dynamic policies show better performance than the single \(n\) key
+policies. The performance and memory utilization for a 100K key cache size is
+the same as the 100K bar in Figure~\ref{fig:methodology-tradeoff-dynamic} but
+the rest reduce the size of the keyspace after the read flash crowd.  We see
+the worst performance when the engine switches to the 10 key limit policy,
+which achieves 94\% of the performance while only using 40KB of memory.
+
+% caveats: it is calculating 90% of the trajectory, memory value reported is final
+\subsubsection*{Caveats}
+
+The results in Figure~\ref{fig:methodology-tradeoff-dynamic} are slightly
+deceiving for three reason: (1) segments take longer to generate later in the
+run, (2) the memory footprint is the value at the end of 2.5 hours, and (3)
+this policy only works well for the 2.5 hour run.  For (1), the curving down of
+the simulation vs. wall-clock time is shown in
+Figure~\ref{fig:methodology-trajectory}; as the nanoparticle grows it takes
+longer to generate segments so by the time we reach 2 hours, over 90\% of the
+trajectory is already generated.  For (2), the memory footprint is around 0.4GB
+until we reach 100K key switch threshold. In
+Figures~\ref{fig:methodology-tradeoff}
+and~\ref{fig:methodology-tradeoff-dynamic} we plot the final value. For (3),
+Figure~\ref{fig:methodology-trajectory} shows that the cache fills up with 100K
+keys at time 7200 seconds and its size is reduced to the size listed in the
+legend.  The curves stay close to ``Unlimited" for up to an hour after the
+cache is reduced but eventually flatten out as the persistent database gets
+overloaded. 10K and 100K follow the ``Unlimited" curve the longest and are
+sufficient policies for the 2.5 hour runs but anything longer would need a
+different dynamic load balancing policy.
+
+\begin{figure}[t]
+  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/methodology-trajectory.png}\\
+  \caption{The rate that the trajectory is computed decays over time (which is
+  expected) but this skews the performance improvements in
+  Figure~\ref{fig:methodology-tradeoff-dynamic}. Our dynamic policy works for 2.5
+  hour jobs but not for 4 hour jobs.  \label{fig:methodology-trajectory}}
+\end{figure}
+
+% but the result is still valid
+Despite these caveats, the result is still valid: we found a dynamic load
+balancing policy that absorbs the cost of a high read throughput on a small
+keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
+show the effectiveness of the load balancing policy engine we integrated into
+ParSplice, not that we were able to identify the best policy for all system
+setups ({\it i.e.} different ParSplice parameters, number of worker tasks, and
+job lengths).  To solve that problem, we need a way to identify what thresholds
+we should use for different job permutations.
+
+\section{Cache Management Using Domain-Specific Application Knowledge}
+\label{sec:cache-management-using-domain-specific-knowledge}
+\begin{figure}[t]
+  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/keyspace-zoomed.png}\\
+
+  \caption{Key activity for a 4 hour run shows groups of accesses to the same
+  subset of keys. Detecting this ``access regimes" leads to a more accurate cache
+  management strategy.\label{fig:keyspace-zoomed}}
+
+\end{figure}
+
+Feeding domain-specific knowledge about the ParSplice application into a policy
+leads to more a accurate cache management strategy.
+Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
+the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
+of the keys, or ``access regimes", occur because molecules are stuck in deep
+trajectories. Recall that the in-memory database stores the molecules' EOM
+minima, which is the smallest effective energy that a molecule observes during
+its trajectory. So molecules stuck in deep trajectories explore the same minima
+until they can escape to a new set of states. This exploration of the same set
+of states is called a superbasin.
+
+Detecting these superbasins can lead to more effective cache management
+strategies because the height of the keyspace accesses is ``how much" of cache
+to evict and the width of the keyspace accesses is ``when" to evict values from
+the cache.  The zoomed portion of Figure~\ref{fig:keyspace-zoomed} shows how a
+single superbasin affects the key accesses. Moving along the \(x\) axis shows
+that the number of unique keys accessed over time grows while moving along the
+\(y\) axis shows that early keys are accessed more often.  Overall, superbasins
+are never re-visited because the simulation only adds molecules; we can never
+reach a state with less molecules. This is why keys are never re-accessed.
+Despite these patterns, the following characteristics of superbasins make it
+hard to detect them:
+
+\begin{itemize}
+
+  \item superbasin key accesses are random and there is no threshold ``minimum distance
+  between key access" that indicates we have moved on to a new superbasin
+
+  \item superbasins change immediately
+
+  \item the number of keys a superbasin accesses differs from other superbasins
+
+\end{itemize}
+
+Below we describe the policies we implemented and deployed using Mantle.
+
+\subsection{Failed Strategies}
+These techniques proliferated more knobs that obfuscated the problem. 
+\begin{itemize}
+  \item Statistics
+  \item Calculus
+  \item K-Means
+  \item DBScan
+  \item Anomaly Detection
+\end{itemize}
+
+\subsection{Regime Detection}
 
@@
 
-\begin{figure}[tb]
-\begin{minted}[linenos]{lua}
-
--- assume that (ts, keyid) are in a table
-local function when()
-
-  if servers[whoami]["load"] > target then
-    if servers[whoami]["load"] > absorb_target then
-      WRstate(1)
-    end
-    if RDstate() == 1 then
-      return true
-    end
-  end
-  return false
-end
-\end{minted}
-\caption{ParSplice cache management policy.\label{lua:parsplice}}
-\end{figure}
-
-\subsubsection{Trajectory Length}
-
-\subsection{Failed, Overcomplicated Brains}
-These techniques proliferated more, less transparent knobs
-\begin{itemize}
-  \item Statistics
-  \item Calculus
-  \item K-Means
-  \item DBScan
-  \item Anomaly Detection
-\end{itemize}
+\subsection{Trajectory Length}
 
@@
 
-How Much: cache policy from past, regime detection
+How Much: cache policy from past, regime detection, Belady's Min

### 2017-11-08 paper/sections/6_conclusion.tex
paper: cleanup intro, pull in evaluation/related work, re-org evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
+\section{Related Work}
+
+Key-value storage organizations for scientific applications is a field gaining
+rapid interest. In particular, the analysis of the ParSplice keyspace and the
+development of an appropriate scheme for load balancing is a direct response to
+a case study for computation caching in scientific
+applications~\cite{jenkins:ipdsw17-mochi}. In that work the authors motivated
+the need for a flexible load balancing \emph{microservice} to efficiently scale
+a memoization microservice. Our work is also heavily influenced by the
+Malacology project~\cite{sevilla:eurosys17-malacology} which seeks to provide
+fundamental services from within the storage system ({\it e.g.}, consensus) to
+the application.
+
+State-of-the-art distributed file systems partition write-heavy workloads and
+replicate read-heavy workloads, similar to the approach we are advocating
+here.  IndexFS~\cite{ren:sc2014-indexfs} partitions directories and clients
+write to different partitions by grabbing leases and caching ancestor metadata
+for path traversal. ShardFS takes the replication approach to the extreme by
+copying all directory state to all nodes. The Ceph file system
+(CephFS)~\cite{weil:sc2004-dyn-metadata, weil:osdi2006-ceph} employs both
+techniques to a lesser extent; directories can be replicated or sharded but the
+caching and replication policies are controlled with tunable parameters.  These
+systems still need to be tuned by hand with {\it ad-hoc} policies designed for
+specific applications.  Setting policies for migrations is arguably more
+difficult than adding the migration mechanisms themselves.  For example,
+IndexFS/CephFS use the GIGA+~\cite{patil:fast2011-giga} technique for
+partitioning directories at a \emph{predefined} threshold. Mantle makes headway
+in this space by providing a framework for exploring these policies, but does
+not attempt anything more sophisticated (e.g., machine learning) to create
+these policies.
+
+% ml and autotuning
+Auto-tuning is a well-known technique used in
+HPC~\cite{behzad:sc2013-autotuning, behzad:techreport2014-io-autotuning}, big
+data systems systems~\cite{herodotou_starfish_2011}, and
+databases~\cite{schnaitter_index_2009}.  Like our work, these systems focus on
+the physical design of the storage ({\it e.g.} cache size) but since we focused
+on a relatively small set of parameters (cache size, migration thresholds), we
+did not need anything as sophisticated as the genetic algorithm used
+in~\cite{behzad:sc2013-autotuning}.  We cannot drop these techniques into
+ParSplice because the magnitude and speed of the workload hotspots/flash crowds
+makes existing approaches less applicable.
+
+Our plan is to use MDHIM~\cite{greenberg:hotstorage2015-mdhim} as our back-end
+key-value store because it was designed for HPC and has the proper mechanisms
+for migration already implemented.
+
 \section{Future Work}
@@
 \section{Conclusion}
-\begin{figure}[tb]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_size.png}\\
-  \caption{\label{fig:keyspace-analysis_size}}
-\end{figure}
-\begin{figure}[tb]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_locality.png}\\
-  \caption{\label{fig:keyspace-analysis_locality}}
-\end{figure}
-\begin{figure}[tb]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_throughput.png}\\
-  \caption{\label{fig:keyspace-analysis_throughput}}
-\end{figure}
-\begin{figure}[tb]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize.png}\\
-  \caption{\label{fig:keyspace-analysis_cachesize}}
-\end{figure}
-\begin{figure}[tb]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize-dynamic.png}\\
-  \caption{\label{fig:keyspace-analysis_cachesize-dynamic}}
-\end{figure}
-\begin{figure}[tb]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize-caveats.png}\\
-  \caption{\label{fig:keyspace-analysis_cachesize-caveats}}
-\end{figure}

### 2017-11-09 paper/sections/1_introduction.tex
paper: clarify Mantle in intro, add metric dicscussion, start hacking on methodology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 spots that offer significant opportunity for optimization. We use the Mantle
-policy engine to dynamically explore a variety of techniques, ranging from
-basic algorithms and heuristics to statistical models, calculus, and machine
-learning. While Mantle was originally designed for distributed file systems, we
-show how it effectively decomposes the problem into manageable policies for a
-different domain and service (in this case, cache management).  Our exploration
-of this space results in a two policy scheme that achieves 96\% efficiency
-while using only 7.6\% of the memory resources required by the base case.
+language and policy engine to dynamically explore a variety of techniques,
+ranging from basic algorithms and heuristics to statistical models, calculus,
+and machine learning. While Mantle was originally designed for distributed file
+systems, we show how the collection of abstractions effectively decomposes the
+problem into manageable policies for a different domain and service (in this
+case, cache management).  Our exploration of this space results in a two policy
+scheme that achieves 96\% efficiency while using only 7.6\% of the memory
+resources required by the base case.
 
@@
 balancing policies for when we switch to a distributed key-value store back-end
-to store EOM minima. We need to know when and how to partition the keyspace: a
-smaller cache hurts performance because key-value pairs need to be retrieved
-from other nodes while a larger cache has higher memory pressure.
-
-For example, Figure~\ref{fig:motivation-regimes} shows the
-cache activity of a ParSplice node and demonstrates that small changes to the
-rates at which new atoms enter the simulation (\(\Delta\)) can have a strong
-effect on the timing and frequency with which new EOM minima are discovered and
-referenced.  The ParSplice caches on each node are trimmed when memory pressure
-reaches a threshold but the number of unique keys accessed (black points
-along bottom of both graphs) in Figure~\ref{fig:motivation-regimes} suggests
-that a small cache of EOM minima should be sufficient.
+to store EOM minima.  For example, Figure~\ref{fig:motivation-regimes} shows
+the cache activity of a ParSplice node and demonstrates that small changes to
+the rates at which new atoms enter the simulation (\(\Delta\)) can have a
+strong effect on the timing and frequency with which new EOM minima are
+discovered and referenced.  The ParSplice caches on each node are trimmed when
+memory pressure reaches a threshold but the number of unique keys accessed
+(black points along bottom of both graphs) in
+Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
+should be sufficient.
 
 % What is Mantle
-To explore the effects of different cache management strategies, we link the
-Mantle policy engine into ParSplice.  Mantle~\cite{sevilla:sc15-mantle} has
-already proven to be a critical control plane for improving file system
-metadata load balancing  and in this work we show its usefulness in cache
-management for the changing key-value workloads generated by ParSplice.
-Developers write policies for ``when" they want data moved, ``where" they want
-data moved, and ``how much" of the data to move and the framework executes
-these policies whenever a decision needs to be made.  This abstraction helps
+We use the Mantle language and policy engine to dynamically explore the effects
+of different cache management strategies for the changing key-value workloads
+generated by ParSplice.  Mantle~\cite{sevilla:sc15-mantle} has already proven
+to be a critical control plane for improving file system metadata load
+balancing  and in this work we show that the collection of abstractions
+provided by Mantle is useful for reasoning about and designing different cache
+management strategies.  Developers write policies for ``when" they want data
+moved and ``how much" of the data to move, then the framework executes these
+policies whenever a decision needs to be made.  These abstraction help
 developers unfamiliar with the domain quickly reason about, develop, and deploy
@@
 
-This last contribution is explored in Section~\S\ref{sec:mantle-brains}, where
-we try a range of policies from different disciplines; but more importantly, in
+This last contribution is explored in
+Sections~\S\ref{sec:cache-management-using-system-archiecture-knowledge}
+and~\S\ref{sec:cache-management-using-domain-specific-knowledge}, where we try
+a range of policies from different disciplines; but more importantly, in
 Section~\S\ref{sec:relation-to-file-systems}, we conclude that the collection

### 2017-11-09 paper/sections/2_parsplice-keyspace-analysis.tex
paper: clarify Mantle in intro, add metric dicscussion, start hacking on methodology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 cores per node.  Each node has 128GB of RAM and our goal is to limit the size
-of the cache to 3\% of RAM\footnote{Empirically, this is a threshold that we
-find to work well for most applications}. Note that this is an addition to the
-30GB that ParSplice uses to manage other ranks on the same node.  A single Cray
-node produced trajectories that are \(5\times\) times longer than our 10 node
-CloudLab clusters and \(25\times\) longer than UCSC's 10 node cluster. As
-a result, it reaches different job phases faster and gives us a more
+of the cache to 3\% of RAM\footnote{Empirically, we find this threshold works
+well for most applications}. Note that this is an addition to the 30GB that
+ParSplice uses to manage other ranks on the same node.  A single Cray node
+produced trajectories that are \(5\times\) times longer than our 10 node
+CloudLab clusters and \(25\times\) longer than UCSC's 10 node cluster. As a
+result, it reaches different job phases faster and gives us a more
 comprehensive view of the workload. The performance gains compared to the

### 2017-11-09 paper/sections/3_methodology.tex
paper: clarify Mantle in intro, add metric dicscussion, start hacking on methodology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\begin{itemize}
-  \item motivation: Mochi load balancer microservice
-  \item background: CephFS implementation
-  \item library architecture, callbacks, environment
-\end{itemize}
+%  \item motivation: Mochi load balancer microservice
+
+%  \item background: CephFS implementation
+
+%  \item library architecture, callbacks, environment
+We extract Mantle as a library that is linked into a service ({\it e.g.}, file
+system, key-value store, application), where administrators program policies
+that are embedded into the runtime.  The service needs to be modified to (1)
+provide environment of metrics (2) identify where policies are set. At runtime,
+Mantle executes the admin's policies for when/where/how much and returns a
+decision.
+
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/mantle.png}\\
+
+  \caption{Extracting Mantle as library.\label{fig:mantle}}
+
+\end{figure}
+
+\begin{table}
+  \centering
+  \begin{tabular}{ r | l | l }
+  Metrics     & Data Structure & Description \\\hline
+  Cluster     & \{server \(\rightarrow\) \{metric \(\rightarrow\) val\}\}
+              & resource util. for servers \\
+  Time Series & [(ts, val), ..., (ts, val)]
+              & accesses by timestamp (ts) \\
+  && \\
+              & Service      & Example \\\hline
+  Cluster     & File Systems & CPU util., Inode RDs \\
+              & ParSplice    & CPU util., Cache Size \\
+  Time Series & File Systems & Accesses to directory \\
+              & ParSplice    & Accesses to key in DB\\
+  \end{tabular}
+  \caption{Types of metrics exposed by the service to the policy engine using Mantle.\label{table:metrics}}
+\end{table}
+
+%of timestamp (ts), value pairs representing accesses over time types of
+%metrics: key value pairs
+Services expose two types of metrics with Mantle: cluster metrics for
+describing resource utilization and time series metrics for describing accesses
+over time. Table~\ref{table:metrics} shows how these metrics are accessed from
+the policies written by administrators. Cluster metrics are read from a
+dictionary indexed by server and metric name, where server is a node identifier
+({e.g.}, MPI Rank, metadata server name), metric is a resource name and val is
+the current reading for that metric.  Examples of \texttt{(metric, value)}
+pairs are CPU or memory pressure, as shown in Figure~\ref{fig:mantle}.  For
+time series metrics, the service passes a pointer to an array of
+\texttt{(timestamp, value)} pairs. We use a pointer so we can pass a large
+number of values, like accesses over time to a database or directory in the
+file system namespace, but this limits the time series metrics to only include
+values from the {\it current } node.
+
+% policies themselves
+Administrators write policies as callbacks that use the metrics from above.
+Figure~\ref{fig:mantle} shows an example policy for the ``when()" callback, where the
+current server (\texttt{whoami}) migrates load if it is has load
+(\texttt{>0.1}) and if its neighbor server (\texttt{whoami + 1}) does not have
+load (\texttt{<0.1}).
 

### 2017-11-09 paper/sections/1_introduction.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 strategies. Baseline measurements show clear keyspace access patterns and hot
-spots that offer significant opportunity for optimization. We use the Mantle
-language and policy engine to dynamically explore a variety of techniques,
-ranging from basic algorithms and heuristics to statistical models, calculus,
-and machine learning. While Mantle was originally designed for distributed file
-systems, we show how the collection of abstractions effectively decomposes the
-problem into manageable policies for a different domain and service (in this
-case, cache management).  Our exploration of this space results in a two policy
-scheme that achieves 96\% efficiency while using only 7.6\% of the memory
-resources required by the base case.
+spots that offer significant opportunity for optimization. We use the data
+management and policy engine from the Mantle system to dynamically explore a
+variety of techniques, ranging from basic algorithms and heuristics to
+statistical models, calculus, and machine learning. While Mantle was originally
+designed for distributed file systems, we show how the collection of
+abstractions effectively decomposes the problem into manageable policies for a
+different domain and service (in this case, cache management).  Our exploration
+of this space results in a two policy scheme that achieves 96\% efficiency
+while using only 7.6\% of the memory resources required by the base case.
 
@@
 hierarchy will quickly saturate both the storage and bandwidth capacity of a
-single node.
+single node, so we need more effective data management techniques, such as
+cache management or load balancing across a cluster.
 
@@
 
-In this paper we present a detailed analysis of how the ParSplice application
-accesses key-value pairs over the course of a long running simulation across a
-variety of initial conditions.  This type of analysis (1) shows the capacity
-and resource requirements of a single node and (2) will help inform our load
-balancing policies for when we switch to a distributed key-value store back-end
-to store EOM minima.  For example, Figure~\ref{fig:motivation-regimes} shows
-the cache activity of a ParSplice node and demonstrates that small changes to
-the rates at which new atoms enter the simulation (\(\Delta\)) can have a
-strong effect on the timing and frequency with which new EOM minima are
-discovered and referenced.  The ParSplice caches on each node are trimmed when
-memory pressure reaches a threshold but the number of unique keys accessed
-(black points along bottom of both graphs) in
-Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
-should be sufficient.
+In this paper, we design cache management policies for the ParSplice worker
+node, driven by a detailed analysis of the application's key-value accesses
+over the course of a long running simulation across a variety of initial
+conditions.  This type of analysis helps the system scale by (1) showing the
+capacity and resource requirements of a single node and (2) informing the
+design of load balancing policies for when we switch to a distributed key-value
+store back-end to store EOM minima.  For example,
+Figure~\ref{fig:motivation-regimes} shows the cache activity of a ParSplice
+node and demonstrates that small changes to the rates at which new atoms enter
+the simulation (\(\Delta\)) can have a strong effect on the timing and
+frequency with which new EOM minima are discovered and referenced.  The
+ParSplice caches on each node are trimmed when memory pressure reaches a
+threshold but the number of unique keys accessed (black points along bottom of
+both graphs) in Figure~\ref{fig:motivation-regimes} suggests that a small cache
+of EOM minima should be sufficient.
 
 % What is Mantle
-We use the Mantle language and policy engine to dynamically explore the effects
-of different cache management strategies for the changing key-value workloads
-generated by ParSplice.  Mantle~\cite{sevilla:sc15-mantle} has already proven
-to be a critical control plane for improving file system metadata load
-balancing  and in this work we show that the collection of abstractions
-provided by Mantle is useful for reasoning about and designing different cache
-management strategies.  Developers write policies for ``when" they want data
-moved and ``how much" of the data to move, then the framework executes these
-policies whenever a decision needs to be made.  These abstraction help
-developers unfamiliar with the domain quickly reason about, develop, and deploy
-new policies that control temporal and spatial locality. We show that Mantle:
+We use the data management language and policy engine from the Mantle
+paper~\cite{sevilla:sc15-mantle} to dynamically explore the effects of
+different cache management strategies for the changing key-value workloads
+generated by ParSplice.  Mantle was originally labeled a programmable file
+system metadata load balancer, but we realize now that the collection of
+abstractions was a control plane that improved metadata access. So in this
+paper we refer to Mantle as a policy engine that injects policies written in
+our data management language directly into a running service ({\it e.g.} file
+system, key-value store, application).  We show how this approach is useful for
+reasoning about and designing different cache management strategies.
+Developers write policies for ``when" they want data moved and ``how much" of
+the data to move, then the framework executes these policies whenever a
+decision needs to be made.  These abstraction help developers unfamiliar with
+the domain quickly reason about, develop, and deploy new policies that control
+temporal and spatial locality. We show that Mantle:
 

### 2017-11-09 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+\begin{figure}[t]
+\noindent\includegraphics[width=0.5\textwidth]{figures/parsplice.png}\\
+\caption{The ParSplice architecture has a storage hierarchy of caches and a
+dedicated cache node (purple boxes) backed by a persistent database (orange). A
+splicer (S) tells worker nodes (W) to generate segments and workers employ
+tasks (T) for more parallelization. We focus on the worker node's cache circled
+in red, which facilitates communication and segment exchange between the worker
+and its \label{fig:parsplice}} \end{figure}
+
 ParSplice~\cite{perez:jctc20150parsplice} is an accelerated molecular dynamics
@@
 Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
-Materials for the Future initiative. Its phases are:
+Materials for the Future initiative. As shown in Figure~\ref{fig:parsplice},
+the phases are:
 
@@
 
-All experiments ran on Trinitite, a Cray XC40 with 32 Intel Haswell 2.3GHz
-cores per node.  Each node has 128GB of RAM and our goal is to limit the size
-of the cache to 3\% of RAM\footnote{Empirically, we find this threshold works
-well for most applications}. Note that this is an addition to the 30GB that
-ParSplice uses to manage other ranks on the same node.  A single Cray node
-produced trajectories that are \(5\times\) times longer than our 10 node
-CloudLab clusters and \(25\times\) longer than UCSC's 10 node cluster. As a
-result, it reaches different job phases faster and gives us a more
-comprehensive view of the workload. The performance gains compared to the
-commodity clusters have more to do with memory/PCI bandwidth than network.
+\subsubsection*{Experimental Setup} All experiments ran on Trinitite, a Cray
+XC40 with 32 Intel Haswell 2.3GHz cores per node.  Each node has 128GB of RAM
+and our goal is to limit the size of the cache to 3\% of
+RAM\footnote{Empirically, we find this threshold works well for most
+applications}. Note that this is an addition to the 30GB that ParSplice uses to
+manage other ranks on the same node.  A single Cray node produced trajectories
+that are \(5\times\) times longer than our 10 node CloudLab clusters and
+\(25\times\) longer than UCSC's 10 node cluster. As a result, it reaches
+different job phases faster and gives us a more comprehensive view of the
+workload. The performance gains compared to the commodity clusters have more to
+do with memory/PCI bandwidth than network.
+
+The scalability experiment uses 1 splicer, persistent database node, and
+in-memory database node and up to 2 workers. We scale up to 1024 tasks, which
+spans 32 nodes and disable hyper-threading because we experience variability in
+performance. For the rest of the experiments, we use 4 nodes, 1 splicer,
+persistent database node, in-memory database node, worker, and up to 512 tasks.
+The keyspace analysis is for the cache on the worker node, which is circled in
+red in Figure~\ref{fig:parsplice}.
 
@@
 state, which is computationally cheap (this motivates
-Section~\S\ref{sec:static-load-balancing}).
+Section~\S\ref{sec:cache-management-using-system-archiecture-knowledge}).
 
@@
 coordinates later in the run (this motivates
-Section~\S\ref{sec:the-need-for-dynamic-load-balancing-policies}).  The growth
-rate, temperature, and number of workers have a predictable effect on the
-structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that the
-number of reads changes with different growth rates, but that the spatial
+Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).  The
+growth rate, temperature, and number of workers have a predictable effect on
+the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
+the number of reads changes with different growth rates, but that the spatial
 locality of key accesses is similar ({\it e.g.}, some keys are still
@@
 structure means that we can learn the regimes and adapt the storage system
-(this motivates Section~\S\ref{sec:ml-for-the-keyspace}).
+(this motivates Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).

### 2017-11-09 paper/sections/3_methodology.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-09-27 paper/sections/3_methodology.tex
@@
-\section{Mantle Engine: Methodology}
+\section{Methodology}
 
-\subsection{Experimental Setup}
+%  \item background: CephFS implementation
+To explore dynamic cache management policies that change during the run, we use
+a data management language and policy engine presented
+in~\cite{sevilla:sc15-mantle}. The prototype in that paper was called Mantle
+and it was originally built on the Ceph file system so administrators could
+control file system metadata load balancing policies. We now refer to Mantle as
+a policy engine that supports our data management language.  The basic premise
+is that load balancing policies can be expressed with a simple API consisting
+of ``when", ``where", and ``how much". ``when" controls how aggressive or
+conservative the decisions are; ``where" lets the policy control how
+distributed or concentrated the data should be; and ``how much" controls the
+amount of data that should be sent.
 
-\subsection{Mantle: Dynamic Load Balancing Policies}
+The succinctness of the API lets users inject multiple, dynamic policies. In
+this work we focus on a single node, so the ``where" policy is not used. When
+we move ParSplice to a distributed key-value store back-end, the ``where" policy
+can be used determine which key-value pairs should be moved to which node.
 
-\begin{itemize}
-  \item motivation: Mochi load balancer microservice
-  \item background: CephFS implementation
-  \item library architecture, callbacks, environment
-\end{itemize}
+\subsection{Extracting Mantle as a Library}
+
+%  \item motivation: Mochi load balancer microservice
+%  \item library architecture, callbacks, environment
+
+We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
+linked into a service ({\it e.g.}, file system, key-value store, application).
+Administrators write policies from different domains, like statistics, machine
+learning, or storage systems, that that are embedded into the runtime. When the
+service makes decisions it executes the administrator-defined policies for
+when/where/how much and returns a decision. To do this, the service needs to be
+modified to (1) provide an environment of metrics and (2) identify where
+policies are set. These modification points are shown by the colored boxes in
+Figure~\ref{fig:mantle} and described below.
+
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/mantle.png}\\
+
+  \caption{Extracting Mantle as library.\label{fig:mantle}}
+
+\end{figure}
+
+\begin{table}
+  \centering
+  \begin{tabular}{ r | l | l }
+  Metrics     & Data Structure & Description \\\hline
+  Cluster     & \{server \(\rightarrow\) \{metric \(\rightarrow\) val\}\}
+              & resource util. for servers \\
+  Time Series & [(ts, val), ..., (ts, val)]
+              & accesses by timestamp (ts) \\
+  && \\
+              & Service      & Example \\\hline
+  Cluster     & File Systems & CPU util., Inode reads \\
+              & ParSplice    & CPU util., Cache size \\
+  Time Series & File Systems & Accesses to directory \\
+              & ParSplice    & Accesses to key in DB\\
+  \end{tabular}
+  \caption{Types of metrics exposed by the service to the policy engine using Mantle.\label{table:metrics}}
+\end{table}
+
+% policies themselves
+\subsubsection{Policies Written as Callbacks} administrators write policies in
+Lua that are executed whenever the service needs to make a data management
+decision. We chose Lua for simplicity, performance, and portability; it is a
+scripting language with simple syntax, which allows administrators to focus on
+the policies themselves; it was designed as an embeddable language, so it is
+lightweight and does less type checking; and it interfaces nicely with C/C++.
+The pink box in Figure~\ref{fig:mantle} shows an example policy for the
+``when()" callback in a distributed service, where the current server
+(\texttt{whoami}) migrates load if it is has load (\texttt{>0.1}) and if its
+neighbor server (\texttt{whoami + 1}) does not have load (\texttt{<0.1}). The
+load is calculated using the metrics provided by the environment.
+
+%of timestamp (ts), value pairs representing accesses over time types of
+\subsubsection{Environment of Metrics} services can expose cluster metrics for
+describing resource utilization and time series metrics for describing accesses
+to some data structure over time. Table~\ref{table:metrics} shows how these
+metrics are accessed from the policies written by administrators.
+
+For cluster metrics, the service passes a dictionary to Mantle, indexed by
+server name and metric. Policies access the cluster metric values by indexing
+into a Lua table using \texttt{server} and \texttt{metric}, where server is a
+node identifier ({e.g.}, MPI Rank, metadata server name) and metric is a
+resource name.  For example, a policy written for an MPI-based service can
+access the CPU utilization of the first rank in a communication group using:
+
+\begin{figure}[h]
+\footnotesize
+\begin{minted}{lua}
+load = servers[0]['cpu']
+\end{minted}
+\end{figure}
+
+Other examples metrics used for file system metadata load balancing, such as
+memory pressure, are shown in Figure~\ref{fig:mantle}. The measurements and
+exchange of metrics between servers is done by the service; Mantle in CephFS
+leverages metrics from other servers collected using CephFS's heartbeats.
+
+For time series metrics, the service passes array of \texttt{(timestamp,
+value)} pairs to Mantle and the policies can iterate over the accesses. For
+example, a policy that uses accesses to a certain directory to make migration
+decisions collects that information using:
+
+\begin{figure}[h]
+\footnotesize
+\begin{minted}{lua}
+ts = array()
+for i=1,arraysize() do
+  for time,value in string.gmatch(ts, (%w+=%w)) do
+    if value == 'mydirectory' then
+      count = count + 1
+    end
+  end
+end
+\end{minted}
+\end{figure}
+
+The service uses a pointer to the time series to facilitate time series with
+many values, like accesses to a database or directory in the file system
+namespace. This decision limits the time series metrics to only include values
+from the {\it current } node, although this is not a limitation of Mantle
+itself.
 
@@
 
-\begin{itemize}
-  \item providing environment of metrics
-  \item identifying where policies are made
-\end{itemize}
+% env of metrics
+Using Mantle cluster metrics, we expose cache size, CPU utilization, memory
+pressure of the worker node to the cache management policies. In
+Section~\S\ref{sec:parsplice-keyspace-analysis} we only end up using the cache
+size although the other metrics proved to be valuable debugging tools. Using
+Mantle time series metrics, we expose accesses to the cache as a list of
+\texttt{timestamp, keyid} pairs. In
+Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}, we
+explore a regime detection algorithm that uses this metric.
+
+% where policies are made
+We link Mantle directly into the worker cache circled in red in
+Figure~\ref{fig:parsplice}. We put the ``when" and ``how much" callbacks
+alongside code that checks for memory pressure. It is executed right before the
+worker processes incoming and outgoing put/get transactions to the cache. As
+stated previously, we do not use the ``where" part of Mantle because we focus
+on a single node, but this part of the API will be used when we move the caches
+and storage nodes to a key-values store back-end that has load balancing and key
+redistribution.

### 2017-11-09 paper/sections/4_dynamic-policies.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 of each strategy; the strategy should be accurate enough so as to sacrifice
-negligible performance and fast tenough to run as often as we want to detect
+negligible performance and fast enough to run as often as we want to detect
 regimes.
@@
 ``system", we mean the hardware and software of the storage hierarchy. We look
-at request rate, unique keys in a sliding window, and bandwidth capabilties. For
+at request rate, unique keys in a sliding window, and bandwidth capabilities. For
 example, we know that LevelDB cannot handle high IO request rates.
@@
   \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
-
   \caption{The performance and resource utilization trade-off for different
@@
   size of the cache to save memory.  \label{fig:methodology-tradeoff}}
-
 \end{figure}
@@
 
-% what is mantle
-To explore dynamic load balancing policies ({\it i.e.} policies that change
-during the run), we use the Mantle approach.  Mantle is a framework built on the
-Ceph file system that lets administrators control file system metadata load
-balancing policies. The basic premise is that load balancing policies can be
-expressed with a simple API consisting of ``when", ``where", and ``how much".
-The succinctness of the API lets users inject multiple, dynamic policies.
-
 % Why is this a good idea
@@
 
-Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using the
-Mantle API to program a dynamic load balancing policy into ParSplice that
-switches between two policies:
+Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using Mantle
+to program a dynamic load balancing policy into ParSplice that switches between
+two policies:
 
@@
   \caption{Key activity for a 4 hour run shows groups of accesses to the same
-  subset of keys. Detecting this ``access regimes" leads to a more accurate cache
+  subset of keys. Detecting these ``access regimes" leads to a more accurate cache
   management strategy.\label{fig:keyspace-zoomed}}

### 2017-11-09 paper/sections/5_file-systems.tex
paper: integrate Carlos feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{Relation to File Systems}
+\label{sec:relation-to-file-systems}
 
@@
 to manage its caches and CephFS uses policies to control load balancing, but
-they both can be expressed with the Mantle API.  From a high-level the
-ParSplice policy trims the cache if the cache reaches a certain size {\it and}
-if it has already absorbed the initial burstiness of the workload; the CephFS
-policy migrates load if the metadata load is higher than the average load {\it
-and} the current load has been overloaded for more than two iterations.
+they both can be expressed with Mantle.  From a high-level the ParSplice policy
+trims the cache if the cache reaches a certain size {\it and} if it has already
+absorbed the initial burstiness of the workload; the CephFS policy migrates
+load if the metadata load is higher than the average load {\it and} the current
+load has been overloaded for more than two iterations.
 
@@
 \textbf{State Persisted Across Decisions} (Fig.~\ref{lua:parsplice}:Lines 5,6;
-Fig~\ref{lua:cephfs}:Lines 3,4,10) - these lines use the Mantle API to
-write/read state from previous decisions.  For ParSplice, we save a boolean
-that indicates whether we have absorbed the workload's initial burstiness. For
-CephFS, we save the number of consecutive instances that the server has been
-overloaded. We also clear the count (Line 10) if the server is no longer
-overloaded. The underlying implementation saves the values to local disk.
+Fig~\ref{lua:cephfs}:Lines 3,4,10) - these lines use Mantle to write/read state
+from previous decisions.  For ParSplice, we save a boolean that indicates
+whether we have absorbed the workload's initial burstiness. For CephFS, we save
+the number of consecutive instances that the server has been overloaded. We
+also clear the count (Line 10) if the server is no longer overloaded. The
+underlying implementation saves the values to local disk.
 
@@
 \begin{figure}[tb]
+\footnotesize
 \begin{minted}[linenos]{lua}
@@
 \end{minted}
-\caption{CephFS file system metadta load balancer.\label{lua:cephfs}}
+\caption{CephFS file system metadata load balancer.\label{lua:cephfs}}
 \end{figure}

### 2017-11-09 paper/sections/6_conclusion.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{Conclusion}
+
+Data management encompasses a wide range of techniques that vary by domain and
+service. Yet, the techniques require policies that shape the decision making
+and finding the best policies is a difficult, multi-dimensional problem. We
+observe that many of the primitives and resulting strategies have enough in
+common that they can be expressed with similar semantics. We present a data
+management language and policy engine, called Mantle that is general enough to
+express complicated, dynamic policies for two different domains and services.
+Rather than attempting to construct a single, complex load balancing policy
+that works for a variety of scenarios, we instead use the Mantle framework to
+enable software-defined storage systems to flexibly change policies as the
+workload changes over time.  In our analysis of the ParSplice key-value
+workload we have detected clear workload regimes that are sensitive to the
+initial conditions and the scale and duration of the simulation. We have also
+demonstrated that changing load balancing policies at runtime in response to
+the current workload is an effective mechanism to providing better load
+distribution.  Finally, we have demonstrated that Mantle is flexible enough to
+support domain-specific knowledge, which lays the groundwork for future work in
+adaptable policies and policy generation.

### 2017-11-09 paper/sections/1_introduction.tex
paper: 1st draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 abstractions effectively decomposes the problem into manageable policies for a
-different domain and service (in this case, cache management).  Our exploration
-of this space results in a two policy scheme that achieves 96\% efficiency
-while using only 7.6\% of the memory resources required by the base case.
+different domain and service.  Our exploration of this space results in a two
+policy scheme that achieves 96\% efficiency while using only 7.6\% of the
+memory resources required by the base case.
 
@@
 ParSplice~\cite{perez:jctc20150parsplice} molecular dynamics simulation, a
-hierarchy of cache nodes and a single node key-value store are used to store
+hierarchy of caches and a single persistent key-value store are used to store
 both observed minima across a molecule's equation of motion (EOM) and the
@@
 
-In this paper, we design cache management policies for the ParSplice worker
-node, driven by a detailed analysis of the application's key-value accesses
-over the course of a long running simulation across a variety of initial
-conditions.  This type of analysis helps the system scale by (1) showing the
-capacity and resource requirements of a single node and (2) informing the
-design of load balancing policies for when we switch to a distributed key-value
-store back-end to store EOM minima.  For example,
-Figure~\ref{fig:motivation-regimes} shows the cache activity of a ParSplice
-node and demonstrates that small changes to the rates at which new atoms enter
-the simulation (\(\Delta\)) can have a strong effect on the timing and
+In this paper, we design cache management policies for the ParSplice worker,
+driven by a detailed analysis of the key-value accesses over the course of a
+long running simulation across a variety of initial conditions.
+Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
+pairs from the worker's cache for two different initial conditions of
+\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
+the read request rate (\(y1\) axis) and the dots along the bottom are the
+number of unique keys accessed (\(y2\) axis).  This figure demonstrates that
+small changes to \(\Delta\) can have a strong effect on the timing and
 frequency with which new EOM minima are discovered and referenced.  The
 ParSplice caches on each node are trimmed when memory pressure reaches a
-threshold but the number of unique keys accessed (black points along bottom of
-both graphs) in Figure~\ref{fig:motivation-regimes} suggests that a small cache
-of EOM minima should be sufficient.
+threshold but the number of unique keys accessed in
+Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
+should be sufficient.
 
@@
 different cache management strategies for the changing key-value workloads
-generated by ParSplice.  Mantle was originally labeled a programmable file
+generated by ParSplice.  Mantle was originally touted as a programmable file
 system metadata load balancer, but we realize now that the collection of
-abstractions was a control plane that improved metadata access. So in this
-paper we refer to Mantle as a policy engine that injects policies written in
-our data management language directly into a running service ({\it e.g.} file
-system, key-value store, application).  We show how this approach is useful for
-reasoning about and designing different cache management strategies.
-Developers write policies for ``when" they want data moved and ``how much" of
-the data to move, then the framework executes these policies whenever a
-decision needs to be made.  These abstraction help developers unfamiliar with
-the domain quickly reason about, develop, and deploy new policies that control
-temporal and spatial locality. We show that Mantle:
+abstractions designed for file systems was a control plane that improved
+metadata access. So in this paper we refer to Mantle as a policy engine that
+injects policies written in our data management language directly into a
+running service ({\it e.g.} file system, key-value store, application) and show
+how this approach is useful for reasoning about and designing different cache
+management strategies in ParSplice. Developers write policies for ``when" they
+want data moved and ``how much" of the data to move, then the framework
+executes these policies whenever a decision needs to be made.  These
+abstraction help developers unfamiliar with the domain quickly reason about,
+develop, and deploy new policies that control temporal and spatial locality. We
+show that Mantle:
 
@@
 a range of policies from different disciplines; but more importantly, in
-Section~\S\ref{sec:relation-to-file-systems}, we conclude that the collection
+Section~\S\ref{sec:relation-to-load-balancing}, we conclude that the collection
 of policies we designed for ParSplice's cache management are very similar to
-the policies in the Ceph file system that are used to load balance metadata,
+the policies used to load balance metadata in the Ceph file system (CephFS)
 suggesting that there is potential for automatically adapting and generating

### 2017-11-09 paper/sections/2_parsplice-keyspace-analysis.tex
paper: 1st draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \caption{The ParSplice architecture has a storage hierarchy of caches and a
-dedicated cache node (purple boxes) backed by a persistent database (orange). A
-splicer (S) tells worker nodes (W) to generate segments and workers employ
-tasks (T) for more parallelization. We focus on the worker node's cache circled
-in red, which facilitates communication and segment exchange between the worker
-and its \label{fig:parsplice}} \end{figure}
+dedicated cache node (boxes) backed by a persistent database (DB). A splicer
+(S) tells worker (W) to generate segments and workers employ tasks (T) for more
+parallelization. We focus on the worker's cache (circled), which facilitates
+communication and segment exchange between the worker and its
+\label{fig:parsplice}}
+\end{figure}
 
@@
 
-The computation can be parallelized by adding more workers or by adding worker
-tasks to parallelize individual workers.  The workers are stateless and read
-initial coordinates from the data store each time they begin generating
-segments. Since worker tasks do not maintain their own history, they can end up
-reading the same coordinates repeatedly. To mitigate the consequences of these
-repeated reads, ParSplice provisions a hierarchy of nodes to act as caches that
-sit in front of a single node persistent database.  Values are written to each
-tier and reads traverse up the hierarchy until they find the data.
+The computation can be parallelized by adding more workers or by adding tasks
+to parallelize individual workers.  The workers are stateless and read initial
+coordinates from the data store each time they begin generating segments. Since
+worker tasks do not maintain their own history, they can end up reading the
+same coordinates repeatedly. To mitigate the consequences of these repeated
+reads, ParSplice provisions a hierarchy of processes to act as caches that sit
+in front of a single persistent database.  Values are written to each tier and
+reads traverse up the hierarchy until they find the data.  Caches also reside
+on the workers to service reads/writes from its tasks.
 
@@
 counters have high overhead we only turn them on for the keyspace analysis.
-The cache hierarchy is unmodified but for the persistent database node, we
-replace BerkeleyDB on NFS with LevelDB on Lustre. Original ParSplice
-experiments showed that BerkeleyDB's syncs caused reads/writes to bottleneck on
-the persistent database. We also use Riak's customized
-LevelDB\footnote{https://github.com/basho/leveldb} version, which comes
-instrumented with its own set of performance counters.
 
@@
 that are \(5\times\) times longer than our 10 node CloudLab clusters and
-\(25\times\) longer than UCSC's 10 node cluster. As a result, it reaches
+\(25\times\) longer than our 10 node cluster at UCSC. As a result, it reaches
 different job phases faster and gives us a more comprehensive view of the
@@
 
-The scalability experiment uses 1 splicer, persistent database node, and
-in-memory database node and up to 2 workers. We scale up to 1024 tasks, which
-spans 32 nodes and disable hyper-threading because we experience variability in
-performance. For the rest of the experiments, we use 4 nodes, 1 splicer,
-persistent database node, in-memory database node, worker, and up to 512 tasks.
-The keyspace analysis is for the cache on the worker node, which is circled in
-red in Figure~\ref{fig:parsplice}.
+The scalability experiment uses 1 splicer, 1 persistent database, 1 cache
+process, and up to 2 workers. We scale up to 1024 tasks, which spans 32 nodes
+and disable hyper-threading because we experience unacceptable variability in
+performance. For the rest of the experiments, we use 4 nodes, 1 splicer, 1
+persistent database, 1 cache process, 1 worker, and up to 512 tasks.  The
+keyspace analysis that follows is for the cache on the worker node, which is
+circled in Figure~\ref{fig:parsplice}.  The cache hierarchy is unmodified but
+for the persistent database node, we replace BerkeleyDB on NFS with LevelDB on
+Lustre. Original ParSplice experiments showed that BerkeleyDB's syncs caused
+reads/writes to bottleneck on the persistent database. We also use Riak's
+customized LevelDB\footnote{https://github.com/basho/leveldb} version, which
+comes instrumented with its own set of performance counters.
 
@@
 \subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
-keyspace size (black annotations) and request load (bars) after a one hour run
+keyspace size (text annotations) and request load (bars) after a one hour run
 with a different number of workers (\(x\) axis). While the keyspace size and
@@
 the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
-the number of reads changes with different growth rates, but that the spatial
-locality of key accesses is similar ({\it e.g.}, some keys are still
-\(5\times\) more popular than others).  Figure~\ref{fig:motivation-regimes}
-shows how entropy for different growth rates has temporal locality, as the
-reads per second for \(\Delta_2\) looks like the reads per second for
-\(\Delta_1\) stretched out along the time axis.  Trends also exist for
-temperature and number of workers but are omitted here for space. This
-structure means that we can learn the regimes and adapt the storage system
-(this motivates Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).
+the number of reads changes with different initial conditions (\(Delta\)), but
+that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
+still \(5\times\) more popular than others).
+Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
+rates has temporal locality, as the reads per second for \(\Delta_2\) looks
+like the reads per second for \(\Delta_1\) stretched out along the time axis.
+Trends also exist for temperature and number of workers but are omitted here
+for space. This structure means that we can learn the regimes and adapt the
+storage system (this motivates
+Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).

### 2017-11-09 paper/sections/3_methodology.tex
paper: 1st draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 To explore dynamic cache management policies that change during the run, we use
-a data management language and policy engine presented
+the data management language and policy engine presented
 in~\cite{sevilla:sc15-mantle}. The prototype in that paper was called Mantle
-and it was originally built on the Ceph file system so administrators could
-control file system metadata load balancing policies. We now refer to Mantle as
-a policy engine that supports our data management language.  The basic premise
-is that load balancing policies can be expressed with a simple API consisting
-of ``when", ``where", and ``how much". ``when" controls how aggressive or
+and it was originally built on the CephFS so administrators could control file
+system metadata data management policies. We now refer to Mantle as a policy
+engine that supports our data management language.  The basic premise is that
+load balancing policies can be expressed with a simple API consisting of
+``when", ``where", and ``how much". ``when" controls how aggressive or
 conservative the decisions are; ``where" lets the policy control how
@@
 We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
-linked into a service ({\it e.g.}, file system, key-value store, application).
-Administrators write policies from different domains, like statistics, machine
-learning, or storage systems, that that are embedded into the runtime. When the
-service makes decisions it executes the administrator-defined policies for
-when/where/how much and returns a decision. To do this, the service needs to be
-modified to (1) provide an environment of metrics and (2) identify where
-policies are set. These modification points are shown by the colored boxes in
-Figure~\ref{fig:mantle} and described below.
+linked into a service.  Administrators write policies from whatever domain they
+choose ({e.g.}, statistics, machine learning, storage system) and the policies
+are embedded into the runtime by Mantle. When the service makes decisions it
+executes the administrator-defined policies for when/where/how much and returns
+a decision.  To do this, the service needs to be modified to (1) provide an
+environment of metrics and (2) identify where policies are set. These
+modification points are shown by the colored boxes in Figure~\ref{fig:mantle}
+and described below.
 
@@
 lightweight and does less type checking; and it interfaces nicely with C/C++.
-The pink box in Figure~\ref{fig:mantle} shows an example policy for the
+The ``callback" box in Figure~\ref{fig:mantle} shows an example policy for the
 ``when()" callback in a distributed service, where the current server
@@
 
-Other examples metrics used for file system metadata load balancing, such as
-memory pressure, are shown in Figure~\ref{fig:mantle}. The measurements and
+Other examples metrics used for file system metadata load balancing are shown
+by the ``environment" box in Figure~\ref{fig:mantle}. The measurements and
 exchange of metrics between servers is done by the service; Mantle in CephFS
@@
 
-For time series metrics, the service passes array of \texttt{(timestamp,
+For time series metrics, the service passes an array of \texttt{(timestamp,
 value)} pairs to Mantle and the policies can iterate over the accesses. For
-example, a policy that uses accesses to a certain directory to make migration
-decisions collects that information using:
+example, a policy that uses accesses to a directory in a file system as a
+metric for load collects that information using:
 
@@
 % env of metrics
-Using Mantle cluster metrics, we expose cache size, CPU utilization, memory
+Using Mantle cluster metrics, we expose cache size, CPU utilization, and memory
 pressure of the worker node to the cache management policies. In
@@
 on a single node, but this part of the API will be used when we move the caches
-and storage nodes to a key-values store back-end that has load balancing and key
-redistribution.
+and storage nodes to a key-values store back-end that uses key load balancing and
+repartitioning.

### 2017-11-09 paper/sections/5_file-systems.tex
paper: 1st draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\section{Relation to File Systems}
-\label{sec:relation-to-file-systems}
+\section{Relation to Load Balancing}
+\label{sec:relation-to-load-balancing}
 
@@
 to manage its caches and CephFS uses policies to control load balancing, but
-they both can be expressed with Mantle.  From a high-level the ParSplice policy
-trims the cache if the cache reaches a certain size {\it and} if it has already
-absorbed the initial burstiness of the workload; the CephFS policy migrates
-load if the metadata load is higher than the average load {\it and} the current
-load has been overloaded for more than two iterations.
+they both can be expressed with our data management language.  From a
+high-level the ParSplice policy trims the cache if the cache reaches a certain
+size {\it and} if it has already absorbed the initial burstiness of the
+workload; the CephFS policy migrates load if the metadata load is higher than
+the average load {\it and} the current load has been overloaded for more than
+two iterations.
 

### 2017-11-09 paper/sections/3_methodology.tex
paper: addresses #10
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+Mantle also provides functions for persisting state across decisions.
+\texttt{WRState(s)} saves state \texttt{s}, which can be a number or boolean
+value, and \texttt{RDState()} returns the state saved by a previous iteration.
+For example, a ``when" policy can avoid trimming a cache or migrating data if
+it had performed that operation in the previous decision using:
+
+\begin{figure}[h]
+\footnotesize
+\begin{minted}{lua}
+if RDState() == 1 then
+  WRState(0)   -- the next decision will return true
+  return false
+else then
+  WRState(1)   -- the next decision will return false
+  return true
+end
+\end{minted}
+\end{figure}
+
 %of timestamp (ts), value pairs representing accesses over time types of

### 2017-11-09 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #9
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-%\section{ParSplice Keyspace Analysis}
 %
@@
 We use ParSplice to simulate the evolution of metallic nanoparticles that grow
-from the vapor phase. As the run progresses, the energy landscape of the system
-becomes more complex. Two domain factors control the number of entries in the
-data store: the growth rate and the temperature. The growth rate controls how
-quickly new atoms are added to the nanoparticle: fast growth rates lead to
-non-equilibrium conditions, and hence increase the number of states that can be
-visited. However, as the particle grows, the simulation slows down because the
-calculations become more expensive, limiting the rate at which new states are
-visited. On the other hand, the temperature controls how easily a trajectory
-can jump from state to state; higher temperatures lead to more frequent
-transitions.
+from the vapor phase.  This simulation stresses the storage hierarchy more than
+other input decks because it uses a cheap potential, has a small number of
+atoms, and operates in a complex energy landscape with many accessible states.
+As the run progresses, the energy landscape of the system becomes more complex
+and more states are visited.  Two domain factors control the number of entries
+in the data store: the growth rate and the temperature. The growth rate
+controls how quickly new atoms are added to the nanoparticle: fast growth rates
+lead to non-equilibrium conditions, and hence increase the number of states
+that can be visited.  However, as the particle grows, the simulation slows down
+because the calculations become more expensive, limiting the rate at which new
+states are visited.  On the other hand, the temperature controls how easily a
+trajectory can jump from state to state; higher temperatures lead to more
+frequent transitions but temperatures that are too high result in meaningless
+simulations because trajectories have so much energy that they are equally
+likely to visit any random state.
 
-The nanoparticle simulation stresses the data store architecture of ParSplice.
-It visits more states than other input decks because the system uses a cheap
-potential, has a small number of atoms, and operates in a complex energy
-landscape with many accessible states. Changing growth rates and temperature
-alters the size, shape, and locality of the data store keyspace. Lower
-temperatures and smaller growth rates create hotter keys with smaller keyspaces
-as many segments are generated in the same set of states before the trajectory
-can escape to a new region of state space.
+Changing growth rates and temperature alters the size, shape, and locality of
+the data store keyspace. Lower temperatures and smaller growth rates create
+hotter keys with smaller keyspaces as many segments are generated in the same
+set of states before the trajectory can escape to a new region of state space.
+In Figure~\ref{fig:motivation-regimes}, the \(\Delta_1\) growth rate adds atoms
+every 100K microseconds while the \(\Delta_2\) growth rate adds atoms every 1
+million microseconds. So \(\Delta_2\) has a smaller growth rate resulting in
+hotter keys (line on \(y1\) axis) and a smaller keyspace (dots on the \(y2\)
+axis).  Values smaller than \(\Delta_2\)'s growth rate or a temperature of 400
+degrees results in very little database activity because state transitions take
+too long. Similarly, values larger than \(\Delta_1\)'s growth rate or a
+temperature of 4000 degrees result in an equally meaningless simulation as
+transitions are unrealistic.
 
@@
 the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
-the number of reads changes with different initial conditions (\(Delta\)), but
+the number of reads changes with different initial conditions (\(\Delta\)), but
 that the spatial locality of key accesses is similar ({\it e.g.}, some keys are

### 2017-11-10 paper/sections/4_dynamic-policies.tex
paper: addresses #9
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #11
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 negligible performance and fast enough to run as often as we want to detect
-regimes.
+regimes. The goal of the following sections is not to find an optimal solution,
+as this can be done with parameter sweeps for thresholds; rather, we try to
+find techniques that work for a range of inputs and system setups.
 
@@
 utilization is measured with memory footprint of just the cache (\(y2\) axis).
-The other graph shares the \(y\) axis and shows the trade-off of constraining
-the cache to different sizes.  The error bars are the standard deviation of 3
-runs.
+The other graph shares the \(y\) axis and shows the trade-off of using a basic
+LRU-style cache for different cache sizes, implemented using the ``when" and
+``how much" policies in Figure~\ref{src:lru}.  The error bars are the standard
+deviation of 3 runs.
+
+\begin{figure}[t]
+\footnotesize
+\centering
+\begin{minted}[xleftmargin=3em,linenos]{lua}
+function when()
+  if server[whoami]['cachesize'] > n then
+    return true
+  end
+  return false
+end
+
+function howmuch()
+  if servers[whoami]['cachesize'] > n
+    return servers[whoami]['cachesize'] - n
+  end
+  return 0
+end
+\end{minted}
+\caption{Policy that implements a basic LRU cache. The performance and memory
+utilization for different values of \(n\) are graphed in
+Figure~\ref{fig:methodology-tradeoff}.
+\label{src:lru}}
+\end{figure}
 
@@
 \end{itemize}
+
+The actual policy is shown in Figure~\ref{src:lru-dyn} where the only
+difference between the code snippet in Figure~\ref{src:lru} is the addition of
+lines 2-9. Note that policy also uses the \texttt{RDState} and \texttt{WRState}
+functions described in Section~\S\ref{sec:extracting}.
+
+\begin{figure}[t]
+\footnotesize
+\centering
+\begin{minted}[xleftmargin=3em,linenos]{lua}
+function when()
+  policy = RDState()
+  if policy == 0
+    if servers[whoami]['cachesize'] > 100K then
+      WRState(1)
+      policy = 1
+    end
+  end
+  if policy == 1
+    if servers[whoami]['cachesize'] > n then
+      return true
+    end
+  end
+  return false
+end
+\end{minted}
+\caption{A more complicated LRU-based cache policy that has two parts. After
+absorbing the initial burstiness of the workload, the policy switches to a more
+constrained cache to limit memory usage. The performance/utilization for
+different values of \texttt{n} is shown in
+Figure~\ref{fig:methodology-tradeoff-dynamic}. \label{src:lru-dyn}}
+\end{figure}
 

### 2017-11-10 paper/sections/5_file-systems.tex
paper: addresses #11
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\begin{figure}[tb]
+\begin{figure}[t]
 \footnotesize
-\begin{minted}[linenos]{lua}
+\begin{minted}[xleftmargin=3em,linenos]{lua}
 local function when()

### 2017-11-10 paper/sections/1_introduction.tex
paper: addressed #8
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 injects policies written in our data management language directly into a
-running service ({\it e.g.} file system, key-value store, application) and show
-how this approach is useful for reasoning about and designing different cache
-management strategies in ParSplice. Developers write policies for ``when" they
-want data moved and ``how much" of the data to move, then the framework
-executes these policies whenever a decision needs to be made.  These
-abstraction help developers unfamiliar with the domain quickly reason about,
-develop, and deploy new policies that control temporal and spatial locality. We
-show that Mantle:
+running service and show how this approach is useful for reasoning about and
+designing different cache management strategies in ParSplice.  By service, we
+mean a system that manages data and responds requests, such as a file system or
+key-value store.  Developers write policies for ``when" they want data moved
+and ``how much" of the data to move, then the framework executes these policies
+whenever a decision needs to be made.  These abstraction help developers
+unfamiliar with the domain quickly reason about, develop, and deploy new
+policies that control temporal and spatial locality. We show that Mantle:
 

### 2017-11-10 paper/sections/4_dynamic-policies.tex
paper: addressed #8
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-% results: raw numbers
+% why is the performance lower for smaller caches?
 Although the keyspace grows to 150K, a 100K key cache achieves 99\% of the
 performance. Decreasing the cache degrades performance and predictability.
-While this result is not unexpected, it nonetheless achieves our goal of
-showing the benefits of load balancing keys across nodes and that smaller
-caches on each node are an effective way to save memory without completely
-sacrificing performance.
-
-% why is the performance lower for smaller caches?
-Despite the memory savings, our results suggest that dynamic load balancing
+Despite the memory savings, our results suggest that dynamic cache management
 policies could save even more memory.  Figure~\ref{fig:methodology-tradeoff}
@@
 should suffice after the initial read flash crowd passes because
-the keyspace is far less active. This suggests a two-part load balancing
+the keyspace is far less active. This suggests a two-part cache management
 policy.
@@
   \caption{The performance and resource utilization trade-off of using a
-  dynamic load balancing policy that switches to a constrained cache policy after absorbing
+  dynamic cache management policy that switches to a constrained cache policy after absorbing
   the initial burstiness of the workload. The sizes of these smaller caches are
@@
 Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using Mantle
-to program a dynamic load balancing policy into ParSplice that switches between
+to program a dynamic cache management policy into ParSplice that switches between
 two policies:
@@
 sufficient policies for the 2.5 hour runs but anything longer would need a
-different dynamic load balancing policy.
+different dynamic cache management policy.
 
@@
 % but the result is still valid
-Despite these caveats, the result is still valid: we found a dynamic load
-balancing policy that absorbs the cost of a high read throughput on a small
+Despite these caveats, the result is still valid: we found a dynamic cache 
+management policy that absorbs the cost of a high read throughput on a small
 keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
-show the effectiveness of the load balancing policy engine we integrated into
+show the effectiveness of the policy engine we integrated into
 ParSplice, not that we were able to identify the best policy for all system

### 2017-11-10 paper/sections/4_dynamic-policies.tex
paper: start fixing the file system section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 difference between the code snippet in Figure~\ref{src:lru} is the addition of
-lines 2-9. Note that policy also uses the \texttt{RDState} and \texttt{WRState}
+lines 2-7. Note that policy also uses the \texttt{RDState} and \texttt{WRState}
 functions described in Section~\S\ref{sec:extracting}.
@@
 function when()
-  policy = RDState()
-  if policy == 0
-    if servers[whoami]['cachesize'] > 100K then
-      WRState(1)
-      policy = 1
+  if server[whoami]['cachesize'] > n then
+    if server[whoami]['cachesize'] > 100K then
+      WRstate(1)
     end
-  end
-  if policy == 1
-    if servers[whoami]['cachesize'] > n then
+    if RDstate() == 1 then
       return true

### 2017-11-10 paper/sections/5_file-systems.tex
paper: start fixing the file system section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-The code snippets in Figures~\ref{lua:parsplice} and~\ref{lua:cephfs} are the
-policies used in ParSplice and CephFS, respectively.  ParSplice uses policies
-to manage its caches and CephFS uses policies to control load balancing, but
-they both can be expressed with our data management language.  From a
-high-level the ParSplice policy trims the cache if the cache reaches a certain
-size {\it and} if it has already absorbed the initial burstiness of the
-workload; the CephFS policy migrates load if the metadata load is higher than
-the average load {\it and} the current load has been overloaded for more than
-two iterations.
+In the previous section, we used our data management language and the Mantle
+policy engine to design effective cache management strategies for a new service
+and domain. In this section, we compare and contrast the policies examined for
+file system metadata load balancing in~\cite{sevilla:sc15-mantle} with the ones
+designed for cache management in ParSplice. The similarities show how the
+``when"/``where"/``how much" abstractions, data anagement language, and policy
+engine may be widely applicable to other data management techniques, such as
+QoS, scheduling, and batching.
 
-\textbf{Condition for ``Overloaded"} (Fig.~\ref{lua:parsplice}:Line 2;
-Fig.~\ref{lua:cephfs}:Line 2) - these lines detect whether the node is
+\subsection{File Systems}
+From a high-level the ParSplice policy trims the cache if the cache reaches a
+certain size {\it and} if it has already absorbed the initial burstiness of the
+workload; the CephFS poicy, which is shown in Figure~\ref{srd:lua-cephfs},
+migrates load if the metadata load is higher than the average load {\it and}
+the current load has been overloaded for more than two iterations.
+
+\textbf{Condition for ``Overloaded"} (Fig.~\ref{src:lru-dyn}: Line 2;
+Fig.~\ref{lua:cephfs}: Line 2) - these lines detect whether the node is
 overloaded using the ``load" calculated in the load callback; while the load
@@
 
-\textbf{State Persisted Across Decisions} (Fig.~\ref{lua:parsplice}:Lines 5,6;
-Fig~\ref{lua:cephfs}:Lines 3,4,10) - these lines use Mantle to write/read state
+\textbf{State Persisted Across Decisions} (Fig.~\ref{src:lru-dyn}: Lines 4,6;
+Fig~\ref{lua:cephfs}: Lines 3,4,10) - these lines use Mantle to write/read state
 from previous decisions.  For ParSplice, we save a boolean that indicates
@@
 
-\textbf{Condition that Switches Policy} (Fig.~\ref{lua:parsplice}:Line 3;
-Fig.~\ref{lua:cephfs}:Line 5) - these lines switch the policies using
+\textbf{Condition that Switches Policy} (Fig.~\ref{src:lru-dyn}: Line 6;
+Fig.~\ref{lua:cephfs}: Line 5) - these lines switch the policies using
 information from previous decisions. ParSplice trims its cache once it eclipses
@@
 policy-switching conditions.
-
-\begin{figure}[t]
-\footnotesize
-\begin{minted}[xleftmargin=3em,linenos]{lua}
-local function when()
-  if servers[whoami]["load"] > target then
-    if servers[whoami]["load"] > absorb_target then
-      WRstate(1)
-    end
-    if RDstate() == 1 then
-      return true
-    end
-  end
-  return false
-end
-\end{minted}
-\caption{ParSplice cache management policy.\label{lua:parsplice}}
-\end{figure}
 
@@
 
-\subsection{Using File System Balancers for ParSplice}
-\subsection{Using ParSplice Balancers for File Systems}
-\subsection{Visualizing File System Traces like ParSplice Keyspace Traces}
+\subsection{Using File System Policies for ParSplice}
+% mantle policies
+\subsection{Using ParSplice Policies for File Systems}
+% burstiness of creates then compile
+
+\subsection{Cross-pollinating Policies}
+% QoS, MRC, RAD - scheduing
+%\subsection{Visualizing File System Traces like ParSplice Keyspace Traces}

### 2017-11-13 paper/sections/5_file-systems.tex
paper: hack on generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\section{Relation to Load Balancing}
+\section{General Data Management Policies}
 \label{sec:relation-to-load-balancing}
+
+%\begin{itemize}
+%  \item Lustre Trace
+%  \item LinkedIn Trace
+%  \item Nathan's Trace
+%\end{itemize}
 
@@
 designed for cache management in ParSplice. The similarities show how the
-``when"/``where"/``how much" abstractions, data anagement language, and policy
+``when"/``where"/``how much" abstractions, data management language, and policy
 engine may be widely applicable to other data management techniques, such as
@@
 
-\subsection{File Systems}
+\subsection{Using File System Policies for ParSplice}
+%- memory utilization/heuristics
 From a high-level the ParSplice policy trims the cache if the cache reaches a
 certain size {\it and} if it has already absorbed the initial burstiness of the
-workload; the CephFS poicy, which is shown in Figure~\ref{srd:lua-cephfs},
+workload; the CephFS poicy, which is shown in Figure~\ref{src:lua-cephfs},
 migrates load if the metadata load is higher than the average load {\it and}
@@
 \textbf{Condition for ``Overloaded"} (Fig.~\ref{src:lru-dyn}: Line 2;
-Fig.~\ref{lua:cephfs}: Line 2) - these lines detect whether the node is
+Fig.~\ref{src:lua-cephfs}: Line 2) - these lines detect whether the node is
 overloaded using the ``load" calculated in the load callback; while the load
@@
 \textbf{State Persisted Across Decisions} (Fig.~\ref{src:lru-dyn}: Lines 4,6;
-Fig~\ref{lua:cephfs}: Lines 3,4,10) - these lines use Mantle to write/read state
+Fig~\ref{src:lua-cephfs}: Lines 3,4,10) - these lines use Mantle to write/read state
 from previous decisions.  For ParSplice, we save a boolean that indicates
@@
 \textbf{Condition that Switches Policy} (Fig.~\ref{src:lru-dyn}: Line 6;
-Fig.~\ref{lua:cephfs}: Line 5) - these lines switch the policies using
+Fig.~\ref{src:lua-cephfs}: Line 5) - these lines switch the policies using
 information from previous decisions. ParSplice trims its cache once it eclipses
@@
 \end{minted}
-\caption{CephFS file system metadata load balancer.\label{lua:cephfs}}
+\caption{CephFS file system metadata load balancer.\label{src:lua-cephfs}}
 \end{figure}
 
-\begin{itemize}
-  \item Lustre Trace
-  \item LinkedIn Trace
-  \item Nathan's Trace
-\end{itemize}
+\subsection{Using ParSplice Policies for File Systems}
 
-\subsection{Using File System Policies for ParSplice}
-% mantle policies
-\subsection{Using ParSplice Policies for File Systems}
+%- how many inodes to keep in mds/client caches
+%- affects LB; increases capacity of single MDS (reduces req/mem pressure)
+%- affects LB; tells us which part of the cache to keep local
+
+% Load balancing in FSs
+Mantle was designed for file system metadata load balancing across a cluster of
+dedicated metadata servers, where spreading requests across servers improves
+performance. But another technique to reduce request load is to keep caches on
+both clients and servers. The caches in CephFS have the same performance and
+utilization trade-off as ParSplice, where large caches improve performance at
+the expense of a large memory footprint. High memory utilization is problematic
+when a single metadata server services many clients with their own caches.
+Using cache management policies has two benefits for load balancing: (1)
+increases the capacity of a single metadata server, and (2) helps identify
+which parts of the cache to keep local to a server. Applying the regime
+detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
+system access pattern has the potential to identify when to migrate keys and
+how many keys to migrate.
+
 % burstiness of creates then compile
+To show how regime detection helps load balance file system metadata across a
+cluster, consider the trace of metadata requests for compiling a large project
+in CephFS shown in Figure~\ref{fig:compile-ops}. That trace shows the number of
+file system metadata requests serviced by the metadata server when
+uncompressing (\texttt{untar}), compiling (\texttt{make}), and deleting
+(\texttt{rm}) the source code for the Linux kernel. If the system knows that
+the job phases would progress from many creates, to many lookups, to many
+deletes, then it could size its caches accordingly. For example, the file
+system could cache none of the inodes from the \texttt{untar} phase and run
+regime detection during the \texttt{make} phase and only cache the inodes that
+are repeatedly used. This will be less than 20K inodes and could result in up
+xMB in memory footprint savings without sacrificing performance.
 
-\subsection{Cross-pollinating Policies}
-% QoS, MRC, RAD - scheduing
+%- strategy: cache all inodes on create
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/compile-ops.png}\\
+\caption{ \label{fig:compile-ops}} \end{figure}
+
+% other use cases caches/tiers 
+Ceph has many other data management techniques that would benefit from the
+caching policies developed in
+Sections~\S\ref{sec:cache-management-using-system-archiecture-knowledge}
+and~\ref{sec:cache-management-using-domain-specific-knowledge}. Administrators
+can use the policies in ParSplice to automatically size and manage cache tiers
+\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/}, caching on
+object storage devices, or in the distributed block
+devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
+Integration with Mantle would be straightforward as it is merged into Ceph's
+mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
+three caching subsystems mentioned above already maintain keyspace access
+traces. We hypothesize that since this is all software defined caching,
+something more clever than LRU would improve cache utilization and performance.
+
+http://docs.ceph.com/docs/master/dev/mds_internals/data-structures/
+
+\subsection{Other Use Cases}
+% we said this is good for LB/Cache management... what about others??
+% QoS: when to move clients, where to move clients, how much of the reservation to move
+% Other cache management: MRU
+% Scheduling: how much of a resource to allocate; more than just LRU (RAD)
 %\subsection{Visualizing File System Traces like ParSplice Keyspace Traces}

### 2017-11-13 paper/sections/1_introduction.tex
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-This last contribution is explored in
-Sections~\S\ref{sec:cache-management-using-system-archiecture-knowledge}
-and~\S\ref{sec:cache-management-using-domain-specific-knowledge}, where we try
-a range of policies from different disciplines; but more importantly, in
-Section~\S\ref{sec:relation-to-load-balancing}, we conclude that the collection
-of policies we designed for ParSplice's cache management are very similar to
-the policies used to load balance metadata in the Ceph file system (CephFS)
-suggesting that there is potential for automatically adapting and generating
-policies dynamically.
+This last contribution is explored in Sections~\S\ref{sec:arch-specific}
+and~\S\ref{sec:dom-specific}, where we try a range of policies from different
+disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
+that the collection of policies we designed for ParSplice's cache management
+are very similar to the policies used to load balance metadata in the Ceph file
+system (CephFS) suggesting that there is potential for automatically adapting
+and generating policies dynamically.
 

### 2017-11-13 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #11
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 and disable hyper-threading because we experience unacceptable variability in
-performance. For the rest of the experiments, we use 4 nodes, 1 splicer, 1
-persistent database, 1 cache process, 1 worker, and up to 512 tasks.  The
+performance. For the rest of the experiments, we use 8 nodes, 1 splicer, 1
+persistent database, 1 cache process, 1 worker, and up to 256 tasks.  The
 keyspace analysis that follows is for the cache on the worker node, which is
@@
 state, which is computationally cheap (this motivates
-Section~\S\ref{sec:cache-management-using-system-archiecture-knowledge}).
+Section~\S\ref{sec:dom-specific}).
 
@@
 (up to \(5\times\)) because worker tasks start generating states with different
-coordinates later in the run (this motivates
-Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).  The
-growth rate, temperature, and number of workers have a predictable effect on
-the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
+coordinates later in the run (this motivates Section~\S\ref{sec:dom-specific}).
+The growth rate, temperature, and number of workers have a predictable effect
+on the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
 the number of reads changes with different initial conditions (\(\Delta\)), but
@@
 for space. This structure means that we can learn the regimes and adapt the
-storage system (this motivates
-Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).
+storage system (this motivates Section~\S\ref{sec:dom-specific}).

### 2017-11-13 paper/sections/3_methodology.tex
paper: addresses #11
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addressed #8
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 engine that supports our data management language.  The basic premise is that
-load balancing policies can be expressed with a simple API consisting of
+data management policies can be expressed with a simple API consisting of
 ``when", ``where", and ``how much". ``when" controls how aggressive or
@@
 \subsection{Extracting Mantle as a Library}
+\label{sec:extracting}
 
@@
 Mantle time series metrics, we expose accesses to the cache as a list of
-\texttt{timestamp, keyid} pairs. In
-Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}, we
+\texttt{timestamp, keyid} pairs. In Section~\S\ref{sec:dom-specific}, we
 explore a regime detection algorithm that uses this metric.

### 2017-11-13 paper/sections/4_dynamic-policies.tex
paper: hack on generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
+%\subsection{Trajectory Length}
+%
+%\subsection{Cloud Techniques: Elastic Search}
+%
+%What if we re-provision resources in response to events outside the
+%application's control, such as a slow Lustre.
+%
+%\subsection{Future work}
+%
+%How Much: cache policy from past, regime detection, Belady's Min
+
 \section{Cache Management Using System Architecture Knowledge}
-\label{sec:cache-management-using-system-archiecture-knowledge}
+\label{sec:arch-specific}
 
@@
 \section{Cache Management Using Domain-Specific Application Knowledge}
-\label{sec:cache-management-using-domain-specific-knowledge}
+\label{sec:dom-specific}
 \begin{figure}[t]
@@
 \subsection{Regime Detection}
+\label{sec:regime-detection}
 
@@
 fan.
-
-\subsection{Trajectory Length}
-
-\subsection{Cloud Techniques: Elastic Search}
-
-What if we re-provision resources in response to events outside the
-application's control, such as a slow Lustre.
-
-\subsection{Future work}
-
-How Much: cache policy from past, regime detection, Belady's Min

### 2017-11-13 paper/sections/5_file-systems.tex
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{General Data Management Policies}
-\label{sec:relation-to-load-balancing}
+\label{sec:scope}
 
@@
 \subsection{Using File System Policies for ParSplice}
+
 %- memory utilization/heuristics
-From a high-level the ParSplice policy trims the cache if the cache reaches a
-certain size {\it and} if it has already absorbed the initial burstiness of the
-workload; the CephFS poicy, which is shown in Figure~\ref{src:lua-cephfs},
-migrates load if the metadata load is higher than the average load {\it and}
-the current load has been overloaded for more than two iterations.
+From a high-level the cache management policy we designed in
+Figure~\ref{src:lru-dyn} trims the cache if the cache reaches a certain size
+{\it and} if it has already absorbed the initial burstiness of the workload.
+Much of this implementation was inspired by the file system metadata load
+balancing policy in Figure~\ref{src:lua-cephfs}, which was presented
+in~\cite{sevilla:sc15-mantle}. That policy migrates load if the metadata load
+is higher than the average load in the cluster {\it and} the current load has
+been overloaded for more than two iterations. The two policies have the
+following in common:
 
@@
 Fig.~\ref{src:lua-cephfs}: Line 2) - these lines detect whether the node is
-overloaded using the ``load" calculated in the load callback; while the load
-calculations and thresholds are different, the actual logic is exactly the
-same.  Recall that this decision is made locally because there is no global
-scheduler or centralized intelligence.
+overloaded using the load calculated in the load callback (not shown). While
+the calculations and thresholds are different, the way the loads are used is
+exactly the same; the ParSplice policy flags the node as overloaded if the
+cache reaches a certain size while the CephFS policy compares the load to other
+nodes in the system.
 
@@
 the number of consecutive instances that the server has been overloaded. We
-also clear the count (Line 10) if the server is no longer overloaded. The
-underlying implementation saves the values to local disk.
+also clear the count (Line 10) if the server is no longer overloaded.
 
-\textbf{Condition that Switches Policy} (Fig.~\ref{src:lru-dyn}: Line 6;
-Fig.~\ref{src:lua-cephfs}: Line 5) - these lines switch the policies using
-information from previous decisions. ParSplice trims its cache once it eclipses
-the ``absorb" threshold while CephFS allows balancing when overloaded for more
-than two iterations. The persistent state is essential for both of these
+\textbf{Two-Part Policies} (Fig.~\ref{src:lru-dyn}: Line 6;
+Fig.~\ref{src:lua-cephfs}: Line 5) - after determining that the node is
+overloaded, these lines add an additional condition before the policy enters a
+data management state.  ParSplice trims its cache once it eclipses the
+``absorb" threshold while CephFS allows balancing when overloaded for more than
+two iterations. The persistent state is essential for both of these
 policy-switching conditions.
@@
 dedicated metadata servers, where spreading requests across servers improves
-performance. But another technique to reduce request load is to keep caches on
-both clients and servers. The caches in CephFS have the same performance and
-utilization trade-off as ParSplice, where large caches improve performance at
-the expense of a large memory footprint. High memory utilization is problematic
-when a single metadata server services many clients with their own caches.
-Using cache management policies has two benefits for load balancing: (1)
-increases the capacity of a single metadata server, and (2) helps identify
-which parts of the cache to keep local to a server. Applying the regime
-detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
-system access pattern has the potential to identify when to migrate keys and
-how many keys to migrate.
+performance. But another technique to reduce request load is caching. If
+clients and servers maintain a consistent cache, the client can do operation
+locally without contacting the metadata server.  The caches in CephFS have the
+same performance and utilization trade-off as ParSplice, where large caches
+improve performance at the expense of a higher memory footprint. High memory
+utilization is problematic when a single metadata server maintains consistent
+caches with many clients .  Using cache management policies from the previous
+two sections has two benefits for load balancing: (1) increases the capacity of
+a single metadata server, and (2) helps identify which parts of the cache to
+keep local to a server. For example, applying the regime detection algorithm from
+Section~\S\ref{sec:regime-detection} to a similar file system access pattern
+has the potential to identify when to migrate keys and how many keys to
+migrate.
 
 % burstiness of creates then compile
-To show how regime detection helps load balance file system metadata across a
-cluster, consider the trace of metadata requests for compiling a large project
-in CephFS shown in Figure~\ref{fig:compile-ops}. That trace shows the number of
-file system metadata requests serviced by the metadata server when
-uncompressing (\texttt{untar}), compiling (\texttt{make}), and deleting
-(\texttt{rm}) the source code for the Linux kernel. If the system knows that
-the job phases would progress from many creates, to many lookups, to many
-deletes, then it could size its caches accordingly. For example, the file
-system could cache none of the inodes from the \texttt{untar} phase and run
-regime detection during the \texttt{make} phase and only cache the inodes that
-are repeatedly used. This will be less than 20K inodes and could result in up
-xMB in memory footprint savings without sacrificing performance.
+To explore this example in more detail, consider the trace of metadata requests
+for compiling code in CephFS shown in Figure~\ref{fig:compile-ops}.  That trace
+shows the number of file system metadata requests serviced by the metadata
+server when uncompressing (\texttt{untar}), compiling (\texttt{make}), and
+deleting (\texttt{rm}) the source code for the Linux kernel. If the system
+knows that the job phases would progress from many creates, to many lookups, to
+many deletes, then it could size its caches accordingly. For example, the file
+system could cache none of the metadatas from the \texttt{untar} phase and run
+regime detection during the \texttt{make} phase, resulting in the metadata
+server/clients only caching metadata that is repeatedly used. For the job in
+Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K inodes
+(the unit of metadata for file systems) instead of 60K, resulting in almost
+40MB (since an inode is about
+1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
+of memory savings without sacrificing performance.
 
@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/compile-ops.png}\\
-\caption{ \label{fig:compile-ops}} \end{figure}
+\caption{The file system metadata requests for a compile job shows distinct
+workload phases, characterized by a dominant request type ({\it e.g.}, creates
+for ``untar", lookups for ``make", etc.). Using a ParSplice cache management
+policy for a workload like this in file systems could reduce memory pressure
+without sacrificing performance.  \label{fig:compile-ops}}
+\end{figure}
 
@@
 Ceph has many other data management techniques that would benefit from the
-caching policies developed in
-Sections~\S\ref{sec:cache-management-using-system-archiecture-knowledge}
-and~\ref{sec:cache-management-using-domain-specific-knowledge}. Administrators
-can use the policies in ParSplice to automatically size and manage cache tiers
-\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/}, caching on
-object storage devices, or in the distributed block
+caching policies developed in Sections~\S\ref{sec:arch-specific}
+and~\ref{sec:dom-specific}. Administrators can use the policies in ParSplice to
+automatically size and manage cache tiers
+\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
+caching on object storage devices, or in the distributed block
 devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
@@
 traces. We hypothesize that since this is all software defined caching,
-something more clever than LRU would improve cache utilization and performance.
-
-http://docs.ceph.com/docs/master/dev/mds_internals/data-structures/
+something more clever than LRU would improve cache utilization and without
+sacrificing too much performance.
 

### 2017-11-14 paper/sections/1_introduction.tex
paper: change first figure
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
-  \caption{The keyspace activity for ParSplice using two different growth
-rates.  The line shows the rate that EOM minima values are retrieved from the
-key-value store (\(y1\) axis) and the points along the bottom show the number
-of keys accessed in a 1 second sliding window (\(y2\) axis).
-\label{fig:motivation-regimes}}
+\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
+
+\caption{Using our data management language and Mantle policy engine, we
+designed a dynamically sized caching policy for the ParSplice application. The
+solution uses knowledge about the application to detect key access patterns and
+adjust the cache accordingly. 
+\label{fig:cache-management}}
 \end{figure}
@@
 driven by a detailed analysis of the key-value accesses over the course of a
-long running simulation across a variety of initial conditions.
-Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
-pairs from the worker's cache for two different initial conditions of
-\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
-the read request rate (\(y1\) axis) and the dots along the bottom are the
-number of unique keys accessed (\(y2\) axis).  This figure demonstrates that
-small changes to \(\Delta\) can have a strong effect on the timing and
-frequency with which new EOM minima are discovered and referenced.  The
-ParSplice caches on each node are trimmed when memory pressure reaches a
-threshold but the number of unique keys accessed in
-Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
-should be sufficient.
+long running simulation across a variety of initial conditions. Our final
+solution, shown by the ``Dynamically Sized Cache" line in
+Figure~\ref{fig:cache-management}, detects keyspace access patterns and
+re-sizes the cache accordingly. Without tuning or parameter sweeps, our
+solution saves more memory than a hand-tuned cache without sacrificing
+performance.  More importantly, it works for a variety initial conditions
+without changing the policy itself.
 
 % What is Mantle
-We use the data management language and policy engine from the Mantle
-paper~\cite{sevilla:sc15-mantle} to dynamically explore the effects of
-different cache management strategies for the changing key-value workloads
-generated by ParSplice.  Mantle was originally touted as a programmable file
-system metadata load balancer, but we realize now that the collection of
-abstractions designed for file systems was a control plane that improved
-metadata access. So in this paper we refer to Mantle as a policy engine that
-injects policies written in our data management language directly into a
-running service and show how this approach is useful for reasoning about and
-designing different cache management strategies in ParSplice.  By service, we
-mean a system that manages data and responds requests, such as a file system or
-key-value store.  Developers write policies for ``when" they want data moved
-and ``how much" of the data to move, then the framework executes these policies
-whenever a decision needs to be made.  These abstraction help developers
-unfamiliar with the domain quickly reason about, develop, and deploy new
-policies that control temporal and spatial locality. We show that Mantle:
+To design more flexible cache management policies, like our ``Dynamically
+Sized Cache" policy, we use the data management language and policy engine
+from the Mantle paper~\cite{sevilla:sc15-mantle}. This framework allows us to
+dynamically explore the effects of different software-defined cache management
+strategies for the changing key-value workloads generated by ParSplice.  Mantle
+was originally touted as a programmable file system metadata load balancer, but
+we realize now that the collection of abstractions designed for file systems
+was a control plane that improved metadata access. So in this paper we refer to
+Mantle as a policy engine that injects policies written in our data management
+language directly into a running service and show how this approach is useful
+for reasoning about and designing different cache management strategies in
+ParSplice.  By service, we mean a system that manages data and responds
+requests, such as a file system or key-value store.  Developers write policies
+for ``when" they want data moved and ``how much" of the data to move, then the
+framework executes these policies whenever a decision needs to be made.  These
+abstractions help developers unfamiliar with the domain quickly reason about,
+develop, and deploy new policies that control temporal and spatial locality. We
+show that Mantle:
 

### 2017-11-14 paper/sections/2_parsplice-keyspace-analysis.tex
paper: change first figure
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\subsubsection*{An active but small keyspace} The bars in
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
+  \caption{The keyspace activity for ParSplice using two different growth
+rates.  The line shows the rate that EOM minima values are retrieved from the
+key-value store (\(y1\) axis) and the points along the bottom show the number
+of keys accessed in a 1 second sliding window (\(y2\) axis).
+\label{fig:motivation-regimes}}
+\end{figure}
+
+\subsubsection*{An active but small keyspace}
+Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
+pairs from the worker's cache for two different initial conditions of
+\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
+the read request rate (\(y1\) axis) and the dots along the bottom are the
+number of unique keys accessed (\(y2\) axis).  This figure demonstrates that
+small changes to \(\Delta\) can have a strong effect on the timing and
+frequency with which new EOM minima are discovered and referenced.  The
+ParSplice caches on each node are trimmed when memory pressure reaches a
+threshold but the number of unique keys accessed in
+Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
+should be sufficient.
+
+The bars in
 Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many reads

### 2017-11-14 paper/sections/1_introduction.tex
paper: cleanup abstract and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 abstractions effectively decomposes the problem into manageable policies for a
-different domain and service.  Our exploration of this space results in a two
-policy scheme that achieves 96\% efficiency while using only 7.6\% of the
-memory resources required by the base case.
+different domain and service.  Our exploration of this space results in a
+dynamically sized cache policy that, for our initial conditions, sacrifices
+negligible performance while using only 28\% of the memory required by our
+hand-tuned cache.
 
@@
 
-In this paper, we design cache management policies for the ParSplice worker,
-driven by a detailed analysis of the key-value accesses over the course of a
-long running simulation across a variety of initial conditions. Our final
-solution, shown by the ``Dynamically Sized Cache" line in
-Figure~\ref{fig:cache-management}, detects keyspace access patterns and
-re-sizes the cache accordingly. Without tuning or parameter sweeps, our
-solution saves more memory than a hand-tuned cache without sacrificing
-performance.  More importantly, it works for a variety initial conditions
-without changing the policy itself.
+In this paper, we design cache management policies for ParSplice, driven by a
+detailed analysis of the key-value accesses over the course of a long running
+simulation across a variety of initial conditions. The default ParSplice
+implementation uses an unlimited sized cache, shown by the ``No Cache
+Management" line Figure~\ref{fig:cache-management}. This solution is
+unacceptable for HPC environments, where memory is precious and a common goal
+is to keep memory for such data structures below 3\%\footnote{Empirically, we
+find this threshold works well for most applications}.  While users can
+configure ParSplice to resize the cache when it reaches a certain threshold,
+this solution requires tuning and parameter sweeps; the dashed ``Cache (too
+small)" curve in Figure~\ref{fig:cache-management} shows how a poorly
+configured cache hurts performance.  Our final solution, shown by the
+``Dynamically Sized Cache" line in Figure~\ref{fig:cache-management}, detects
+keyspace access patterns and re-sizes the cache accordingly.  Without tuning or
+parameter sweeps, our solution saves more memory than a hand-tuned cache
+without sacrificing performance.  More importantly, it works for a variety
+initial conditions without changing the policy itself.
 

### 2017-11-14 paper/sections/2_parsplice-keyspace-analysis.tex
paper: cleanup abstract and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 XC40 with 32 Intel Haswell 2.3GHz cores per node.  Each node has 128GB of RAM
-and our goal is to limit the size of the cache to 3\% of
-RAM\footnote{Empirically, we find this threshold works well for most
-applications}. Note that this is an addition to the 30GB that ParSplice uses to
-manage other ranks on the same node.  A single Cray node produced trajectories
-that are \(5\times\) times longer than our 10 node CloudLab clusters and
-\(25\times\) longer than our 10 node cluster at UCSC. As a result, it reaches
-different job phases faster and gives us a more comprehensive view of the
-workload. The performance gains compared to the commodity clusters have more to
-do with memory/PCI bandwidth than network.
+and our goal is to limit the size of the cache to 3\% of RAM . Note that this
+is an addition to the 30GB that ParSplice uses to manage other ranks on the
+same node.  A single Cray node produced trajectories that are \(5\times\) times
+longer than our 10 node CloudLab clusters and \(25\times\) longer than our 10
+node cluster at UCSC. As a result, it reaches different job phases faster and
+gives us a more comprehensive view of the workload. The performance gains
+compared to the commodity clusters have more to do with memory/PCI bandwidth
+than network.
 

### 2017-11-14 paper/sections/1_introduction.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 simulation across a variety of initial conditions. The default ParSplice
-implementation uses an unlimited sized cache, shown by the ``No Cache
-Management" line Figure~\ref{fig:cache-management}. This solution is
-unacceptable for HPC environments, where memory is precious and a common goal
-is to keep memory for such data structures below 3\%\footnote{Empirically, we
-find this threshold works well for most applications}.  While users can
-configure ParSplice to resize the cache when it reaches a certain threshold,
-this solution requires tuning and parameter sweeps; the dashed ``Cache (too
-small)" curve in Figure~\ref{fig:cache-management} shows how a poorly
-configured cache hurts performance.  Our final solution, shown by the
-``Dynamically Sized Cache" line in Figure~\ref{fig:cache-management}, detects
-keyspace access patterns and re-sizes the cache accordingly.  Without tuning or
-parameter sweeps, our solution saves more memory than a hand-tuned cache
-without sacrificing performance.  More importantly, it works for a variety
-initial conditions without changing the policy itself.
+implementation uses an unlimited sized cache, as shown by the ``No Cache
+Management" line Figure~\ref{fig:cache-management}. Using this much memory for
+one cache is unacceptable for HPC environments, where memory is precious and a
+common goal is to keep memory for such data structures below
+3\%\footnote{Empirically, we find this threshold works well for most
+applications}.  While users can configure ParSplice to resize the cache when it
+reaches a certain threshold, this solution requires tuning and parameter sweeps
+to find the optimal cache size that sacrifices negligible performance. The
+dashed ``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows
+how a poorly configured cache hurts performance.  Our final solution, shown by
+the ``Dynamically Sized Cache" line in Figure~\ref{fig:cache-management},
+detects keyspace access patterns and re-sizes the cache accordingly.  Without
+tuning or parameter sweeps, our solution saves more memory than a hand-tuned
+cache without sacrificing performance.  More importantly, it works for a
+variety initial conditions without changing the policy itself.
 
@@
 for reasoning about and designing different cache management strategies in
-ParSplice.  By service, we mean a system that manages data and responds
+ParSplice.  By service, we mean a system that manages data and responds to
 requests, such as a file system or key-value store.  Developers write policies

### 2017-11-14 paper/sections/4_dynamic-policies.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 of each strategy; the strategy should be accurate enough so as to sacrifice
-negligible performance and fast enough to run as often as we want to detect
-regimes. The goal of the following sections is not to find an optimal solution,
-as this can be done with parameter sweeps for thresholds; rather, we try to
-find techniques that work for a range of inputs and system setups.
+negligible performance and fast enough to run as often as we want to detect key
+access patterns. The goal of the following sections is not to find an optimal
+solution, as this can be done with parameter sweeps for thresholds; rather, we
+try to find techniques that work for a range of inputs and system setups.
 
@@
   \caption{Key activity for a 4 hour run shows groups of accesses to the same
-  subset of keys. Detecting these ``access regimes" leads to a more accurate cache
+  subset of keys. Detecting these access patterns leads to a more accurate cache
   management strategy.\label{fig:keyspace-zoomed}}
@@
 the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
-of the keys, or ``access regimes", occur because molecules are stuck in deep
-trajectories. Recall that the in-memory database stores the molecules' EOM
-minima, which is the smallest effective energy that a molecule observes during
-its trajectory. So molecules stuck in deep trajectories explore the same minima
-until they can escape to a new set of states. This exploration of the same set
-of states is called a superbasin.
+of keys, which we call an access ``fan" in subsequent sections, occurs because
+molecules are stuck in deep trajectories. Recall that the in-memory database
+stores the molecules' EOM minima, which is the smallest effective energy that a
+molecule observes during its trajectory. So molecules stuck in deep
+trajectories explore the same minima until they can escape to a new set of
+states. This exploration of the same set of states is called a superbasin.
 
@@
 \subsection{Failed Strategies}
-These techniques proliferated more knobs that obfuscated the problem. 
+
+To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
+variety of techniques using Mantle. Unfortunately, we found that they all
+proliferated more parameters that need to be tuned. Furthermore, we find that
+many of the metrics do not signal a new fans, that our metrics are not accurate
+enough, and that even if we tune all parameters they still will not work for
+all initial conditions.  Below, we indicate which parameters we need to add for
+each technique and the value we find to work best, via tuning and parameter
+sweeps, for one set of initial conditions.
+
 \begin{itemize}
-  \item Statistics
-  \item Calculus
-  \item K-Means
-  \item DBScan
-  \item Anomaly Detection
+
+  \item Statistics: decay on each key counts down until 0 0-valued keys are
+evicted.  ``history-of-keyspace", set to 10 seconds, to evict keys.
+
+  \item Calculus: use derivative to strip away magnitudes; use large positive
+slopes followed large negative slope as signal for new fan. ``Zero-crossing"
+for distance between small/large spikes to avoid false positives, set to 40
+seconds; `window size", set to 200 seconds, for the size of the moving average
+
+  \item K-Means Clustering fails because ``K" is not known {\it a-priori} and
+fans are different size. ``K", set to 4, for the number of clusters in the data
+using the sum of the distances to the centroid
+
+  \item DBScan: finds clusters using density as a metric. ``Eps", set to 20, for
+max distance between 2 samples in same neighborhood; ``Min", set to 5, for the
+number of samples per core
+
+  \item Anomaly Detection: size of the image is too big and bottom edges are
+not thick enough
+
 \end{itemize}
 
-\subsection{Regime Detection}
+\subsection{Dynamically Sized Cache: Access Pattern Detection}
 \label{sec:regime-detection}
 
-At each time step, we find the lowest ID and compare against the local minimum,
-which is the smallest ID we have seen thus far. If we move left to right, the
-local minimum never changes because the local minimum will start small. If we
-move from right to left, the local minimum changes at each access regime.  For
-points \(z\), \(y\), and \(x\), if the local minimum is the same we are in a
-regime. Processing \(y\), we set the local minimum to be \(min(y, m_l)\), where
-\(m_l\) is the local minimum of the previous time step of \(z\). The algorithm
-incorrectly detects a regime change if the local minimum of \(y\) is lower than
-local minimum of \(z\), since \(y\) may have points {\it within} \(z\); recall
-that we are trying to detect the whole fan, not just the bottom edge of each
-fan.
+Figure~\ref{src:dyn-cache} shows the algorithm we use to detect the key access
+patterns in Figure~\ref{fig:keyspace-zoomed}. At each time step, we find the
+lowest ID and compare against the local minimum, which is the smallest ID we
+have seen thus far (Line 7). We also maintain the top and bottom of each fan
+(Line 13) so we can tell the ``how much" policy the number of keys to evict
+(Line 19).  If we move left to right, the local minimum never changes because
+the local minimum will start small. If we move from right to left, the local
+minimum changes at each access regime.  For points \(z\), \(y\), and \(x\), if
+the local minimum is the same we are in a regime.  Processing \(y\), we set the
+local minimum to be \(min(y, m_l)\), where \(m_l\) is the local minimum of the
+previous time step of \(z\).  The algorithm incorrectly detects a regime change
+if the local minimum of \(y\) is lower than local minimum of \(z\), since \(y\)
+may have points {\it within} \(z\); recall that we are trying to detect the
+whole fan, not just the bottom edge of each fan. We exchange runtime, in the
+form of an O(n) algorithm, for simplicity. This approach avoids adding new
+thresholds for new fan detection, such as space between key accesses, space
+between key IDs, and window size of consecutive key accesses.
+
+\begin{figure}[h]
+\footnotesize
+\begin{minted}[xleftmargin=3em,linenos]{lua}
+d = timeseries()          -- data (d) provided by mantle
+ts, id = d:get(d:size())  -- functions provided by mantle
+fan  = {start=nil, finish=ts, top=0, bottom=id}
+fans = {}
+for i=d:size(),1,-1 do -- detect possible fans
+  ts, id = d:get(i)
+  if id < fan['bottom'] then
+    fan['start']  = ts
+    fans[#fans+1] = fan 
+    fan = {start=nil, finish=ts, top=0, bottom=id}
+  end
+
+  if id > fan['top'] then fan['top'] = id end 
+end
+fan['start'] = 0 
+fans[#fans+1] = fan
+
+if #fans < 2 then return false else
+  WRstate(fans[#fans-1]['top']-fans[1]['bottom'])
+  return true
+end
+\end{minted}
+\caption{The ``Dynamically Sized Cache" policy from
+Figure~\ref{fig:cache-management} iterates backwards over timestamp-keyID pairs
+and detects when key accesses moves on to a new subset of
+keys.\label{src:dyn-cache}}
+\end{figure}

### 2017-11-14 paper/sections/5_file-systems.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\section{General Data Management Policies}
+\section{Towards General Data Management Policies}
 \label{sec:scope}
@@
 balancing policy in Figure~\ref{src:lua-cephfs}, which was presented
-in~\cite{sevilla:sc15-mantle}. That policy migrates load if the metadata load
-is higher than the average load in the cluster {\it and} the current load has
-been overloaded for more than two iterations. The two policies have the
-following in common:
+in~\cite{sevilla:sc15-mantle}. That policy migrates file system metadata if the
+load is higher than the average load in the cluster {\it and} the current
+server has been overloaded for more than two iterations. The two policies have
+the following in common:
 
@@
 performance. But another technique to reduce request load is caching. If
-clients and servers maintain a consistent cache, the client can do operation
+clients and servers maintain a consistent cache, the client can do operations
 locally without contacting the metadata server.  The caches in CephFS have the
@@
 utilization is problematic when a single metadata server maintains consistent
-caches with many clients .  Using cache management policies from the previous
+caches with many clients.  Using cache management policies from the previous
 two sections has two benefits for load balancing: (1) increases the capacity of
 a single metadata server, and (2) helps identify which parts of the cache to
-keep local to a server. For example, applying the regime detection algorithm from
-Section~\S\ref{sec:regime-detection} to a similar file system access pattern
-has the potential to identify when to migrate keys and how many keys to
-migrate.
+keep local to a server. For example, applying the keyspace access pattern
+detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
+system access pattern has the potential to identify when to migrate keys and
+how many keys to migrate.
 
@@
 many deletes, then it could size its caches accordingly. For example, the file
-system could cache none of the metadatas from the \texttt{untar} phase and run
+system could cache none of the metadata from the \texttt{untar} phase and run
 regime detection during the \texttt{make} phase, resulting in the metadata
@@
 (the unit of metadata for file systems) instead of 60K, resulting in almost
-40MB (since an inode is about
+40MB of memory savings (since an inode is about
 1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
-of memory savings without sacrificing performance.
+without sacrificing performance.
 
@@
 and~\ref{sec:dom-specific}. Administrators can use the policies in ParSplice to
-automatically size and manage cache tiers
-\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
+automatically size and manage cache
+tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
 caching on object storage devices, or in the distributed block
@@
 traces. We hypothesize that since this is all software defined caching,
-something more clever than LRU would improve cache utilization and without
+something more clever than LRU would improve cache utilization without
 sacrificing too much performance.

### 2017-11-14 paper/sections/1_introduction.tex
paper: clear up figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 detailed analysis of the key-value accesses over the course of a long running
-simulation across a variety of initial conditions. The default ParSplice
-implementation uses an unlimited sized cache, as shown by the ``No Cache
-Management" line Figure~\ref{fig:cache-management}. Using this much memory for
-one cache is unacceptable for HPC environments, where memory is precious and a
-common goal is to keep memory for such data structures below
-3\%\footnote{Empirically, we find this threshold works well for most
-applications}.  While users can configure ParSplice to resize the cache when it
-reaches a certain threshold, this solution requires tuning and parameter sweeps
-to find the optimal cache size that sacrifices negligible performance. The
-dashed ``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows
-how a poorly configured cache hurts performance.  Our final solution, shown by
-the ``Dynamically Sized Cache" line in Figure~\ref{fig:cache-management},
-detects keyspace access patterns and re-sizes the cache accordingly.  Without
-tuning or parameter sweeps, our solution saves more memory than a hand-tuned
-cache without sacrificing performance.  More importantly, it works for a
-variety initial conditions without changing the policy itself.
+simulation across a variety of initial conditions.
+
+The memory usage for a ParSplice cache that stores molecule coordinates is
+shown in Figure~\ref{fig:cache-management}. The solid lines are the existing
+cache management policies, exposed as configurations in ParSplice.  The default
+ParSplice configuration uses an unlimited sized cache, shown by the ``No Cache
+Management" line, but using this much memory for one cache is unacceptable for
+HPC environments, where memory is precious and a common goal is to keep memory
+for such data structures below 3\%\footnote{Empirically, we find this threshold
+works well for most applications}.  While users can configure ParSplice to
+resize the cache when it reaches a certain threshold, this solution requires
+tuning and parameter sweeps to find the optimal cache size that sacrifices
+negligible performance.  Furthermore, this threshold changes with different
+initial configurations and clusters, as shown by our keyspace analysis in
+Section~\S\ref{sec:parsplice-keyspace-analysis}.  The ``Cache (too small)"
+curve in Figure~\ref{fig:cache-management} shows how a poorly configured cache
+can save memory but at the expense of performance.  Our dynamically sized
+cache, shown by the dashed line in Figure~\ref{fig:cache-management} detects
+keyspace access patterns and re-sizes the cache accordingly.  Without tuning or
+parameter sweeps, our solution saves more memory than a hand-tuned cache
+without sacrificing performance.  More importantly, it works for a variety
+initial conditions without changing the policy itself.
 
 % What is Mantle
-To design more flexible cache management policies, like our ``Dynamically
-Sized Cache" policy, we use the data management language and policy engine
-from the Mantle paper~\cite{sevilla:sc15-mantle}. This framework allows us to
-dynamically explore the effects of different software-defined cache management
-strategies for the changing key-value workloads generated by ParSplice.  Mantle
-was originally touted as a programmable file system metadata load balancer, but
-we realize now that the collection of abstractions designed for file systems
-was a control plane that improved metadata access. So in this paper we refer to
-Mantle as a policy engine that injects policies written in our data management
-language directly into a running service and show how this approach is useful
-for reasoning about and designing different cache management strategies in
-ParSplice.  By service, we mean a system that manages data and responds to
-requests, such as a file system or key-value store.  Developers write policies
-for ``when" they want data moved and ``how much" of the data to move, then the
-framework executes these policies whenever a decision needs to be made.  These
-abstractions help developers unfamiliar with the domain quickly reason about,
-develop, and deploy new policies that control temporal and spatial locality. We
-show that Mantle:
+To design more flexible cache management policies, like our dynamically sized
+cache solution in Figure~\ref{fig:cache-management}, we use the data management
+language and policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.
+This framework allows us to dynamically explore the effects of different
+software-defined cache management strategies for the changing key-value
+workloads generated by ParSplice.  Mantle was originally touted as a
+programmable file system metadata load balancer, but we realize now that the
+collection of abstractions designed for file systems was a control plane that
+improved metadata access. So in this paper we refer to Mantle as a policy
+engine that injects policies written in our data management language directly
+into a running service and show how this approach is useful for reasoning about
+and designing different cache management strategies in ParSplice.  By service,
+we mean a system that manages data and responds to requests, such as a file
+system or key-value store.  Developers write policies for ``when" they want
+data moved and ``how much" of the data to move, then the framework executes
+these policies whenever a decision needs to be made.  These abstractions help
+developers unfamiliar with the domain quickly reason about, develop, and deploy
+new policies that control temporal and spatial locality. We show that Mantle:
 

### 2017-11-14 paper/sections/5_file-systems.tex
paper: addresses #13 and #15
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/compile-ops.png}\\
-\caption{The file system metadata requests for a compile job shows distinct
-workload phases, characterized by a dominant request type ({\it e.g.}, creates
-for ``untar", lookups for ``make", etc.). Using a ParSplice cache management
-policy for a workload like this in file systems could reduce memory pressure
-without sacrificing performance.  \label{fig:compile-ops}}
+\caption{File system metadata requests for a compile job show workload phases
+characterized by a dominant request type ({\it e.g.}, creates for ``untar").
+Using ParSplice cache management policies for these types of workloads would
+reduce memory pressure without sacrificing performance.
+\label{fig:compile-ops}}
 \end{figure}

### 2017-11-14 paper/sections/1_introduction.tex
paper: Reza comments - intro: address cache management as system challenge in ParSplice - intro: this lets us deploy existing techniques from other domains - intro: clarify that Fig1 shows general problem (with preview of solution) - scope: pseudo code side by side
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-In this paper, we design cache management policies for ParSplice, driven by a
+The biggest challenge for ParSplice is properly sizing the caches in the
+storage hierarchy.  The memory usage for a ParSplice cache that stores molecule
+coordinates is shown in Figure~\ref{fig:cache-management}. The solid lines are
+the existing cache management policies, exposed as configurations in ParSplice.
+The default ParSplice configuration uses an unlimited sized cache, shown by the
+``No Cache Management" line, but using this much memory for one cache is
+unacceptable for HPC environments, where memory is precious and a common goal
+is to keep memory for such data structures below 3\%\footnote{Empirically, we
+find this threshold works well for most applications}.  While users can
+configure ParSplice to resize the cache when it reaches a certain threshold,
+this solution requires tuning and parameter sweeps to find the optimal cache
+size that sacrifices negligible performance.  Furthermore, this threshold
+changes with different initial configurations and clusters, as shown by our
+keyspace analysis in Section~\S\ref{sec:parsplice-keyspace-analysis}.  The
+``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
+poorly configured cache can save memory but at the expense of performance.  Our
+dynamically sized cache, shown by the dashed line in
+Figure~\ref{fig:cache-management}, detects keyspace access patterns and
+re-sizes the cache accordingly. The design of this policy is driven by a
 detailed analysis of the key-value accesses over the course of a long running
-simulation across a variety of initial conditions.
-
-The memory usage for a ParSplice cache that stores molecule coordinates is
-shown in Figure~\ref{fig:cache-management}. The solid lines are the existing
-cache management policies, exposed as configurations in ParSplice.  The default
-ParSplice configuration uses an unlimited sized cache, shown by the ``No Cache
-Management" line, but using this much memory for one cache is unacceptable for
-HPC environments, where memory is precious and a common goal is to keep memory
-for such data structures below 3\%\footnote{Empirically, we find this threshold
-works well for most applications}.  While users can configure ParSplice to
-resize the cache when it reaches a certain threshold, this solution requires
-tuning and parameter sweeps to find the optimal cache size that sacrifices
-negligible performance.  Furthermore, this threshold changes with different
-initial configurations and clusters, as shown by our keyspace analysis in
-Section~\S\ref{sec:parsplice-keyspace-analysis}.  The ``Cache (too small)"
-curve in Figure~\ref{fig:cache-management} shows how a poorly configured cache
-can save memory but at the expense of performance.  Our dynamically sized
-cache, shown by the dashed line in Figure~\ref{fig:cache-management} detects
-keyspace access patterns and re-sizes the cache accordingly.  Without tuning or
-parameter sweeps, our solution saves more memory than a hand-tuned cache
-without sacrificing performance.  More importantly, it works for a variety
-initial conditions without changing the policy itself.
+simulation across a variety of initial conditions.  Without tuning or parameter
+sweeps, our solution saves more memory than a hand-tuned cache without
+sacrificing performance.  More importantly, it works for a variety initial
+conditions without changing the policy itself.
 
@@
 engine that injects policies written in our data management language directly
-into a running service and show how this approach is useful for reasoning about
-and designing different cache management strategies in ParSplice.  By service,
-we mean a system that manages data and responds to requests, such as a file
-system or key-value store.  Developers write policies for ``when" they want
-data moved and ``how much" of the data to move, then the framework executes
-these policies whenever a decision needs to be made.  These abstractions help
-developers unfamiliar with the domain quickly reason about, develop, and deploy
-new policies that control temporal and spatial locality. We show that Mantle:
+into a running service, such as a file system or key-value store, and show how
+this approach is useful for reasoning about and designing different cache
+management strategies in ParSplice.  Developers write policies for ``when" they
+want data moved and ``how much" of the data to move, then the framework
+executes these policies whenever a decision needs to be made.  These
+abstractions help developers unfamiliar with the domain quickly deploy new
+policies {\it or} well-understood techniques from other domains.  We show that
+Mantle:
 

### 2017-11-14 paper/sections/4_dynamic-policies.tex
paper: Reza comments - intro: address cache management as system challenge in ParSplice - intro: this lets us deploy existing techniques from other domains - intro: clarify that Fig1 shows general problem (with preview of solution) - scope: pseudo code side by side
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-The actual policy is shown in Figure~\ref{src:lru-dyn} where the only
-difference between the code snippet in Figure~\ref{src:lru} is the addition of
-lines 2-7. Note that policy also uses the \texttt{RDState} and \texttt{WRState}
-functions described in Section~\S\ref{sec:extracting}.
-
-\begin{figure}[t]
-\footnotesize
-\centering
-\begin{minted}[xleftmargin=3em,linenos]{lua}
-function when()
-  if server[whoami]['cachesize'] > n then
-    if server[whoami]['cachesize'] > 100K then
-      WRstate(1)
-    end
-    if RDstate() == 1 then
-      return true
-    end
-  end
-  return false
-end
-\end{minted}
-\caption{A more complicated LRU-based cache policy that has two parts. After
-absorbing the initial burstiness of the workload, the policy switches to a more
-constrained cache to limit memory usage. The performance/utilization for
-different values of \texttt{n} is shown in
-Figure~\ref{fig:methodology-tradeoff-dynamic}. \label{src:lru-dyn}}
-\end{figure}
-
-We trigger the policy switch at 100K keys to absorb the flash crowd at the
-beginning of the run. Once triggered, keys are evicted to bring the size of the
-cache down to the threshold.  In the bar chart, the cache sizes for the \(n\)
-key limit policy are along the \(x\) axis.
+The actual policy is shown and described in more detail in the next section in
+Figure~\ref{src:lru-dyn}, but the only difference between the code snippet in
+Figure~\ref{src:lru} is the addition of lines 2-7.  We trigger the policy
+switch at 100K keys to absorb the flash crowd at the beginning of the run. Once
+triggered, keys are evicted to bring the size of the cache down to the
+threshold.  In the bar chart, the cache sizes for the \(n\) key limit policy
+are along the \(x\) axis.
 

### 2017-11-14 paper/sections/5_file-systems.tex
paper: Reza comments - intro: address cache management as system challenge in ParSplice - intro: this lets us deploy existing techniques from other domains - intro: clarify that Fig1 shows general problem (with preview of solution) - scope: pseudo code side by side
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \label{sec:scope}
+
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[t]{0.48\textwidth}
+        \centering
+        \footnotesize
+        \centering
+        \begin{minted}[xleftmargin=3em,linenos]{lua}
+function when()
+  if server[whoami]['cachesize'] > n then
+    if server[whoami]['cachesize'] > 100K then
+      WRstate(1)
+    end
+    if RDstate() == 1 then
+      return true
+    end
+  end
+  return false
+end
+        \end{minted}
+        \caption{ParSplice cache management policy that absorbs the burstiness of the
+        workload before switching to a constrained cache.  The
+        performance/utilization for different  \texttt{n} is in
+        Figure~\ref{fig:methodology-tradeoff-dynamic}. \label{src:lru-dyn}}
+    \end{subfigure}%
+    ~ 
+    \begin{subfigure}[t]{0.48\textwidth}
+        \centering
+        \footnotesize
+        \begin{minted}[xleftmargin=3em,linenos]{lua}
+local function when()
+  if servers[whoami]["load"] > target then
+    overloaded = RDstate() + 1
+    WRstate(overloaded)
+    if overloaded > 2 then
+      return true
+    end
+  end
+  else then WRstate(0) end
+  return false
+end
+        \end{minted}
+	\caption{CephFS file system metadata load balancer, designed in 2004
+        in~\cite{weil:sc2004-dyn-metadata}, reimplemented in Lua
+        in~\cite{sevilla:sc15-mantle}. This policy has many similarities to the
+        ParSplice cache management policy.\label{src:lua-cephfs}}
+    \end{subfigure}
+    \caption{ParSplice's cache management policy has the same components as
+    CephFS's load balancing policy.}
+\end{figure*}
 
@@
 \textbf{State Persisted Across Decisions} (Fig.~\ref{src:lru-dyn}: Lines 4,6;
-Fig~\ref{src:lua-cephfs}: Lines 3,4,10) - these lines use Mantle to write/read state
+Fig~\ref{src:lua-cephfs}: Lines 3,4,9) - these lines use Mantle to write/read state
 from previous decisions.  For ParSplice, we save a boolean that indicates
@@
 policy-switching conditions.
-
-\begin{figure}[t]
-\footnotesize
-\begin{minted}[xleftmargin=3em,linenos]{lua}
-local function when()
-  if servers[whoami]["load"] > target then
-    overloaded = RDstate() + 1
-    WRstate(overloaded)
-    if overloaded > 2 then
-      return true
-    end
-  end
-  else then
-    WRstate(0)
-  end
-  return false
-end
-\end{minted}
-\caption{CephFS file system metadata load balancer.\label{src:lua-cephfs}}
-\end{figure}
 

### 2017-11-14 paper/sections/4_dynamic-policies.tex
paper: addresses #14 and #15
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-% results: cache size trade-offs
-\begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
-  \caption{The performance and resource utilization trade-off for different
-  cache sizes. ``Baseline" is
-  ParSplice unmodified and the ``Policy: Constrain Cache Size" graph limits the
-  size of the cache to save memory.  \label{fig:methodology-tradeoff}}
-\end{figure}
-
 The results for different cache sizes for a growth rate of \(\Delta_1\) over a
@@
 The other graph shares the \(y\) axis and shows the trade-off of using a basic
-LRU-style cache for different cache sizes, implemented using the ``when" and
-``how much" policies in Figure~\ref{src:lru}.  The error bars are the standard
+LRU-style cache for different cache sizes, implemented using a ``when" policy
+of \texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
+\texttt{servers[whoami]['cachesize']-n}. The error bars are the standard
 deviation of 3 runs. 
+%\begin{figure}[h] \footnotesize \begin{minted}{lua}
+%server[whoami]['cachesize'] > n \end{minted} \end{figure}
+
+%and ``how much":
+%\begin{figure}[h]
+%\footnotesize
+%\begin{minted}{lua}
+%servers[whoami]['cachesize'] - n
+%\end{minted}
+%\end{figure}
 
 \begin{figure}[t]
-\footnotesize
-\centering
-\begin{minted}[xleftmargin=3em,linenos]{lua}
-function when()
-  if server[whoami]['cachesize'] > n then
-    return true
-  end
-  return false
-end
-
-function howmuch()
-  if servers[whoami]['cachesize'] > n
-    return servers[whoami]['cachesize'] - n
-  end
-  return 0
-end
-\end{minted}
-\caption{Policy that implements a basic LRU cache. The performance and memory
-utilization for different values of \(n\) are graphed in
-Figure~\ref{fig:methodology-tradeoff}.
-\label{src:lru}}
+\noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
+\caption{The performance and resource utilization trade-off for different cache
+sizes. ``Baseline" is ParSplice unmodified and the ``Policy: Constrain Cache
+Size" graph limits the size of the cache to save memory.
+\label{fig:methodology-tradeoff}}
 \end{figure}
+%
+%\footnotesize
+%\centering
+%\begin{minted}[xleftmargin=3em,linenos]{lua}
+%function when()
+%  if server[whoami]['cachesize'] > n then
+%    return true
+%  end
+%  return false
+%end
+%
+%function howmuch()
+%  if servers[whoami]['cachesize'] > n
+%    return servers[whoami]['cachesize'] - n
+%  end
+%  return 0
+%end
+%\end{minted}
+%\caption{Policy that implements a basic LRU cache. The performance and memory
+%utilization for different values of \(n\) are graphed in
+%Figure~\ref{fig:methodology-tradeoff}.
+%\label{src:lru}}
+%\end{figure}
 
@@
 
-% Why is this a good idea
-Although ParSplice does not use a distributed file system, its workload is very
-similar because the minima key-value store responds to small and frequent
-requests, which results in hot spots and flash crowds.  Modern distributed file
-systems have found efficient ways to measure, migrate, and partition metadata
-load and have shown large performance gains and
-better scalability~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs,
-grider:pdsw2015-marfs, ren:sc2014-indexfs, patil:fast2011-giga+,
-brandt:msst2003-lh}.  Previous work quantified the speedups achieved with
-Mantle and formalized balancers that were good for file systems.
+%% Why is this a good idea
+%Although ParSplice does not use a distributed file system, its workload is very
+%similar because the minima key-value store responds to small and frequent
+%requests, which results in hot spots and flash crowds.  Modern distributed file
+%systems have found efficient ways to measure, migrate, and partition metadata
+%load and have shown large performance gains and
+%better scalability~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs,
+%grider:pdsw2015-marfs, ren:sc2014-indexfs, patil:fast2011-giga+,
+%brandt:msst2003-lh}.  Previous work quantified the speedups achieved with
+%Mantle and formalized balancers that were good for file systems.
 
@@
 The actual policy is shown and described in more detail in the next section in
-Figure~\ref{src:lru-dyn}, but the only difference between the code snippet in
-Figure~\ref{src:lru} is the addition of lines 2-7.  We trigger the policy
-switch at 100K keys to absorb the flash crowd at the beginning of the run. Once
-triggered, keys are evicted to bring the size of the cache down to the
-threshold.  In the bar chart, the cache sizes for the \(n\) key limit policy
-are along the \(x\) axis.
+Figure~\ref{src:lru-dyn}.  We trigger the policy switch at 100K keys to absorb
+the flash crowd at the beginning of the run. Once triggered, keys are evicted
+to bring the size of the cache down to the threshold.  In the bar chart, the
+cache sizes for the \(n\) key limit policy are along the \(x\) axis.
 

### 2017-11-14 paper/sections/4_dynamic-policies.tex
paper: addresses #7
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Figure~\ref{src:dyn-cache} shows the algorithm we use to detect the key access
-patterns in Figure~\ref{fig:keyspace-zoomed}. At each time step, we find the
-lowest ID and compare against the local minimum, which is the smallest ID we
-have seen thus far (Line 7). We also maintain the top and bottom of each fan
-(Line 13) so we can tell the ``how much" policy the number of keys to evict
-(Line 19).  If we move left to right, the local minimum never changes because
-the local minimum will start small. If we move from right to left, the local
-minimum changes at each access regime.  For points \(z\), \(y\), and \(x\), if
-the local minimum is the same we are in a regime.  Processing \(y\), we set the
-local minimum to be \(min(y, m_l)\), where \(m_l\) is the local minimum of the
-previous time step of \(z\).  The algorithm incorrectly detects a regime change
-if the local minimum of \(y\) is lower than local minimum of \(z\), since \(y\)
-may have points {\it within} \(z\); recall that we are trying to detect the
-whole fan, not just the bottom edge of each fan. We exchange runtime, in the
-form of an O(n) algorithm, for simplicity. This approach avoids adding new
-thresholds for new fan detection, such as space between key accesses, space
-between key IDs, and window size of consecutive key accesses.
+patterns. Iterating backwards through the keyspace trace, at each time step we
+find the lowest key ID and compare against the lowest key ID we have seen so
+far (Line 7). We also maintain the top and bottom of each fan (Line 13) so we
+can tell the ``how much" policy the number of keys to evict (Line 19).  The
+algorithm iterates backwards over the keyspace trace because a change in the
+minimum value signals a new fan. No signal exists iterating left to right, as
+the maximum value always increases and the minimum values at the bottom of each
+fan are sparse.  For example, the maximum distance between values along the
+bottom edge of the zoomed fan in Figure~\ref{fig:keyspace-zoomed} is 125
+seconds, while the maximum distance between minimum values for the fan before
+is 0 seconds. As a result of this sparseness, iterating left to right requires
+a ``window size" parameter to determine when we think a minimum value will not
+show up again.  We exchange runtime, in the form of an O(\(n\)) algorithm where
+\(n\) is the number events, for simplicity because this approach avoids adding
+new thresholds for fan detection ({\it e.g.}, space between key accesses, space
+between key IDs, and window size of consecutive key accesses).
+
+%For points \(z\), \(y\), and \(x\), if the
+%local minimum is the same we are in a regime.  Processing \(y\), we set the
+%local minimum to be \(min(y, m_l)\), where \(m_l\) is the local minimum of the
+%previous time step of \(z\).  The algorithm incorrectly detects a regime change
+%if the local minimum of \(y\) is lower than local minimum of \(z\), since \(y\)
+%may have points {\it within} \(z\); recall that we are trying to detect the
+%whole fan, not just the bottom edge of each fan.
 

### 2017-11-15 paper/sections/1_introduction.tex
paper: provide motivation for this work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-The fine-grained data annotation capabilities provided by key-value storage is
-a natural match for many types of scientific simulation. Simulations relying on
-a mesh-based decomposition of a physical region may result in millions or
-billions of mesh cells. Each cell contains materials, pressures, temperatures
-and other characteristics that are required to accurately simulate phenomena of
-interest. In our target application, the
-ParSplice~\cite{perez:jctc20150parsplice} molecular dynamics simulation, a
-hierarchy of caches and a single persistent key-value store are used to store
-both observed minima across a molecule's equation of motion (EOM) and the
-hundreds or thousands of partial trajectories calculated each second during a
-parallel job. Unfortunately, if we scale the system the IO to the storage
-hierarchy will quickly saturate both the storage and bandwidth capacity of a
-single node, so we need more effective data management techniques, such as
-cache management or load balancing across a cluster.
+Storage systems use software-based caches extensively to improve performance
+and resource utilization but the policies that guide what data to evict and
+when to evict vary with the application or use case. For example, caching file
+system metadata on both clients and servers reduces the number of remote
+procedure calls and drastically improves the performance of create-heavy
+workloads common in HPC~\cite{ren:sc2014-indexfs, patil:fast2011-giga+}. But
+these techniques are specific to the domain and application's behavior and a
+new use case may prove to be poor a match for the selected caching
+policy~\cite{sevilla:sc15-mantle, weil:sc2004-dyn-metadata,
+weil:osdi2006-ceph}. In this paper, we evaluate a variety of caching policies
+using a general framework that supports policy switching and arrive at a
+customized policy that works well for our target application and storage
+system.  Trying general policies and iterating to a very customized solution
+confirms that our prototype can adapt to different types of workloads; but more
+importantly, this is the second system and domain that we have successfully used
+this policy language and engine in. This breadth helps us reason about more
+general data management policies by giving us a framework to compare and
+contrast policies from other domains and services.
+
+Our target application is ParSplice~\cite{perez:jctc20150parsplice}, a
+molecular dynamics simulation that uses a hierarchy of caches and a single
+persistent key-value store to store both observed minima across a molecule's
+equation of motion (EOM) and the hundreds or thousands of partial trajectories
+calculated each second during a parallel job.  The fine-grained data annotation
+capabilities provided by key-value storage is a natural match for scientific
+simulations like ParSplice. But these simulations rely on a mesh-based
+decomposition of a physical region and result in millions or billions of mesh
+cells, where each cell contains materials, pressures, temperatures and other
+characteristics that are required to accurately simulate phenomena of interest.
+Unfortunately, this simulations of this size saturate the capcity and bandwidth
+capabilities of a single node so we need more effective data management
+techniques, such as cache management or load balancing across a cluster.
 

### 2017-11-15 paper/sections/4_dynamic-policies.tex
paper: provide motivation for this work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
+
 \caption{The performance and resource utilization trade-off for different cache
-sizes. ``Baseline" is ParSplice unmodified and the ``Policy: Constrain Cache
-Size" graph limits the size of the cache to save memory.
-\label{fig:methodology-tradeoff}}
+management policies. ``None" is ParSplice unmodified, ``Fixed Sized Cache"
+limits the size of the cache by evicting the least recently used items, and
+``Multi-Policy Cache" switches to a fixed sized cache after absorbing the
+initial burstiness of the workload.  The size of the fixed sized cache for each
+experiment is on the \(x\) axis.  \label{fig:methodology-tradeoff}}
 \end{figure}
 %
-%\footnotesize
-%\centering
-%\begin{minted}[xleftmargin=3em,linenos]{lua}
-%function when()
-%  if server[whoami]['cachesize'] > n then
-%    return true
-%  end
-%  return false
+%\footnotesize \centering \begin{minted}[xleftmargin=3em,linenos]{lua} function
+%when() if server[whoami]['cachesize'] > n then return true end return false
 %end
 %
-%function howmuch()
-%  if servers[whoami]['cachesize'] > n
-%    return servers[whoami]['cachesize'] - n
-%  end
-%  return 0
-%end
-%\end{minted}
-%\caption{Policy that implements a basic LRU cache. The performance and memory
-%utilization for different values of \(n\) are graphed in
-%Figure~\ref{fig:methodology-tradeoff}.
-%\label{src:lru}}
-%\end{figure}
+%function howmuch() if servers[whoami]['cachesize'] > n return
+%servers[whoami]['cachesize'] - n end return 0 end \end{minted} \caption{Policy
+%that implements a basic LRU cache. The performance and memory utilization for
+%different values of \(n\) are graphed in
+%Figure~\ref{fig:methodology-tradeoff}.  \label{src:lru}} \end{figure}
 
@@
 Despite the memory savings, our results suggest that dynamic cache management
-policies could save even more memory.  Figure~\ref{fig:methodology-tradeoff}
-show that a 100K key cache is sufficient as a static policy but the top graph
-in Figure~\ref{fig:motivation-regimes} indicates that the cache size could be
-much smaller. That graph shows that the beginning of the run is characterized
-by many reads to a small set of keys and the end sees much lower reads per
-second to a larger keyspace. Specifically, it shows only about 100 keys as
-active in the latter half of the run.
+policies could save even more memory.  The middle plot in
+Figure~\ref{fig:methodology-tradeoff} show that a 100K key cache is sufficient
+as a static policy but the top graph in Figure~\ref{fig:motivation-regimes}
+indicates that the cache size could be much smaller. That graph shows that the
+beginning of the run is characterized by many reads to a small set of keys and
+the end sees much lower reads per second to a larger keyspace. Specifically, it
+shows only about 100 keys as active in the latter half of the run.
 
@@
 
-\begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff-dynamic.png}\\
-  \caption{The performance and resource utilization trade-off of using a
-  dynamic cache management policy that switches to a constrained cache policy after absorbing
-  the initial burstiness of the workload. The sizes of these smaller caches are
-  on the \(x\) axis.  \label{fig:methodology-tradeoff-dynamic}}
-\end{figure}
-
-Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using Mantle
+The plot on the right side of Figure~\ref{fig:methodology-tradeoff} shows the results of using Mantle
 to program a dynamic cache management policy into ParSplice that switches between
@@
 % results: same level of performance can be achieved 
-The dynamic policies show better performance than the single \(n\) key
+The multi-part policies show better performance than the single \(n\) key
 policies. The performance and memory utilization for a 100K key cache size is
-the same as the 100K bar in Figure~\ref{fig:methodology-tradeoff-dynamic} but
+the same as the 100K bar in the ``Multi-Part Policy" graph in  Figure~\ref{fig:methodology-tradeoff} but
 the rest reduce the size of the keyspace after the read flash crowd.  We see
@@
 
-The results in Figure~\ref{fig:methodology-tradeoff-dynamic} are slightly
-deceiving for three reason: (1) segments take longer to generate later in the
-run, (2) the memory footprint is the value at the end of 2.5 hours, and (3)
-this policy only works well for the 2.5 hour run.  For (1), the curving down of
-the simulation vs. wall-clock time is shown in
-Figure~\ref{fig:methodology-trajectory}; as the nanoparticle grows it takes
-longer to generate segments so by the time we reach 2 hours, over 90\% of the
-trajectory is already generated.  For (2), the memory footprint is around 0.4GB
-until we reach 100K key switch threshold. In
-Figures~\ref{fig:methodology-tradeoff}
-and~\ref{fig:methodology-tradeoff-dynamic} we plot the final value. For (3),
+The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
+three reason: (1) segments take longer to generate later in the run, (2) the
+memory footprint is the value at the end of 2.5 hours, and (3) this policy only
+works well for the 2.5 hour run.  For (1), the curving down of the simulation
+vs. wall-clock time is shown in Figure~\ref{fig:methodology-trajectory}; as the
+nanoparticle grows it takes longer to generate segments so by the time we reach
+2 hours, over 90\% of the trajectory is already generated.  For (2), the memory
+footprint is around 0.4GB until we reach 100K key switch threshold. In
+Figures~\ref{fig:methodology-tradeoff} we plot the final value. For (3),
 Figure~\ref{fig:methodology-trajectory} shows that the cache fills up with 100K
@@
   \caption{The rate that the trajectory is computed decays over time (which is
-  expected) but this skews the performance improvements in
-  Figure~\ref{fig:methodology-tradeoff-dynamic}. Our dynamic policy works for 2.5
+  expected) but this skews the performance improvements in the ``Multi-Part Policy" graph in 
+  Figure~\ref{fig:methodology-tradeoff}. Our dynamic policy works for 2.5
   hour jobs but not for 4 hour jobs.  \label{fig:methodology-trajectory}}

### 2017-11-15 paper/sections/1_introduction.tex
paper: addresses #22
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 new use case may prove to be poor a match for the selected caching
-policy~\cite{sevilla:sc15-mantle, weil:sc2004-dyn-metadata,
-weil:osdi2006-ceph}. In this paper, we evaluate a variety of caching policies
-using a general framework that supports policy switching and arrive at a
-customized policy that works well for our target application and storage
-system.  Trying general policies and iterating to a very customized solution
-confirms that our prototype can adapt to different types of workloads; but more
-importantly, this is the second system and domain that we have successfully used
-this policy language and engine in. This breadth helps us reason about more
-general data management policies by giving us a framework to compare and
-contrast policies from other domains and services.
+policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
+weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
+variety of caching policies using a general framework that supports policy
+switching and arrive at a customized policy that works well for our target
+application and storage system.  Trying general policies and iterating to a
+very customized solution confirms that our prototype can adapt to different
+types of workloads; but more importantly, this is the second system and domain
+that we have successfully used this policy language and engine in. This breadth
+helps us reason about more general data management policies by giving us a
+framework to compare and contrast policies from other domains and services.
 

### 2017-11-15 paper/sections/1_introduction.tex
paper: addresses #20 and #21
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
-
-\caption{Using our data management language and Mantle policy engine, we
-designed a dynamically sized caching policy for the ParSplice application. The
-solution uses knowledge about the application to detect key access patterns and
-adjust the cache accordingly. 
+\caption{Using our data management language and policy engine, we designed a
+dynamically sized caching policy (dashed line) for ParSplice. The solution uses
+knowledge about the application to detect key access patterns and adjust the
+cache accordingly. Compared to existing ParSplice configurations (solid lines),
+our solution saves the most memory without sacrificing performance.
 \label{fig:cache-management}}

### 2017-11-15 paper/sections/2_parsplice-keyspace-analysis.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #20 and #21
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 parallelization. We focus on the worker's cache (circled), which facilitates
-communication and segment exchange between the worker and its
+communication and segment exchange between the worker and its tasks.
 \label{fig:parsplice}}
@@
 Trends also exist for temperature and number of workers but are omitted here
-for space. This structure means that we can learn the regimes and adapt the
+for space. This structure means that we can learn the key access patterns and adapt the
 storage system (this motivates Section~\S\ref{sec:dom-specific}).

### 2017-11-15 paper/sections/3_methodology.tex
paper: addresses #14 and #15
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #20 and #21
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 %  \item background: CephFS implementation
-To explore dynamic cache management policies that change during the run, we use
+To explore software-defined cache management, we use
 the data management language and policy engine presented
@@
 distributed or concentrated the data should be; and ``how much" controls the
-amount of data that should be sent.
+amount of data that should be sent. There is also a ``load" policy that lets
+administrators specify how to collapse many metrics into a single load metric
+({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).
 
-The succinctness of the API lets users inject multiple, dynamic policies. In
+The succinctness of the API lets users inject multiple, possibly dynamic, policies. In
 this work we focus on a single node, so the ``where" policy is not used. When
@@
 \texttt{timestamp, keyid} pairs. In Section~\S\ref{sec:dom-specific}, we
-explore a regime detection algorithm that uses this metric.
+explore a key access pattern detection algorithm that uses this metric.
 

### 2017-11-15 paper/sections/4_dynamic-policies.tex
paper: addresses #20 and #21
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 performance. Decreasing the cache degrades performance and predictability.
-Despite the memory savings, our results suggest that dynamic cache management
-policies could save even more memory.  The middle plot in
+Despite the memory savings, our results suggest that a multi-policy cache management strategy
+could save even more memory.  The middle plot in
 Figure~\ref{fig:methodology-tradeoff} show that a 100K key cache is sufficient
@@
 should suffice after the initial read flash crowd passes because
-the keyspace is far less active. This suggests a two-part cache management
-policy.
+the keyspace is far less active. This suggests a multi-policy cache management
+strategy.
 
@@
 The plot on the right side of Figure~\ref{fig:methodology-tradeoff} shows the results of using Mantle
-to program a dynamic cache management policy into ParSplice that switches between
+to program a multi-policy cache management strategy into ParSplice that switches between
 two policies:
@@
 sufficient policies for the 2.5 hour runs but anything longer would need a
-different dynamic cache management policy.
+different multi-policy cache management strategy.
 
@@
   expected) but this skews the performance improvements in the ``Multi-Part Policy" graph in 
-  Figure~\ref{fig:methodology-tradeoff}. Our dynamic policy works for 2.5
+  Figure~\ref{fig:methodology-tradeoff}. Our multi-policy cache management strategy works for 2.5
   hour jobs but not for 4 hour jobs.  \label{fig:methodology-trajectory}}
@@
 % but the result is still valid
-Despite these caveats, the result is still valid: we found a dynamic cache 
-management policy that absorbs the cost of a high read throughput on a small
+Despite these caveats, the result is still valid: we found multi-policy cache 
+management strategy that absorbs the cost of a high read throughput on a small
 keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
@@
 \end{minted}
-\caption{The ``Dynamically Sized Cache" policy from
-Figure~\ref{fig:cache-management} iterates backwards over timestamp-keyID pairs
-and detects when key accesses moves on to a new subset of
+\caption{The dynamically sized cache policy, whose performance is shown in 
+Figure~\ref{fig:cache-management}, iterates backwards over timestamp-keyID pairs
+and detects when key accesses move on to a new subset of
 keys.\label{src:dyn-cache}}

### 2017-11-15 paper/sections/5_file-systems.tex
paper: provide motivation for this work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #20 and #21
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
         performance/utilization for different  \texttt{n} is in
-        Figure~\ref{fig:methodology-tradeoff-dynamic}. \label{src:lru-dyn}}
+        Figure~\ref{fig:methodology-tradeoff}. \label{src:lru-dyn}}
     \end{subfigure}%
@@
 
-\textbf{Two-Part Policies} (Fig.~\ref{src:lru-dyn}: Line 6;
+\textbf{Multi-Policy Strategy} (Fig.~\ref{src:lru-dyn}: Line 6;
 Fig.~\ref{src:lua-cephfs}: Line 5) - after determining that the node is
@@
 system could cache none of the metadata from the \texttt{untar} phase and run
-regime detection during the \texttt{make} phase, resulting in the metadata
+key access pattern detection during the \texttt{make} phase, resulting in the metadata
 server/clients only caching metadata that is repeatedly used. For the job in

### 2017-11-15 paper/sections/6_conclusion.tex
paper: addresses #20 and #21
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 management language and policy engine, called Mantle that is general enough to
-express complicated, dynamic policies for two different domains and services.
+express complicated, multi-policy strategies for two different domains and services.
 Rather than attempting to construct a single, complex load balancing policy
@@
 workload changes over time.  In our analysis of the ParSplice key-value
-workload we have detected clear workload regimes that are sensitive to the
+workload we have detected clear workload access patterns that are sensitive to the
 initial conditions and the scale and duration of the simulation. We have also

### 2017-11-15 paper/sections/1_introduction.tex
paper: addresses #23
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 dynamics simulation demonstrates the need for more complex cache management
-strategies. Baseline measurements show clear keyspace access patterns and hot
+strategies. Baseline measurements show clear key access patterns and hot
 spots that offer significant opportunity for optimization. We use the data
@@
 dynamically sized cache, shown by the dashed line in
-Figure~\ref{fig:cache-management}, detects keyspace access patterns and
+Figure~\ref{fig:cache-management}, detects key access patterns and
 re-sizes the cache accordingly. The design of this policy is driven by a
@@
   development. Changing the policy in use is critical in applications such as
-  ParSplice that have alternating stable and chaotic keyspace access patterns
+  ParSplice that have alternating stable and chaotic key access patterns
   over the course of a long-running simulation.

### 2017-11-15 paper/sections/4_dynamic-policies.tex
paper: addresses #23
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Detecting these superbasins can lead to more effective cache management
-strategies because the height of the keyspace accesses is ``how much" of cache
-to evict and the width of the keyspace accesses is ``when" to evict values from
+strategies because the height of the groups of key accesses is ``how much" of cache
+to evict and the width of the groups of key accesses is ``when" to evict values from
 the cache.  The zoomed portion of Figure~\ref{fig:keyspace-zoomed} shows how a
@@
   \item Statistics: decay on each key counts down until 0 0-valued keys are
-evicted.  ``history-of-keyspace", set to 10 seconds, to evict keys.
+evicted.  ``history-of-key-accesses", set to 10 seconds, to evict keys.
 
@@
 Figure~\ref{src:dyn-cache} shows the algorithm we use to detect the key access
-patterns. Iterating backwards through the keyspace trace, at each time step we
+patterns. Iterating backwards through the key access trace, at each time step we
 find the lowest key ID and compare against the lowest key ID we have seen so
@@
 can tell the ``how much" policy the number of keys to evict (Line 19).  The
-algorithm iterates backwards over the keyspace trace because a change in the
+algorithm iterates backwards over the key access trace because a change in the
 minimum value signals a new fan. No signal exists iterating left to right, as

### 2017-11-15 paper/sections/5_file-systems.tex
paper: addresses #23
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 a single metadata server, and (2) helps identify which parts of the cache to
-keep local to a server. For example, applying the keyspace access pattern
+keep local to a server. For example, applying the key access pattern
 detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
@@
 mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
-three caching subsystems mentioned above already maintain keyspace access
+three caching subsystems mentioned above already maintain key access
 traces. We hypothesize that since this is all software defined caching,

### 2017-11-16 paper/sections/1_introduction.tex
paper: fix logic flaw in intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 workloads common in HPC~\cite{ren:sc2014-indexfs, patil:fast2011-giga+}. But
-these techniques are specific to the domain and application's behavior and a
-new use case may prove to be poor a match for the selected caching
+these techniques have policies, such as when to evict and what data to evict,
+that are specific to the domain and application's behavior, so a new use case
+may prove to be poor a match for the selected caching
 policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
 weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
-variety of caching policies using a general framework that supports policy
-switching and arrive at a customized policy that works well for our target
-application and storage system.  Trying general policies and iterating to a
-very customized solution confirms that our prototype can adapt to different
-types of workloads; but more importantly, this is the second system and domain
-that we have successfully used this policy language and engine in. This breadth
-helps us reason about more general data management policies by giving us a
-framework to compare and contrast policies from other domains and services.
+variety of caching policies using a data management language/policy engine and
+arrive at a customized policy that works well for our target application.
+Trying general policies and iterating to a very customized solution confirms
+that our prototype can adapt to different types of workloads, but more
+importantly, the collection of policies we developed were similar to policies
+we designed for file system metadata load balancing, which suggests that there
+are generalities in data management policy design that can be used across
+domains.
+
+%very customized solution confirms that our prototype can adapt to different
+%types of workloads; but more importantly, this is the second system and domain
+%that we have successfully used this policy language and engine in. This breadth
+%helps us reason about more general data management policies by giving us a
+%framework to compare and contrast policies from other domains and services.
 

### 2017-11-16 paper/sections/1_introduction.tex
paper: add parameter sweeps for dynamically sized cache policy
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Storage systems use software-based caches extensively to improve performance
-and resource utilization but the policies that guide what data to evict and
-when to evict vary with the application or use case. For example, caching file
-system metadata on both clients and servers reduces the number of remote
-procedure calls and drastically improves the performance of create-heavy
-workloads common in HPC~\cite{ren:sc2014-indexfs, patil:fast2011-giga+}. But
-these techniques have policies, such as when to evict and what data to evict,
-that are specific to the domain and application's behavior, so a new use case
-may prove to be poor a match for the selected caching
+but the policies that guide what data to evict and when to evict vary with the
+application or use case. For example, caching file system metadata on both
+clients and servers reduces the number of remote procedure calls and
+drastically improves the performance of create-heavy workloads common in
+HPC~\cite{ren:sc2014-indexfs, patil:fast2011-giga+}. But these techniques have
+policies for when to evict and what data to evict that are specific to the
+application's behavior, so a new use case may prove to be poor a match for the
+selected caching
 policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,

### 2017-11-16 paper/sections/4_dynamic-policies.tex
paper: add parameter sweeps for dynamically sized cache policy
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-%\subsection{Trajectory Length}
+a\subsection{Trajectory Length}
 %
@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
-
 \caption{The performance and resource utilization trade-off for different cache
@@
 The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
-three reason: (1) segments take longer to generate later in the run, (2) the
-memory footprint is the value at the end of 2.5 hours, and (3) this policy only
-works well for the 2.5 hour run.  For (1), the curving down of the simulation
-vs. wall-clock time is shown in Figure~\ref{fig:methodology-trajectory}; as the
-nanoparticle grows it takes longer to generate segments so by the time we reach
-2 hours, over 90\% of the trajectory is already generated.  For (2), the memory
-footprint is around 0.4GB until we reach 100K key switch threshold. In
-Figures~\ref{fig:methodology-tradeoff} we plot the final value. For (3),
-Figure~\ref{fig:methodology-trajectory} shows that the cache fills up with 100K
-keys at time 7200 seconds and its size is reduced to the size listed in the
-legend.  The curves stay close to ``Unlimited" for up to an hour after the
-cache is reduced but eventually flatten out as the persistent database gets
-overloaded. 10K and 100K follow the ``Unlimited" curve the longest and are
-sufficient policies for the 2.5 hour runs but anything longer would need a
-different multi-policy cache management strategy.
-
-\begin{figure}[t]
-  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/methodology-trajectory.png}\\
-  \caption{The rate that the trajectory is computed decays over time (which is
-  expected) but this skews the performance improvements in the ``Multi-Part Policy" graph in 
-  Figure~\ref{fig:methodology-tradeoff}. Our multi-policy cache management strategy works for 2.5
-  hour jobs but not for 4 hour jobs.  \label{fig:methodology-trajectory}}
-\end{figure}
+two reasons: (1) segments take longer to generate later in the run and (2) the
+memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
+length vs vs.  wall-clock time curves down over time; as the nanoparticle grows
+it takes longer to generate segments so by the time we reach 2 hours, over 90\%
+of the trajectory is already generated.  For (2), the memory footprint is rises
+until we reach 100K key switch threshold at 0.4GB, as shown by the
+``\(\Delta_1\), Multi-Policy" curve in Figure~\ref{fig:memory-vs-time} but in
+Figure~\ref{fig:methodology-tradeoff} we plot the final value. As a result, we
+can only conclude that this policy only works well for this 2.5 hour run.
 
@@
 
+\end{figure}
+
+\begin{figure}[t]
+\noindent\includegraphics[width=0.5\textwidth]{figures/dscache-vs-none.png}\\
+
+\caption{The performance and resource utilization trade-off for the dynamically
+sized cache (DSCache) policy compared to no cache management. With a negligible
+performance impact, DSCache adjusts to different initial conditions
+(\(\Delta\)) automatically and saves over 2\(\times\) memory in the best case.
+\label{fig:dscache-vs-none}}
+\end{figure}
+
+\begin{figure}[t]
+\noindent\includegraphics[width=0.5\textwidth]{figures/memory-vs-time.png}\\
+
+\caption{Memory utilization of different cache management policies.  ``No Cache
+Management" allows unlimited cache growth, ``Multi-Policy" absorbs the initial
+burstiness of the workload before switching to a more constrained cache, and
+the ``Dynamic Policy" curves size the cache according to key access patterns.
+We show a \(\Delta_2\) growth rate to demonstrate the dynamic policy's ability
+to adjust to a different set of initial conditions.\label{fig:memory-vs-time}}
 \end{figure}
@@
 
+The performance and memory utilization is shown by the ``DSCache" bars in
+Figure~\ref{dscache-vs-none.png}. Without sacrificing performance (trajectory
+length), the dynamically sized cache policy uses between 32\%-66\% less memory
+than the default ParSplice configuration (no cache management) for the 3
+initial conditions we test. The memory usage is shown by the ``Dynamic Policy"
+curves in Figure~\ref{figures/memory-vs-time.png}, where the behavior mirrors
+the key access patterns in Figure~\ref{fig:keyspace-zoomed}.
+
 %For points \(z\), \(y\), and \(x\), if the

### 2017-11-16 experiments/regime-detection-parmsweep/compress.sh

@@ -1,0 +1,25 @@
+cale-workers-8-d1M-164573!/bin/bash
+
+set -x
+jobs=`find . -type d -maxdepth 1 | grep scale`
+for j in $jobs; do
+  rm -r $j/cache-policies/semanticPerf.*
+#  rm -rf scale-*/db0
+#  mv $j.out $j
+#  mv $j.err $j
+#
+#  cd $j/out;
+#  mkdir perf;
+#  find . -type f -name 'perf.*' -maxdepth 1 -exec mv {} perf/ \;
+#  for i in `ls`; do
+#    tar czf $i.tar.gz $i
+#    rm -r $i
+#  done
+#  cd -
+#
+#  cd $j
+#  tar czf times.tar.gz times.out
+#  tar czf traj.tar.gz traj.out
+#  rm -f times.out traj.out
+#  cd -
+done

### 2017-11-16 paper/sections/3_methodology.tex
paper: addresses #17, #25, and #31
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 For example, a ``when" policy can avoid trimming a cache or migrating data if
-it had performed that operation in the previous decision using:
+it had performed that operation in the previous decision.
 
-\begin{figure}[h]
-\footnotesize
-\begin{minted}{lua}
-if RDState() == 1 then
-  WRState(0)   -- the next decision will return true
-  return false
-else then
-  WRState(1)   -- the next decision will return false
-  return true
-end
-\end{minted}
-\end{figure}
+%\begin{figure}[h]
+%\footnotesize
+%\begin{minted}{lua}
+%if RDState() == 1 then
+%  WRState(0)   -- the next decision will return true
+%  return false
+%else then
+%  WRState(1)   -- the next decision will return false
+%  return true
+%end
+%\end{minted}
+%\end{figure}
 

### 2017-11-16 paper/sections/4_dynamic-policies.tex
paper: addresses #17, #25, and #31
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-a\subsection{Trajectory Length}
+\subsection{Trajectory Length}
 %
@@
 \label{sec:dom-specific}
-\begin{figure}[t]
-  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/keyspace-zoomed.png}\\
 
+\begin{figure*}[t!]
+    \begin{subfigure}[t]{0.32\textwidth}
+	\includegraphics[width=0.9\textwidth]{figures/keyspace-zoomed.png}\\
 \caption{Key activity for a 4 hour run shows groups of accesses to the same
 subset of keys. Detecting these access patterns leads to a more accurate cache
-  management strategy.\label{fig:keyspace-zoomed}}
+management strategy, which is discussed in
+Section~\S\ref{sec:regime-detection}.\label{fig:keyspace-zoomed}}
+    \end{subfigure}%
+    ~ 
+    \begin{subfigure}[t]{0.32\textwidth}
+        \includegraphics[width=1\textwidth]{figures/memory-vs-time.png}\\
+	\caption{Memory utilization for ``No Cache Management" (unlimited cache
+growth), ``Multi-Policy" (absorbs initial burstiness of workload) , and
+``Dynamic Policy" (sizes cache according to key access patterns).
+\label{fig:memory-vs-time}}
+    \end{subfigure}%
+    ~
+    \begin{subfigure}[t]{0.32\textwidth}
+        \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
+	\caption{Performance/Utilization for dynamically sized cache (DSCache)
+policy. With negligible performance degradation, DSCache adjusts to different
+\(\Delta\)s over 2\(\times\) memory in the best case.
+\label{fig:dscache-vs-none}}
+    \end{subfigure}%
+    \caption{Different cache management policies tested over the Mantle policy engine.}
+\end{figure*}
 
-\end{figure}
-
-\begin{figure}[t]
-\noindent\includegraphics[width=0.5\textwidth]{figures/dscache-vs-none.png}\\
-
-\caption{The performance and resource utilization trade-off for the dynamically
-sized cache (DSCache) policy compared to no cache management. With a negligible
-performance impact, DSCache adjusts to different initial conditions
-(\(\Delta\)) automatically and saves over 2\(\times\) memory in the best case.
-\label{fig:dscache-vs-none}}
-\end{figure}
-
-\begin{figure}[t]
-\noindent\includegraphics[width=0.5\textwidth]{figures/memory-vs-time.png}\\
-
-\caption{Memory utilization of different cache management policies.  ``No Cache
-Management" allows unlimited cache growth, ``Multi-Policy" absorbs the initial
-burstiness of the workload before switching to a more constrained cache, and
-the ``Dynamic Policy" curves size the cache according to key access patterns.
-We show a \(\Delta_2\) growth rate to demonstrate the dynamic policy's ability
-to adjust to a different set of initial conditions.\label{fig:memory-vs-time}}
-\end{figure}
+%\begin{figure}[t]
+%  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/keyspace-zoomed.png}\\
+%
+%  \caption{Key activity for a 4 hour run shows groups of accesses to the same
+%  subset of keys. Detecting these access patterns leads to a more accurate cache
+%  management strategy.\label{fig:keyspace-zoomed}}
+%
+%\end{figure}
+%
+%\begin{figure}[t]
+%\noindent\includegraphics[width=0.5\textwidth]{figures/dscache-vs-none.png}\\
+%
+%\caption{The performance and resource utilization trade-off for the dynamically
+%sized cache (DSCache) policy compared to no cache management. With a negligible
+%performance impact, DSCache adjusts to different initial conditions
+%(\(\Delta\)) automatically and saves over 2\(\times\) memory in the best case.
+%\label{fig:dscache-vs-none}}
+%\end{figure}
+%
+%\begin{figure}[t]
+%\noindent\includegraphics[width=0.5\textwidth]{figures/memory-vs-time.png}\\
+%
+%\caption{Memory utilization of different cache management policies.  ``No Cache
+%Management" allows unlimited cache growth, ``Multi-Policy" absorbs the initial
+%burstiness of the workload before switching to a more constrained cache, and
+%the ``Dynamic Policy" curves size the cache according to key access patterns.
+%We show a \(\Delta_2\) growth rate to demonstrate the dynamic policy's ability
+%to adjust to a different set of initial conditions.\label{fig:memory-vs-time}}
+%\end{figure}
 
@@
 The performance and memory utilization is shown by the ``DSCache" bars in
-Figure~\ref{dscache-vs-none.png}. Without sacrificing performance (trajectory
+Figure~\ref{fig:dscache-vs-none}. Without sacrificing performance (trajectory
 length), the dynamically sized cache policy uses between 32\%-66\% less memory
@@
 initial conditions we test. The memory usage is shown by the ``Dynamic Policy"
-curves in Figure~\ref{figures/memory-vs-time.png}, where the behavior mirrors
-the key access patterns in Figure~\ref{fig:keyspace-zoomed}.
+curves in Figure~\ref{fig:memory-vs-time}, where the behavior mirrors the key
+access patterns in Figure~\ref{fig:keyspace-zoomed}. We also show a
+\(\Delta_2\) growth rate to demonstrate the dynamic policy's ability to adjust
+to a different set of initial conditions.
 
@@
 \begin{minted}[xleftmargin=3em,linenos]{lua}
-d = timeseries()          -- data (d) provided by mantle
-ts, id = d:get(d:size())  -- functions provided by mantle
+d = timeseries()          -- provided by mantle
+ts, id = d:get(d:size())  -- provided by mantle
 fan  = {start=nil, finish=ts, top=0, bottom=id}
 fans = {}
-for i=d:size(),1,-1 do -- detect possible fans
+for i=d:size(),1,-1 do    -- detect all fans
   ts, id = d:get(i)

### 2017-11-17 paper/sections/1_introduction.tex
paper: addresses #27 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 but the policies that guide what data to evict and when to evict vary with the
-application or use case. For example, caching file system metadata on both
-clients and servers reduces the number of remote procedure calls and
-drastically improves the performance of create-heavy workloads common in
-HPC~\cite{ren:sc2014-indexfs, patil:fast2011-giga+}. But these techniques have
-policies for when to evict and what data to evict that are specific to the
-application's behavior, so a new use case may prove to be poor a match for the
-selected caching
+use case. For example, caching file system metadata on clients and servers
+reduces the number of remote procedure calls and improves the performance of
+create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
+patil:fast2011-giga+}. But these the policies for when to evict and what data
+to evict are specific to the application's behavior so a new use case may prove
+to be poor a match for the selected caching
 policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
@@
 variety of caching policies using a data management language/policy engine and
-arrive at a customized policy that works well for our target application.
-Trying general policies and iterating to a very customized solution confirms
-that our prototype can adapt to different types of workloads, but more
-importantly, the collection of policies we developed were similar to policies
-we designed for file system metadata load balancing, which suggests that there
-are generalities in data management policy design that can be used across
-domains.
+arrive at a customized policy that works well for our target application.  This
+process of trying general policies and quickly iterating to a customized
+solution shows that our prototype can adapt to different of workloads.
 
@@
 characteristics that are required to accurately simulate phenomena of interest.
-Unfortunately, this simulations of this size saturate the capcity and bandwidth
+Unfortunately, this simulations of this size saturate the capacity and bandwidth
 capabilities of a single node so we need more effective data management
@@
 \caption{Using our data management language and policy engine, we designed a
-dynamically sized caching policy (dashed line) for ParSplice. The solution uses
-knowledge about the application to detect key access patterns and adjust the
-cache accordingly. Compared to existing ParSplice configurations (solid lines),
-our solution saves the most memory without sacrificing performance.
+dynamically sized caching policy (thick line) for ParSplice.  Compared to
+existing configurations (thin lines with \(\times\)'s), our solution saves the most
+memory without sacrificing performance and works for a variety of inputs.
 \label{fig:cache-management}}
@@
 The biggest challenge for ParSplice is properly sizing the caches in the
-storage hierarchy.  The memory usage for a ParSplice cache that stores molecule
-coordinates is shown in Figure~\ref{fig:cache-management}. The solid lines are
-the existing cache management policies, exposed as configurations in ParSplice.
-The default ParSplice configuration uses an unlimited sized cache, shown by the
-``No Cache Management" line, but using this much memory for one cache is
-unacceptable for HPC environments, where memory is precious and a common goal
-is to keep memory for such data structures below 3\%\footnote{Empirically, we
-find this threshold works well for most applications}.  While users can
-configure ParSplice to resize the cache when it reaches a certain threshold,
-this solution requires tuning and parameter sweeps to find the optimal cache
-size that sacrifices negligible performance.  Furthermore, this threshold
-changes with different initial configurations and clusters, as shown by our
-keyspace analysis in Section~\S\ref{sec:parsplice-keyspace-analysis}.  The
-``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
-poorly configured cache can save memory but at the expense of performance.  Our
-dynamically sized cache, shown by the dashed line in
-Figure~\ref{fig:cache-management}, detects key access patterns and
-re-sizes the cache accordingly. The design of this policy is driven by a
-detailed analysis of the key-value accesses over the course of a long running
-simulation across a variety of initial conditions.  Without tuning or parameter
-sweeps, our solution saves more memory than a hand-tuned cache without
-sacrificing performance.  More importantly, it works for a variety initial
-conditions without changing the policy itself.
+storage hierarchy.  The memory usage for a single cache that stores molecule
+coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
+lines marked with \(\times\)'s are the existing configurations in ParSplice.
+The default configuration uses an unlimited sized cache, shown by the ``No
+Cache Management" line, but using this much memory for one cache is
+unacceptable for HPC environments, where a common goal is to keep memory for
+such data structures below 3\%\footnote{Anecdotally, we find this threshold
+works well for most applications.  For reference, a 1GB cache for a distributed
+file system is too large in LANL deployments.}. Since ParSplice deploys a cache
+per 300 worker processes, large simulations will not scale. Users can configure
+ParSplice to evict data when the cache reaches a threshold but this solution
+requires tuning and parameter sweeps; the ``Cache (too small)" curve in
+Figure~\ref{fig:cache-management} shows how a poorly configured cache can save
+memory but at the expense of performance.  Even worse, this threshold changes
+with different initial configurations and clusters so tuning needs to be done
+for all system permutations.  Our dynamically sized cache, shown by the think
+line in Figure~\ref{fig:cache-management}, detects key access patterns and
+re-sizes the cache accordingly.  Without tuning or parameter sweeps, our
+solution saves more memory than a hand-tuned cache with a negligible
+performance degradation (within the baseline's standard deviation) and works
+for a variety of initial conditions.
 
 % What is Mantle
-To design more flexible cache management policies, like our dynamically sized
-cache solution in Figure~\ref{fig:cache-management}, we use the data management
-language and policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.
-This framework allows us to dynamically explore the effects of different
-software-defined cache management strategies for the changing key-value
-workloads generated by ParSplice.  Mantle was originally touted as a
-programmable file system metadata load balancer, but we realize now that the
-collection of abstractions designed for file systems was a control plane that
+To design more flexible cache management policies, like our solution in
+Figure~\ref{fig:cache-management}, we use the data management language and
+policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  Mantle was
+originally touted as a programmable file system metadata load balancer, but we
+realize now that the collection of abstractions was a control plane that
 improved metadata access. So in this paper we refer to Mantle as a policy
@@
 this approach is useful for reasoning about and designing different cache
-management strategies in ParSplice.  Developers write policies for ``when" they
-want data moved and ``how much" of the data to move, then the framework
-executes these policies whenever a decision needs to be made.  These
-abstractions help developers unfamiliar with the domain quickly deploy new
-policies {\it or} well-understood techniques from other domains.  We show that
-Mantle:
+management strategies in ParSplice.  We show that Mantle:
 
@@
   \item decomposes cache management into independent policies that can be
-  dynamically changed, making the problem more manageable and facilitating rapid
-  development. Changing the policy in use is critical in applications such as
-  ParSplice that have alternating stable and chaotic key access patterns
-  over the course of a long-running simulation.
+  dynamically changed, making the problem more manageable. Even developers
+  unfamiliar with the domain can quickly deploy policies.
 
-  \item can be used to quickly deploy a variety of cache management strategies,
-  ranging from basic algorithms and heuristics to statistical models and machine
+  \item can be used to deploy a variety of cache management strategies ranging
+  from basic algorithms and heuristics to statistical models and machine
   learning.

### 2017-11-17 paper/sections/1_introduction.tex
paper: addresses #37 and #38
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 engine that injects policies written in our data management language directly
-into a running service, such as a file system or key-value store, and show how
-this approach is useful for reasoning about and designing different cache
-management strategies in ParSplice.  We show that Mantle:
+into a running service, such as a file system or key-value store.  Rather than
+co-designing a policy directly into the service, which requires domain-specific
+knowledge to find and change policies, we expose the policies in a general way
+so even developers unfamiliar with the domain can deploy quickly deploy solutions.
+We show that our framework:
 
@@
   \item decomposes cache management into independent policies that can be
-  dynamically changed, making the problem more manageable. Even developers
-  unfamiliar with the domain can quickly deploy policies.
+  dynamically changed, making the problem more manageable and easier to reason
+  about.
 
-  \item can be used to deploy a variety of cache management strategies ranging
-  from basic algorithms and heuristics to statistical models and machine
-  learning.
+  \item can deploy a variety of cache management strategies ranging from basic
+  algorithms and heuristics to statistical models and machine learning.
 
   \item has useful primitives that, while designed for file system metadata
-  load balancing, turn out to also be effective for cache management. This
-  finding shows how the policy engine generalizes to different domains and
-  enables control of policies by machines instead of administrators.
+  load balancing, turn out to also be effective for cache management.
 

### 2017-11-17 paper/sections/4_dynamic-policies.tex
paper: addresses #37 and #38
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
-of keys, which we call an access ``fan" in subsequent sections, occurs because
-molecules are stuck in deep trajectories. Recall that the in-memory database
-stores the molecules' EOM minima, which is the smallest effective energy that a
-molecule observes during its trajectory. So molecules stuck in deep
-trajectories explore the same minima until they can escape to a new set of
-states. This exploration of the same set of states is called a superbasin.
+of keys occurs because molecules are stuck in deep trajectories. Recall that
+the in-memory database stores the molecules' EOM minima, which is the smallest
+effective energy that a molecule observes during its trajectory. So molecules
+stuck in deep trajectories explore the same minima until they can escape to a
+new set of states. This exploration of the same set of states is called a
+superbasin.
 
@@
 proliferated more parameters that need to be tuned. Furthermore, we find that
-many of the metrics do not signal a new fans, that our metrics are not accurate
-enough, and that even if we tune all parameters they still will not work for
-all initial conditions.  Below, we indicate which parameters we need to add for
-each technique and the value we find to work best, via tuning and parameter
-sweeps, for one set of initial conditions.
+many of the metrics do not signal a new set of key accesses, that our metrics
+are not accurate enough, and that even if we tune all parameters they still
+will not work for all initial conditions.  Below, we indicate which parameters
+we need to add for each technique and the value we find to work best, via
+tuning and parameter sweeps, for one set of initial conditions.
 
@@
   \item Calculus: use derivative to strip away magnitudes; use large positive
-slopes followed large negative slope as signal for new fan. ``Zero-crossing"
-for distance between small/large spikes to avoid false positives, set to 40
-seconds; `window size", set to 200 seconds, for the size of the moving average
+slopes followed large negative slope as signal for new set of key accesses.
+``Zero-crossing" for distance between small/large spikes to avoid false
+positives, set to 40 seconds; `window size", set to 200 seconds, for the size
+of the moving average
 
   \item K-Means Clustering fails because ``K" is not known {\it a-priori} and
-fans are different size. ``K", set to 4, for the number of clusters in the data
+groups of key accesses are different size. ``K", set to 4, for the number of clusters in the data
 using the sum of the distances to the centroid
@@
 
-Figure~\ref{src:dyn-cache} shows the algorithm we use to detect the key access
-patterns. Iterating backwards through the key access trace, at each time step we
-find the lowest key ID and compare against the lowest key ID we have seen so
-far (Line 7). We also maintain the top and bottom of each fan (Line 13) so we
-can tell the ``how much" policy the number of keys to evict (Line 19).  The
-algorithm iterates backwards over the key access trace because a change in the
-minimum value signals a new fan. No signal exists iterating left to right, as
-the maximum value always increases and the minimum values at the bottom of each
-fan are sparse.  For example, the maximum distance between values along the
-bottom edge of the zoomed fan in Figure~\ref{fig:keyspace-zoomed} is 125
-seconds, while the maximum distance between minimum values for the fan before
-is 0 seconds. As a result of this sparseness, iterating left to right requires
-a ``window size" parameter to determine when we think a minimum value will not
-show up again.  We exchange runtime, in the form of an O(\(n\)) algorithm where
-\(n\) is the number events, for simplicity because this approach avoids adding
-new thresholds for fan detection ({\it e.g.}, space between key accesses, space
-between key IDs, and window size of consecutive key accesses).
+Figure~\ref{src:dyn-cache} shows the algorithm we use to detect groups of key
+access patterns, which we call ``fans". Iterating backwards through the key
+access trace, at each time ste`p we find the lowest key ID and compare against
+the lowest key ID we have seen so far (Line 7). We also maintain the top and
+bottom of each group of key accesses (Line 13) so we can tell the ``how much"
+policy the number of keys to evict (Line 19).  The algorithm iterates backwards
+over the key access trace because a change in the minimum value signals a new
+group of key accesses. No signal exists iterating left to right, as the maximum
+value always increases and the minimum values at the bottom of each group of
+key accesses are sparse.  For example, the maximum distance between values
+along the bottom edge of the zoomed group of key accesses in
+Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the maximum distance
+between minimum values for the group of key accesses before is 0 seconds. As a
+result of this sparseness, iterating left to right requires a ``window size"
+parameter to determine when we think a minimum value will not show up again.
+We exchange runtime, in the form of an O(\(n\)) algorithm where \(n\) is the
+number events, for simplicity because this approach avoids adding new
+thresholds for key access pattern detection ({\it e.g.}, space between key
+accesses, space between key IDs, and window size of consecutive key accesses).
 
@@
 to a different set of initial conditions.
-
-%For points \(z\), \(y\), and \(x\), if the
-%local minimum is the same we are in a regime.  Processing \(y\), we set the
-%local minimum to be \(min(y, m_l)\), where \(m_l\) is the local minimum of the
-%previous time step of \(z\).  The algorithm incorrectly detects a regime change
-%if the local minimum of \(y\) is lower than local minimum of \(z\), since \(y\)
-%may have points {\it within} \(z\); recall that we are trying to detect the
-%whole fan, not just the bottom edge of each fan.
 
@@
 and detects when key accesses move on to a new subset of
-keys.\label{src:dyn-cache}}
+keys. We refer to groups of key accesses as ``fans".\label{src:dyn-cache}}
 \end{figure}

### 2017-11-17 paper/sections/6_conclusion.tex
paper: addresses #17, #25, and #31
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #37 and #38
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-11-08 paper/sections/6_conclusion.tex
@@
 fundamental services from within the storage system ({\it e.g.}, consensus) to
-the application.
+the application.  Our plan is to use
+MDHIM~\cite{greenberg:hotstorage2015-mdhim} as our back-end key-value store
+because it was designed for HPC and has the proper mechanisms for migration
+already implemented.
 
@@
 
-Our plan is to use MDHIM~\cite{greenberg:hotstorage2015-mdhim} as our back-end
-key-value store because it was designed for HPC and has the proper mechanisms
-for migration already implemented.
-
 \section{Future Work}
@@
 constructed automatically by examining the collection of existing policies.
-Such work is made possible with Mantle's ability to dynamically change
-policies.
+Ultimately, we hope that this automation enables control of policies by
+machines instead of administrators.
 
 \section{Conclusion}
+
+Data management encompasses a wide range of techniques that vary by domain and
+service. Yet, the techniques require policies that shape the decision making
+and finding the best policies is a difficult, multi-dimensional problem.  We
+iterate to a custom solution for our target application that uses workload
+access patterns to size its caches. Without tuning or parameter sweeps, our
+solution saves memory without sacrificing performance for a variety of initial
+conditions, including the scale, duration, configuration, and hardware of the
+simulation. More importantly, rather than attempting to construct a single,
+complex policy that works for a variety of scenarios, we instead use the Mantle
+framework to enable software-defined storage systems to flexibly change
+policies as the workload changes.  We also observe that many of the primitives
+and strategies have enough in common with data management in file systems that
+they both can be expressed with similar semantics and a general policy engine.
+Our prototype lays the groundwork for future work in adaptable policies and
+policy generation, which ushers in an era where policies are controlled by
+machines instead of administrators.

### 2017-11-22 jupyter.sh
paper: fix section 6 headings, add lustre trace
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -2,4 +2,4 @@
 
 set -e -x
 #docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:xkcd
-docker run --rm -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:ea9927c3f043
+docker run --rm -v ~/Downloads:/root -v `pwd`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:ea9927c3f043

### 2017-11-26 paper/sections/4_dynamic-policies.tex
paper: addresses #41, add lustre keyspace reconstruction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
     \begin{subfigure}[t]{0.32\textwidth}
-	\includegraphics[width=0.9\textwidth]{figures/keyspace-zoomed.png}\\
+        \centering
+	\includegraphics[width=0.89\textwidth]{figures/keyspace-zoomed.png}\\
 \caption{Key activity for a 4 hour run shows groups of accesses to the same
@@
     \begin{subfigure}[t]{0.32\textwidth}
+        \centering
         \includegraphics[width=1\textwidth]{figures/memory-vs-time.png}\\
@@
     \begin{subfigure}[t]{0.32\textwidth}
-        \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
+        \centering
+        \includegraphics[width=0.86\textwidth]{figures/dscache-vs-none.png}\\
 	\caption{Performance/Utilization for dynamically sized cache (DSCache)
@@
 initial conditions we test. The memory usage is shown by the ``Dynamic Policy"
-curves in Figure~\ref{fig:memory-vs-time}, where the behavior mirrors the key
-access patterns in Figure~\ref{fig:keyspace-zoomed}. We also show a
-\(\Delta_2\) growth rate to demonstrate the dynamic policy's ability to adjust
-to a different set of initial conditions.
+curves in Figure~\ref{fig:memory-vs-time}, where the behavior resembles the key
+access patterns in Figure~\ref{fig:keyspace-zoomed}\footnote{The memory usage
+is not {\it exactly} the same because these are two different runs;
+Figure~\ref{fig:keyspace-zoomed} has key activity tracing turned on, which
+reduces performance.}.  We also show a \(\Delta_2\) growth rate to demonstrate
+the dynamic policy's ability to adjust to a different set of initial
+conditions.
 

### 2017-11-26 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #23
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresss #36
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
-Materials for the Future initiative. As shown in Figure~\ref{fig:parsplice},
-the phases are:
+Materials for the Future initiative.
+
+\subsection{Background}
+
+As shown in Figure~\ref{fig:parsplice}, the phases are:
 
@@
 Changing growth rates and temperature alters the size, shape, and locality of
-the data store keyspace. Lower temperatures and smaller growth rates create
+the keyspace. Lower temperatures and smaller growth rates create
 hotter keys with smaller keyspaces as many segments are generated in the same
@@
 
-\subsection{ParSplice Keyspace Analysis}
+\subsection{Results and Conclusions}
 \label{sec:parsplice-keyspace-analysis}
@@
 small changes to \(\Delta\) can have a strong effect on the timing and
-frequency with which new EOM minima are discovered and referenced.  The
-ParSplice caches on each node are trimmed when memory pressure reaches a
-threshold but the number of unique keys accessed in
-Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
-should be sufficient.
+frequency with which new EOM minima are discovered and referenced. This finding
+suggests that we need a flexible policy language and engine.
 
-The bars in
-Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many reads
-(\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same key
-for extended periods because the trajectory can remain stuck in so-called
+The bars in Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many
+reads (\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same
+key for extended periods because the trajectory can remain stuck in so-called
 superbasins composed of tightly connected sets of states. In this case, many
@@
 for space. This structure means that we can learn the key access patterns and adapt the
-storage system (this motivates Section~\S\ref{sec:dom-specific}).
+storage system accordingly (this motivates Section~\S\ref{sec:dom-specific}).

### 2017-11-26 paper/sections/1_introduction.tex
paper: addresses #39
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 patil:fast2011-giga+}. But these the policies for when to evict and what data
-to evict are specific to the application's behavior so a new use case may prove
-to be poor a match for the selected caching
+to evict are specific to the application's behavior and the system's
+configuration (hardware, settings, etc.) so a new use case may prove to be a
+poor match for the selected caching
 policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
@@
 process of trying general policies and quickly iterating to a customized
-solution shows that our prototype can adapt to different of workloads.
+solution shows that our prototype can adapt to different workloads.
 
@@
 
-Our target application is ParSplice~\cite{perez:jctc20150parsplice}, a
-molecular dynamics simulation that uses a hierarchy of caches and a single
-persistent key-value store to store both observed minima across a molecule's
-equation of motion (EOM) and the hundreds or thousands of partial trajectories
-calculated each second during a parallel job.  The fine-grained data annotation
+Our target application, ParSplice~\cite{perez:jctc20150parsplice}, is
+representative of these types of applications. ParSplice is a molecular
+dynamics simulation that uses a hierarchy of caches and a single persistent
+key-value store to store both observed minima across a molecule's equation of
+motion (EOM) and the hundreds or thousands of partial trajectories calculated
+each second during a parallel job.  The fine-grained data annotation
 capabilities provided by key-value storage is a natural match for scientific
@@
 characteristics that are required to accurately simulate phenomena of interest.
-Unfortunately, this simulations of this size saturate the capacity and bandwidth
+Unfortunately, simulations of this size saturate the capacity and bandwidth
 capabilities of a single node so we need more effective data management
@@
 such data structures below 3\%\footnote{Anecdotally, we find this threshold
-works well for most applications.  For reference, a 1GB cache for a distributed
+works well for HPC applications.  For reference, a 1GB cache for a distributed
 file system is too large in LANL deployments.}. Since ParSplice deploys a cache
-per 300 worker processes, large simulations will not scale. Users can configure
-ParSplice to evict data when the cache reaches a threshold but this solution
-requires tuning and parameter sweeps; the ``Cache (too small)" curve in
-Figure~\ref{fig:cache-management} shows how a poorly configured cache can save
-memory but at the expense of performance.  Even worse, this threshold changes
-with different initial configurations and clusters so tuning needs to be done
-for all system permutations.  Our dynamically sized cache, shown by the think
-line in Figure~\ref{fig:cache-management}, detects key access patterns and
-re-sizes the cache accordingly.  Without tuning or parameter sweeps, our
+per 300 worker processes, large simulations will not scale.  Users can
+configure ParSplice to evict data when the cache reaches a threshold but this
+solution requires tuning and parameter sweeps; the ``Cache (too small)" curve
+in Figure~\ref{fig:cache-management} shows how a poorly configured cache can
+save memory but at the expense of performance.  Even worse, this threshold
+changes with different initial configurations and clusters so tuning needs to
+be done for all system permutations.  Our dynamically sized cache, shown by the
+think line in Figure~\ref{fig:cache-management}, detects key access patterns
+and re-sizes the cache accordingly.  Without tuning or parameter sweeps, our
 solution saves more memory than a hand-tuned cache with a negligible

### 2017-11-26 paper/sections/5_file-systems.tex
paper: fix section 6 headings, add lustre trace
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #39
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 file system metadata load balancing in~\cite{sevilla:sc15-mantle} with the ones
-designed for cache management in ParSplice. The similarities show how the
-``when"/``where"/``how much" abstractions, data management language, and policy
-engine may be widely applicable to other data management techniques, such as
-QoS, scheduling, and batching.
+we designed and evaluated above for cache management in ParSplice. The
+similarities show how the ``when"/``where"/``how much" abstractions, data
+management language, and policy engine may be widely applicable to other data
+management techniques, such as QoS, scheduling, and batching.
 
-\subsection{Using File System Policies for ParSplice}
+\subsection{Using Load Balancing Policies for Cache Management}
 
@@
 
-\subsection{Using ParSplice Policies for File Systems}
+\subsection{Using Cache Management Policies for Load Balancing}
 
@@
 % Load balancing in FSs
-Mantle was designed for file system metadata load balancing across a cluster of
-dedicated metadata servers, where spreading requests across servers improves
-performance. But another technique to reduce request load is caching. If
-clients and servers maintain a consistent cache, the client can do operations
-locally without contacting the metadata server.  The caches in CephFS have the
-same performance and utilization trade-off as ParSplice, where large caches
-improve performance at the expense of a higher memory footprint. High memory
-utilization is problematic when a single metadata server maintains consistent
-caches with many clients.  Using cache management policies from the previous
-two sections has two benefits for load balancing: (1) increases the capacity of
-a single metadata server, and (2) helps identify which parts of the cache to
-keep local to a server. For example, applying the key access pattern
-detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
-system access pattern has the potential to identify when to migrate keys and
-how many keys to migrate.
+The abstractions in Mantle were designed for file system metadata load
+balancing across a cluster of dedicated metadata servers, where spreading
+requests across servers improves performance. But another technique to reduce
+request load is caching. If clients and servers maintain a consistent cache,
+the client can do operations locally without contacting the metadata server.
+The caches in CephFS have the same performance and utilization trade-off as
+ParSplice, where large caches improve performance at the expense of a higher
+memory footprint. High memory utilization is problematic when a single metadata
+server maintains consistent caches with many clients.  Using cache management
+policies from the previous two sections has two benefits for load balancing:
+(1) increases the capacity of a single metadata server, and (2) helps identify
+which parts of the cache to keep local to a server. For example, applying the
+key access pattern detection algorithm from
+Section~\S\ref{sec:regime-detection} to a similar file system access pattern
+has the potential to identify when to migrate keys and how many keys to
+migrate.
 
@@
 % other use cases caches/tiers 
-Ceph has many other data management techniques that would benefit from the
-caching policies developed in Sections~\S\ref{sec:arch-specific}
-and~\ref{sec:dom-specific}. Administrators can use the policies in ParSplice to
-automatically size and manage cache
+Storage systems have many other data management techniques that would benefit
+from the caching policies developed in Sections~\S\ref{sec:arch-specific}
+and~\ref{sec:dom-specific}. For example, Ceph administrators can use the
+policies in ParSplice to automatically size and manage cache
 tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
@@
 mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
-three caching subsystems mentioned above already maintain key access
-traces. We hypothesize that since this is all software defined caching,
-something more clever than LRU would improve cache utilization without
-sacrificing too much performance.
+three caching subsystems mentioned above already maintain key access traces. We
+hypothesize that since this is all software defined caching, something more
+clever than LRU would improve cache utilization without sacrificing too much
+performance.
 

### 2017-11-26 paper/sections/1_introduction.tex
paper: addresses #40
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 in Figure~\ref{fig:cache-management} shows how a poorly configured cache can
-save memory but at the expense of performance.  Even worse, this threshold
-changes with different initial configurations and clusters so tuning needs to
-be done for all system permutations.  Our dynamically sized cache, shown by the
-think line in Figure~\ref{fig:cache-management}, detects key access patterns
-and re-sizes the cache accordingly.  Without tuning or parameter sweeps, our
-solution saves more memory than a hand-tuned cache with a negligible
-performance degradation (within the baseline's standard deviation) and works
-for a variety of initial conditions.
+save memory but at the expense of performance, which is shown by the
+annotation.  Even worse, this threshold changes with different initial
+configurations and clusters so tuning needs to be done for all system
+permutations.  Our dynamically sized cache, shown by the think line in
+Figure~\ref{fig:cache-management}, detects key access patterns and re-sizes the
+cache accordingly.  Without tuning or parameter sweeps, our solution saves more
+memory than a hand-tuned cache with a negligible performance degradation
+(within the baseline's standard deviation) and works for a variety of initial
+conditions.
 
@@
 Figure~\ref{fig:cache-management}, we use the data management language and
-policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  Mantle was
-originally touted as a programmable file system metadata load balancer, but we
-realize now that the collection of abstractions was a control plane that
-improved metadata access. So in this paper we refer to Mantle as a policy
-engine that injects policies written in our data management language directly
-into a running service, such as a file system or key-value store.  Rather than
-co-designing a policy directly into the service, which requires domain-specific
-knowledge to find and change policies, we expose the policies in a general way
-so even developers unfamiliar with the domain can deploy quickly deploy solutions.
-We show that our framework:
+policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  While Mantle
+was designed for the narrow purpose of filesystem metadata load balancing, this
+paper provides evidence that the approach is more broad.  Specifically, the
+collection of abstractions in Mantle provides a general control plane that
+improves the performance of metadata access. So in this paper we refer to
+Mantle as a policy engine that injects policies written in our data management
+language directly into a running service, such as a file system or key-value
+store.  Rather than co-designing a policy directly into the service, which
+requires domain-specific knowledge to find and change policies, we expose the
+policies in a general way so even developers unfamiliar with the domain can
+deploy quickly deploy solutions.  We show that our framework:
 

### 2017-11-26 paper/sections/5_file-systems.tex
paper: addresses #40
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+These similarities among effective policies for two very different domains
+suggests that the heuristics and techniques in other load balancers can be used
+for cache management. This generalization may reduce the work that needs to be
+done for cache management as ideas may have already been explored and could
+work ``out-of-the-box".
+
 \subsection{Using Cache Management Policies for Load Balancing}

### 2017-11-26 paper/sections/2_parsplice-keyspace-analysis.tex
paper: spellcheck
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 longer than our 10 node CloudLab clusters and \(25\times\) longer than our 10
-node cluster at UCSC. As a result, it reaches different job phases faster and
+node cluster at UC Santa Cruz. As a result, it reaches different job phases faster and
 gives us a more comprehensive view of the workload. The performance gains

### 2017-11-26 paper/sections/4_dynamic-policies.tex
paper: spellcheck
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 access patterns, which we call ``fans". Iterating backwards through the key
-access trace, at each time ste`p we find the lowest key ID and compare against
+access trace, at each time step we find the lowest key ID and compare against
 the lowest key ID we have seen so far (Line 7). We also maintain the top and
@@
 \caption{The dynamically sized cache policy, whose performance is shown in 
-Figure~\ref{fig:cache-management}, iterates backwards over timestamp-keyID pairs
+Figure~\ref{fig:cache-management}, iterates backwards over timestamp-key pairs
 and detects when key accesses move on to a new subset of

### 2017-11-26 paper/sections/1_introduction.tex
paper: spellcheck
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #45
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 dynamics simulation demonstrates the need for more complex cache management
-strategies. Baseline measurements show clear key access patterns and hot
-spots that offer significant opportunity for optimization. We use the data
-management and policy engine from the Mantle system to dynamically explore a
+strategies. Baseline measurements show clear key access patterns and hot spots
+that offer significant opportunity for optimization. We use the data management
+language and policy engine from the Mantle system to dynamically explore a
 variety of techniques, ranging from basic algorithms and heuristics to
@@
 
-Storage systems use software-based caches extensively to improve performance
-but the policies that guide what data to evict and when to evict vary with the
-use case. For example, caching file system metadata on clients and servers
-reduces the number of remote procedure calls and improves the performance of
+Storage systems use software-based caches to improve performance but the
+policies that guide what data to evict and when to evict vary with the use
+case. For example, caching file system metadata on clients and servers reduces
+the number of remote procedure calls and improves the performance of
 create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
-patil:fast2011-giga+}. But these the policies for when to evict and what data
-to evict are specific to the application's behavior and the system's
-configuration (hardware, settings, etc.) so a new use case may prove to be a
-poor match for the selected caching
+patil:fast2011-giga+}. But the policies for when to evict and what data to
+evict are specific to the application's behavior and the system's configuration
+(hardware, settings, etc.) so a new use case may prove to be a poor match for
+the selected caching
 policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
@@
 Our target application, ParSplice~\cite{perez:jctc20150parsplice}, is
-representative of these types of applications. ParSplice is a molecular
-dynamics simulation that uses a hierarchy of caches and a single persistent
-key-value store to store both observed minima across a molecule's equation of
-motion (EOM) and the hundreds or thousands of partial trajectories calculated
-each second during a parallel job.  The fine-grained data annotation
-capabilities provided by key-value storage is a natural match for scientific
-simulations like ParSplice. But these simulations rely on a mesh-based
-decomposition of a physical region and result in millions or billions of mesh
-cells, where each cell contains materials, pressures, temperatures and other
-characteristics that are required to accurately simulate phenomena of interest.
-Unfortunately, simulations of this size saturate the capacity and bandwidth
-capabilities of a single node so we need more effective data management
-techniques, such as cache management or load balancing across a cluster.
+representative of applications that extensively use software-based caches.
+ParSplice is a molecular dynamics simulation that uses a hierarchy of caches
+and a single persistent key-value store to store both observed minima across a
+molecule's equation of motion (EOM) and the hundreds or thousands of partial
+trajectories calculated each second during a parallel job.  The fine-grained
+data annotation capabilities provided by key-value storage is a natural match
+for scientific simulations like ParSplice. But these simulations rely on a
+mesh-based decomposition of a physical region and result in millions or
+billions of mesh cells, where each cell contains materials, pressures,
+temperatures and other characteristics that are required to accurately simulate
+phenomena of interest.  Unfortunately, simulations of this size saturate the
+capacity and bandwidth capabilities of a single node so we need more effective
+data management techniques, such as cache management or load balancing across a
+cluster.
 
@@
 works well for HPC applications.  For reference, a 1GB cache for a distributed
-file system is too large in LANL deployments.}. Since ParSplice deploys a cache
-per 300 worker processes, large simulations will not scale.  Users can
-configure ParSplice to evict data when the cache reaches a threshold but this
-solution requires tuning and parameter sweeps; the ``Cache (too small)" curve
-in Figure~\ref{fig:cache-management} shows how a poorly configured cache can
-save memory but at the expense of performance, which is shown by the
-annotation.  Even worse, this threshold changes with different initial
-configurations and clusters so tuning needs to be done for all system
-permutations.  Our dynamically sized cache, shown by the think line in
-Figure~\ref{fig:cache-management}, detects key access patterns and re-sizes the
-cache accordingly.  Without tuning or parameter sweeps, our solution saves more
-memory than a hand-tuned cache with a negligible performance degradation
-(within the baseline's standard deviation) and works for a variety of initial
-conditions.
+file system is too large in LANL deployments.}. ParSplice deploys a cache per
+300 worker processes, so large simulations will not scale because they need
+multiple caches.  Users can configure ParSplice to evict data when the cache
+reaches a threshold but this solution requires tuning and parameter sweeps; the
+``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
+poorly configured cache can save memory but at the expense of performance,
+which is shown by the annotation to the right.  Even worse, this threshold
+changes with different initial configurations and cluster setups so tuning
+needs to be done for all system permutations.  Our dynamically sized cache,
+shown by the thick line in Figure~\ref{fig:cache-management}, detects key
+access patterns and re-sizes the cache accordingly.  Without tuning or
+parameter sweeps, our solution saves more memory than a hand-tuned cache with a
+negligible performance degradation (within the baseline's standard deviation)
+and works for a variety of initial conditions.
 
@@
 policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  While Mantle
-was designed for the narrow purpose of filesystem metadata load balancing, this
-paper provides evidence that the approach is more broad.  Specifically, the
-collection of abstractions in Mantle provides a general control plane that
+was designed for the narrow purpose of file system metadata load balancing,
+this paper provides evidence that the approach is more broad.  Specifically,
+the collection of abstractions in Mantle provides a general control plane that
 improves the performance of metadata access. So in this paper we refer to
@@
 policies in a general way so even developers unfamiliar with the domain can
-deploy quickly deploy solutions.  We show that our framework:
+quickly deploy solutions.  We show that our framework:
 
@@
 disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
-that the collection of policies we designed for ParSplice's cache management
+that the collection of policies we designed for cache management in ParSplice
 are very similar to the policies used to load balance metadata in the Ceph file

### 2017-11-26 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #45
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/parsplice.png}\\
-\caption{The ParSplice architecture has a storage hierarchy of caches and a
-dedicated cache node (boxes) backed by a persistent database (DB). A splicer
-(S) tells worker (W) to generate segments and workers employ tasks (T) for more
+\caption{The ParSplice architecture has a storage hierarchy of caches (boxes) and a
+dedicated cache process (large box) backed by a persistent database (DB). A splicer
+(S) tells workers (W) to generate segments and workers employ tasks (T) for more
 parallelization. We focus on the worker's cache (circled), which facilitates
@@
 
-Changing growth rates and temperature alters the size, shape, and locality of
-the keyspace. Lower temperatures and smaller growth rates create
-hotter keys with smaller keyspaces as many segments are generated in the same
-set of states before the trajectory can escape to a new region of state space.
-In Figure~\ref{fig:motivation-regimes}, the \(\Delta_1\) growth rate adds atoms
-every 100K microseconds while the \(\Delta_2\) growth rate adds atoms every 1
-million microseconds. So \(\Delta_2\) has a smaller growth rate resulting in
-hotter keys (line on \(y1\) axis) and a smaller keyspace (dots on the \(y2\)
-axis).  Values smaller than \(\Delta_2\)'s growth rate or a temperature of 400
-degrees results in very little database activity because state transitions take
-too long. Similarly, values larger than \(\Delta_1\)'s growth rate or a
-temperature of 4000 degrees result in an equally meaningless simulation as
-transitions are unrealistic.
-
 Our evaluation uses the total ``trajectory length" as the goodness metric. This
@@
 \begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
+  \caption{The keyspace activity for ParSplice using two different growth
+rates.  The line shows the rate that EOM minima values are retrieved from the
+key-value store (\(y1\) axis) and the points along the bottom show the number
+of keys accessed in a 1 second sliding window (\(y2\) axis).
+\label{fig:motivation-regimes}}
+\end{figure}
+
+\begin{figure}[t]
   \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keys.png}\\
@@
 keyspace size (text annotations) and request load (bars) after a one hour run
-with a different number of workers (\(x\) axis). While the keyspace size and
+with a different number of tasks(\(x\) axis). While the keyspace size and
 capacity is relatively modest the memory usage scales linearly with the number
-of workers. This is a problem if we want to scale to Trinitite's 6000 cores.
+of tasks. This is a problem if we want to scale to Trinitite's 3000 cores.
 Furthermore, the size of the keyspace also increases linearly with the length
@@
 Trinitite nodes at 20GB for the cache.  This memory utilization easily eclipses
-the 3\% threshold we set earlier, even without factoring in the memory usage
-from other workers.
-
-\begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
-  \caption{The keyspace activity for ParSplice using two different growth
-rates.  The line shows the rate that EOM minima values are retrieved from the
-key-value store (\(y1\) axis) and the points along the bottom show the number
-of keys accessed in a 1 second sliding window (\(y2\) axis).
-\label{fig:motivation-regimes}}
-\end{figure}
+the 3\% memory usage per node threshold we set earlier, even without factoring
+in the usage from other workers.
 
 \subsubsection*{An active but small keyspace}
+
 Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
@@
 the read request rate (\(y1\) axis) and the dots along the bottom are the
-number of unique keys accessed (\(y2\) axis).  This figure demonstrates that
-small changes to \(\Delta\) can have a strong effect on the timing and
-frequency with which new EOM minima are discovered and referenced. This finding
-suggests that we need a flexible policy language and engine.
+number of unique keys accessed (\(y2\) axis).  The \(\Delta_1\) growth rate
+adds atoms every 100K microseconds while the \(\Delta_2\) growth rate adds
+atoms every 1 million microseconds. So \(\Delta_2\) has a smaller growth rate
+resulting in hotter keys (line on \(y1\) axis) and a smaller keyspace (dots on
+the \(y2\) axis).  Values smaller than \(\Delta_2\)'s growth rate or a
+temperature of 400 degrees results in very little database activity because
+state transitions take too long. Similarly, values larger than \(\Delta_1\)'s
+growth rate or a temperature of 4000 degrees result in an equally meaningless
+simulation as transitions are unrealistic.  This figure demonstrates that small
+changes to \(\Delta\) can have a strong effect on the timing and frequency with
+which new EOM minima are discovered and referenced. This finding suggests that
+we need a flexible policy language and engine to explore these trade-offs.
 
@@
 moves on.  Writes only occur for the final state of segments generated by
-worker tasks; their magnitude is smaller than reads because the caches ignore
+tasks; their magnitude is smaller than reads because the caches ignore
 redundant write requests. The number of read and write requests are highest at
-the beginning of the run when worker tasks generate segments for the same
-state, which is computationally cheap (this motivates
-Section~\S\ref{sec:dom-specific}).
+the beginning of the run when tasks generate segments for the same state, which
+is computationally cheap (this motivates Section~\S\ref{sec:arch-specific}).
 
@@
 (up to \(5\times\)) because worker tasks start generating states with different
-coordinates later in the run (this motivates Section~\S\ref{sec:dom-specific}).
-The growth rate, temperature, and number of workers have a predictable effect
-on the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
-the number of reads changes with different initial conditions (\(\Delta\)), but
-that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
-still \(5\times\) more popular than others).
-Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
-rates has temporal locality, as the reads per second for \(\Delta_2\) looks
-like the reads per second for \(\Delta_1\) stretched out along the time axis.
-Trends also exist for temperature and number of workers but are omitted here
-for space. This structure means that we can learn the key access patterns and adapt the
-storage system accordingly (this motivates Section~\S\ref{sec:dom-specific}).
+coordinates later in the run.  The growth rate, temperature, and number of
+workers have a predictable effect on the structure of the keyspace.
+Figure~\ref{fig:methodology-keys} shows that the number of reads changes with
+different initial conditions (\(\Delta\)), but that the spatial locality of key
+accesses is similar ({\it e.g.}, some keys are still \(5\times\) more popular
+than others).  Figure~\ref{fig:motivation-regimes} shows how entropy for
+different growth rates has temporal locality, as the reads per second for
+\(\Delta_2\) look like the reads per second for \(\Delta_1\) stretched out
+along the time axis.  Trends also exist for temperature and number of workers
+but are omitted here for space. This structure means that we can learn the key
+access patterns and adapt the storage system accordingly (this motivates
+Section~\S\ref{sec:dom-specific}).

### 2017-11-26 paper/sections/3_methodology.tex
paper: spellcheck
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #45
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 %  \item background: CephFS implementation
-To explore software-defined cache management, we use
-the data management language and policy engine presented
-in~\cite{sevilla:sc15-mantle}. The prototype in that paper was called Mantle
-and it was originally built on the CephFS so administrators could control file
-system metadata data management policies. We now refer to Mantle as a policy
-engine that supports our data management language.  The basic premise is that
-data management policies can be expressed with a simple API consisting of
-``when", ``where", and ``how much". ``when" controls how aggressive or
-conservative the decisions are; ``where" lets the policy control how
-distributed or concentrated the data should be; and ``how much" controls the
-amount of data that should be sent. There is also a ``load" policy that lets
-administrators specify how to collapse many metrics into a single load metric
-({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).
+To explore software-defined cache management, we use the data management
+language and policy engine presented in~\cite{sevilla:sc15-mantle}. The
+prototype in that paper was called Mantle and it was originally built on CephFS
+so administrators could control file system metadata data management policies.
+We now refer to Mantle as a policy engine that supports our data management
+language.  The basic premise is that data management policies can be expressed
+with a simple API consisting of ``when", ``where", and ``how much". ``when"
+controls how aggressive or conservative the decisions are; ``where" controls
+how distributed or concentrated the data should be; and ``how much" controls
+the amount of data that should be sent. There is also a ``load" policy that
+lets administrators specify how to collapse many metrics into a single load
+metric ({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).
 
@@
 we move ParSplice to a distributed key-value store back-end, the ``where" policy
-can be used determine which key-value pairs should be moved to which node.
+will be used determine which key-value pairs should be moved to which node.
 
@@
 and described below.
+
+Administrators write policies in
+Lua that are executed whenever the service needs to make a data management
+decision. We chose Lua for simplicity, performance, and portability; it is a
+scripting language with simple syntax, which allows administrators to focus on
+the policies themselves; it was designed as an embeddable language, so it is
+lightweight and does less type checking; and it interfaces nicely with C/C++.
 
@@
 
-% policies themselves
-\subsubsection{Policies Written as Callbacks} administrators write policies in
-Lua that are executed whenever the service needs to make a data management
-decision. We chose Lua for simplicity, performance, and portability; it is a
-scripting language with simple syntax, which allows administrators to focus on
-the policies themselves; it was designed as an embeddable language, so it is
-lightweight and does less type checking; and it interfaces nicely with C/C++.
-The ``callback" box in Figure~\ref{fig:mantle} shows an example policy for the
-``when()" callback in a distributed service, where the current server
-(\texttt{whoami}) migrates load if it is has load (\texttt{>0.1}) and if its
-neighbor server (\texttt{whoami + 1}) does not have load (\texttt{<0.1}). The
-load is calculated using the metrics provided by the environment.
-
-Mantle also provides functions for persisting state across decisions.
-\texttt{WRState(s)} saves state \texttt{s}, which can be a number or boolean
-value, and \texttt{RDState()} returns the state saved by a previous iteration.
-For example, a ``when" policy can avoid trimming a cache or migrating data if
-it had performed that operation in the previous decision.
-
-%\begin{figure}[h]
-%\footnotesize
-%\begin{minted}{lua}
-%if RDState() == 1 then
-%  WRState(0)   -- the next decision will return true
-%  return false
-%else then
-%  WRState(1)   -- the next decision will return false
-%  return true
-%end
-%\end{minted}
-%\end{figure}
-
-%of timestamp (ts), value pairs representing accesses over time types of
-\subsubsection{Environment of Metrics} services can expose cluster metrics for
+\subsubsection{Environment of Metrics} services expose cluster metrics for
 describing resource utilization and time series metrics for describing accesses
@@
 
+% policies themselves
+\subsubsection{Policies Written as Callbacks} the ``callback" box in
+Figure~\ref{fig:mantle} shows an example policy for ``when()", where the
+current server (\texttt{whoami}) migrates load if it is has load
+(\texttt{>0.1}) and if its neighbor server (\texttt{whoami + 1}) does not have
+load (\texttt{<0.1}). The load is calculated using the metrics provided by the
+environment.
+
+Mantle also provides functions for persisting state across decisions.
+\texttt{WRState(s)} saves state \texttt{s}, which can be a number or boolean
+value, and \texttt{RDState()} returns the state saved by a previous iteration.
+For example, a ``when" policy can avoid trimming a cache or migrating data if
+it had performed that operation in the previous decision.
+
+%\begin{figure}[h]
+%\footnotesize
+%\begin{minted}{lua}
+%if RDState() == 1 then
+%  WRState(0)   -- the next decision will return true
+%  return false
+%else then
+%  WRState(1)   -- the next decision will return false
+%  return true
+%end
+%\end{minted}
+%\end{figure}
+
+%of timestamp (ts), value pairs representing accesses over time types of
 \subsection{Integrating Mantle into ParSplice}
@@
 Mantle time series metrics, we expose accesses to the cache as a list of
-\texttt{timestamp, keyid} pairs. In Section~\S\ref{sec:dom-specific}, we
+\texttt{timestamp, key} pairs. In Section~\S\ref{sec:dom-specific}, we
 explore a key access pattern detection algorithm that uses this metric.

### 2017-11-26 paper/sections/4_dynamic-policies.tex
paper: addresses #45
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\subsection{Trajectory Length}
+%\subsection{Trajectory Length}
 %
@@
 Using the Mantle policy engine, we test a variety of cache management tools and
-algorithms on a single in-memory database node using the keyspace analysis in
-Section~\S\ref{sec:parsplice-keyspace-analysis}. These strategies are
-implemented as the ``when" and ``how much" policies from the Mantle API; the
-``where" callback does not make sense for a single node (segments are either in
-the cache or they are not). The evaluation metric is the accuracy and runtime
-of each strategy; the strategy should be accurate enough so as to sacrifice
-negligible performance and fast enough to run as often as we want to detect key
-access patterns. The goal of the following sections is not to find an optimal
-solution, as this can be done with parameter sweeps for thresholds; rather, we
-try to find techniques that work for a range of inputs and system setups.
-
-First we sized the cache according to our system specific knowledge. By
-``system", we mean the hardware and software of the storage hierarchy. We look
-at request rate, unique keys in a sliding window, and bandwidth capabilities. For
-example, we know that LevelDB cannot handle high IO request rates.
+algorithms on the worker using the keyspace analysis in
+Section~\S\ref{sec:parsplice-keyspace-analysis}.  The evaluation metric is the
+accuracy and runtime of each strategy; the policy should be accurate enough so
+as to sacrifice negligible performance and fast enough to run as often as we
+want to detect key access patterns.  First we sized the cache according to our
+system specific knowledge. By ``system", we mean the hardware and software of
+the storage hierarchy. We look at request rate, unique keys in a sliding
+window, and bandwidth capabilities. For example, we know that LevelDB cannot
+handle high IO request rates.
 
 % technical details
-In the original ParSplice implementation, each cache node uses an unlimited
-amount of memory to store segment coordinates. We limit the size of the cache
-using an LRU eviction policy, where the penalty for a cache miss is retrieving
-the data from the persistent database.  We evict keys (if necessary) at every
-operation instead of when segments complete because the cache fills up too
-quickly otherwise.
-
 The results for different cache sizes for a growth rate of \(\Delta_1\) over a
@@
 utilization is measured with memory footprint of just the cache (\(y2\) axis).
-The other graph shares the \(y\) axis and shows the trade-off of using a basic
-LRU-style cache for different cache sizes, implemented using a ``when" policy
-of \texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
+The middle graph shares the \(y\) axis and shows the trade-off of using a basic
+LRU-style cache for different cache sizes, where the penalty for a cache miss
+is retrieving the data from the persistent database.  We evict keys (if
+necessary) at every operation instead of when segments complete because the
+cache fills up too quickly otherwise. This is implemented using a ``when"
+policy of \texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
 \texttt{servers[whoami]['cachesize']-n}.  The error bars are the standard
-deviation of 3 runs. 
+deviation of 3 runs.  Although the keyspace grows to 150K, a 100K key cache
+achieves 99\% of the performance. Decreasing the cache degrades performance and
+predictability.
+
 %\begin{figure}[h] \footnotesize \begin{minted}{lua}
@@
 % why is the performance lower for smaller caches?
-Although the keyspace grows to 150K, a 100K key cache achieves 99\% of the
-performance. Decreasing the cache degrades performance and predictability.
-Despite the memory savings, our results suggest that a multi-policy cache management strategy
-could save even more memory.  The middle plot in
-Figure~\ref{fig:methodology-tradeoff} show that a 100K key cache is sufficient
-as a static policy but the top graph in Figure~\ref{fig:motivation-regimes}
-indicates that the cache size could be much smaller. That graph shows that the
-beginning of the run is characterized by many reads to a small set of keys and
-the end sees much lower reads per second to a larger keyspace. Specifically, it
-shows only about 100 keys as active in the latter half of the run.
-
-After analyzing traces, we see that the 100 key cache is insufficient because
-the persistent database cannot service the read-write traffic.  According to
-Figure~\ref{fig:motivation-regimes}, the read requests arrive at 750 reads per
-second in addition to the writes that land in each tier (about 300 puts/second,
-some redundant). This traffic triggers a LevelDB compaction and reads block,
-resulting in very slow progress.  Traces verify this hypothesis and show reads
-getting backed up as the read/write ratio increases. To recap, small caches
-incur too much load on the persistent database  at the beginning of the run but
-should suffice after the initial read flash crowd passes because
-the keyspace is far less active. This suggests a multi-policy cache management
-strategy.
-
-%% Why is this a good idea
-%Although ParSplice does not use a distributed file system, its workload is very
-%similar because the minima key-value store responds to small and frequent
-%requests, which results in hot spots and flash crowds.  Modern distributed file
-%systems have found efficient ways to measure, migrate, and partition metadata
-%load and have shown large performance gains and
-%better scalability~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs,
-%grider:pdsw2015-marfs, ren:sc2014-indexfs, patil:fast2011-giga+,
-%brandt:msst2003-lh}.  Previous work quantified the speedups achieved with
-%Mantle and formalized balancers that were good for file systems.
-
-The plot on the right side of Figure~\ref{fig:methodology-tradeoff} shows the results of using Mantle
-to program a multi-policy cache management strategy into ParSplice that switches between
-two policies:
+But the top graph in Figure~\ref{fig:motivation-regimes} suggests that a
+smaller cache size should suffice, as only 100 keys seem to be active at any
+one time.  It turns out that the unique keys plotted in
+Figure~\ref{fig:motivation-regimes} are per second and are not representative
+of the actual active keyspace; the number of active keys is larger than 100, as
+some keys may be accessed at time \(t_0\), not in \(t_1\), and then again in
+\(t_2\). Because the cache is too small, reads and writes hit the persistent
+database, where the excessive traffic triggers a LevelDB compaction and reads
+block.  To avoid overloading the storage hierarchy with requests, we design a
+multi-part policy that switches between:
 
@@
 
-The actual policy is shown and described in more detail in the next section in
-Figure~\ref{src:lru-dyn}.  We trigger the policy switch at 100K keys to absorb
-the flash crowd at the beginning of the run. Once triggered, keys are evicted
-to bring the size of the cache down to the threshold.  In the bar chart, the
-cache sizes for the \(n\) key limit policy are along the \(x\) axis.
-
 % results: same level of performance can be achieved 
-The multi-part policies show better performance than the single \(n\) key
-policies. The performance and memory utilization for a 100K key cache size is
-the same as the 100K bar in the ``Multi-Part Policy" graph in  Figure~\ref{fig:methodology-tradeoff} but
-the rest reduce the size of the keyspace after the read flash crowd.  We see
-the worst performance when the engine switches to the 10 key limit policy,
-which achieves 94\% of the performance while only using 40KB of memory.
+The key observation is that small caches incur too much load on the persistent
+database at the beginning of the run but should suffice after the initial read
+flash crowd passes because the keyspace is far less active.  We program Mantle
+to trigger the policy switch at 100K keys to absorb the flash crowd at the
+beginning of the run. Once triggered, keys are evicted to bring the size of the
+cache down to the threshold.  The actual policy is shown and described in more
+detail in the next section in Figure~\ref{src:lru-dyn}.  The plot on the right
+side of Figure~\ref{fig:methodology-tradeoff} shows the performance/utilization
+trade-off of the multi-part policy, where the cache sizes for the \(n\) key
+limit policy are along the \(x\) axis.  They show better performance than the
+single \(n\) key policies. The performance and memory utilization for a 100K
+key cache size is the same as the 100K bar in the ``Multi-Part Policy" graph in
+Figure~\ref{fig:methodology-tradeoff} but the rest reduce the size of the
+keyspace after the read flash crowd.  We see the worst performance when the
+engine switches to the 10 key limit policy, which achieves 94\% of the
+performance while only using 40KB of memory.
 
@@
 memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
-length vs vs.  wall-clock time curves down over time; as the nanoparticle grows
+length vs.  wall-clock time curves down over time; as the nanoparticle grows
 it takes longer to generate segments so by the time we reach 2 hours, over 90\%
-of the trajectory is already generated.  For (2), the memory footprint is rises
+of the trajectory is already generated.  For (2), the memory footprint rises
 until we reach 100K key switch threshold at 0.4GB, as shown by the
@@
 % but the result is still valid
-Despite these caveats, the result is still valid: we found multi-policy cache 
+Despite these caveats, the result is still valid: we found a multi-policy cache 
 management strategy that absorbs the cost of a high read throughput on a small
@@
 Feeding domain-specific knowledge about the ParSplice application into a policy
-leads to more a accurate cache management strategy.
+leads to more a accurate cache management strategy.  The goal of the following
+sections is not to find an optimal solution, as this can be done with parameter
+sweeps for thresholds; rather, we try to find techniques that work for a range
+of inputs and system setups.
+
 Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
-the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
-of keys occurs because molecules are stuck in deep trajectories. Recall that
-the in-memory database stores the molecules' EOM minima, which is the smallest
-effective energy that a molecule observes during its trajectory. So molecules
-stuck in deep trajectories explore the same minima until they can escape to a
-new set of states. This exploration of the same set of states is called a
-superbasin.
+tasks over time (\(x\) axis). The groups of accesses to a subset of keys occurs
+because molecules are stuck in deep trajectories. Recall that the cache stores
+the molecules' EOM minima, which is the smallest effective energy that a
+molecule observes during its trajectory. So molecules stuck in deep
+trajectories explore the same minima until they can escape to a new set of
+states. This exploration of the same set of states is called a superbasin.  In
+Figure~\ref{fig:keyspace-zoomed}, superbasins are never re-visited because the
+simulation only adds molecules; we can never reach a state with less molecules.
+This is why keys are never re-accessed.  Despite these patterns, the following
+characteristics of superbasins make it hard to detect them:
 
 Detecting these superbasins can lead to more effective cache management
-strategies because the height of the groups of key accesses is ``how much" of cache
-to evict and the width of the groups of key accesses is ``when" to evict values from
-the cache.  The zoomed portion of Figure~\ref{fig:keyspace-zoomed} shows how a
-single superbasin affects the key accesses. Moving along the \(x\) axis shows
-that the number of unique keys accessed over time grows while moving along the
-\(y\) axis shows that early keys are accessed more often.  Overall, superbasins
-are never re-visited because the simulation only adds molecules; we can never
-reach a state with less molecules. This is why keys are never re-accessed.
-Despite these patterns, the following characteristics of superbasins make it
-hard to detect them:
+strategies because the height of the groups of key accesses is ``how much" of
+the cache to evict and the width of the groups of key accesses is ``when" to
+evict values from the cache.  The zoomed portion of
+Figure~\ref{fig:keyspace-zoomed} shows how a single superbasin affects the key
+accesses. Moving along the \(x\) axis shows that the number of unique keys
+accessed over time grows while moving along the \(y\) axis shows that early
+keys are accessed more often.
 
@@
 To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
-variety of techniques using Mantle. Unfortunately, we found that they all
-proliferated more parameters that need to be tuned. Furthermore, we find that
-many of the metrics do not signal a new set of key accesses, that our metrics
-are not accurate enough, and that even if we tune all parameters they still
-will not work for all initial conditions.  Below, we indicate which parameters
-we need to add for each technique and the value we find to work best, via
-tuning and parameter sweeps, for one set of initial conditions.
+variety of techniques using Mantle. Unfortunately, we found that the following
+techniques proliferate more parameters that need to be tuned per
+hardware/software configuration. Furthermore, we find that many of the metrics
+do not signal a new set of key accesses.  Below, we indicate with quotes which
+parameters we need to add for each technique with and the value we find to work
+best, via tuning and parameter sweeps, for one set of initial conditions.
 
@@
 
-  \item Statistics: decay on each key counts down until 0 0-valued keys are
+  \item Statistics: decay on each key counts down until 0; 0-valued keys are
 evicted.  ``history-of-key-accesses", set to 10 seconds, to evict keys.
@@
   \item Calculus: use derivative to strip away magnitudes; use large positive
-slopes followed large negative slope as signal for new set of key accesses.
-``Zero-crossing" for distance between small/large spikes to avoid false
-positives, set to 40 seconds; `window size", set to 200 seconds, for the size
-of the moving average
+slopes followed by large negative slope as signal for new set of key accesses.
+``Zero-crossing", set to 40 seconds, for distance between small/large spikes to
+avoid false positives; `window size", set to 200 seconds, for the size of the
+moving average
 
@@
 
-Figure~\ref{src:dyn-cache} shows the algorithm we use to detect groups of key
-access patterns, which we call ``fans". Iterating backwards through the key
-access trace, at each time step we find the lowest key ID and compare against
-the lowest key ID we have seen so far (Line 7). We also maintain the top and
-bottom of each group of key accesses (Line 13) so we can tell the ``how much"
-policy the number of keys to evict (Line 19).  The algorithm iterates backwards
-over the key access trace because a change in the minimum value signals a new
-group of key accesses. No signal exists iterating left to right, as the maximum
-value always increases and the minimum values at the bottom of each group of
-key accesses are sparse.  For example, the maximum distance between values
-along the bottom edge of the zoomed group of key accesses in
+After trying these techniques we found that the basic algorithm in
+Figure~\ref{src:dyn-cache} works best. The algorithm detects groups of key
+access patterns, which we call ``fans", by iterating backwards through the key
+access trace, finding the lowest key ID, and comparing against the lowest key
+ID we have seen so far (Line 7). We also maintain the top and bottom of each
+group of key accesses (Line 13) so we can tell the ``how much" policy the
+number of keys to evict (Line 19).  The algorithm iterates backwards over the
+key access trace because a change in the minimum value signals a new group of
+key accesses. No signal exists iterating left to right, as the maximum value
+always increases and the minimum values at the bottom of each group of key
+accesses are sparse.  For example, the maximum distance between values along
+the bottom edge of the zoomed group of key accesses in
 Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the maximum distance

### 2017-11-26 paper/sections/6_conclusion.tex
paper: addresses #45
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 in this space by providing a framework for exploring these policies, but does
-not attempt anything more sophisticated (e.g., machine learning) to create
+not attempt anything more sophisticated ({\it e.g.}, machine learning) to create
 these policies.
@@
 the development cycle, such as (1) adaptable policies that automatically switch
-to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
+to new strategies when the current strategy behaves poorly ({\it e.g.}, thrashing,
 making no progress, etc.), and (2) policy generation, where new policies are

### 2017-11-27 experiments/lustre-trace/graph_namespace.sh
benchmark: more lustre trace visualizations
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+
+if [ "$#" -ne 1 ]; then
+  echo "USAGE: $0 <filename>"
+  exit 1
+fi
+
+docker run --rm -it \
+  -v `pwd`:/root/ \
+  -w /root \
+  graphviz -Tpng $1 -o $1.png

### 2017-11-27 paper/sections/5_file-systems.tex
paper: outline for #43
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 file system metadata load balancing in~\cite{sevilla:sc15-mantle} with the ones
-we designed and evaluated above for cache management in ParSplice. The
-similarities show how the ``when"/``where"/``how much" abstractions, data
-management language, and policy engine may be widely applicable to other data
-management techniques, such as QoS, scheduling, and batching.
+we designed and evaluated above for cache management in ParSplice.
 
@@
 
+\begin{figure}[t]
+\noindent\includegraphics[width=0.5\textwidth]{figures/trace-atime.png}\\
+\caption{File system metadata reads for a Lustre trace collected at LANL. The
+vertical lines are the access patterns detected by the ParSplice cache
+management policy from Section~\S\ref{sec:dom-specific}.
+\label{fig:trace-atime}}
+\end{figure}
+
 %- how many inodes to keep in mds/client caches
@@
 
-% Load balancing in FSs
-The abstractions in Mantle were designed for file system metadata load
-balancing across a cluster of dedicated metadata servers, where spreading
-requests across servers improves performance. But another technique to reduce
-request load is caching. If clients and servers maintain a consistent cache,
-the client can do operations locally without contacting the metadata server.
-The caches in CephFS have the same performance and utilization trade-off as
-ParSplice, where large caches improve performance at the expense of a higher
-memory footprint. High memory utilization is problematic when a single metadata
-server maintains consistent caches with many clients.  Using cache management
-policies from the previous two sections has two benefits for load balancing:
-(1) increases the capacity of a single metadata server, and (2) helps identify
-which parts of the cache to keep local to a server. For example, applying the
-key access pattern detection algorithm from
-Section~\S\ref{sec:regime-detection} to a similar file system access pattern
-has the potential to identify when to migrate keys and how many keys to
-migrate.
+The cache management policies we developed earlier can be used by load
+balancing policies to effectively spread load across a cluster. For example,
+many distributed file systems spread the hierarchical namespace across a
+dedicated metadata cluster and caching policies could be used to determine what
+metadata to move and when to move it. To demonstrate this idea, consider
+Figure~\ref{fig:trace-atime}, where the dots are the file system metadata reads
+in a 1 hour window of a 3-day Lustre namespace trace and the vertical lines are
+the groups of accesses detected by our policy from
+Section~\S\ref{sec:regime-detection}.  The \(x\) axis is time and the \(y\)
+axis is the file ID, where the file IDs are listed in the order that they
+appear in the trace.  The trace is anonymized so all file names are replaced
+with a unique identifier and we do not know what clients are doing.  Although
+other access pattern detection algorithms are possible, we use the one designed
+for cache management in Section~\S\ref{sec:regime-detection} without
+modification and it successfully detects the largest group of key accesses that
+starts at time 1000 seconds and ends at time 2200 seconds. File systems that
+load balance file system metadata across a cluster would want to keep metadata
+in that group of key accesses on the same server to reduce RPCs.
 
 % burstiness of creates then compile
-To explore this example in more detail, consider the trace of metadata requests
-for compiling code in CephFS shown in Figure~\ref{fig:compile-ops}.  That trace
-shows the number of file system metadata requests serviced by the metadata
+Cache management can also increase the capacity of a single server, which
+affects how load should be distributed across a cluster.  For example, CephFS
+clients and servers maintain a consistent cache so that the client can do
+operations locally without contacting the metadata server.  But large caches
+use more memory, so having smart policies that detect key access patterns would
+greatly reduce memory footprints. For example, Figure~\ref{fig:compile-ops}
+shows the trace of metadata requests for compiling code in CephFS.  The \(y\)
+axis is  the number of file system metadata requests serviced by the metadata
 server when uncompressing (\texttt{untar}), compiling (\texttt{make}), and
@@
 system could cache none of the metadata from the \texttt{untar} phase and run
-key access pattern detection during the \texttt{make} phase, resulting in the metadata
-server/clients only caching metadata that is repeatedly used. For the job in
-Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K inodes
-(the unit of metadata for file systems) instead of 60K, resulting in almost
-40MB of memory savings (since an inode is about
+key access pattern detection during the \texttt{make} phase, resulting in the
+metadata server/clients only caching metadata that is repeatedly used. For the
+job in Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K
+inodes (the unit of metadata for file systems) instead of 60K, resulting in
+almost 40MB of memory savings (since an inode is about
 1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
@@
 \end{figure}
+
+\subsection{Other Use Cases}
 
@@
 
-\subsection{Other Use Cases}
-% we said this is good for LB/Cache management... what about others??
-% QoS: when to move clients, where to move clients, how much of the reservation to move
-% Other cache management: MRU
-% Scheduling: how much of a resource to allocate; more than just LRU (RAD)
-%\subsection{Visualizing File System Traces like ParSplice Keyspace Traces}
+More generally, the similarities between load balancing and cache management
+show how the ``when"/``where"/``how much" abstractions, data management
+language, and policy engine may be widely applicable to other data management
+techniques, such as:
+
+\begin{itemize}
+
+  \item QoS: when to move clients, where to move clients, how much of the
+reservation to move
+
+  \item Scheduling: how much of a resource to allocate; more than just LRU
+(RAD)
+
+  \item Batching: how many operations to group together and when to send large
+batches of updates
+
+  \item Prefetching: how much to prefetch and how to select data to reterive
+(forward/backward/stride detection algorithms for prefetching,
+
+\end{itemize}

### 2017-11-27 paper/sections/1_introduction.tex
paper: addresses #35, #44, #46
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 different domain and service.  Our exploration of this space results in a
-dynamically sized cache policy that, for our initial conditions, sacrifices
-negligible performance while using only 28\% of the memory required by our
-hand-tuned cache.
+dynamically sized cache policy that sacrifices negligible performance while
+using 32-66\% less memory than the default ParSplice configuration.
 
@@
 create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
-patil:fast2011-giga+}. But the policies for when to evict and what data to
+patil:fast2011-giga+}. But the policies for what data to evict and when to
 evict are specific to the application's behavior and the system's configuration
@@
 weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
-variety of caching policies using a data management language/policy engine and
+variety of caching policies using our data management language/policy engine and
 arrive at a customized policy that works well for our target application.  This
@@
 capacity and bandwidth capabilities of a single node so we need more effective
-data management techniques, such as cache management or load balancing across a
-cluster.
+data management techniques.
 
@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
-\caption{Using our data management language and policy engine, we designed a
+\caption{Using our data management language and policy engine, we design a
 dynamically sized caching policy (thick line) for ParSplice.  Compared to
@@
 works well for HPC applications.  For reference, a 1GB cache for a distributed
-file system is too large in LANL deployments.}. ParSplice deploys a cache per
-300 worker processes, so large simulations will not scale because they need
-multiple caches.  Users can configure ParSplice to evict data when the cache
-reaches a threshold but this solution requires tuning and parameter sweeps; the
-``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
-poorly configured cache can save memory but at the expense of performance,
-which is shown by the annotation to the right.  Even worse, this threshold
-changes with different initial configurations and cluster setups so tuning
-needs to be done for all system permutations.  Our dynamically sized cache,
-shown by the thick line in Figure~\ref{fig:cache-management}, detects key
-access patterns and re-sizes the cache accordingly.  Without tuning or
+file system is too large in LANL deployments.}. Furthermore, ParSplice deploys
+a cache per 300 worker processes, so large simulations need more caches and
+will use even more memory.  Users can configure ParSplice to evict data when
+the cache reaches a threshold but this solution requires tuning and parameter
+sweeps; the ``Cache (too small)" curve in Figure~\ref{fig:cache-management}
+shows how a poorly configured cache can save memory but at the cost of
+performance, which is shown by the text annotation to the right.  Even worse, this
+threshold changes with different initial configurations and cluster setups so
+tuning needs to be done for all system permutations.  Our dynamically sized
+cache, shown by the thick line in Figure~\ref{fig:cache-management}, detects
+key access patterns and re-sizes the cache accordingly.  Without tuning or
 parameter sweeps, our solution saves more memory than a hand-tuned cache with a
@@
 this paper provides evidence that the approach is more broad.  Specifically,
-the collection of abstractions in Mantle provides a general control plane that
+the collection of abstractions in Mantle provide a general control plane that
 improves the performance of metadata access. So in this paper we refer to
@@
 disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
-that the collection of policies we designed for cache management in ParSplice
+that the collection of policies we design for cache management in ParSplice
 are very similar to the policies used to load balance metadata in the Ceph file

### 2017-11-27 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #35, #44, #46
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 same coordinates repeatedly. To mitigate the consequences of these repeated
-reads, ParSplice provisions a hierarchy of processes to act as caches that sit
-in front of a single persistent database.  Values are written to each tier and
-reads traverse up the hierarchy until they find the data.  Caches also reside
-on the workers to service reads/writes from its tasks.
+reads, ParSplice provisions a hierarchy of caches that sit in front of a single
+persistent database.  Values are written to each tier and reads traverse up the
+hierarchy until they find the data.
 
@@
 likely to visit any random state.
-
-Our evaluation uses the total ``trajectory length" as the goodness metric. This
-value is the duration of the overall trajectory produced by ParSplice. At
-ideal efficiency, the trajectory length should increase with the square root of
-the wall-clock time, since the wall-clock cost of time-stepping the system by
-one simulation time unit increases in proportion of the total number of atoms.
 
@@
 of the run.  Extrapolating these results puts an 8 hour run across all 100
-Trinitite nodes at 20GB for the cache.  This memory utilization easily eclipses
+Trinitite nodes at 8GB for one cache.  This memory utilization easily eclipses
 the 3\% memory usage per node threshold we set earlier, even without factoring
@@
 atoms every 1 million microseconds. So \(\Delta_2\) has a smaller growth rate
-resulting in hotter keys (line on \(y1\) axis) and a smaller keyspace (dots on
-the \(y2\) axis).  Values smaller than \(\Delta_2\)'s growth rate or a
-temperature of 400 degrees results in very little database activity because
-state transitions take too long. Similarly, values larger than \(\Delta_1\)'s
-growth rate or a temperature of 4000 degrees result in an equally meaningless
-simulation as transitions are unrealistic.  This figure demonstrates that small
-changes to \(\Delta\) can have a strong effect on the timing and frequency with
-which new EOM minima are discovered and referenced. This finding suggests that
-we need a flexible policy language and engine to explore these trade-offs.
+resulting in hotter keys and a smaller keyspace.  Values smaller than
+\(\Delta_2\)'s growth rate or a temperature of 400 degrees results in very
+little database activity because state transitions take too long. Similarly,
+values larger than \(\Delta_1\)'s growth rate or a temperature of 4000 degrees
+result in an equally meaningless simulation as transitions are unrealistic.
+This figure demonstrates that small changes to \(\Delta\) can have a strong
+effect on the timing and frequency with which new EOM minima are discovered and
+referenced. This finding suggests that we need a flexible policy language and
+engine to explore these trade-offs.
 
@@
 (up to \(5\times\)) because worker tasks start generating states with different
-coordinates later in the run.  The growth rate, temperature, and number of
-workers have a predictable effect on the structure of the keyspace.
-Figure~\ref{fig:methodology-keys} shows that the number of reads changes with
-different initial conditions (\(\Delta\)), but that the spatial locality of key
-accesses is similar ({\it e.g.}, some keys are still \(5\times\) more popular
-than others).  Figure~\ref{fig:motivation-regimes} shows how entropy for
-different growth rates has temporal locality, as the reads per second for
-\(\Delta_2\) look like the reads per second for \(\Delta_1\) stretched out
-along the time axis.  Trends also exist for temperature and number of workers
-but are omitted here for space. This structure means that we can learn the key
-access patterns and adapt the storage system accordingly (this motivates
+coordinates later in the run.  Figure~\ref{fig:methodology-keys} shows that the
+number of reads changes with different initial conditions (\(\Delta\)), but
+that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
+still \(5\times\) more popular than others).
+Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
+rates has temporal locality, as the reads per second for \(\Delta_2\) look like
+the reads per second for \(\Delta_1\) stretched out along the time axis.
+Trends also exist for temperature and number of workers but are omitted here
+for space. This structure means that we can learn the key access patterns and
+adapt the storage system accordingly (this motivates
 Section~\S\ref{sec:dom-specific}).

### 2017-11-27 paper/sections/3_methodology.tex
paper: addresses #35, #44, #46
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 prototype in that paper was called Mantle and it was originally built on CephFS
-so administrators could control file system metadata data management policies.
+so administrators could control file system metadata load balancing policies.
 We now refer to Mantle as a policy engine that supports our data management
 language.  The basic premise is that data management policies can be expressed
-with a simple API consisting of ``when", ``where", and ``how much". ``when"
+with a simple API consisting of ``when", ``where", and ``how much". The ``when" policy
 controls how aggressive or conservative the decisions are; ``where" controls
@@
 We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
-linked into a service.  Administrators write policies from whatever domain they
-choose ({e.g.}, statistics, machine learning, storage system) and the policies
-are embedded into the runtime by Mantle. When the service makes decisions it
-executes the administrator-defined policies for when/where/how much and returns
-a decision.  To do this, the service needs to be modified to (1) provide an
-environment of metrics and (2) identify where policies are set. These
-modification points are shown by the colored boxes in Figure~\ref{fig:mantle}
-and described below.
-
-Administrators write policies in
-Lua that are executed whenever the service needs to make a data management
-decision. We chose Lua for simplicity, performance, and portability; it is a
-scripting language with simple syntax, which allows administrators to focus on
-the policies themselves; it was designed as an embeddable language, so it is
-lightweight and does less type checking; and it interfaces nicely with C/C++.
+linked into a service.  Administrators write policies in Lua from whatever
+domain they choose ({\it e.g.}, statistics, machine learning, storage system)
+and the policies are embedded into the runtime by Mantle.  We chose Lua for
+simplicity, performance, and portability; it is a scripting language with
+simple syntax, which allows administrators to focus on the policies themselves;
+it was designed as an embeddable language, so it is lightweight and does less
+type checking; and it interfaces nicely with C/C++.  When the service makes
+decisions it executes the administrator-defined policies for when/where/how
+much and returns a decision.  To do this, the service needs to be modified to
+(1) provide an environment of metrics and (2) identify where policies are set.
+These modification points are shown by the colored boxes in
+Figure~\ref{fig:mantle} and described below.
 
@@
 
-\subsubsection{Environment of Metrics} services expose cluster metrics for
-describing resource utilization and time series metrics for describing accesses
-to some data structure over time. Table~\ref{table:metrics} shows how these
-metrics are accessed from the policies written by administrators.
+\subsubsection{Environment of Metrics} services expose \textbf{cluster} metrics
+for describing resource utilization and \textbf{time series} metrics for
+describing accesses to some data structure over time. Table~\ref{table:metrics}
+shows how these metrics are accessed from the policies written by
+administrators.
 
-For cluster metrics, the service passes a dictionary to Mantle, indexed by
-server name and metric. Policies access the cluster metric values by indexing
-into a Lua table using \texttt{server} and \texttt{metric}, where server is a
-node identifier ({e.g.}, MPI Rank, metadata server name) and metric is a
-resource name.  For example, a policy written for an MPI-based service can
-access the CPU utilization of the first rank in a communication group using:
+For cluster metrics, the service passes a dictionary to Mantle. Policies access
+the cluster metric values by indexing into a Lua table using \texttt{server}
+and \texttt{metric}, where server is a node identifier ({\it e.g.}, MPI Rank,
+metadata server name) and metric is a resource name.  Metrics used for file
+system metadata load balancing are shown by the ``environment" box in
+Figure~\ref{fig:mantle}. The measurements and exchange of metrics between
+servers is done by the service; Mantle in CephFS leverages metrics from other
+servers collected using CephFS's heartbeats.  For example, a policy written for
+an MPI-based service can access the CPU utilization of the first rank in a
+communication group using:
 
@@
 
-Other examples metrics used for file system metadata load balancing are shown
-by the ``environment" box in Figure~\ref{fig:mantle}. The measurements and
-exchange of metrics between servers is done by the service; Mantle in CephFS
-leverages metrics from other servers collected using CephFS's heartbeats.
-
 For time series metrics, the service passes an array of \texttt{(timestamp,
-value)} pairs to Mantle and the policies can iterate over the accesses. For
+value)} pairs to Mantle and the policies can iterate over the accesses.  The
+service uses a pointer to the time series to facilitate time series with many
+values, like accesses to a database or directory in the file system namespace.
+This decision limits the time series metrics to only include values from the
+{\it current } node, although this is not a limitation of Mantle itself.  For
 example, a policy that uses accesses to a directory in a file system as a
@@
 \begin{minted}{lua}
-ts = array()
-for i=1,arraysize() do
-  for time,value in string.gmatch(ts, (%w+=%w)) do
+d = timeseries()       -- get data (d) from service
+for i=1,d:size() do    -- iterate over timeseries
+  ts, value = d:get(i) -- index into timeseries 
   if value == 'mydirectory' then
@@
 end
-end
 \end{minted}
 \end{figure}
-
-The service uses a pointer to the time series to facilitate time series with
-many values, like accesses to a database or directory in the file system
-namespace. This decision limits the time series metrics to only include values
-from the {\it current } node, although this is not a limitation of Mantle
-itself.
 
@@
 pressure of the worker node to the cache management policies. In
-Section~\S\ref{sec:parsplice-keyspace-analysis} we only end up using the cache
-size although the other metrics proved to be valuable debugging tools. Using
-Mantle time series metrics, we expose accesses to the cache as a list of
-\texttt{timestamp, key} pairs. In Section~\S\ref{sec:dom-specific}, we
-explore a key access pattern detection algorithm that uses this metric.
+Section~\S\ref{sec:arch-specific} we only end up using the cache size although
+the other metrics proved to be valuable debugging tools. Using Mantle time
+series metrics, we expose accesses to the cache as a list of \texttt{timestamp,
+key} pairs. In Section~\S\ref{sec:dom-specific}, we explore a key access
+pattern detection algorithm that uses this metric.
 

### 2017-11-27 paper/sections/4_dynamic-policies.tex
paper: addresses #35, #44, #46
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 algorithms on the worker using the keyspace analysis in
-Section~\S\ref{sec:parsplice-keyspace-analysis}.  The evaluation metric is the
-accuracy and runtime of each strategy; the policy should be accurate enough so
-as to sacrifice negligible performance and fast enough to run as often as we
-want to detect key access patterns.  First we sized the cache according to our
-system specific knowledge. By ``system", we mean the hardware and software of
-the storage hierarchy. We look at request rate, unique keys in a sliding
-window, and bandwidth capabilities. For example, we know that LevelDB cannot
-handle high IO request rates.
+Section~\S\ref{sec:parsplice-keyspace-analysis}.  Our evaluation uses the total
+``trajectory length" as the goodness metric. This value is the duration of the
+overall trajectory produced by ParSplice. At ideal efficiency, the trajectory
+length should increase with the square root of the wall-clock time, since the
+wall-clock cost of time-stepping the system by one simulation time unit
+increases in proportion of the total number of atoms.  The policy should avoid
+reducing the trajectory length and be fast enough to run as often as we want to
+detect key access patterns.  First we sized the cache according to our system
+specific knowledge, {\it i.e.} the hardware and software of the storage
+hierarchy such as request rate, unique keys in a sliding window, and bandwidth
+capabilities.
 
@@
 utilization is measured with memory footprint of just the cache (\(y2\) axis).
-The middle graph shares the \(y\) axis and shows the trade-off of using a basic
-LRU-style cache for different cache sizes, where the penalty for a cache miss
-is retrieving the data from the persistent database.  We evict keys (if
-necessary) at every operation instead of when segments complete because the
-cache fills up too quickly otherwise. This is implemented using a ``when"
-policy of \texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
-\texttt{servers[whoami]['cachesize']-n}.  The error bars are the standard
-deviation of 3 runs.  Although the keyspace grows to 150K, a 100K key cache
-achieves 99\% of the performance. Decreasing the cache degrades performance and
-predictability.
+The middle graph labeled "Fixed Cache Size" shares the \(y\) axis and shows the
+trade-off of using a basic LRU-style cache for different cache sizes, where the
+penalty for a cache miss is retrieving the data from the persistent database.
+We evict keys (if necessary) at every operation instead of when segments
+complete because the cache fills up too quickly otherwise. This is implemented
+using a ``when" policy of:
+
+\begin{figure}[h]
+\footnotesize
+\begin{minted}{lua}
+server[whoami]['cachesize']>n
+\end{minted}
+\end{figure}
+
+and a ``how much" policy of:
+
+\begin{figure}[h]
+\footnotesize
+\begin{minted}{lua}
+servers[whoami]['cachesize']-n
+\end{minted}
+\end{figure}
+
+The error bars are the standard deviation of 3 runs.  Although the keyspace
+grows to 150K, a 100K key cache achieves 99\% of the performance. Decreasing
+the cache degrades performance and predictability.
 
@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
-\caption{The performance and resource utilization trade-off for different cache
-management policies. ``None" is ParSplice unmodified, ``Fixed Sized Cache"
-limits the size of the cache by evicting the least recently used items, and
-``Multi-Policy Cache" switches to a fixed sized cache after absorbing the
-initial burstiness of the workload.  The size of the fixed sized cache for each
-experiment is on the \(x\) axis.  \label{fig:methodology-tradeoff}}
+\caption{Policy performance/utilization: ``None" is ParSplice unmodified,
+``Fixed Sized Cache" evicts keys using LRU , and ``Multi-Policy Cache" switches
+to fixed sized cache after absorbing the workload's initial burstiness.  Fixed
+sized caches are on the \(x\) axis.\label{fig:methodology-tradeoff}}
 \end{figure}
+
+\begin{figure}[t]
+        \centering
+        \includegraphics[width=0.5\textwidth]{figures/memory-vs-time.png}\\
+	\caption{Memory utilization for ``No Cache Management" (unlimited cache
+growth), ``Multi-Policy" (absorbs initial burstiness of workload) , and
+``Dynamic Policy" (sizes cache according to key access patterns).
+\label{fig:memory-vs-time}}
+\end{figure}%
+
 %
@@
 some keys may be accessed at time \(t_0\), not in \(t_1\), and then again in
-\(t_2\). Because the cache is too small, reads and writes hit the persistent
-database, where the excessive traffic triggers a LevelDB compaction and reads
-block.  To avoid overloading the storage hierarchy with requests, we design a
-multi-part policy that switches between:
+\(t_2\). Because the cache is too small, reads and writes hit rest of the
+storage hierarchy and the excessive traffic triggers a LevelDB compaction on
+the persistent database.  To avoid overloading the storage hierarchy with
+requests, we design a multi-part policy that switches between:
 
@@
 trade-off of the multi-part policy, where the cache sizes for the \(n\) key
-limit policy are along the \(x\) axis.  They show better performance than the
-single \(n\) key policies. The performance and memory utilization for a 100K
-key cache size is the same as the 100K bar in the ``Multi-Part Policy" graph in
-Figure~\ref{fig:methodology-tradeoff} but the rest reduce the size of the
-keyspace after the read flash crowd.  We see the worst performance when the
-engine switches to the 10 key limit policy, which achieves 94\% of the
+limit policy are along the \(x\) axis.  The performance and memory utilization
+for a 100K key cache size is the same as the 100K bar in the ``Fixed Cache
+Size" graph in Figure~\ref{fig:methodology-tradeoff} but the rest reduce the
+size of the keyspace after the read flash crowd.  We see the worst performance
+when the engine switches to the 10 key limit policy, which achieves 94\% of the
 performance while only using 40KB of memory.
@@
 memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
-length vs.  wall-clock time curves down over time; as the nanoparticle grows
-it takes longer to generate segments so by the time we reach 2 hours, over 90\%
-of the trajectory is already generated.  For (2), the memory footprint rises
-until we reach 100K key switch threshold at 0.4GB, as shown by the
-``\(\Delta_1\), Multi-Policy" curve in Figure~\ref{fig:memory-vs-time} but in
+length vs.  wall-clock time curves down over time; as the nanoparticle grows it
+takes longer to generate segments so by the time we reach 2 hours, over 90\% of
+the trajectory is already generated.  For (2), the memory footprint rises until
+we reach 100K key switch threshold at 0.4GB, as shown by the ``\(\Delta_1\),
+Multi-Policy" curve in Figure~\ref{fig:memory-vs-time} but in
 Figure~\ref{fig:methodology-tradeoff} we plot the final value. As a result, we
-can only conclude that this policy only works well for this 2.5 hour run.
+conclude that this policy only works well for this 2.5 hour run.
 
@@
 job lengths).  To solve that problem, we need a way to identify what thresholds
-we should use for different job permutations.
+we should use for different configurations.
 
@@
 \begin{figure*}[t!]
-    \begin{subfigure}[t]{0.32\textwidth}
+    \begin{subfigure}[t]{0.5\textwidth}
         \centering
-	\includegraphics[width=0.89\textwidth]{figures/keyspace-zoomed.png}\\
+	\includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\
 \caption{Key activity for a 4 hour run shows groups of accesses to the same
@@
     ~ 
-    \begin{subfigure}[t]{0.32\textwidth}
+    \begin{subfigure}[t]{0.5\textwidth}
         \centering
-        \includegraphics[width=1\textwidth]{figures/memory-vs-time.png}\\
-	\caption{Memory utilization for ``No Cache Management" (unlimited cache
-growth), ``Multi-Policy" (absorbs initial burstiness of workload) , and
-``Dynamic Policy" (sizes cache according to key access patterns).
-\label{fig:memory-vs-time}}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[width=0.86\textwidth]{figures/dscache-vs-none.png}\\
+        \includegraphics[width=0.9\textwidth]{figures/dscache-vs-none.png}\\
 	\caption{Performance/Utilization for dynamically sized cache (DSCache)
@@
 simulation only adds molecules; we can never reach a state with less molecules.
-This is why keys are never re-accessed.  Despite these patterns, the following
-characteristics of superbasins make it hard to detect them:
+This is why keys are never re-accessed.
 
@@
 accessed over time grows while moving along the \(y\) axis shows that early
-keys are accessed more often.
+keys are accessed more often.  Despite these patterns, the following
+characteristics of superbasins make it hard to detect them:
 
@@
 \end{itemize}
-
-Below we describe the policies we implemented and deployed using Mantle.
 
@@
 ``Zero-crossing", set to 40 seconds, for distance between small/large spikes to
-avoid false positives; `window size", set to 200 seconds, for the size of the
-moving average
+avoid false positives; ``window size", set to 200 seconds, for the size of the
+moving average.
 
@@
 groups of key accesses are different size. ``K", set to 4, for the number of clusters in the data
-using the sum of the distances to the centroid
+using the sum of the distances to the centroid.
 
@@
 max distance between 2 samples in same neighborhood; ``Min", set to 5, for the
-number of samples per core
+samples per core.
 
   \item Anomaly Detection: size of the image is too big and bottom edges are
-not thick enough
+not thick enough.
 
@@
 Figure~\ref{src:dyn-cache} works best. The algorithm detects groups of key
-access patterns, which we call ``fans", by iterating backwards through the key
-access trace, finding the lowest key ID, and comparing against the lowest key
-ID we have seen so far (Line 7). We also maintain the top and bottom of each
-group of key accesses (Line 13) so we can tell the ``how much" policy the
-number of keys to evict (Line 19).  The algorithm iterates backwards over the
-key access trace because a change in the minimum value signals a new group of
-key accesses. No signal exists iterating left to right, as the maximum value
-always increases and the minimum values at the bottom of each group of key
-accesses are sparse.  For example, the maximum distance between values along
-the bottom edge of the zoomed group of key accesses in
-Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the maximum distance
-between minimum values for the group of key accesses before is 0 seconds. As a
-result of this sparseness, iterating left to right requires a ``window size"
-parameter to determine when we think a minimum value will not show up again.
-We exchange runtime, in the form of an O(\(n\)) algorithm where \(n\) is the
-number events, for simplicity because this approach avoids adding new
+accesses, which we call ``fans", by iterating backwards through the key access
+trace, finding the lowest key ID, and comparing against the lowest key ID we
+have seen so far (Line 7). We also maintain the top and bottom of each group of
+key accesses (Line 13) so we can tell the ``how much" policy the number of keys
+to evict (Line 19).  The algorithm iterates backwards over the key access trace
+because a change in the minimum value signals a new group of key accesses. No
+signal exists iterating left to right, as the maximum value always increases
+and the minimum values at the bottom of each group of key accesses are sparse.
+For example, the maximum distance between values along the bottom edge of the
+zoomed group of key accesses in Figure~\ref{fig:keyspace-zoomed} is 125
+seconds, while the maximum distance between minimum values for the group of key
+accesses before is 0 seconds. As a result of this sparseness, iterating left to
+right requires a ``window size" parameter to determine when we think a minimum
+value will not show up again.  The algorithm is O(\(n\)), where \(n\) is the
+number events, but the benefit is that the approach avoids adding new
 thresholds for key access pattern detection ({\it e.g.}, space between key
@@
 \begin{minted}[xleftmargin=3em,linenos]{lua}
-d = timeseries()          -- provided by mantle
-ts, id = d:get(d:size())  -- provided by mantle
-fan  = {start=nil, finish=ts, top=0, bottom=id}
+d = timeseries()
+ts, id = d:get(d:size())
+fan  = {start=nil, finish=ts, top=0, bot=id}
 fans = {}
-for i=d:size(),1,-1 do    -- detect all fans
+for i=d:size(),1,-1 do     -- iterate backwards
   ts, id = d:get(i)
-  if id < fan['bottom'] then
+  if id < fan['bot'] then  -- found a new fan!
     fan['start']  = ts
     fans[#fans+1] = fan 
-    fan = {start=nil, finish=ts, top=0, bottom=id}
+    fan = {start=nil, finish=ts, top=0, bot=id}
   end
 
-  if id > fan['top'] then fan['top'] = id end 
+  if id > fan['top'] then  -- track top of fan
+    fan['top'] = id 
+  end 
 end
@@
 
-if #fans < 2 then return false else
-  WRstate(fans[#fans-1]['top']-fans[1]['bottom'])
+if #fans < 2 then -- do not evict current fan
+  return false 
+else
+  WRstate(fans[#fans-1]['top']-fans[1]['bot'])
   return true
@@
 and detects when key accesses move on to a new subset of
-keys. We refer to groups of key accesses as ``fans".\label{src:dyn-cache}}
+keys ({\it i.e.} ``fans"). \label{src:dyn-cache}}
 \end{figure}

### 2017-11-27 paper/sections/5_file-systems.tex
paper: addresses #35, #44, #46
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 axis is the file ID, where the file IDs are listed in the order that they
-appear in the trace.  The trace is anonymized so all file names are replaced
-with a unique identifier and we do not know what clients are doing.  Although
-other access pattern detection algorithms are possible, we use the one designed
-for cache management in Section~\S\ref{sec:regime-detection} without
-modification and it successfully detects the largest group of key accesses that
-starts at time 1000 seconds and ends at time 2200 seconds. File systems that
-load balance file system metadata across a cluster would want to keep metadata
-in that group of key accesses on the same server to reduce RPCs.
+appear in the trace.
 
-% burstiness of creates then compile
-Cache management can also increase the capacity of a single server, which
-affects how load should be distributed across a cluster.  For example, CephFS
-clients and servers maintain a consistent cache so that the client can do
-operations locally without contacting the metadata server.  But large caches
-use more memory, so having smart policies that detect key access patterns would
-greatly reduce memory footprints. For example, Figure~\ref{fig:compile-ops}
-shows the trace of metadata requests for compiling code in CephFS.  The \(y\)
-axis is  the number of file system metadata requests serviced by the metadata
-server when uncompressing (\texttt{untar}), compiling (\texttt{make}), and
-deleting (\texttt{rm}) the source code for the Linux kernel. If the system
-knows that the job phases would progress from many creates, to many lookups, to
-many deletes, then it could size its caches accordingly. For example, the file
-system could cache none of the metadata from the \texttt{untar} phase and run
-key access pattern detection during the \texttt{make} phase, resulting in the
-metadata server/clients only caching metadata that is repeatedly used. For the
-job in Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K
-inodes (the unit of metadata for file systems) instead of 60K, resulting in
-almost 40MB of memory savings (since an inode is about
-1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
-without sacrificing performance.
+The trace is anonymized so all file names are replaced with a unique identifier
+and we do not know what clients are doing.  Although other access pattern
+detection algorithms are possible, we use the one designed for cache management
+in Section~\S\ref{sec:regime-detection} with slight modifications based on our
+knowledge of file systems~\footnote{We filtered requests key ids less than
+2000, as these are most likely path traversal requests to higher parts of the
+namespace, {\it i.e.} root directories.} and it successfully detects the
+largest group of key accesses that starts at time 1000 seconds and ends at time
+2200 seconds. File systems that load balance file system metadata across a
+cluster would want to keep metadata in that group of key accesses on the same
+server to reduce RPCs.
 
-%- strategy: cache all inodes on create
-\begin{figure}[t]
-\noindent\includegraphics[width=0.5\textwidth]{figures/compile-ops.png}\\
-\caption{File system metadata requests for a compile job show workload phases
-characterized by a dominant request type ({\it e.g.}, creates for ``untar").
-Using ParSplice cache management policies for these types of workloads would
-reduce memory pressure without sacrificing performance.
-\label{fig:compile-ops}}
-\end{figure}
+%% burstiness of creates then compile
+%Cache management can also increase the capacity of a single server, which
+%affects how load should be distributed across a cluster.  For example, CephFS
+%clients and servers maintain a consistent cache so that the client can do
+%operations locally without contacting the metadata server.  But large caches
+%use more memory, so having smart policies that detect key access patterns would
+%greatly reduce memory footprints. For example, Figure~\ref{fig:compile-ops}
+%shows the trace of metadata requests for compiling code in CephFS.  The \(y\)
+%axis is  the number of file system metadata requests serviced by the metadata
+%server when uncompressing (\texttt{untar}), compiling (\texttt{make}), and
+%deleting (\texttt{rm}) the source code for the Linux kernel. If the system
+%knows that the job phases would progress from many creates, to many lookups, to
+%many deletes, then it could size its caches accordingly. For example, the file
+%system could cache none of the metadata from the \texttt{untar} phase and run
+%key access pattern detection during the \texttt{make} phase, resulting in the
+%metadata server/clients only caching metadata that is repeatedly used. For the
+%job in Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K
+%inodes (the unit of metadata for file systems) instead of 60K, resulting in
+%almost 40MB of memory savings (since an inode is about
+%1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
+%without sacrificing performance.
+%
+%%- strategy: cache all inodes on create
+%\begin{figure}[t]
+%\noindent\includegraphics[width=0.5\textwidth]{figures/compile-ops.png}\\
+%\caption{File system metadata requests for a compile job show workload phases
+%characterized by a dominant request type ({\it e.g.}, creates for ``untar").
+%Using ParSplice cache management policies for these types of workloads would
+%reduce memory pressure without sacrificing performance.
+%\label{fig:compile-ops}}
+%\end{figure}
 
@@
   \item QoS: when to move clients, where to move clients, how much of the
-reservation to move
+reservation to move. We could use Mantle to implement something like the
+reservation algorithms based on utilization and period in
+Fahrrad~\cite{povzner_horizon_2010} to achieve better guarantees without
+sacrificing performance.
 
-  \item Scheduling: how much of a resource to allocate; more than just LRU
-(RAD)
+  \item Scheduling: when to yield computation cycles to another process, how
+much of a resource to allocate. We could use Mantle to implement something like the RBED
+scheduler~\cite{brandt:rtss2003-rad}, that separates resource allocation from
+timing of delivery to enable real-time and non-real-time-scheduling classes.
 
   \item Batching: how many operations to group together and when to send large
-batches of updates
+batches of updates. We could use Mantle to shape the batching policies in the
+shared resource interface in  Malacology~\cite{sevilla:eurosys17-malacology}.
 
-  \item Prefetching: how much to prefetch and how to select data to reterive
-(forward/backward/stride detection algorithms for prefetching,
+  \item Prefetching: how much to prefetch and how to select data to retrieve.
+We could use Mantle to implement forward/backward/stride detection algorithms
+for prefetching in RAID arrays or something more complicated, like the time
+series algorithms for adaptive I/O prefetching presented
+in~\cite{tran:sc01-arima}.
 

### 2017-11-27 paper/sections/1_introduction.tex
paper: addresses #47
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 evict are specific to the application's behavior and the system's configuration
-(hardware, settings, etc.) so a new use case may prove to be a poor match for
+(hardware, settings, etc.) so a new workload may prove to be a poor match for
 the selected caching
@@
 process of trying general policies and quickly iterating to a customized
-solution shows that our prototype can adapt to different workloads.
+solution shows that our prototype can be adapted to different workloads.
 
@@
 
-Our target application, ParSplice~\cite{perez:jctc20150parsplice}, is
-representative of applications that extensively use software-based caches.
-ParSplice is a molecular dynamics simulation that uses a hierarchy of caches
-and a single persistent key-value store to store both observed minima across a
-molecule's equation of motion (EOM) and the hundreds or thousands of partial
-trajectories calculated each second during a parallel job.  The fine-grained
-data annotation capabilities provided by key-value storage is a natural match
-for scientific simulations like ParSplice. But these simulations rely on a
-mesh-based decomposition of a physical region and result in millions or
-billions of mesh cells, where each cell contains materials, pressures,
-temperatures and other characteristics that are required to accurately simulate
-phenomena of interest.  Unfortunately, simulations of this size saturate the
-capacity and bandwidth capabilities of a single node so we need more effective
-data management techniques.
+Our example application, ParSplice~\cite{perez:jctc20150parsplice}, is
+representative of an important class applications with similar working set
+behaviors that extensively use software-based caches.  ParSplice is a molecular
+dynamics simulation that uses a hierarchy of caches and a single persistent
+key-value store to store both observed minima across a molecule's equation of
+motion (EOM) and the hundreds or thousands of partial trajectories calculated
+each second during a parallel job.  This workload is pervasive across
+simulations that rely on a mesh-based decomposition of a physical region and
+result in millions or billions of mesh cells, where each cell contains
+materials, pressures, temperatures and other characteristics that are required
+to accurately simulate phenomena of interest.  The fine-grained data annotation
+capabilities provided by key-value storage is a natural match for these types
+of scientific simulations.  Unfortunately, simulations of this size saturate
+the capacity and bandwidth capabilities of a single node so we need more
+effective data management techniques.
 
@@
 shows how a poorly configured cache can save memory but at the cost of
-performance, which is shown by the text annotation to the right.  Even worse, this
-threshold changes with different initial configurations and cluster setups so
-tuning needs to be done for all system permutations.  Our dynamically sized
+performance, which is shown by the text annotation to the right.  Even worse,
+this threshold changes with different initial configurations and cluster setups
+so tuning needs to be done for all system permutations.  Our dynamically sized
 cache, shown by the thick line in Figure~\ref{fig:cache-management}, detects
@@
 parameter sweeps, our solution saves more memory than a hand-tuned cache with a
-negligible performance degradation (within the baseline's standard deviation)
-and works for a variety of initial conditions.
+negligible performance degradation (within the baseline's standard deviation),
+works for a variety of initial conditions, and could generalize to similar
+applications.
 
@@
 store.  Rather than co-designing a policy directly into the service, which
-requires domain-specific knowledge to find and change policies, we expose the
-policies in a general way so even developers unfamiliar with the domain can
-quickly deploy solutions.  We show that our framework:
+requires (1) domain-specific knowledge about the application to design an
+effective policy and (2) service-specific knowledge to find and change
+policies, we separate policies from mechanisms in a general way. By exposing
+these policies (and not the mechanism) developers unfamiliar with the service
+that their application uses can quickly deploy solutions based on their
+domain-specific knowledge.  We show that our framework:
 

### 2017-11-27 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #26
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 keyspace analysis that follows is for the cache on the worker node, which is
-circled in Figure~\ref{fig:parsplice}.  The cache hierarchy is unmodified but
-for the persistent database node, we replace BerkeleyDB on NFS with LevelDB on
-Lustre. Original ParSplice experiments showed that BerkeleyDB's syncs caused
-reads/writes to bottleneck on the persistent database. We also use Riak's
-customized LevelDB\footnote{https://github.com/basho/leveldb} version, which
-comes instrumented with its own set of performance counters.
+circled in Figure~\ref{fig:parsplice}.
 

### 2017-11-27 paper/sections/3_methodology.tex
paper: addresses #26
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 alongside code that checks for memory pressure. It is executed right before the
-worker processes incoming and outgoing put/get transactions to the cache. As
-stated previously, we do not use the ``where" part of Mantle because we focus
-on a single node, but this part of the API will be used when we move the caches
-and storage nodes to a key-values store back-end that uses key load balancing and
-repartitioning.
+worker processes incoming and outgoing put/get transactions to the cache. We
+only do cache management once every second to avoid maintaining the cache for
+every request. We expected to have to increase this polling interval to
+accommodate more complex policies but even our most complicated policy in
+Section~\S\ref{sec:dom-specific} had a negligible effect on performance when
+executed every second (within the standard deviation for multiple runs when
+compared against a policy that returns immediately). This may be because the
+worker is not overloaded and the bottleneck is somewhere else in the system.
+As stated previously, we do not use the ``where" part of Mantle because we
+focus on a single node, but this part of the API will be used when we move the
+caches and storage nodes to a key-values store back-end that uses key load
+balancing and repartitioning.

### 2017-11-27 paper/sections/1_introduction.tex
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 abstractions effectively decomposes the problem into manageable policies for a
-different domain and service.  Our exploration of this space results in a
+different application and storage system.  Our exploration of this space results in a
 dynamically sized cache policy that sacrifices negligible performance while
@@
 solution shows that our prototype can be adapted to different workloads.
-
-%very customized solution confirms that our prototype can adapt to different
-%types of workloads; but more importantly, this is the second system and domain
-%that we have successfully used this policy language and engine in. This breadth
-%helps us reason about more general data management policies by giving us a
-%framework to compare and contrast policies from other domains and services.
 
@@
 Mantle as a policy engine that injects policies written in our data management
-language directly into a running service, such as a file system or key-value
-store.  Rather than co-designing a policy directly into the service, which
-requires (1) domain-specific knowledge about the application to design an
-effective policy and (2) service-specific knowledge to find and change
-policies, we separate policies from mechanisms in a general way. By exposing
-these policies (and not the mechanism) developers unfamiliar with the service
-that their application uses can quickly deploy solutions based on their
-domain-specific knowledge.  We show that our framework:
+language directly into a running storage system, such as a file system or key-value
+store.  Rather than implementing a policy directly into the storage system,
+which requires domain-specific knowledge to find and change policies in the
+storage system, we separate policies from mechanisms in a general way.  By
+exposing just the policies developers unfamiliar with the storage system that
+their application uses can quickly deploy efficient solutions based on their
+application-specific knowledge.  We show that our framework:
 

### 2017-11-27 paper/sections/3_methodology.tex
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-%  \item motivation: Mochi load balancer microservice
-%  \item library architecture, callbacks, environment
-
 We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
-linked into a service.  Administrators write policies in Lua from whatever
+linked into a storage system.  Administrators write policies in Lua from whatever
 domain they choose ({\it e.g.}, statistics, machine learning, storage system)
@@
 it was designed as an embeddable language, so it is lightweight and does less
-type checking; and it interfaces nicely with C/C++.  When the service makes
+type checking; and it interfaces nicely with C/C++.  When the storage system makes
 decisions it executes the administrator-defined policies for when/where/how
-much and returns a decision.  To do this, the service needs to be modified to
+much and returns a decision.  To do this, the storage system needs to be modified to
 (1) provide an environment of metrics and (2) identify where policies are set.
@@
   && \\
-              & Service      & Example \\\hline
+              & Storage System & Example \\\hline
   Cluster     & File Systems & CPU util., Inode reads \\
@@
   \end{tabular}
-  \caption{Types of metrics exposed by the service to the policy engine using Mantle.\label{table:metrics}}
+  \caption{Types of metrics exposed by the storage system to the policy engine using Mantle.\label{table:metrics}}
 \end{table}
 
-\subsubsection{Environment of Metrics} services expose \textbf{cluster} metrics
+\subsubsection{Environment of Metrics} storage systems expose \textbf{cluster} metrics
 for describing resource utilization and \textbf{time series} metrics for
@@
 
-For cluster metrics, the service passes a dictionary to Mantle. Policies access
+For cluster metrics, the storage system passes a dictionary to Mantle. Policies access
 the cluster metric values by indexing into a Lua table using \texttt{server}
@@
 Figure~\ref{fig:mantle}. The measurements and exchange of metrics between
-servers is done by the service; Mantle in CephFS leverages metrics from other
+servers is done by the storage system; Mantle in CephFS leverages metrics from other
 servers collected using CephFS's heartbeats.  For example, a policy written for
-an MPI-based service can access the CPU utilization of the first rank in a
+an MPI-based storage system can access the CPU utilization of the first rank in a
 communication group using:
@@
 
-For time series metrics, the service passes an array of \texttt{(timestamp,
+For time series metrics, the storage system passes an array of \texttt{(timestamp,
 value)} pairs to Mantle and the policies can iterate over the accesses.  The
-service uses a pointer to the time series to facilitate time series with many
+storage system uses a pointer to the time series to facilitate time series with many
 values, like accesses to a database or directory in the file system namespace.
@@
 \begin{minted}{lua}
-d = timeseries()       -- get data (d) from service
+d = timeseries()       -- data (d) from storage system
 for i=1,d:size() do    -- iterate over timeseries

### 2017-11-27 paper/sections/4_dynamic-policies.tex
paper: addresses #47
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\section{Cache Management Using System Architecture Knowledge}
+\section{Cache Management Using Storage System Architecture Knowledge}
 \label{sec:arch-specific}
@@
 utilization is measured with memory footprint of just the cache (\(y2\) axis).
-The middle graph labeled "Fixed Cache Size" shares the \(y\) axis and shows the
+The middle graph labeled ``Fixed Cache Size" shares the \(y\) axis and shows the
 trade-off of using a basic LRU-style cache for different cache sizes, where the
@@
 the trajectory is already generated.  For (2), the memory footprint rises until
-we reach 100K key switch threshold at 0.4GB, as shown by the ``\(\Delta_1\),
-Multi-Policy" curve in Figure~\ref{fig:memory-vs-time} but in
-Figure~\ref{fig:methodology-tradeoff} we plot the final value. As a result, we
-conclude that this policy only works well for this 2.5 hour run.
+it reaches the 100K key switch threshold at 0.4GB and then reduces to the final
+value after switching policies. The memory usage over time for this policy is
+shown by the ``\(\Delta_1\), Multi-Policy" curve in
+Figure~\ref{fig:memory-vs-time} but in Figure~\ref{fig:methodology-tradeoff} we
+plot the final value.  Despite these caveats, the result is still valid: we
+found a multi-policy cache management strategy that absorbs the cost of a high
+read throughput on a small keyspace and reduces the memory pressure for a 2.5
+hour run. Our experiments show the effectiveness of the policy engine we
+integrated into ParSplice, not that we were able to identify the best policy
+for all system setups ({\it i.e.} different ParSplice parameters, number of
+worker tasks, and job lengths).  To solve that problem, we need a way to
+identify what thresholds we should use for different configurations.
 
-% but the result is still valid
-Despite these caveats, the result is still valid: we found a multi-policy cache 
-management strategy that absorbs the cost of a high read throughput on a small
-keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
-show the effectiveness of the policy engine we integrated into
-ParSplice, not that we were able to identify the best policy for all system
-setups ({\it i.e.} different ParSplice parameters, number of worker tasks, and
-job lengths).  To solve that problem, we need a way to identify what thresholds
-we should use for different configurations.
-
-\section{Cache Management Using Domain-Specific Application Knowledge}
+\section{Cache Management Using Application-Specific Knowledge}
 \label{sec:dom-specific}
@@
 \begin{figure*}[t!]
-    \begin{subfigure}[t]{0.5\textwidth}
+    \begin{subfigure}[t]{0.48\textwidth}
         \centering
@@
     ~ 
-    \begin{subfigure}[t]{0.5\textwidth}
+    \begin{subfigure}[t]{0.48\textwidth}
         \centering
-        \includegraphics[width=0.9\textwidth]{figures/dscache-vs-none.png}\\
+        \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
 	\caption{Performance/Utilization for dynamically sized cache (DSCache)
@@
 
-Feeding domain-specific knowledge about the ParSplice application into a policy
-leads to more a accurate cache management strategy.  The goal of the following
+Feeding application-specific knowledge about the ParSplice application into a policy
+leads to a more accurate cache management strategy.  The goal of the following
 sections is not to find an optimal solution, as this can be done with parameter

### 2017-11-27 paper/sections/5_file-systems.tex
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 In the previous section, we used our data management language and the Mantle
-policy engine to design effective cache management strategies for a new service
-and domain. In this section, we compare and contrast the policies examined for
-file system metadata load balancing in~\cite{sevilla:sc15-mantle} with the ones
-we designed and evaluated above for cache management in ParSplice.
+policy engine to design effective cache management strategies for a new
+application and storage system. In this section, we compare and contrast the
+policies examined for file system metadata load balancing
+in~\cite{sevilla:sc15-mantle} with the ones we designed and evaluated above for
+cache management in ParSplice.
 
@@
 balancing policies to effectively spread load across a cluster. For example,
-many distributed file systems spread the hierarchical namespace across a
-dedicated metadata cluster and caching policies could be used to determine what
-metadata to move and when to move it. To demonstrate this idea, consider
-Figure~\ref{fig:trace-atime}, where the dots are the file system metadata reads
-in a 1 hour window of a 3-day Lustre namespace trace and the vertical lines are
-the groups of accesses detected by our policy from
-Section~\S\ref{sec:regime-detection}.  The \(x\) axis is time and the \(y\)
-axis is the file ID, where the file IDs are listed in the order that they
-appear in the trace.
+distributed file systems that load balance file system metadata across a
+dedicated metadata cluster could use the caching policies to determine what
+metadata to move and when to move it.  To demonstrate this idea, we analyze a
+3-day Lustre file system metadata trace, collected at LANL.  The trace is
+anonymized so all file names are replaced with a unique identifier and we do
+not know which applications are running. We visualize a 1 hour window of the
+trace in Figure~\ref{fig:trace-atime}, where the dots are the file system
+metadata reads in a 1 hour window.  The \(x\) axis is time and the \(y\) axis
+is the file ID, listed in the order that file IDs appear in the trace.  The
+groups of accesses look similar to the ParSplice key accesses in
+Figure~\ref{fig:keyspace-zoomed}.
 
-The trace is anonymized so all file names are replaced with a unique identifier
-and we do not know what clients are doing.  Although other access pattern
-detection algorithms are possible, we use the one designed for cache management
-in Section~\S\ref{sec:regime-detection} with slight modifications based on our
-knowledge of file systems~\footnote{We filtered requests key ids less than
-2000, as these are most likely path traversal requests to higher parts of the
-namespace, {\it i.e.} root directories.} and it successfully detects the
-largest group of key accesses that starts at time 1000 seconds and ends at time
-2200 seconds. File systems that load balance file system metadata across a
-cluster would want to keep metadata in that group of key accesses on the same
-server to reduce RPCs.
+Although other access pattern detection algorithms are possible, we use the one
+designed for cache management in Section~\S\ref{sec:regime-detection} with
+slight modifications based on our knowledge of file systems\footnote{We
+filtered out requests for key IDs less than 2000, as these are most likely path
+traversal requests to higher parts of the namespace.}. The vertical lines in
+Figure~\ref{fig:trace-atime} are the groups of accesses identified by the
+algorithm; it successfully detects the largest group of key accesses that
+starts at time 1000 seconds and ends at time 2200 seconds. File systems that
+load balance file system metadata across a cluster would want to keep metadata
+in that group of key accesses on the same server for locality and would want to
+avoid migrating metadata to a different server until the group of key accesses
+completes.
 
@@
 mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
-three caching subsystems mentioned above already maintain key access traces. We
-hypothesize that since this is all software defined caching, something more
-clever than LRU would improve cache utilization without sacrificing too much
-performance.
+three caching subsystems mentioned above already maintain key access traces.
 

### 2017-11-27 paper/sections/6_conclusion.tex
paper: outline for #43
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\section{Future Work}
+\section{Conclusion}
+
+Data management encompasses a wide range of techniques that vary by application
+and storage system.  Yet, the techniques require policies that shape the
+decision making and finding the best policies is a difficult, multi-dimensional
+problem.  We iterate to a custom solution for our target application that uses
+workload access patterns to size its caches. Without tuning or parameter
+sweeps, our solution saves memory without sacrificing performance for a variety
+of initial conditions, including the scale, duration, configuration, and
+hardware of the simulation. More importantly, rather than attempting to
+construct a single, complex policy that works for a variety of scenarios, we
+instead use the Mantle framework to enable software-defined storage systems to
+flexibly change policies as the workload changes.  We also observe that many of
+the primitives and strategies have enough in common with data management in
+file systems that they both can be expressed with similar semantics and a
+general policy engine.
 
 This lays the foundation for future work, where we will focus on formalizing a
-collection of general data management policies that can be used across domains
-and services. The value of such a collection eases the burden of policy
-development and paves the way for solutions that remove the administrator from
-the development cycle, such as (1) adaptable policies that automatically switch
-to new strategies when the current strategy behaves poorly ({\it e.g.}, thrashing,
-making no progress, etc.), and (2) policy generation, where new policies are
-constructed automatically by examining the collection of existing policies.
-Ultimately, we hope that this automation enables control of policies by
-machines instead of administrators.
-
-\section{Conclusion}
-
-Data management encompasses a wide range of techniques that vary by domain and
-service. Yet, the techniques require policies that shape the decision making
-and finding the best policies is a difficult, multi-dimensional problem.  We
-iterate to a custom solution for our target application that uses workload
-access patterns to size its caches. Without tuning or parameter sweeps, our
-solution saves memory without sacrificing performance for a variety of initial
-conditions, including the scale, duration, configuration, and hardware of the
-simulation. More importantly, rather than attempting to construct a single,
-complex policy that works for a variety of scenarios, we instead use the Mantle
-framework to enable software-defined storage systems to flexibly change
-policies as the workload changes.  We also observe that many of the primitives
-and strategies have enough in common with data management in file systems that
-they both can be expressed with similar semantics and a general policy engine.
-Our prototype lays the groundwork for future work in adaptable policies and
-policy generation, which ushers in an era where policies are controlled by
-machines instead of administrators.
+collection of general data management policies that can be used across
+applications and storage systems.  The value of such a collection eases the
+burden of policy development and paves the way for solutions that remove the
+administrator from the development cycle, such as (1) adaptable policies that
+automatically switch to new strategies when the current strategy behaves poorly
+({\it e.g.}, thrashing, making no progress, etc.), and (2) policy generation,
+where new policies are constructed automatically by examining the collection of
+existing policies.  Ultimately, we hope that this automation enables control of
+policies by machines instead of administrators.

### 2017-11-28 paper/sections/5_file-systems.tex
paper: addresses #49
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 suggests that the heuristics and techniques in other load balancers can be used
-for cache management. This generalization may reduce the work that needs to be
-done for cache management as ideas may have already been explored and could
-work ``out-of-the-box".
+for cache management. The result suggests that the concepts and problems that
+architects grapple with are transcendent and the solutions they design can be
+re-used in different code bases.
 
@@
 completes.
+
+Before we showed how policies designed for load balancing heavily influence our
+cache management in a different application and storage system. But in this
+section we show how an {\it unmodified} cache management policy can be used in
+a load balancing strategy.  This generalization may reduce the work that needs to
+be done for load balancing as ideas may have already been explored in other
+domains and could work ``out-of-the-box".
 

### 2017-11-28 paper/sections/5_file-systems.tex
paper: update author list, addresses #51
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   \item Scheduling: when to yield computation cycles to another process, how
-much of a resource to allocate. We could use Mantle to implement something like the RBED
-scheduler~\cite{brandt:rtss2003-rad}, that separates resource allocation from
-timing of delivery to enable real-time and non-real-time-scheduling classes.
+much of a resource to allocate. We could use Mantle to implement the
+fairness/priority models used in the Mesos~\cite{hindman_mesos_2011} ``how
+many" policies.
 
-  \item Batching: how many operations to group together and when to send large
-batches of updates. We could use Mantle to shape the batching policies in the
-shared resource interface in  Malacology~\cite{sevilla:eurosys17-malacology}.
+  \item Batching: how many operations to group together, when to send large
+batches of updates. We could use Mantle to implement to the pathname leases from
+IndexFS~\cite{ren:sc2014-indexfs} or the capabilities from
+CephFS\footnote{http://docs.ceph.com/docs/master/cephfs/capabilities/}.
 
-  \item Prefetching: how much to prefetch and how to select data to retrieve.
-We could use Mantle to implement forward/backward/stride detection algorithms
-for prefetching in RAID arrays or something more complicated, like the time
-series algorithms for adaptive I/O prefetching presented
-in~\cite{tran:sc01-arima}.
+  \item Prefetching: how much to prefetch, how to select data.  We could use
+Mantle to implement forward/backward/stride detection algorithms for
+prefetching in RAID arrays or something more complicated, like the time series
+algorithms for adaptive I/O prefetching from~\cite{tran:sc01-arima}.
 

### 2017-11-28 paper/sections/1_introduction.tex
Changed paragraph before list of contribs

@@
 % What is Mantle
-To design more flexible cache management policies, like our solution in
-Figure~\ref{fig:cache-management}, we use the data management language and
-policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  While Mantle
-was designed for the narrow purpose of file system metadata load balancing,
-this paper provides evidence that the approach is more broad.  Specifically,
-the collection of abstractions in Mantle provide a general control plane that
-improves the performance of metadata access. So in this paper we refer to
-Mantle as a policy engine that injects policies written in our data management
-language directly into a running storage system, such as a file system or key-value
-store.  Rather than implementing a policy directly into the storage system,
-which requires domain-specific knowledge to find and change policies in the
-storage system, we separate policies from mechanisms in a general way.  By
-exposing just the policies developers unfamiliar with the storage system that
-their application uses can quickly deploy efficient solutions based on their
-application-specific knowledge.  We show that our framework:
+In this paper we are presenting the successful use of the Mantle policy language and engine to manage the dynamically-sized cache. Mantle provides a control plane that
+injects policies into a running storage system, such as a file system or key-value store. While Mantle
+was originally designed for file system metadata load balancing~\cite{sevilla:sc15-mantle},
+we found that it works surprisingly well for specifying cache management policies without requiring users to possess extensive knowledge about the internals of storage systems.
+We show that our framework:
 

### 2017-11-28 paper/sections/1_introduction.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 different application and storage system.  Our exploration of this space results in a
-dynamically sized cache policy that sacrifices negligible performance while
+dynamically sized cache policy that does not sacrifice any performance while
 using 32-66\% less memory than the default ParSplice configuration.
@@
 Our example application, ParSplice~\cite{perez:jctc20150parsplice}, is
-representative of an important class applications with similar working set
+representative of an important class of applications with similar working set
 behaviors that extensively use software-based caches.  ParSplice is a molecular
@@
 unacceptable for HPC environments, where a common goal is to keep memory for
-such data structures below 3\%\footnote{Anecdotally, we find this threshold
-works well for HPC applications.  For reference, a 1GB cache for a distributed
-file system is too large in LANL deployments.}. Furthermore, ParSplice deploys
-a cache per 300 worker processes, so large simulations need more caches and
-will use even more memory.  Users can configure ParSplice to evict data when
-the cache reaches a threshold but this solution requires tuning and parameter
-sweeps; the ``Cache (too small)" curve in Figure~\ref{fig:cache-management}
-shows how a poorly configured cache can save memory but at the cost of
-performance, which is shown by the text annotation to the right.  Even worse,
-this threshold changes with different initial configurations and cluster setups
-so tuning needs to be done for all system permutations.  Our dynamically sized
-cache, shown by the thick line in Figure~\ref{fig:cache-management}, detects
-key access patterns and re-sizes the cache accordingly.  Without tuning or
-parameter sweeps, our solution saves more memory than a hand-tuned cache with a
-negligible performance degradation (within the baseline's standard deviation),
-works for a variety of initial conditions, and could generalize to similar
-applications.
+such data structures below 3\%\footnote{Anecdotally, this threshold works well
+for HPC applications.  For reference, a 1GB cache for a distributed file system
+is too large in LANL deployments.}. Furthermore, ParSplice deploys a cache per
+300 worker processes, so large simulations need more caches and will use even
+more memory.  Users can configure ParSplice to evict data when the cache
+reaches a threshold but this solution requires tuning and parameter sweeps; the
+``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
+poorly configured cache can save memory but at the cost of performance, which
+is shown by the text annotation to the right.  Even worse, this threshold
+changes with different initial configurations and cluster setups so tuning
+needs to be done for all system permutations.  Our dynamically sized cache,
+shown by the thick line in Figure~\ref{fig:cache-management}, detects key
+access patterns and re-sizes the cache accordingly.  Without tuning or
+parameter sweeps, our solution saves more memory than a hand-tuned cache
+without any performance degradation, works for a variety of initial conditions,
+and could generalize to similar applications.
 
@@
 disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
-that the collection of policies we design for cache management in ParSplice
-are very similar to the policies used to load balance metadata in the Ceph file
-system (CephFS) suggesting that there is potential for automatically adapting
-and generating policies dynamically.
+that the collection of policies we design for cache management in ParSplice are
+very similar to the policies used to load balance metadata in the Ceph file
+system (CephFS~\cite{weil:osdi2006-ceph}) suggesting that there is potential
+for automatically adapting and generating policies dynamically.
 

### 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 is an addition to the 30GB that ParSplice uses to manage other ranks on the
-same node.  A single Cray node produced trajectories that are \(5\times\) times
-longer than our 10 node CloudLab clusters and \(25\times\) longer than our 10
-node cluster at UC Santa Cruz. As a result, it reaches different job phases faster and
-gives us a more comprehensive view of the workload. The performance gains
-compared to the commodity clusters have more to do with memory/PCI bandwidth
-than network.
-
-The scalability experiment uses 1 splicer, 1 persistent database, 1 cache
-process, and up to 2 workers. We scale up to 1024 tasks, which spans 32 nodes
-and disable hyper-threading because we experience unacceptable variability in
-performance. For the rest of the experiments, we use 8 nodes, 1 splicer, 1
-persistent database, 1 cache process, 1 worker, and up to 256 tasks.  The
-keyspace analysis that follows is for the cache on the worker node, which is
-circled in Figure~\ref{fig:parsplice}.
+same node.  The scalability experiment uses 1 splicer, 1 persistent database, 1
+cache process, and up to 2 workers. We scale up to 1024 tasks, which spans 32
+nodes and disable hyper-threading because we experience unacceptable
+variability in performance. For the rest of the experiments, we use 8 nodes, 1
+splicer, 1 persistent database, 1 cache process, 1 worker, and up to 256 tasks.
+The keyspace analysis that follows is for the cache on the worker node, which
+is circled in Figure~\ref{fig:parsplice}.
 
@@
   \caption{The keyspace size is small but must satisfy many reads as workers
-  calculate new segments. It is likely that we will need
+  calculate new segments. Memory usage scales linearly, so It is likely that we will need
   more than one node to manage segment coordinates when we scale the system or jobs up.
@@
   \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
-  \caption{The keyspace activity for ParSplice using two different growth
+  \caption{The key activity for ParSplice using two different growth
 rates.  The line shows the rate that EOM minima values are retrieved from the
 key-value store (\(y1\) axis) and the points along the bottom show the number
-of keys accessed in a 1 second sliding window (\(y2\) axis).
-\label{fig:motivation-regimes}}
+of keys accessed in a 1 second sliding window (\(y2\) axis).\label{fig:motivation-regimes}}
 \end{figure}
@@
 keyspace size (text annotations) and request load (bars) after a one hour run
-with a different number of tasks(\(x\) axis). While the keyspace size and
-capacity is relatively modest the memory usage scales linearly with the number
-of tasks. This is a problem if we want to scale to Trinitite's 3000 cores.
+with a different number of tasks (\(x\) axis). While the keyspace size and
+capacity is modest the memory usage scales linearly with the number
+of tasks, which is a problem if we want to scale to Trinitite's 3000 cores.
 Furthermore, the size of the keyspace also increases linearly with the length
@@
 The bars in Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many
-reads (\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same
-key for extended periods because the trajectory can remain stuck in so-called
+reads (\texttt{get()}) as writes (\texttt{put()}).  Tasks read the same
+key for extended periods because the trajectory gets stuck in so-called
 superbasins composed of tightly connected sets of states. In this case, many

### 2017-11-28 paper/sections/3_methodology.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 language and policy engine presented in~\cite{sevilla:sc15-mantle}. The
-prototype in that paper was called Mantle and it was originally built on CephFS
-so administrators could control file system metadata load balancing policies.
-We now refer to Mantle as a policy engine that supports our data management
-language.  The basic premise is that data management policies can be expressed
-with a simple API consisting of ``when", ``where", and ``how much". The ``when" policy
-controls how aggressive or conservative the decisions are; ``where" controls
-how distributed or concentrated the data should be; and ``how much" controls
-the amount of data that should be sent. There is also a ``load" policy that
-lets administrators specify how to collapse many metrics into a single load
-metric ({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).
+prototype in that paper, Mantle, was built on CephFS and lets administrators
+control file system metadata load balancing policies.  We now refer to Mantle
+as a policy engine that supports our data management language.  The basic
+premise is that data management policies can be expressed with a simple API
+consisting of ``when", ``where", and ``how much". The ``when" policy controls
+how aggressive or conservative the decisions are; ``where" controls how
+distributed or concentrated the data should be; and ``how much" controls the
+amount of data that should be sent. There is also a ``load" policy that lets
+administrators specify how to collapse many metrics into a single load metric
+({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).
 
@@
 we move ParSplice to a distributed key-value store back-end, the ``where" policy
-will be used determine which key-value pairs should be moved to which node.
+will be used to determine which key-value pairs should be moved to which node.
 
@@
 We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
-linked into a storage system.  Administrators write policies in Lua from whatever
+linked into a storage system service.  Administrators write policies in Lua from whatever
 domain they choose ({\it e.g.}, statistics, machine learning, storage system)
@@
 \subsubsection{Environment of Metrics} storage systems expose \textbf{cluster} metrics
-for describing resource utilization and \textbf{time series} metrics for
+for describing resource utilizations and \textbf{time series} metrics for
 describing accesses to some data structure over time. Table~\ref{table:metrics}
@@
 For time series metrics, the storage system passes an array of \texttt{(timestamp,
-value)} pairs to Mantle and the policies can iterate over the accesses.  The
+value)} pairs to Mantle and the policies can iterate over the values. The
 storage system uses a pointer to the time series to facilitate time series with many
@@
 % where policies are made
-We link Mantle directly into the worker cache circled in red in
-Figure~\ref{fig:parsplice}. We put the ``when" and ``how much" callbacks
-alongside code that checks for memory pressure. It is executed right before the
-worker processes incoming and outgoing put/get transactions to the cache. We
-only do cache management once every second to avoid maintaining the cache for
-every request. We expected to have to increase this polling interval to
-accommodate more complex policies but even our most complicated policy in
-Section~\S\ref{sec:dom-specific} had a negligible effect on performance when
-executed every second (within the standard deviation for multiple runs when
-compared against a policy that returns immediately). This may be because the
-worker is not overloaded and the bottleneck is somewhere else in the system.
-As stated previously, we do not use the ``where" part of Mantle because we
-focus on a single node, but this part of the API will be used when we move the
-caches and storage nodes to a key-values store back-end that uses key load
-balancing and repartitioning.
+We link Mantle into all caches in the system and put the ``when" and ``how
+much" callbacks alongside code that checks for memory pressure. It is executed
+right before the worker processes incoming and outgoing put/get transactions to
+the cache. We only do cache management once every second to avoid maintaining
+the cache for every request. We expected to have to increase this polling
+interval to accommodate more complex policies but even our most complicated
+policy in Section~\S\ref{sec:dom-specific} had a negligible effect on
+performance when executed every second (within the standard deviation for
+multiple runs when compared against a policy that returns immediately). This
+may be because the worker is not overloaded and the bottleneck is somewhere
+else in the system.  As stated previously, we do not use the ``where" part of
+Mantle because we focus on a single node, but this part of the API will be used
+when we move the caches and storage nodes to a key-values store back-end that
+uses key load balancing and repartitioning.

### 2017-11-28 paper/sections/4_dynamic-policies.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-Using the Mantle policy engine, we test a variety of cache management tools and
+Using the Mantle policy engine, we test a variety of cache management 
 algorithms on the worker using the keyspace analysis in
@@
 reducing the trajectory length and be fast enough to run as often as we want to
-detect key access patterns.  First we sized the cache according to our system
+detect key access patterns.  First we size the cache according to our system
 specific knowledge, {\it i.e.} the hardware and software of the storage
-hierarchy such as request rate, unique keys in a sliding window, and bandwidth
-capabilities.
+hierarchy.
 
@@
 Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
-unmodified ParSplice  measured in trajectory duration (\(y\) axis) and
+unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
 utilization is measured with memory footprint of just the cache (\(y2\) axis).
-The middle graph labeled ``Fixed Cache Size" shares the \(y\) axis and shows the
-trade-off of using a basic LRU-style cache for different cache sizes, where the
+The middle graph labeled ``Fixed Cache Size" shares the \(y\) axes and shows the
+trade-off of using a basic LRU-style cache of different sizes, where the
 penalty for a cache miss is retrieving the data from the persistent database.
@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
-\caption{Policy performance/utilization: ``None" is ParSplice unmodified,
-``Fixed Sized Cache" evicts keys using LRU , and ``Multi-Policy Cache" switches
-to fixed sized cache after absorbing the workload's initial burstiness.  Fixed
-sized caches are on the \(x\) axis.\label{fig:methodology-tradeoff}}
+\caption{Policy performance/utilization shows the trade-offs of different sized caches
+(\(x\) axis).  ``None" is ParSplice unmodified, ``Fixed Sized Cache" evicts
+keys using LRU, and ``Multi-Policy Cache" switches to fixed sized cache after
+absorbing the workload's initial burstiness.  \label{fig:methodology-tradeoff}}
 \end{figure}
@@
 	\caption{Memory utilization for ``No Cache Management" (unlimited cache
-growth), ``Multi-Policy" (absorbs initial burstiness of workload) , and
-``Dynamic Policy" (sizes cache according to key access patterns).
+growth), ``Multi-Policy" (absorbs initial burstiness of workload), and
+``Dynamic Policy" (sizes cache according to key access patterns). The ``Dynamic
+Policy" saves the most memory without sacrificing performance.
 \label{fig:memory-vs-time}}
@@
 read throughput on a small keyspace and reduces the memory pressure for a 2.5
-hour run. Our experiments show the effectiveness of the policy engine we
-integrated into ParSplice, not that we were able to identify the best policy
-for all system setups ({\it i.e.} different ParSplice parameters, number of
-worker tasks, and job lengths).  To solve that problem, we need a way to
-identify what thresholds we should use for different configurations.
+hour run.  To improve the policy even more, we need a way to identify what
+thresholds to use for different system setups ({\it e.g.}, different ParSplice
+parameters, number of worker tasks, and job lengths).
 
@@
 	\includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\
+
 \caption{Key activity for a 4 hour run shows groups of accesses to the same
 subset of keys. Detecting these access patterns leads to a more accurate cache
-management strategy, which is discussed in
-Section~\S\ref{sec:regime-detection}.\label{fig:keyspace-zoomed}}
+management strategy, which is discussed in Section~\S\ref{sec:regime-detection}
+and the results are in Figure~\ref{fig:dscache-vs-none}.
+\label{fig:keyspace-zoomed}}
     \end{subfigure}%
@@
         \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
+
 	\caption{Performance/Utilization for dynamically sized cache (DSCache)
 policy. With negligible performance degradation, DSCache adjusts to different
-\(\Delta\)s over 2\(\times\) memory in the best case.
+\(\Delta\)'s and saves 3\(\times\) as much memory in the best case.
 \label{fig:dscache-vs-none}}
+
     \end{subfigure}%
@@
 \end{figure*}
+
+\begin{figure}[tb]
+\footnotesize
+\begin{minted}[xleftmargin=3em,linenos]{lua}
+d = timeseries()
+ts, id = d:get(d:size())
+fan  = {start=nil, finish=ts, top=0, bot=id}
+fans = {}
+for i=d:size(),1,-1 do     -- iterate backwards
+  ts, id = d:get(i)
+  if id < fan['bot'] then  -- found a new fan!
+    fan['start']  = ts
+    fans[#fans+1] = fan 
+    fan = {start=nil, finish=ts, top=0, bot=id}
+  end
+
+  if id > fan['top'] then  -- track top of fan
+    fan['top'] = id 
+  end 
+end
+fan['start'] = 0 
+fans[#fans+1] = fan
+
+if #fans < 2 then -- do not evict current fan
+  return false 
+else
+  WRstate(fans[#fans-1]['top']-fans[1]['bot'])
+  return true
+end
+\end{minted}
+\caption{The dynamically sized cache policy, whose performance is shown in 
+Figure~\ref{fig:cache-management}, iterates backwards over timestamp-key pairs
+and detects when key accesses move on to a new subset of
+keys ({\it i.e.} ``fans"). \label{src:dyn-cache}}
+\end{figure}
 
@@
 
-Feeding application-specific knowledge about the ParSplice application into a policy
+Feeding application-specific knowledge about ParSplice into a policy
 leads to a more accurate cache management strategy.  The goal of the following
-sections is not to find an optimal solution, as this can be done with parameter
+section is not to find an optimal solution, as this can be done with parameter
 sweeps for thresholds; rather, we try to find techniques that work for a range
@@
 techniques proliferate more parameters that need to be tuned per
-hardware/software configuration. Furthermore, we find that many of the metrics
-do not signal a new set of key accesses.  Below, we indicate with quotes which
-parameters we need to add for each technique with and the value we find to work
-best, via tuning and parameter sweeps, for one set of initial conditions.
+hardware/software configuration. Furthermore, many of the metrics do not signal
+a new set of key accesses.  Below, we indicate with quotes which parameters we
+need to add for each technique and the value we find to work best, via
+tuning and parameter sweeps, for one set of initial conditions.
 
@@
 
-  \item Anomaly Detection: size of the image is too big and bottom edges are
+  \item Edge Detection: size of the image is too big and bottom edges are
 not thick enough.
@@
 key accesses (Line 13) so we can tell the ``how much" policy the number of keys
-to evict (Line 19).  The algorithm iterates backwards over the key access trace
-because a change in the minimum value signals a new group of key accesses. No
-signal exists iterating left to right, as the maximum value always increases
-and the minimum values at the bottom of each group of key accesses are sparse.
-For example, the maximum distance between values along the bottom edge of the
-zoomed group of key accesses in Figure~\ref{fig:keyspace-zoomed} is 125
-seconds, while the maximum distance between minimum values for the group of key
-accesses before is 0 seconds. As a result of this sparseness, iterating left to
-right requires a ``window size" parameter to determine when we think a minimum
-value will not show up again.  The algorithm is O(\(n\)), where \(n\) is the
-number events, but the benefit is that the approach avoids adding new
-thresholds for key access pattern detection ({\it e.g.}, space between key
-accesses, space between key IDs, and window size of consecutive key accesses).
+to evict (Line 19).  The algorithm is O(\(n\)), where \(n\) is the number
+events, but the benefit is that the approach avoids adding new thresholds for
+key access pattern detection ({\it e.g.}, space between key accesses, space
+between key IDs, and window size of consecutive key accesses).
+
+The algorithm iterates backwards over the key access trace because a change in
+the minimum value signals a new group of key accesses. No signal exists
+iterating left to right, as the maximum value always increases and the minimum
+values at the bottom of each group of key accesses are sparse.  For example,
+the maximum distance between values along the bottom edge of the zoomed group
+of key accesses in Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the
+maximum distance between minimum values for the group of key accesses before is
+0 seconds. As a result of this sparseness, iterating left to right requires a
+``window size" parameter to determine when we think a minimum value will not
+show up again.
 
@@
 than the default ParSplice configuration (no cache management) for the 3
-initial conditions we test. The memory usage is shown by the ``Dynamic Policy"
+initial conditions we test. The memory usage over time is shown by the ``Dynamic Policy"
 curves in Figure~\ref{fig:memory-vs-time}, where the behavior resembles the key
@@
 conditions.
-
-\begin{figure}[h]
-\footnotesize
-\begin{minted}[xleftmargin=3em,linenos]{lua}
-d = timeseries()
-ts, id = d:get(d:size())
-fan  = {start=nil, finish=ts, top=0, bot=id}
-fans = {}
-for i=d:size(),1,-1 do     -- iterate backwards
-  ts, id = d:get(i)
-  if id < fan['bot'] then  -- found a new fan!
-    fan['start']  = ts
-    fans[#fans+1] = fan 
-    fan = {start=nil, finish=ts, top=0, bot=id}
-  end
-
-  if id > fan['top'] then  -- track top of fan
-    fan['top'] = id 
-  end 
-end
-fan['start'] = 0 
-fans[#fans+1] = fan
-
-if #fans < 2 then -- do not evict current fan
-  return false 
-else
-  WRstate(fans[#fans-1]['top']-fans[1]['bot'])
-  return true
-end
-\end{minted}
-\caption{The dynamically sized cache policy, whose performance is shown in 
-Figure~\ref{fig:cache-management}, iterates backwards over timestamp-key pairs
-and detects when key accesses move on to a new subset of
-keys ({\it i.e.} ``fans"). \label{src:dyn-cache}}
-\end{figure}

### 2017-11-28 paper/sections/5_file-systems.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 {\it and} if it has already absorbed the initial burstiness of the workload.
-Much of this implementation was inspired by the file system metadata load
+Much of this implementation was inspired by the CephFS metadata load
 balancing policy in Figure~\ref{src:lua-cephfs}, which was presented
@@
 the number of consecutive instances that the server has been overloaded. We
-also clear the count (Line 10) if the server is no longer overloaded.
+also clear the count (Line 9) if the server is no longer overloaded.
 
@@
 These similarities among effective policies for two very different domains
-suggests that the heuristics and techniques in other load balancers can be used
-for cache management. The result suggests that the concepts and problems that
-architects grapple with are transcendent and the solutions they design can be
-re-used in different code bases.
+suggest that the heuristics and techniques in other load balancers can be used
+for cache management. The result support the notion that concepts and problems
+that architects grapple with are transcendent and the solutions they design can
+be re-used in different code bases.
 
@@
 vertical lines are the access patterns detected by the ParSplice cache
-management policy from Section~\S\ref{sec:dom-specific}.
+management policy from Section~\S\ref{sec:dom-specific}. The policy works
+``out-of-the-box", which suggests that data management policies are general and
+can be used across domains.
 \label{fig:trace-atime}}
@@
   \item Batching: how many operations to group together, when to send large
-batches of updates. We could use Mantle to implement to the pathname leases from
+batches of updates. We could use Mantle to implement pathname leases from
 IndexFS~\cite{ren:sc2014-indexfs} or the capabilities from

### 2017-11-28 paper/sections/6_conclusion.tex
add prefetching reference
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 State-of-the-art distributed file systems partition write-heavy workloads and
-replicate read-heavy workloads, similar to the approach we are advocating
-here.  IndexFS~\cite{ren:sc2014-indexfs} partitions directories and clients
-write to different partitions by grabbing leases and caching ancestor metadata
-for path traversal. ShardFS takes the replication approach to the extreme by
-copying all directory state to all nodes. The Ceph file system
-(CephFS)~\cite{weil:sc2004-dyn-metadata, weil:osdi2006-ceph} employs both
+replicate read-heavy workloads, similar to the approach we are advocating here.
+IndexFS~\cite{ren:sc2014-indexfs} partitions directories and clients write to
+different partitions by grabbing leases and caching ancestor metadata for path
+traversal. ShardFS~\cite{xiao:socc15-shardfs} takes the replication approach to
+the extreme by copying all directory state to all nodes.
+CephFS~\cite{weil:sc2004-dyn-metadata, weil:osdi2006-ceph} employs both
 techniques to a lesser extent; directories can be replicated or sharded but the
@@
 in this space by providing a framework for exploring these policies, but does
-not attempt anything more sophisticated ({\it e.g.}, machine learning) to create
-these policies.
+not attempt anything more sophisticated ({\it e.g.}, machine learning) to
+create these policies.
 
@@
 the primitives and strategies have enough in common with data management in
-file systems that they both can be expressed with similar semantics and a
-general policy engine.
+file systems that they both can be expressed with similar semantics.
 
@@
 applications and storage systems.  The value of such a collection eases the
-burden of policy development and paves the way for solutions that remove the
-administrator from the development cycle, such as (1) adaptable policies that
-automatically switch to new strategies when the current strategy behaves poorly
-({\it e.g.}, thrashing, making no progress, etc.), and (2) policy generation,
-where new policies are constructed automatically by examining the collection of
-existing policies.  Ultimately, we hope that this automation enables control of
-policies by machines instead of administrators.
+burden of policy development and paves the way for automated solutions such as
+(1) adaptable policies that switch to new strategies when the current strategy
+behaves poorly ({\it e.g.}, thrashing, making no progress, etc.), and (2)
+policy generation, where new policies are constructed by examining the
+collection of existing policies.  Ultimately, we hope that this automation
+enables control of policies by machines instead of administrators.

### 2017-11-28 paper/sections/1_introduction.tex
paper: fix typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
-weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
+weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. We evaluate a
 variety of caching policies using our data management language/policy engine and
@@
 % What is Mantle
-In this paper we are presenting the successful use of the Mantle policy language and engine to manage the dynamically-sized cache. Mantle provides a control plane that
-injects policies into a running storage system, such as a file system or key-value store. While Mantle
-was originally designed for file system metadata load balancing~\cite{sevilla:sc15-mantle},
-we found that it works surprisingly well for specifying cache management policies without requiring users to possess extensive knowledge about the internals of storage systems.
-We show that our framework:
+
+In this paper we are presenting the successful use of our data management
+language and Mantle policy engine to control the behavior of ParSplice's
+caches. Mantle provides a control plane that injects policies into a running
+storage system, such as a file system or key-value store. While Mantle was
+originally designed for file system metadata load
+balancing~\cite{sevilla:sc15-mantle}, we find that it works surprisingly well
+for specifying cache management policies without requiring users to possess
+extensive knowledge about the internals of storage systems. We show that our
+framework:
 

### 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\
-  \caption{The keyspace size is small but must satisfy many reads as workers
-  calculate new segments. Memory usage scales linearly, so It is likely that we will need
+  \caption{The keyspace is small but must satisfy many reads as workers
+  calculate segments. Memory usage scales linearly, so it is likely that we will need
   more than one node to manage segment coordinates when we scale the system or jobs up.
@@
   \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
-  \caption{The key activity for ParSplice using two different growth
-rates.  The line shows the rate that EOM minima values are retrieved from the
-key-value store (\(y1\) axis) and the points along the bottom show the number
-of keys accessed in a 1 second sliding window (\(y2\) axis).\label{fig:motivation-regimes}}
+  \caption{Key activity for ParSplice starts with many reads to a small set of
+keys and progresses to less reads to a larger set of keys.  The line shows the
+rate that EOM minima values are retrieved from the key-value store (\(y1\)
+axis) and the points along the bottom show the number of keys accessed in a 1
+second sliding window (\(y2\) axis). Despite having different growth rates
+(\(\Delta\)), the structure and behavior of the key activities are similar.
+\label{fig:motivation-regimes}}
 \end{figure}
@@
   \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keys.png}\\
-  \caption{The keyspace imbalance is due to workers generating deep
-  trajectories and reading the same coordinates. Over time, the accesses get
-  dispersed across different coordinates resulting in some keys being more
-  popular than others.\label{fig:methodology-keys}}
+  \caption{ Over time, tasks start to access a larger set of keys resulting in
+some keys being more popular than others.  Despite different growth rates
+(\(\Delta\)), the spatial locality of key accesses is similar between the two
+runs.  ({\it e.g.}, some keys are still read 5 times as many times others).
+\label{fig:methodology-keys}}
 \end{figure}

### 2017-11-28 paper/sections/3_methodology.tex
paper: addresses #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \begin{minted}{lua}
-d = timeseries()       -- data (d) from storage system
+d = timeseries()       -- d(ata) from storage system
 for i=1,d:size() do    -- iterate over timeseries

### 2017-11-28 paper/sections/4_dynamic-policies.tex
paper: addresses #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
-\caption{Policy performance/utilization shows the trade-offs of different sized caches
-(\(x\) axis).  ``None" is ParSplice unmodified, ``Fixed Sized Cache" evicts
-keys using LRU, and ``Multi-Policy Cache" switches to fixed sized cache after
-absorbing the workload's initial burstiness.  \label{fig:methodology-tradeoff}}
+\caption{Policy performance/utilization shows the trade-offs of different sized
+caches (\(x\) axis).  ``None" is ParSplice unmodified, ``Fixed Sized Cache"
+evicts keys using LRU, and ``Multi-Policy Cache" switches to fixed sized cache
+after absorbing the workload's initial burstiness. This parameter sweep
+identifies the ``Multi-Policy Cache" of 1K keys as the best solution but this
+only works for this system setup and initial configurations.
+ \label{fig:methodology-tradeoff}}
 \end{figure}
@@
 growth), ``Multi-Policy" (absorbs initial burstiness of workload), and
-``Dynamic Policy" (sizes cache according to key access patterns). The ``Dynamic
-Policy" saves the most memory without sacrificing performance.
+``Dynamic Policy" (sizes cache according to key access patterns). The dynamic
+policies saves the most memory without sacrificing performance.
 \label{fig:memory-vs-time}}
@@
 
-	\caption{Performance/Utilization for dynamically sized cache (DSCache)
-policy. With negligible performance degradation, DSCache adjusts to different
-\(\Delta\)'s and saves 3\(\times\) as much memory in the best case.
+	\caption{The performance/utilization for the dynamically sized cache (DSCache)
+policy. With negligible performance degradation, DSCache adjusts to different initial configurations
+(\(\Delta\)'s) and saves 3\(\times\) as much memory in the best case.
 \label{fig:dscache-vs-none}}

### 2017-11-28 paper/sections/5_file-systems.tex
paper: add to Fig12 to address #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 vertical lines are the access patterns detected by the ParSplice cache
-management policy from Section~\S\ref{sec:dom-specific}. The policy works
-``out-of-the-box", which suggests that data management policies are general and
-can be used across domains.
+management policy from Section~\S\ref{sec:dom-specific}. A load balancing
+policy can use the same pattern detection to make decisions, such as avoiding
+migration when the workload is accessing the same subset of keys or keeping
+groups of accesses local to a single server.
 \label{fig:trace-atime}}

### 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #5
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-%
-%We demonstrate these contributions by changing the input to our molecular
-%dynamics simulation, which changes the keyspace access patterns. 
-%
-%\subsection{Structured Access Regimes}
-%
-%Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
-%%shows ``access regimes", or groups of accesses
-%%to the same subset of keys:
-%\begin{itemize}
-%  \item change immediately, different sizes
-%  \item monotonically increasing
-%  \item random access to a single key
-%  \item early keys accessed more
-%\end{itemize}
-%
-%\noindent\textbf{Conclusion}: unique patterns of a real HPC application
-%
-%\subsection{Cache Size Analysis}
-%
-%Figure~\ref{fig:keyspace-analysis_cachesize}.
-%
-%\noindent\textbf{Conclusion}: workload phases (high request rate to a small
-%number of keys and then low request rate to a large number keys) need a dynamic
-%load balancing policy.
-%
-%\subsection{Overfitting Policies}
-%
-%Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.
-%
-%\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
-%sensitive to: max usage or usage over time.\\
-%
-%\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
-%request rate for a 2.5 hour run, but it is infeasible to do this for every
-%combination of system setup, job lengths, parsplice parameters.
-%
 %\subsubsection*{Discussion}
@@
 
-\subsection{Results and Conclusions}
-\label{sec:parsplice-keyspace-analysis}
 %IT HAS 8K keys!  How did we take these measurements
+\subsection{Experimental Setup}
 
@@
 
-\subsubsection*{Experimental Setup} All experiments ran on Trinitite, a Cray
-XC40 with 32 Intel Haswell 2.3GHz cores per node.  Each node has 128GB of RAM
-and our goal is to limit the size of the cache to 3\% of RAM . Note that this
-is an addition to the 30GB that ParSplice uses to manage other ranks on the
-same node.  The scalability experiment uses 1 splicer, 1 persistent database, 1
-cache process, and up to 2 workers. We scale up to 1024 tasks, which spans 32
-nodes and disable hyper-threading because we experience unacceptable
-variability in performance. For the rest of the experiments, we use 8 nodes, 1
-splicer, 1 persistent database, 1 cache process, 1 worker, and up to 256 tasks.
-The keyspace analysis that follows is for the cache on the worker node, which
-is circled in Figure~\ref{fig:parsplice}.
+All experiments ran on Trinitite, a Cray XC40 with 32 Intel Haswell 2.3GHz
+cores per node.  Each node has 128GB of RAM and our goal is to limit the size
+of the cache to 3\% of RAM . Note that this is an addition to the 30GB that
+ParSplice uses to manage other ranks on the same node.  The scalability
+experiment uses 1 splicer, 1 persistent database, 1 cache process, and up to 2
+workers. We scale up to 1024 tasks, which spans 32 nodes and disable
+hyper-threading because we experience unacceptable variability in performance.
+For the rest of the experiments, we use 8 nodes, 1 splicer, 1 persistent
+database, 1 cache process, 1 worker, and up to 256 tasks.  The keyspace
+analysis that follows is for the cache on the worker node, which is circled in
+Figure~\ref{fig:parsplice}.
+
+\subsection{Results and Observations}
+\label{sec:parsplice-keyspace-analysis}
 
@@
 
-\subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
+Our analysis shows that ParSplice accesses keys in a structured and predictable
+way. The following 3 observations shape the policies we design later in the
+paper.
+
+\subsubsection{\underline{Scalability}} Figure~\ref{fig:methodology-keyspace} shows the
 keyspace size (text annotations) and request load (bars) after a one hour run
@@
 
-\subsubsection*{An active but small keyspace}
+\subsubsection{\underline{An active but small keyspace}}
 
@@
 
-\subsubsection*{Entropy increases over time} The reads per second in
+\subsubsection{\underline{Entropy increases over time}} The reads per second in
 Figure~\ref{fig:motivation-regimes} show that the number of requests decreases

### 2017-11-28 paper/sections/1_introduction.tex
paper: more typos and final read through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
-patil:fast2011-giga+}. But the policies for what data to evict and when to
-evict are specific to the application's behavior and the system's configuration
-(hardware, settings, etc.) so a new workload may prove to be a poor match for
-the selected caching
+patil:fast2011-giga+, weil:sc2004-dyn-metadata}. But the policies for what data
+to evict and when to evict are specific to the application's behavior and the
+hardware configuration so a new workload may prove to be a poor match for the
+selected caching
 policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
-weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. We evaluate a
-variety of caching policies using our data management language/policy engine and
-arrive at a customized policy that works well for our target application.  This
-process of trying general policies and quickly iterating to a customized
-solution shows that our prototype can be adapted to different workloads.
+weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. We evaluate a variety of caching
+policies using our data management language/policy engine and arrive at a
+customized policy that works well for our example application,
+ParSplice~\cite{perez:jctc20150parsplice}.
 
-Our example application, ParSplice~\cite{perez:jctc20150parsplice}, is
-representative of an important class of applications with similar working set
-behaviors that extensively use software-based caches.  ParSplice is a molecular
-dynamics simulation that uses a hierarchy of caches and a single persistent
-key-value store to store both observed minima across a molecule's equation of
-motion (EOM) and the hundreds or thousands of partial trajectories calculated
-each second during a parallel job.  This workload is pervasive across
-simulations that rely on a mesh-based decomposition of a physical region and
-result in millions or billions of mesh cells, where each cell contains
-materials, pressures, temperatures and other characteristics that are required
-to accurately simulate phenomena of interest.  The fine-grained data annotation
-capabilities provided by key-value storage is a natural match for these types
-of scientific simulations.  Unfortunately, simulations of this size saturate
-the capacity and bandwidth capabilities of a single node so we need more
-effective data management techniques.
+ParSplice, the molecular dynamics simulation, is representative of an important
+class of HPC applications with similar working set behaviors that extensively
+use software-based caches. It uses a hierarchy of caches and a single
+persistent key-value store to store both observed minima across a molecule's
+equation of motion (EOM) and the hundreds or thousands of partial trajectories
+calculated each second during a parallel job.  This workload is pervasive
+across simulations that (1) rely on a mesh-based decomposition of a physical
+region and (2) result in millions or billions of mesh cells, where each cell
+contains materials, pressures, temperatures and other characteristics that are
+required to accurately simulate phenomena of interest.  The fine-grained data
+annotation capabilities provided by key-value storage is a natural match for
+these types of scientific simulations.  Unfortunately, simulations of this size
+saturate the capacity and bandwidth capabilities of a single node so we need
+more effective data management techniques.
 

### 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
paper: more typos and final read through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-%\subsubsection*{Discussion}
-%Why don't we just an LRU cache:
-%\begin{itemize}
-%  \item we can do better (workload is structured and has locality)
-%  \item finding the size of the cache is hard
-%  \item hotspots dissipate too quickly
-%\end{itemize}
-
 \section{ParSplice Keyspace Analysis}
@@
 cores per node.  Each node has 128GB of RAM and our goal is to limit the size
-of the cache to 3\% of RAM . Note that this is an addition to the 30GB that
+of the cache to 3\% of RAM. Note that this is an addition to the 30GB that
 ParSplice uses to manage other ranks on the same node.  The scalability
@@
 
-\begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
-  \caption{Key activity for ParSplice starts with many reads to a small set of
-keys and progresses to less reads to a larger set of keys.  The line shows the
-rate that EOM minima values are retrieved from the key-value store (\(y1\)
-axis) and the points along the bottom show the number of keys accessed in a 1
-second sliding window (\(y2\) axis). Despite having different growth rates
-(\(\Delta\)), the structure and behavior of the key activities are similar.
-\label{fig:motivation-regimes}}
-\end{figure}
-
-\begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keys.png}\\
-  \caption{ Over time, tasks start to access a larger set of keys resulting in
-some keys being more popular than others.  Despite different growth rates
-(\(\Delta\)), the spatial locality of key accesses is similar between the two
-runs.  ({\it e.g.}, some keys are still read 5 times as many times others).
-\label{fig:methodology-keys}}
-\end{figure}
-
 Our analysis shows that ParSplice accesses keys in a structured and predictable
-way. The following 3 observations shape the policies we design later in the
+way. The following 4 observations shape the policies we design later in the
 paper.
@@
 
+The bars in Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many
+reads (\texttt{get()}) as writes (\texttt{put()}).  Tasks read the same key for
+extended periods because the trajectory gets stuck in so-called superbasins
+composed of tightly connected sets of states.  Writes only occur for the final
+state of segments generated by tasks; their magnitude is smaller than reads
+because the caches ignore redundant write requests.
+
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
+  \caption{Key activity for ParSplice starts with many reads to a small set of
+keys and progresses to less reads to a larger set of keys.  The line shows the
+rate that EOM minima values are retrieved from the key-value store (\(y1\)
+axis) and the points along the bottom show the number of unique keys accessed in a 1
+second sliding window (\(y2\) axis). Despite having different growth rates
+(\(\Delta\)), the structure and behavior of the key activities are similar.
+\label{fig:motivation-regimes}}
+\end{figure}
+
+\subsubsection{\underline{Initial conditions influence key activity}}
 Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
@@
 resulting in hotter keys and a smaller keyspace.  Values smaller than
-\(\Delta_2\)'s growth rate or a temperature of 400 degrees results in very
+\(\Delta_2\)'s growth rate or a temperature of 400 degrees result in very
 little database activity because state transitions take too long. Similarly,
@@
 
-The bars in Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many
-reads (\texttt{get()}) as writes (\texttt{put()}).  Tasks read the same
-key for extended periods because the trajectory gets stuck in so-called
-superbasins composed of tightly connected sets of states. In this case, many
-trajectory segments with the same coordinates are needed before the trajectory
-moves on.  Writes only occur for the final state of segments generated by
-tasks; their magnitude is smaller than reads because the caches ignore
-redundant write requests. The number of read and write requests are highest at
-the beginning of the run when tasks generate segments for the same state, which
-is computationally cheap (this motivates Section~\S\ref{sec:arch-specific}).
+\begin{figure}[t]
+  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keys.png}\\
+  \caption{Over time, tasks start to access a larger set of keys resulting in
+some keys being more popular than others.  Despite different growth rates
+(\(\Delta\)), the spatial locality of key accesses is similar between the two
+runs.  ({\it e.g.}, some keys are still read 5 times as many times others).
+\label{fig:methodology-keys}}
+\end{figure}
 
@@
 Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
-and the number of active keys increases over time. The resulting key access
-imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
-shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
-unique state, or key, along the \(x\) axis. Keys are more popular than others
-(up to \(5\times\)) because worker tasks start generating states with different
+and the number of active keys increases over time.  The number of read and
+write requests are highest at the beginning of the run when tasks generate
+segments for the same state, which is computationally cheap (this motivates
+Section~\S\ref{sec:arch-specific}).  The resulting key access imbalance for the
+two growth rates in Figure~\ref{fig:motivation-regimes} are shown in
+Figure~\ref{fig:methodology-keys}, where reads are plotted for each unique
+state, or key, along the \(x\) axis. Keys are more popular than others (up to
+\(5\times\)) because worker tasks start generating states with different
 coordinates later in the run.  Figure~\ref{fig:methodology-keys} shows that the
@@
 still \(5\times\) more popular than others).
+
 Figure~\ref{fig:motivation-regimes} shows how entropy for different growth

### 2017-11-28 paper/sections/4_dynamic-policies.tex
paper: more typos and final read through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % technical details
-The results for different cache sizes for a growth rate of \(\Delta_1\) over a
-2.5 hour run across 256 workers is shown in
-Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
-unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
-utilization is measured with memory footprint of just the cache (\(y2\) axis).
-The middle graph labeled ``Fixed Cache Size" shares the \(y\) axes and shows the
-trade-off of using a basic LRU-style cache of different sizes, where the
-penalty for a cache miss is retrieving the data from the persistent database.
-We evict keys (if necessary) at every operation instead of when segments
-complete because the cache fills up too quickly otherwise. This is implemented
-using a ``when" policy of:
+We implement a basic LRU cache using a ``when" policy of:
 
@@
 
+The results for different cache sizes for a growth rate of \(\Delta_1\) over a
+2.5 hour run across 256 workers is shown in
+Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
+unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
+utilization is measured with memory footprint of just the cache (\(y2\) axis).
+The middle graph labeled ``Fixed Cache Size" shares the \(y\) axes and shows
+the trade-off of using a basic LRU-style cache of different sizes, where the
+penalty for a cache miss is retrieving the data from the persistent database.
 The error bars are the standard deviation of 3 runs.  Although the keyspace
@@
 some keys may be accessed at time \(t_0\), not in \(t_1\), and then again in
-\(t_2\). Because the cache is too small, reads and writes hit rest of the
-storage hierarchy and the excessive traffic triggers a LevelDB compaction on
-the persistent database.  To avoid overloading the storage hierarchy with
-requests, we design a multi-part policy that switches between:
+\(t_2\). Because the cache is too small, reads and writes fall through to the
+rest of the storage hierarchy and the excessive traffic triggers a LevelDB
+compaction on the persistent database.  To avoid these compactions, which
+temporarily block operations, we design a multi-policy cache that switches
+between:
 
@@
 cache down to the threshold.  The actual policy is shown and described in more
-detail in the next section in Figure~\ref{src:lru-dyn}.  The plot on the right
-side of Figure~\ref{fig:methodology-tradeoff} shows the performance/utilization
-trade-off of the multi-part policy, where the cache sizes for the \(n\) key
-limit policy are along the \(x\) axis.  The performance and memory utilization
-for a 100K key cache size is the same as the 100K bar in the ``Fixed Cache
-Size" graph in Figure~\ref{fig:methodology-tradeoff} but the rest reduce the
-size of the keyspace after the read flash crowd.  We see the worst performance
-when the engine switches to the 10 key limit policy, which achieves 94\% of the
-performance while only using 40KB of memory.
+detail in Section~\S\ref{sec:scope} in Figure~\ref{src:lru-dyn}.  The plot on
+the right side of Figure~\ref{fig:methodology-tradeoff} shows the
+performance/utilization trade-off of the multi-policy cache, where the cache
+sizes for the \(n\) key limit policy are along the \(x\) axis.  The performance
+and memory utilization for a 100K key cache size is the same as the 100K bar in
+the ``Fixed Cache Size" graph in Figure~\ref{fig:methodology-tradeoff} but the
+rest reduce the size of the keyspace after the read flash crowd.  We see the
+worst performance when the policy switches to the 10 key limit policy, which
+achieves 94\% of the performance while only using 40KB of memory.
 
@@
 \end{minted}
-\caption{The dynamically sized cache policy, whose performance is shown in 
-Figure~\ref{fig:cache-management}, iterates backwards over timestamp-key pairs
-and detects when key accesses move on to a new subset of
-keys ({\it i.e.} ``fans"). \label{src:dyn-cache}}
+\caption{The dynamically sized cache policy iterates backwards over
+timestamp-key pairs and detects when accesses move on to a new subset of keys
+({\it i.e.} ``fans"). The performance and total memory usage is in
+Figure~\ref{fig:dscache-vs-none} and the memory usage over time is in
+Figure~\ref{fig:memory-vs-time}.
+\label{src:dyn-cache}}
 \end{figure}
@@
 keys are accessed more often.  Despite these patterns, the following
-characteristics of superbasins make it hard to detect them:
+characteristics of superbasins make them hard to detect:
 
@@
 
-After trying these techniques we found that the basic algorithm in
+After trying these techniques we found that the basic O(\(n\)) algorithm in
 Figure~\ref{src:dyn-cache} works best. The algorithm detects groups of key
@@
 key accesses (Line 13) so we can tell the ``how much" policy the number of keys
-to evict (Line 19).  The algorithm is O(\(n\)), where \(n\) is the number
+to evict (Line 23).  The algorithm is O(\(n\)), where \(n\) is the number
 events, but the benefit is that the approach avoids adding new thresholds for

### 2017-11-28 paper/sections/5_file-systems.tex
paper: more typos and final read through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-11-28 paper/sections/5_file-systems.tex
@@
         workload before switching to a constrained cache.  The
-        performance/utilization for different  \texttt{n} is in
+        performance/utilization for different  \(n\) is in
         Figure~\ref{fig:methodology-tradeoff}. \label{src:lru-dyn}}
@@
 \end{figure*}
-
-%\begin{itemize}
-%  \item Lustre Trace
-%  \item LinkedIn Trace
-%  \item Nathan's Trace
-%\end{itemize}
 
@@
 suggest that the heuristics and techniques in other load balancers can be used
-for cache management. The result support the notion that concepts and problems
-that architects grapple with are transcendent and the solutions they design can
-be re-used in different code bases.
+for cache management. The result supports the notion that concepts and problems
+that architects grapple with are transcendent across domains and the solutions
+they design can be re-used in different code bases.
 
@@
 vertical lines are the access patterns detected by the ParSplice cache
-management policy from Section~\S\ref{sec:dom-specific}. The policy works
-``out-of-the-box", which suggests that data management policies are general and
-can be used across domains.
-\label{fig:trace-atime}}
+management policy from Section~\S\ref{sec:dom-specific}. A file system that
+load balances metadata across a cluster of servers could use the same pattern
+detection to make migration decisions, such as avoiding migration when the
+workload is accessing the same subset of keys or keeping groups of accesses
+local to a server.  \label{fig:trace-atime}}
 \end{figure}
@@
 from the caching policies developed in Sections~\S\ref{sec:arch-specific}
-and~\ref{sec:dom-specific}. For example, Ceph administrators can use the
+and~\S\ref{sec:dom-specific}. For example, Ceph administrators can use the
 policies in ParSplice to automatically size and manage cache

### 2017-11-28 paper/sections/1_introduction.tex
paper: final draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-ParSplice, the molecular dynamics simulation, is representative of an important
+The ParSplice molecular dynamics simulation is representative of an important
 class of HPC applications with similar working set behaviors that extensively

### 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
paper: final draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 the read request rate (\(y1\) axis) and the dots along the bottom are the
-number of unique keys accessed (\(y2\) axis).  The \(\Delta_1\) growth rate
-adds atoms every 100K microseconds while the \(\Delta_2\) growth rate adds
-atoms every 1 million microseconds. So \(\Delta_2\) has a smaller growth rate
-resulting in hotter keys and a smaller keyspace.  Values smaller than
-\(\Delta_2\)'s growth rate or a temperature of 400 degrees result in very
-little database activity because state transitions take too long. Similarly,
-values larger than \(\Delta_1\)'s growth rate or a temperature of 4000 degrees
-result in an equally meaningless simulation as transitions are unrealistic.
+number of unique keys accessed (\(y2\) axis).  The access patterns for
+different growth rates have temporal locality, as the reads per second for
+\(\Delta_2\) look like the reads per second for \(\Delta_1\) stretched out
+along the time axis.  The \(\Delta_1\) growth rate adds atoms every 100K
+microseconds while the \(\Delta_2\) growth rate adds atoms every 1 million
+microseconds. So \(\Delta_2\) has a smaller growth rate resulting in hotter
+keys and a smaller keyspace.  Values smaller than \(\Delta_2\)'s growth rate or
+a temperature of 400 degrees result in very little database activity because
+state transitions take too long. Similarly, values larger than \(\Delta_1\)'s
+growth rate or a temperature of 4000 degrees result in an equally meaningless
+simulation as transitions are unrealistic.
+
 This figure demonstrates that small changes to \(\Delta\) can have a strong
 effect on the timing and frequency with which new EOM minima are discovered and
-referenced. This finding suggests that we need a flexible policy language and
-engine to explore these trade-offs.
+referenced.  Trends also exist for temperature and number of workers but are
+omitted here for space.  This finding suggests that we need a flexible policy
+language and engine to explore these trade-offs.
 
@@
 \(5\times\)) because worker tasks start generating states with different
-coordinates later in the run.  Figure~\ref{fig:methodology-keys} shows that the
+coordinates later in the run.  Figure~\ref{fig:methodology-keys} also shows that the
 number of reads changes with different initial conditions (\(\Delta\)), but
@@
 still \(5\times\) more popular than others).
-
-Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
-rates has temporal locality, as the reads per second for \(\Delta_2\) look like
-the reads per second for \(\Delta_1\) stretched out along the time axis.
-Trends also exist for temperature and number of workers but are omitted here
-for space. This structure means that we can learn the key access patterns and
-adapt the storage system accordingly (this motivates
-Section~\S\ref{sec:dom-specific}).

### 2018-02-20 paper/sections/1_introduction.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\begin{abstract}
-
-Our analysis of the key-value activity generated by the ParSplice molecular
-dynamics simulation demonstrates the need for more complex cache management
-strategies. Baseline measurements show clear key access patterns and hot spots
-that offer significant opportunity for optimization. We use the data management
-language and policy engine from the Mantle system to dynamically explore a
-variety of techniques, ranging from basic algorithms and heuristics to
-statistical models, calculus, and machine learning. While Mantle was originally
-designed for distributed file systems, we show how the collection of
-abstractions effectively decomposes the problem into manageable policies for a
-different application and storage system.  Our exploration of this space results in a
-dynamically sized cache policy that does not sacrifice any performance while
-using 32-66\% less memory than the default ParSplice configuration.
-
-\end{abstract}
-
 \section{Introduction}
@@
 
-The ParSplice molecular dynamics simulation is representative of an important
+ParSplice molecular dynamics simulations are representative of an important
 class of HPC applications with similar working set behaviors that extensively
-use software-based caches. It uses a hierarchy of caches and a single
-persistent key-value store to store both observed minima across a molecule's
-equation of motion (EOM) and the hundreds or thousands of partial trajectories
-calculated each second during a parallel job.  This workload is pervasive
+use software-based caches. ParSplice uses a hierarchy of caches and a single
+persistent key-value store to store the atomic coordinates corresponding
+to local energy minima (referred to simply as minima) encountered
+by a large number of independent dynamical trajectories.
+This workload is also pervasive
 across simulations that (1) rely on a mesh-based decomposition of a physical
@@
 memory without sacrificing performance and works for a variety of inputs.
+Solutions labeled ``no performance degradation" are in comparison to the
+baseline of using a cache of unlimited size (``No Cache Management").
 \label{fig:cache-management}}
@@
 
-The biggest challenge for ParSplice is properly sizing the caches in the
-storage hierarchy.  The memory usage for a single cache that stores molecule
+A challenge for ParSplice is properly sizing the caches in the
+storage hierarchy.  The memory usage for a single cache that stores atomic
 coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
@@
 without any performance degradation, works for a variety of initial conditions,
-and could generalize to similar applications.
+and could generalize to similar applications. Triggering key eviction at a
+certain memory pressure ({\it e.g.} 3\%) requires {\it a priori} system
+knowledge while our approach saves the most memory because we model the
+behavior of keyspace accesses.
 
@@
 and~\S\ref{sec:dom-specific}, where we try a range of policies from different
-disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
-that the collection of policies we design for cache management in ParSplice are
-very similar to the policies used to load balance metadata in the Ceph file
-system (CephFS~\cite{weil:osdi2006-ceph}) suggesting that there is potential
-for automatically adapting and generating policies dynamically.
+disciplines. We analyze why our early policies fail for our HPC hardware and
+software and show how we iterated to the final solution.  More importantly, in
+Section~\S\ref{sec:scope}, we conclude that the collection of policies we
+design for cache management in ParSplice are very similar to the policies used
+to load balance metadata in the Ceph file system
+(CephFS~\cite{weil:osdi2006-ceph}) suggesting that there is potential for
+automatically adapting and generating policies dynamically. This paper does not
+thoroughly study cache management policies for file systems, but our findings
+for this specific application and related storage systems work suggests that
+this approach is a strong avenue for future work.
 

### 2018-02-20 paper/sections/2_parsplice-keyspace-analysis.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 (MD) simulation package developed at LANL. It is part of the Exascale Computing
-Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
+Project\footnote{https://www.exascaleproject.org/} and is important to LANL's
 Materials for the Future initiative.
@@
 reads, ParSplice provisions a hierarchy of caches that sit in front of a single
-persistent database.  Values are written to each tier and reads traverse up the
-hierarchy until they find the data.
+persistent database.  Values corresponding to new keys are written to each tier
+and reads traverse up the hierarchy until they find the data.
 
@@
 trajectory can jump from state to state; higher temperatures lead to more
-frequent transitions but temperatures that are too high result in meaningless
-simulations because trajectories have so much energy that they are equally
-likely to visit any random state.
+frequent transitions but temperatures that are too high lead to the melting of the
+nanoparticle and hence to a complete change in the physics of the system.
 
@@
   \caption{The keyspace is small but must satisfy many reads as workers
-  calculate segments. Memory usage scales linearly, so it is likely that we will need
-  more than one node to manage segment coordinates when we scale the system or jobs up.
+  calculate segments. Memory usage scales linearly, so it is likely that we will
+  need more than one node to manage segment coordinates when we scale the system
+  or jobs up. \(\Delta_1\) is the growth rate, an application specific parameter
+  described in Section~\S\ref{sec:delta}.
   \label{fig:methodology-keyspace}}
@@
 \subsubsection{\underline{Initial conditions influence key activity}}
+\label{sec:delta}
 Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
@@
 along the time axis.  The \(\Delta_1\) growth rate adds atoms every 100K
-microseconds while the \(\Delta_2\) growth rate adds atoms every 1 million
-microseconds. So \(\Delta_2\) has a smaller growth rate resulting in hotter
+picoseconds while the \(\Delta_2\) growth rate adds atoms every 1 million
+picoseconds. So \(\Delta_2\) has a smaller growth rate resulting in hotter
 keys and a smaller keyspace.  Values smaller than \(\Delta_2\)'s growth rate or
 a temperature of 400 degrees result in very little database activity because
-state transitions take too long. Similarly, values larger than \(\Delta_1\)'s
-growth rate or a temperature of 4000 degrees result in an equally meaningless
-simulation as transitions are unrealistic.
+state transitions take very long. Similarly, growth rate values larger than \(\Delta_1\)
+lead to far out-of-equilibrium growth, while temperatures in excess of 800 degrees
+result in the melting of the nanoparticle, which is not relevant in this application.
 
 This figure demonstrates that small changes to \(\Delta\) can have a strong
-effect on the timing and frequency with which new EOM minima are discovered and
+effect on the timing and frequency with which new minima are discovered and
 referenced.  Trends also exist for temperature and number of workers but are

### 2018-02-20 paper/sections/4_dynamic-policies.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 We implement a basic LRU cache using a ``when" policy of:
-
-\begin{figure}[h]
-\footnotesize
-\begin{minted}{lua}
-server[whoami]['cachesize']>n
-\end{minted}
-\end{figure}
-
-and a ``how much" policy of:
-
-\begin{figure}[h]
-\footnotesize
-\begin{minted}{lua}
-servers[whoami]['cachesize']-n
-\end{minted}
-\end{figure}
-
-The results for different cache sizes for a growth rate of \(\Delta_1\) over a
-2.5 hour run across 256 workers is shown in
-Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
-unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
+\texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
+\texttt{servers[whoami]['cachesize']-n}.  The results for different cache sizes
+for a growth rate of \(\Delta_1\) over a 2.5 hour run across 256 workers is
+shown in Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance
+of unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
 utilization is measured with memory footprint of just the cache (\(y2\) axis).
@@
         \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
-
 	\caption{The performance/utilization for the dynamically sized cache (DSCache)
 policy. With negligible performance degradation, DSCache adjusts to different initial configurations
-(\(\Delta\)'s) and saves 3\(\times\) as much memory in the best case.
+(\(\Delta\)s) and saves 3\(\times\) as much memory in the best case.
 \label{fig:dscache-vs-none}}
-
     \end{subfigure}%
@@
 tasks over time (\(x\) axis). The groups of accesses to a subset of keys occurs
-because molecules are stuck in deep trajectories. Recall that the cache stores
-the molecules' EOM minima, which is the smallest effective energy that a
-molecule observes during its trajectory. So molecules stuck in deep
-trajectories explore the same minima until they can escape to a new set of
-states. This exploration of the same set of states is called a superbasin.  In
-Figure~\ref{fig:keyspace-zoomed}, superbasins are never re-visited because the
-simulation only adds molecules; we can never reach a state with less molecules.
-This is why keys are never re-accessed.
+because the system is stuck in so-called superbasins, {\it i.e.} coarse regions of space from
+which it is difficult to escape, but within which it is easy to move. Systems
+stuck in superbasins will explore the same set of minima for a
+long time, leading to the same keys being accessed repeatedly. In
+Figure~\ref{fig:keyspace-zoomed}, superbasins are typically not re-visited because the
+simulation only adds atoms; we can never revisit a superbasin that contains less atoms
+than the simulation currently contains.
+This is why keys are never re-accessed after a given amount of time.
 

### 2018-02-20 paper/sections/6_conclusion.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Auto-tuning is a well-known technique used in
-HPC~\cite{behzad:sc2013-autotuning, behzad:techreport2014-io-autotuning}, big
-data systems systems~\cite{herodotou_starfish_2011}, and
+HPC~\cite{behzad:sc2013-autotuning}, big data systems
+systems~\cite{herodotou_starfish_2011}, and
 databases~\cite{schnaitter_index_2009}.  Like our work, these systems focus on
@@
 did not need anything as sophisticated as the genetic algorithm used
-in~\cite{behzad:sc2013-autotuning}.  We cannot drop these techniques into
-ParSplice because the magnitude and speed of the workload hotspots/flash crowds
-makes existing approaches less applicable.
+in~\cite{behzad:sc2013-autotuning}.  
+%We cannot drop these techniques into
+%ParSplice because the magnitude and speed of the workload hotspots/flash crowds
+%makes existing approaches less applicable.
 

### 2018-02-20 paper/sections/2_parsplice-keyspace-analysis.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
@@
   \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\
+
   \caption{The keyspace is small but must satisfy many reads as workers
-  calculate segments. Memory usage scales linearly, so it is likely that we will need
-  more than one node to manage segment coordinates when we scale the system or jobs up.
+  calculate segments. Memory usage scales linearly, so it is likely that we will
+  need more than one node to manage segment coordinates when we scale the system
+  or jobs up. \(\Delta_1\) is the growth rate, an application specific parameter
+  described in Section~\S\ref{sec:delta}.
+
   \label{fig:methodology-keyspace}}

### 2018-02-20 paper/sections/3_methodology.tex
paper: more typos and final read through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 an MPI-based storage system can access the CPU utilization of the first rank in a
-communication group using:
-
-\begin{figure}[h]
-\footnotesize
-\begin{minted}{lua}
-load = servers[0]['cpu']
-\end{minted}
-\end{figure}
+communication group using: \texttt{servers[0]['cpu']}.
 
@@
 This decision limits the time series metrics to only include values from the
-{\it current } node, although this is not a limitation of Mantle itself.  For
+{\it current} node, although this is not a limitation of Mantle itself.  For
 example, a policy that uses accesses to a directory in a file system as a
@@
 Figure~\ref{fig:mantle} shows an example policy for ``when()", where the
-current server (\texttt{whoami}) migrates load if it is has load
-(\texttt{>0.1}) and if its neighbor server (\texttt{whoami + 1}) does not have
-load (\texttt{<0.1}). The load is calculated using the metrics provided by the
-environment.
+current server migrates work if it is has load and if its neighbor does not
+have load; \texttt{whoami} is the current server, its neighbor is
+\texttt{whoami+1}, and the load threshold is 0.1. The load is calculated using
+the metrics provided by the environment.
 

### 2018-02-21 paper/sections/1_introduction.tex
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Solutions labeled ``no performance degradation" are in comparison to the
-baseline of using a cache of unlimited size (``No Cache Management").
+baseline of a cache of unlimited size.
 \label{fig:cache-management}}
@@
 and could generalize to similar applications. Triggering key eviction at a
-certain memory pressure ({\it e.g.} 3\%) requires {\it a priori} system
+certain memory pressure ({\it e.g.}, 3\%) requires {\it a priori} system
 knowledge while our approach saves the most memory because we model the

### 2018-02-21 paper/sections/2_parsplice-keyspace-analysis.tex
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \begin{figure}[t]
-\noindent\includegraphics[width=0.5\textwidth]{figures/parsplice.png}\\
-\caption{The ParSplice architecture has a storage hierarchy of caches (boxes) and a
+\centering
+\noindent\includegraphics[width=0.35\textwidth]{figures/parsplice.png}\\
+\caption{ParSplice has a storage hierarchy of caches (boxes) and a
 dedicated cache process (large box) backed by a persistent database (DB). A splicer
-(S) tells workers (W) to generate segments and workers employ tasks (T) for more
+(S) tells workers (W) to generate segments. Workers employ tasks (T) for
 parallelization. We focus on the worker's cache (circled), which facilitates
-communication and segment exchange between the worker and its tasks.
+segment exchange.
 \label{fig:parsplice}}
@@
 
-\begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\
-
-  \caption{The keyspace is small but must satisfy many reads as workers
-  calculate segments. Memory usage scales linearly, so it is likely that we will
-  need more than one node to manage segment coordinates when we scale the system
-  or jobs up. \(\Delta_1\) is the growth rate, an application specific parameter
-  described in Section~\S\ref{sec:delta}.
-
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[b]{.32\linewidth}
+    \centering
+    \includegraphics[width=1\textwidth]{figures/methodology-keyspace.png}\\
+    \caption{Keyspace size and shape.
     \label{fig:methodology-keyspace}}
-\end{figure}
+  \end{subfigure}
+  \begin{subfigure}[b]{.32\linewidth}
+    \includegraphics[width=0.975\textwidth]{figures/motivation-regimes.png}\\
+    \caption{Key accesses over time.
+    \label{fig:motivation-regimes}}
+  \end{subfigure}
+  \begin{subfigure}[b]{.32\linewidth}
+    \includegraphics[width=1.05\textwidth]{figures/methodology-keys.png}\\
+    \caption{Redundant key accesses.
+    \label{fig:methodology-keys}}
+  \end{subfigure}
+  \caption{Keyspace analysis for ParSplice.
+    (a) shows that the keyspace is small but must satisfy many reads. Memory
+    usage scales linearly, so we will eventually need more than one node to manage
+    segment coordinates for larger jobs. \(\Delta_1\) is the growth rate (see
+    Section~\S\ref{sec:delta}).
+    (b) shows that key accesses start with many reads to a small set of
+    keys and progresses to less reads to a larger set of keys.  The line shows the
+    rate that EOM minima values are retrieved from the key-value store (\(y1\));
+    points along the bottom show the unique keys accessed in a 1
+    second window (\(y2\)). Despite having different
+    \(\Delta\)s, the structure of key accesses are similar.
+    (c) shows that over time, tasks start accessing a larger set of keys resulting in
+    some keys being more popular than others.  Despite different
+    \(\Delta\)s, in both runs some keys are \(5\times\) more popular than others.
+  }
+\end{figure*}
 
@@
 
-\begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/motivation-regimes.png}\\
-  \caption{Key activity for ParSplice starts with many reads to a small set of
-keys and progresses to less reads to a larger set of keys.  The line shows the
-rate that EOM minima values are retrieved from the key-value store (\(y1\)
-axis) and the points along the bottom show the number of unique keys accessed in a 1
-second sliding window (\(y2\) axis). Despite having different growth rates
-(\(\Delta\)), the structure and behavior of the key activities are similar.
-\label{fig:motivation-regimes}}
-\end{figure}
-
 \subsubsection{\underline{Initial conditions influence key activity}}
@@
 
-\begin{figure}[t]
-  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keys.png}\\
-  \caption{Over time, tasks start to access a larger set of keys resulting in
-some keys being more popular than others.  Despite different growth rates
-(\(\Delta\)), the spatial locality of key accesses is similar between the two
-runs.  ({\it e.g.}, some keys are still read 5 times as many times others).
-\label{fig:methodology-keys}}
-\end{figure}
-
 \subsubsection{\underline{Entropy increases over time}} The reads per second in

### 2018-02-21 paper/sections/4_dynamic-policies.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-11-28 paper/sections/4_dynamic-policies.tex
@@
 % technical details
-We implement a basic LRU cache using a ``when" policy of:
+We implement an LRU cache using a ``when" policy of:\\
+\noindent\texttt{server[whoami]['cachesize']>n}\\
+and a ``how much" policy of:\\
+\noindent\texttt{servers[whoami]['cachesize']-n}.
 
-\begin{figure}[h]
-\footnotesize
-\begin{minted}{lua}
-server[whoami]['cachesize']>n
-\end{minted}
-\end{figure}
-
-and a ``how much" policy of:
-
-\begin{figure}[h]
-\footnotesize
-\begin{minted}{lua}
-servers[whoami]['cachesize']-n
-\end{minted}
-\end{figure}
-
-The results for different cache sizes for a growth rate of \(\Delta_1\) over a
-2.5 hour run across 256 workers is shown in
-Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
-unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
+The results for different cache sizes
+for a growth rate of \(\Delta_1\) over a 2.5 hour run across 256 workers is
+shown in Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance
+of unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
 utilization is measured with memory footprint of just the cache (\(y2\) axis).
@@
 
-\begin{figure}[t]
-\noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
-\caption{Policy performance/utilization shows the trade-offs of different sized
-caches (\(x\) axis).  ``None" is ParSplice unmodified, ``Fixed Sized Cache"
-evicts keys using LRU, and ``Multi-Policy Cache" switches to fixed sized cache
-after absorbing the workload's initial burstiness. This parameter sweep
-identifies the ``Multi-Policy Cache" of 1K keys as the best solution but this
-only works for this system setup and initial configurations.
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[t]{0.32\textwidth}
+    \centering
+    \includegraphics[width=0.97\textwidth]{figures/methodology-tradeoff.png}\\
+    \caption{Policy with system knowledge (\S\ref{sec:arch-specific}).
     \label{fig:methodology-tradeoff}}
-\end{figure}
+  \end{subfigure}
+  \begin{subfigure}[t]{0.32\textwidth}
+    \centering
+    \includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\
+    \caption{Key accesses (application knowledge).
+    \label{fig:keyspace-zoomed}}
+  \end{subfigure}
+  \begin{subfigure}[t]{0.32\textwidth}
+    \centering
+    \includegraphics[width=0.97\textwidth]{figures/dscache-vs-none.png}\\
+    \caption{Policy with application knowledge (\S\ref{sec:dom-specific}).
+    \label{fig:dscache-vs-none}}
+  \end{subfigure}
+  \caption{Cache management policies tested over the Mantle policy engine.
+    (a) shows performance/utilization trade-offs of different cache
+    sizes (\(x\) axis).  ``None" is unmodified, ``Fixed Sized Cache" evicts keys
+    using LRU, and ``Multi-Policy Cache" switches to a fixed sized cache after
+    absorbing the workload's initial burstiness.  ``Multi-Policy Cache" with 1K
+    keys is the best policy but only works for our hardware/configurations.
+    (b) shows key accesses for a 4 hour run. Detecting these groups of accesses
+    leads to a more accurate cache management strategy.
+    (c) shows performance/utilization for the dynamically sized cache
+    (DSCache) policy. DSCache adjusts to different initial configurations
+    (\(\Delta\)'s) and saves 3\(\times\) as much memory in the best case.
+  }
+\end{figure*}
 
@@
 \label{sec:dom-specific}
-
-\begin{figure*}[t!]
-    \begin{subfigure}[t]{0.48\textwidth}
-        \centering
-	\includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\
-
-\caption{Key activity for a 4 hour run shows groups of accesses to the same
-subset of keys. Detecting these access patterns leads to a more accurate cache
-management strategy, which is discussed in Section~\S\ref{sec:regime-detection}
-and the results are in Figure~\ref{fig:dscache-vs-none}.
-\label{fig:keyspace-zoomed}}
-    \end{subfigure}%
-    ~ 
-    \begin{subfigure}[t]{0.48\textwidth}
-        \centering
-        \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
-
-	\caption{The performance/utilization for the dynamically sized cache (DSCache)
-policy. With negligible performance degradation, DSCache adjusts to different initial configurations
-(\(\Delta\)'s) and saves 3\(\times\) as much memory in the best case.
-\label{fig:dscache-vs-none}}
-
-    \end{subfigure}%
-    \caption{Different cache management policies tested over the Mantle policy engine.}
-\end{figure*}
 
@@
 curves in Figure~\ref{fig:memory-vs-time}, where the behavior resembles the key
-access patterns in Figure~\ref{fig:keyspace-zoomed}\footnote{The memory usage
+access patterns in Figure~\ref{fig:keyspace-zoomed}. The memory usage
 is not {\it exactly} the same because these are two different runs;
 Figure~\ref{fig:keyspace-zoomed} has key activity tracing turned on, which
-reduces performance.}.  We also show a \(\Delta_2\) growth rate to demonstrate
+reduces performance.  We also show a \(\Delta_2\) growth rate to demonstrate
 the dynamic policy's ability to adjust to a different set of initial

### 2018-02-21 paper/sections/5_file-systems.tex
paper: address reviewer feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \subsection{Using Cache Management Policies for Load Balancing}
-
+\label{sec:using-cache-management-policies-for-load-balancing}
 \begin{figure}[t]
@@
 \caption{File system metadata reads for a Lustre trace collected at LANL. The
-vertical lines are the access patterns detected by the ParSplice cache
-management policy from Section~\S\ref{sec:dom-specific}. A file system that
-load balances metadata across a cluster of servers could use the same pattern
-detection to make migration decisions, such as avoiding migration when the
-workload is accessing the same subset of keys or keeping groups of accesses
-local to a server.  \label{fig:trace-atime}}
+vertical lines are the results of running the pattern detection algorithm
+used for cache management in Section~\S\ref{sec:dom-specific}. File system
+cases are in
+Section~\S\ref{sec:using-cache-management-policies-for-load-balancing}.
+\label{fig:trace-atime}}
 \end{figure}
+%A file system that
+%load balances metadata across a cluster of servers could use the same pattern
+%detection to make migration decisions, such as avoiding migration when the
+%workload is accessing the same subset of keys or keeping groups of accesses
+%local to a server.
 
@@
 designed for cache management in Section~\S\ref{sec:regime-detection} with
-slight modifications based on our knowledge of file systems\footnote{We
+slight modifications based on our knowledge of file systems. We
 filtered out requests for key IDs less than 2000, as these are most likely path
-traversal requests to higher parts of the namespace.}. The vertical lines in
+traversal requests to higher parts of the namespace. The vertical lines in
 Figure~\ref{fig:trace-atime} are the groups of accesses identified by the
@@
 and~\S\ref{sec:dom-specific}. For example, Ceph administrators can use the
-policies in ParSplice to automatically size and manage cache
-tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
-caching on object storage devices, or in the distributed block
-devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
-Integration with Mantle would be straightforward as it is merged into Ceph's
-mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
+policies in ParSplice to automatically size and manage cache tiers, caching on
+object storage devices, or in the distributed block device.  Integration with
+Mantle would be straightforward as it is merged into Ceph's mainline and the
 three caching subsystems mentioned above already maintain key access traces.
@@
 IndexFS~\cite{ren:sc2014-indexfs} or the capabilities from
-CephFS\footnote{http://docs.ceph.com/docs/master/cephfs/capabilities/}.
+CephFS.
 

### 2018-02-23 paper/sections/1_introduction.tex
paper: parsplice corrections from danny
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-02-20 paper/sections/1_introduction.tex
@@
 class of HPC applications with similar working set behaviors that extensively
-use software-based caches. ParSplice uses a hierarchy of caches and a single
+use software-based caches. It uses a hierarchy of caches and a single
 persistent key-value store to store the atomic coordinates corresponding

### 2018-02-23 paper/sections/2_parsplice-keyspace-analysis.tex
paper: parsplice corrections from danny
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 along the time axis.  The \(\Delta_1\) growth rate adds atoms every 100K
-microseconds while the \(\Delta_2\) growth rate adds atoms every 1 million
-microseconds. So \(\Delta_2\) has a smaller growth rate resulting in hotter
+picoseconds while the \(\Delta_2\) growth rate adds atoms every 1 million
+picoseconds. So \(\Delta_2\) has a smaller growth rate resulting in hotter
 keys and a smaller keyspace.  Values smaller than \(\Delta_2\)'s growth rate or
 a temperature of 400 degrees result in very little database activity because
-state transitions take too long. Similarly, values larger than \(\Delta_1\)'s
-growth rate or a temperature of 4000 degrees result in an equally meaningless
-simulation as transitions are unrealistic.
+state transitions take very long. Similarly, growth rate values larger than \(\Delta_1\)'s
+lead to far out-of-equilibrium growth, while temperatures in excess of 800 degrees
+result in the melting of the nanoparticle, which is not relevant in this application.
 
 This figure demonstrates that small changes to \(\Delta\) can have a strong
-effect on the timing and frequency with which new EOM minima are discovered and
+effect on the timing and frequency with which new minima are discovered and
 referenced.  Trends also exist for temperature and number of workers but are

### 2018-02-23 paper/sections/4_dynamic-policies.tex
paper: fix title and minor typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: parsplice corrections from danny
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
     (DSCache) policy. DSCache adjusts to different initial configurations
-    (\(\Delta\)'s) and saves 3\(\times\) as much memory in the best case.
+    (\(\Delta\)s) and saves 3\(\times\) as much memory in the best case.
   }

### 2018-02-23 paper/sections/1_introduction.tex
paper: parsplice corrections from danny
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
+\begin{abstract}
+
+Our analysis of the key-value activity generated by the ParSplice molecular
+dynamics simulation demonstrates the need for more complex cache management
+strategies. Baseline measurements show clear key access patterns and hot spots
+that offer significant opportunity for optimization. We use the data management
+language and policy engine from the Mantle system to dynamically explore a
+variety of techniques, ranging from basic algorithms and heuristics to
+statistical models, calculus, and machine learning. While Mantle was originally
+designed for distributed file systems, we show how the collection of
+abstractions effectively decomposes the problem into manageable policies for a
+different application and storage system.  Our exploration of this space results in a
+dynamically sized cache policy that does not sacrifice any performance while
+using 32-66\% less memory than the default ParSplice configuration.
+
+\end{abstract}
+
 \section{Introduction}
@@
 memory without sacrificing performance and works for a variety of inputs.
-Solutions labeled ``no performance degradation" are in comparison to the
-baseline of a cache of unlimited size.
 \label{fig:cache-management}}
@@
 without any performance degradation, works for a variety of initial conditions,
-and could generalize to similar applications. Triggering key eviction at a
-certain memory pressure ({\it e.g.}, 3\%) requires {\it a priori} system
-knowledge while our approach saves the most memory because we model the
-behavior of keyspace accesses.
+and could generalize to similar applications.
 
@@
 and~\S\ref{sec:dom-specific}, where we try a range of policies from different
-disciplines. We analyze why our early policies fail for our HPC hardware and
-software and show how we iterated to the final solution.  More importantly, in
-Section~\S\ref{sec:scope}, we conclude that the collection of policies we
-design for cache management in ParSplice are very similar to the policies used
-to load balance metadata in the Ceph file system
-(CephFS~\cite{weil:osdi2006-ceph}) suggesting that there is potential for
-automatically adapting and generating policies dynamically. This paper does not
-thoroughly study cache management policies for file systems, but our findings
-for this specific application and related storage systems work suggests that
-this approach is a strong avenue for future work.
+disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
+that the collection of policies we design for cache management in ParSplice are
+very similar to the policies used to load balance metadata in the Ceph file
+system (CephFS~\cite{weil:osdi2006-ceph}) suggesting that there is potential
+for automatically adapting and generating policies dynamically.
 

### 2018-02-23 paper/sections/2_parsplice-keyspace-analysis.tex
paper: parsplice corrections from danny
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-02-20 paper/sections/2_parsplice-keyspace-analysis.tex
@@
   \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\
-
   \caption{The keyspace is small but must satisfy many reads as workers
-  calculate segments. Memory usage scales linearly, so it is likely that we will
-  need more than one node to manage segment coordinates when we scale the system
-  or jobs up. \(\Delta_1\) is the growth rate, an application specific parameter
-  described in Section~\S\ref{sec:delta}.
-
+  calculate segments. Memory usage scales linearly, so it is likely that we will need
+  more than one node to manage segment coordinates when we scale the system or jobs up.
   \label{fig:methodology-keyspace}}
@@
 \subsubsection{\underline{Initial conditions influence key activity}}
-\label{sec:delta}
 Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value

### 2018-02-23 paper/sections/4_dynamic-policies.tex
paper: parsplice corrections from danny
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-02-20 paper/sections/4_dynamic-policies.tex
@@
 We implement a basic LRU cache using a ``when" policy of:
-\texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
-\texttt{servers[whoami]['cachesize']-n}.  The results for different cache sizes
-for a growth rate of \(\Delta_1\) over a 2.5 hour run across 256 workers is
-shown in Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance
-of unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
+
+\begin{figure}[h]
+\footnotesize
+\begin{minted}{lua}
+server[whoami]['cachesize']>n
+\end{minted}
+\end{figure}
+
+and a ``how much" policy of:
+
+\begin{figure}[h]
+\footnotesize
+\begin{minted}{lua}
+servers[whoami]['cachesize']-n
+\end{minted}
+\end{figure}
+
+The results for different cache sizes for a growth rate of \(\Delta_1\) over a
+2.5 hour run across 256 workers is shown in
+Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
+unmodified ParSplice  measured in trajectory duration (\(y1\) axis) and
 utilization is measured with memory footprint of just the cache (\(y2\) axis).
@@
         \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
+
 	\caption{The performance/utilization for the dynamically sized cache (DSCache)
 policy. With negligible performance degradation, DSCache adjusts to different initial configurations
-(\(\Delta\)s) and saves 3\(\times\) as much memory in the best case.
+(\(\Delta\)'s) and saves 3\(\times\) as much memory in the best case.
 \label{fig:dscache-vs-none}}
+
     \end{subfigure}%

### 2018-02-23 paper/sections/1_introduction.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-02-23 paper/sections/1_introduction.tex
@@
 class of HPC applications with similar working set behaviors that extensively
-use software-based caches. It uses a hierarchy of caches and a single
+use software-based caches. ParSplice uses a hierarchy of caches and a single
 persistent key-value store to store the atomic coordinates corresponding

### 2018-02-23 paper/sections/2_parsplice-keyspace-analysis.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-02-23 paper/sections/2_parsplice-keyspace-analysis.tex
@@
     keys and progresses to less reads to a larger set of keys.  The line shows the
-    rate that EOM minima values are retrieved from the key-value store (\(y1\));
+    rate that minima values are retrieved from the key-value store (\(y1\));
     points along the bottom show the unique keys accessed in a 1
@@
 a temperature of 400 degrees result in very little database activity because
-state transitions take very long. Similarly, growth rate values larger than \(\Delta_1\)'s
+state transitions take very long. Similarly, growth rate values larger than \(\Delta_1\)
 lead to far out-of-equilibrium growth, while temperatures in excess of 800 degrees

### 2018-02-23 paper/sections/5_file-systems.tex
paper: fix title and minor typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: add acknowledgements
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 used for cache management in Section~\S\ref{sec:dom-specific}. File system
-cases are in
+use cases are in
 Section~\S\ref{sec:using-cache-management-policies-for-load-balancing}.
@@
 
-\begin{itemize}
-
-  \item QoS: when to move clients, where to move clients, how much of the
+\textbf{QoS}: when to move clients, where to move clients, how much of the
 reservation to move. We could use Mantle to implement something like the
@@
 
-  \item Scheduling: when to yield computation cycles to another process, how
+\textbf{Scheduling}: when to yield computation cycles to another process, how
 much of a resource to allocate. We could use Mantle to implement the
@@
 
-  \item Batching: how many operations to group together, when to send large
+\textbf{Batching}: how many operations to group together, when to send large
 batches of updates. We could use Mantle to implement pathname leases from
@@
 
-  \item Prefetching: how much to prefetch, how to select data.  We could use
+\textbf{Prefetching}: how much to prefetch, how to select data.  We could use
 Mantle to implement forward/backward/stride detection algorithms for
@@
 algorithms for adaptive I/O prefetching from~\cite{tran:sc01-arima}.
-
-\end{itemize}

### 2018-02-23 paper/sections/6_conclusion.tex
paper: camera-ready format; squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: add acknowledgements
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 CephFS~\cite{weil:sc2004-dyn-metadata, weil:osdi2006-ceph} employs both
-techniques to a lesser extent; directories can be replicated or sharded but the
-caching and replication policies are controlled with tunable parameters.  These
-systems still need to be tuned by hand with {\it ad-hoc} policies designed for
-specific applications.  Setting policies for migrations is arguably more
-difficult than adding the migration mechanisms themselves.  For example,
-IndexFS/CephFS use the GIGA+~\cite{patil:fast2011-giga} technique for
-partitioning directories at a \emph{predefined} threshold. Mantle makes headway
+techniques to a lesser extent.  These systems still need to be tuned by hand
+with {\it ad-hoc} policies designed for specific applications.  For example,
+IndexFS and CephFS partition directories at \emph{predefined} thresholds,
+which are difficult to define at the beginning of the job. Mantle makes headway
 in this space by providing a framework for exploring these policies, but does
@@
 sweeps, our solution saves memory without sacrificing performance for a variety
-of initial conditions, including the scale, duration, configuration, and
-hardware of the simulation. More importantly, rather than attempting to
+of initial conditions. More importantly, rather than attempting to
 construct a single, complex policy that works for a variety of scenarios, we
@@
 collection of general data management policies that can be used across
-applications and storage systems.  The value of such a collection eases the
+applications and storage systems.  This eases the
 burden of policy development and paves the way for automated solutions such as
 (1) adaptable policies that switch to new strategies when the current strategy
-behaves poorly ({\it e.g.}, thrashing, making no progress, etc.), and (2)
+behaves poorly ({\it e.g.}, thrashing), and (2)
 policy generation, where new policies are constructed by examining the
-collection of existing policies.  Ultimately, we hope that this automation
+collection of existing policies.  We hope that this automation
 enables control of policies by machines instead of administrators.

### 2018-02-26 paper/sections/1_introduction.tex
paper: minor typos, add acks, shorten to 10pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-02-23 paper/sections/1_introduction.tex
@@
 the number of remote procedure calls and improves the performance of
-create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
-patil:fast2011-giga+, weil:sc2004-dyn-metadata}. But the policies for what data
+create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata}. But the policies for what data
 to evict and when to evict are specific to the application's behavior and the

### 2018-02-26 paper/sections/2_parsplice-keyspace-analysis.tex
paper: minor typos, add acks, shorten to 10pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
     keys and progresses to less reads to a larger set of keys.  The line shows the
-    rate that minima values are retrieved from the key-value store (\(y1\));
+    rate that minima values are retrieved from the key-value store (\(y1\) axis);
     points along the bottom show the unique keys accessed in a 1
-    second window (\(y2\)). Despite having different
+    second window (\(y2\) axis). Despite having different
     \(\Delta\)s, the structure of key accesses are similar.

### 2018-02-26 paper/sections/4_dynamic-policies.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: minor typos, add acks, shorten to 10pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-02-23 paper/sections/4_dynamic-policies.tex
@@
         \includegraphics[width=0.5\textwidth]{figures/memory-vs-time.png}\\
-	\caption{Memory utilization for ``No Cache Management" (unlimited cache
+	\caption{Memory utilization for ``Default Policy" (unlimited cache
 growth), ``Multi-Policy" (absorbs initial burstiness of workload), and
 ``Dynamic Policy" (sizes cache according to key access patterns). The dynamic
-policies saves the most memory without sacrificing performance.
+policies save the most memory without sacrificing performance.
 \label{fig:memory-vs-time}}

### 2018-02-26 paper/sections/6_conclusion.tex
paper: minor typos, add acks, shorten to 10pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-Key-value storage organizations for scientific applications is a field gaining
-rapid interest. In particular, the analysis of the ParSplice keyspace and the
+The analysis of the ParSplice keyspace and the
 development of an appropriate scheme for load balancing is a direct response to
